### 2019-02-06 03b_R-data.Rmd
Added package data summary

---
title: "Data in R packages"
author: "Karthik"
date: "2/6/2019"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
suppressPackageStartupMessages(library(tidyverse))
df <- suppressMessages(read_csv("package_files_with_metadata.csv"))
data_folder <-  df %>% filter(location == "data") %>% pull(package) %>% unique %>% length
extdata_folder   <- df %>% filter(location == "extdata") %>% pull(package) %>% unique %>% length
packages_with_data <- df %>% filter(grepl("data|extdata", location)) %>% pull(package) %>% unique %>% length
unique_packages <- df %>% pull(package) %>% unique %>% length

#  Do some summaries
only_data_extdata <- df %>% filter(grepl("data|extdata", location))
types <- only_data_extdata %>% count(extension) %>% arrange(desc(n)) %>% mutate(percentage = n/nrow(only_data_extdata))

percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}

```

Of the 15539 packages on Central R Archive Network (CRAN),  `r packages_with_data` contain datasets either as binary data (`r data_folder` packages) or as external datasets (`r extdata_folder`). Binary files comprise a bulk of the data shipped in the `data` folder (`r types[1:2, ]$percentage %>% sum %>% percent`) with other plain text formats such as txt, csv, dat, json comprising less than one percent of data shipped in packages.

### 2019-02-09 03b_R-data.Rmd

@@
 knitr::opts_chunk$set(echo = TRUE)
+# TODO
+# Can the raw numbers have the %s provided with them? Or perhaps we can refer to a table of results for this as well.
+# A bit meta, but should we include metadata, data dictionary, README, for this data used in the paper?
+# scales::percent(0.1)
+# ------------------------------------------------------------------------------------------------------
 ```

### 2019-05-09 paper.Rmd
merge everything into one rmd

---
title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility
output: 
  bookdown::pdf_book:
    base_format: rmarkdown::pdf_document
bibliography: references.bib
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE,
                      out.width = "100%")
```

> “Data! data! data!" he cried impatiently. "I can't make bricks without clay.” - Sherlock Holmes (PBS Series, 2010-2017)

# Introduction

Data are a fundamental currency upon which scientific discoveries are made. Without access to good data, it becomes extremely difficult, if not impossible, to advance science. Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors (citation).  Sharing datasets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work. But there is also a component of benefit to the broader scientific community. This primarily comes in the form of potential for reuse in other contexts along use for training and teaching (@McKiernan2016). Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing doesn't put the authors at a competitive disadvantage, sharing data will always have a net positive benefit. First and foremost, sharing data along with other artifacts can make it easier for a reader to independently reproduce the results, thereby increasing transparency and trust in the work. The act of easy data sharing can also improves model training, as many different models can be tried and tested on latest data sources, closing the loop on research and application of statistical techniques. Existing datasets can be combined or linked with new or existing data, fostering the development and synthesis of new ideas and research areas. The biggest of these benefits is the overall increase in reproducibility.

For nearly two decades, researchers who work in areas related to computational science have pushed for better standards to verify scientific claims, especially in areas where a full replication of the study would be prohibitively expensive. To meet these minimal standards, there must be easy access to the data, models, and code. Among the different areas with a computational bent, the bioinformatics community in particular has a strong culture around open source [@Gentleman2004], and has made releasing code and associated software a recognized mainstream activity. Many journals in these fields have also pushed authors to submit code (model implementations) alongside their papers, with a few journals going as far as providing a "reproducibility review" [@Peng2011].

**In this paper we focus on the practical side of sharing data for the purpose of reproducibility.** Our goal is to describe various methods in which an analyst can share their data with minimal friction. We steer clear of theoretical ideas such as the FAIR data principles [@Wilkinson2016] since they still do not help a researcher share their data. We also skip the discussion around citation and credit because data citations are still poorly tracked and there is no standardized metric or a h-index equivalent for data as of this writing.

For a piece of computational research to be minimally reproducible, it requires three distinct elements:

1. Code  
2. Computing environment  
3. Data

The first two of these challenges have largely been solved.

Although code sharing in science had a rocky start [@Barnes2010], more and more code writing by scientists is being shared, partly due to the rapid increase in training opportunities made available by organizations like The Carpentries, combined with the rapid adoption of Github by scientists [@Ram2013]. The bigger driver for this may also be connected to the rise in popularity of data science as a discipline distinct from statistics [@Donoho2017]. This rapid growth in data science has largely been fueled by easy access to open source software tools. Programming languages such as Python, R and Julia help scientists implement and share new methods to work with data. Each of these languages is supported by thriving communities of researchers and software developers who contribute many of the building blocks that make them popular. As of this writing, Python, Julia, and R have  167k packages (https://pypi.org/), ~ 14k packages (https://cran.r-project.org/) and ~2k  packages (https://pkg.julialang.org/) respectively. These packages form the building blocks of a data scientists daily workflow.

A typical data scientist loads a dozen or two of these open source libraries at the top of a notebook and then rely on existing routines to rapidly read, transform, visualize, and model data. Each of these packages individually depend on a complex web of other packages, building upon existing work rather than reimplementing everything from scratch. Working from script and a list of such dependencies, a data analyst can easily install all the necessary tools in any local or remote environment and reproduce the computation. When new functionality is developed, it is packaged into a separate entity and added to a language's package manager.

The computing environment is also easily captured with modern tools such as Docker [@Boettiger2015; @Jupyter2018]. Modern tools such as Binder (citation) can parse Docker files and dependency trees to provide on demand, live notebooks in R and Python that a reader can immediately execute in the browser without dealing with the challenges of local installation. This makes it simple to load a specific environment to run any analysis. Code is handled by version control with tools like Git and GitHub [@git; @github], paired with archiving such as Zenodo provide access to code (particularly model implementations)[@zenodo]. All the necessary software is available from various package managers (and their numerous geographic mirrors and archives) making it easy to install any version of a software package. However, the biggest challenge, even today, remains easy and repeatable access to data in a data analysis.

Although there are numerous public and private data repositories, none  of them function as package managers. Datasets are also far more diverse in complexity, size, and formats, making them particularly challenging to standardize or easily "install" where the code is running. As a result, data used in an analysis is often read from various locations (local, network, or cloud), various formats, varying levels of tidiness [@Wickham2014]. There is also the overhead associated with data publishing, the act of archiving data in a repository that also mints permanent identifiers, that are not required of all projects due to the effort and time involved. It is worth drawing the distinction between data sharing (making the data available with little effort) and data publishing (archiving the data in a long-term repository, with our without curated metadata).

What aspects of data make them particularly challenging to share from a reproducibility perspective? This is the question we tackle in this paper.  While there are several papers that serve as best-practice guides for formatting data and getting them ready for sharing, the aims of this paper are a bit different. Our aim to address the issue of data in the context of reproducibility in data science workflows. In particular we discuss the various tradeoffs one has to consider when documenting and sharing data, when it is worth publishing and how this would change depending on the use case, perceived impact, and potential audience.

We discuss how to share and/or publish data and cover various tradeoffs when deciding how much to do. We skip detailed discussions of the minutiae of data preparation (covered elsewhere in various articles), licenses (there are not many options when it comes to data), or citation (as of this writing, data citation is still in its infancy). We also analyze the state of data contained inside software packages, shared and made available as part of modern data analysis. How often are data shipped inside software packages? How does data format and size impact availability?

```{r fig-data-sharing-workflow, fig.cap = "The steps required compared to effort required for data sharing for reproducibility."}
knitr::include_graphics("figures/data_sharing_workflow.png")
```

# Challenges in documenting your dataset

Providing a well documented dataset alongside research requires meeting several challenges. This includes understanding all the pieces and features of documentation such as metadata, licenses, and raw data; where to literally share the data; how to ensure long term provenance of the data.

One challenge is that better documentation requires more time, and that higher impact data needs better documentation.

Sharing data with colleagues: binary data is quick to share, but not as good as something like a simpler format like CSV (ask Laura DeCicco re this)

Providing the right amount of documentation for your data is challenging. Too much is overwhelming; too little and you don't know anything. To illustrate this, imagine two datasets. The first is contained in a folder with dozens of data files and several papers on the topic. The second contains only a single csv file. The first is too much; understanding it requires a high time cost. The second is too little; we don't know anything about what the data is or why it was collected. The challenge is to provide the right level of detail, somewhere between these two extremes.

It takes more effort to prepare data to make it easy to understand and share. The challenge here is understanding how much time to spend to make it easier to understand before the gains in understanding are outweighed by the time taken to improve them.

There is also a challenge in understanding the impact of data. Data with high potential for impact should be made easier to understand and share. This can be seen in the research created from the 50 hectare plot.

There are many features to consider when documenting data, such as a README file with human-readable descriptions of data and variables, data licenses, whether to include raw data, data cleaning scripts, and machine readable metadata. Together these all help make the data easier to understand. These features are not clearly described in the scientific literature in the context of including them alongside research.

To frame discussion around the challenges of data documenting, we can think of two features: "Effort to prepare", and "Ease of understanding". We can plot these on two axes and place datasets along each to indicate the ease of understanding and preparation (see Figure \@ref(fig:fig-effort-underrstanding)).

```{r fig-effort-underrstanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to udnerstand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
library(ggplot2)

# There are two Axis:
#   X axis: <- Robustness (brittle - robust)
#   Y axis: <- Difficulty (easy - hard)

library(tibble)

dat_prep <- tibble(prep = c(15,20,25,95, 96),
               understanding = c(15,20,95,95,96),
                  label = c("Data Dump",
                            "Most studies",
                            "Sensor",
                            "50 hectare",
                            "RCT"
                            ))

library(ggrepel)

ggplot(dat_prep,
       aes(x = prep,
           y = understanding,
           label = label)) +
  geom_point() +
  geom_label_repel(
    segment.size = 0.2,
    segment.color = "grey50") +
  theme_minimal() +
  labs(x = "Effort to Prepare",
       y = "Ease of Understanding") +
  xlim(0,100) + 
  ylim(0, 100)

```

## Challenges

**Challenges: higher potential impact means more time**

To improve data impact it needs more time to prepare so it is easier to understand. Accordingly, data with higher potential for impact and reuse should be made as easy to understand as possible. Therefore data with high potential impact requires more time and effort to prepare. For example, data from medical randomized control trials (RCTs) on cancer treatment can have high impact, so requires a lot of time and effort document. Comparatively, a small survey on simple demographics can have low impact. So it might not have a worthwhile tradeoff for ease of understanding and ease of preparation. This means effort put into preparing data documentation for a small survey to be made easy to use should be kept relatively simple and not overcomplicated.

Tradeoffs may change over time as the cost of making them easy to prepare and understand change. If it becomes technologically cheaper and easier to document data more rigorously in the same time as it takes to do poorly now, then there is no reason why more people cannot do this, even when the benefits are not immediately apparent. So if something might appear to have low impact, it should still be made easy to understand, if it doesn't take long to prepare that documentation.

Data that is easy to use should avoid being in a proprietary format that requires special software to read. For example, a CSV should be preferred over proprietary data formats, such as SPSS or SAS custom data types.

## Understanding tradeoffs

To help explain these descriptive qualities of the data sharing, we discuss three datasets and their relationship of **Effort to prepare**, and **Ease of understanding** in section 5.

One of the easiest solutions to making data available alongside an analysis is to include it in the same version controlled repository as the analysis scripts. This works well with the data file sizes are in the order of bytes to megabytes. When a reader clones the repository, the scripts and notebooks will be able to read all data files from relative paths with no further modifications to the code itself. Assuming the data would be useful beyond the context of the analysis, it would be beneficial to also archive it in a long-term repository.

# Releasing data in R

One low cost and easy way to distribute data alongside compute would be to package the datasets as a data only package or as part of something larger where the methods are being implemented. The R ecosystem has many noteworthy examples of data-only packages. One exemplar.  One exemplar is the `nycflights13` package by Hadley Wickham (https://cran.r-project.org/web/packages/nycflights13/index.html). This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes. The package not only provides ready to use binary data but also ships raw data (in a folder called `data-raw`) along with scripts used to clean them. The package was originally created as a way to provide example data to teach tidy data principles and serves as a great model for how one can publish a data package in R.

A more common use case is to include data as part of a regular package where analytical methods are being implemented and shared. This serves the dual purpose of exposing methods and data together, making it extremely easy for a researcher to simply install and load the library at the top of their script. CRAN's distributed network (along with the global content distribution network maintained by Rstudio) ensure that the package is quickly accessible to anyone in the R ecosystem. A second critical advantage in this approach is that one could also benefit from R's package documentation syntax to generate useful metadata for fast reference. This approach can also easily be adapted to other languages. Python for example, is far more flexible with respect to including arbitrary files as part of a package.

**Other benefits of packaging data in R**

1. Packaging data can be a very powerful pedagogical tool to help researchers understand how to transform data and prepare it for further analysis. To do so, one can package raw data alongside scripts. Long form documentation such as Vignettes can provide further detailed discussion on the process. Advanced users can simply skip the raw data and scripts and proceed directly to the fast binary data which can hold a lot of data when heavily compressed.
2. When the primary motivation for shipping data is to illustrate visualization techniques or to run examples, one can skip the raw data and processing scripts and only include the binary data. As long as the total package size doesn't exceed 5 megabytes, it would be acceptable as part of CRAN. For cases when this size is hard to maintain, CRAN recommends data-only packages that will be rarely updated. For a detailed discussion on this issue and alternative approaches, see [@Brooke_Anderson2017]

One major disadvantage of packaging data inside R is that it makes the data availability very language centric. Non R users are unlikely to download and export data out of a package. This is why we recommend, as a rule, that researchers also archive data in a long-term data repository. These include domain specific repositories (see SECTION) or more general purpose ones such as Zenodo or figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.

Of the 15539 packages on Central R Archive Network (CRAN), 6278 contain
datasets either as binary data (5903 packages) or as external datasets
(766). Binary files comprise a bulk of the data shipped in the `data`
folder (68.06%) with other plain text formats such as txt, csv, dat,
json comprising less than one percent of data shipped in packages.

# Data Sharing in Julia and Python

Sharing data in a programming language happens to some extent in other languages. We discuss how Julia and Python share and distribute data in the following section.

# Julia

Julia was established in 2012, and is a high level language with great performance. Searching METADATA.jl [@METADATA-jl] for packages mentioning data revealed three main categories of data packages in Julia:

1. Database access [@data-access] for services like MySQL, Hive, and ODBC, similar to R’s DBI organisation [@R-DBI]
2. Accessing data online via APIs, such as accessing the weather source NOAA [@NOAA-jl]
3. Data is provided directly. For example, accessing Econometric datasets [@econ-data-jl], face datasets [@face-data-jl], Market data [@market-data-jl], and even R datasets [@r-data-jl]. It is telling that the datasets provided in R are useful, as there is even a Julia package, RDatasets that provides over 700 datasets from R packages. There are also packages for Accessing datasets used in vega [@vega-data-jl], for accessing common machine learning datasets [@ml-data-jl]. There are also packages for generating fake data [@fake-data-jl], for which there are at least two R packages that do this, charlatan [@charlatan-data-r], and [@wakefield-data-r].

# Python

Python is a popular general purpose language, used in engineering, statistics, and computer science. One language agnostic solution is to use the Open Knowledge Foundations Frictionless data spec. This is implemented in both R and Python.

Including data in a python package: https://python-packaging.readthedocs.io/en/latest/non-code-files.html
A user can typically include data (or any other arbitrary files) inside a python package by declaring it in the manifest.

# Dealing with medium to large data

Another common situation that researchers face is in dealing with data that fall somewhere between small (tabular data such as csvs that are in the order of bytes to megabytes to gigabytes that can be easily compressed) to large (what falls under the umbrella of `big data`). The happy medium is generally data that are too big to fit on memory of most standard machines, but can successfully fit on disk (https://ropensci.github.io/arkdb/articles/articles/noapi.html). In this case, users who do not have the support to maintain resource intensive database servers can instead rely on light serverless databases such as MonetDB or SQLite. These databases provide disk based storage and using language agnostic interfaces, a analyst can easily query these data in manageable chunks that don't overrun memory limitations. Using software packages such as arkdb (https://ropensci.github.io/arkdb/index.html) one could easily chunk large data from flat text files to these lite databases without running into memory limitations.

As for making these files available alongside compute, one ingenious but short-term solution is to use the GitHub release feature to attach such large database dumps. GitHub releases are designed to serve software releases and provide the opportunity to attach binary files. This mechanism can also be used to attach arbitrary files such as large data files, binary data, and database dumps as long as each file does not exceed 2gb. The R package `piggyback` allows for uploading and downloading such files to GitHub releases, which would make it easy for anyone to access data files wherever the script is being run. It's worth emphasizing again that this is a short-term solution that is dependent on GitHub maintaining this service.

# Publishing and repositories

It is worth distinguishing between sharing data and publishing data. One can share data in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify. Data can be shared in numerous ways including by placing it in a repository, packaging it with methods, or by using various free tiers of commercial services. However, one must publish data when appropriate.

There are a three common options for publishing data in research:

1. **Least Moving Parts**
2. **Domain Specific Venue**
3. **Long Term Archive**

We discuss each of these options and provide recommendations on how to publish data in these areas.

In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databaseses. This approach **gets the data somewhere rather than nowhere**. It's minimal features means it is simple to maintain. A downsides is that it does not scale to larger data. Self hosting the data is an option, but we discourage this, as it may succumb to bit rot.

In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data. For example, in astronomy the data is hosted at the SDSS (SDSS), Genetic data can be hosted at GenBank (@GenBank), and so on.  Some other places to put data include the open science framework: https://osf.io/, dataverse, https://dataverse.org/, and data.world, https://data.world/. Other datasets might be produced

The purpose, use, and origin of the data is an important component to consider. Data for research has a different domain compared to data collected by governments, or by journalists. Many governments or civil organisations are now making their own data available through a government website interface. Media and journalism groups are also making their data available either through places like GitHub (https://github.com/the-pudding/data), or organisations such as data.world, or may self host their data (https://data.fivethirtyeight.com/).

This is a good option when the data is appropriate for the domain. This is often decided by community standards. We recommend adhering to community standards for a location to publish your data, as this will likely improve data reuse. The guidelines suggested in this paper for sharing the data should be included.

A **Long Term Archive** is the best option to share the data. Long term archives provide information such as DOI (Digital Object Identifier) that make it easier to cite. Data can be placed in a long term archive and a DOI can be minted. This DOI can then be used in other venues, such as domain specific or even self hosted, and will ensure that the projects refer back appropriately.

If the dataset you are shipping has a research application, the most relevant home for it would be Zenodo, a research data repository. Launched in 2013 in a joint collaboration between openAIRE and CERN, Zenodo provides a free, archival location for any researcher to depositor their datasets. The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases. Zenodo is able to accommodate larger file sizes upon request.

**Principles for all three**

The data should contain information on metadata, data dictionaries, the README, and data used in analysis (ref section 2). No matter where data is submitted, there should ideally be a canonical data repo in one long term archive that links to others with a DOI

## Publishing data in a data repository

## Publishing data through a data journal

Data used in publications are often shared in the supplementary materials of articles, or served on repositories such as the Dryad Digital Repository (@dryad). Dryad makes data from scientific publications discoverable, reusable, and citable. It is well funded through grants from the NSF, European Commission.

To provide better context around the data used in research and better expose data for reuse, journals are now adding "data papers". These are specifically designed for publishing articles about the data, and sharing it. This benefits both researchers and readers. Researchers receive credit for data they have collected or produced. Readers get more context about the data.

Data papers are similar to research articles, they have titles, authors, affiliations, abstract, and references.  They generally require an explanation of why the data is useful to others, a direct link to the data, description of the design, materials, and methods. Other information on the subject area, data type, format, and related articles are usually required.

Whilst useful, these requirements do not tell the author how to actually structure the data and folders. Instead providing ideas on what they should include. This is a useful step towards improving data reuse, but it _lacks some minimal structure that allows a researcher to have a predictable way to access and interpret the data_.

Other journals operating in this space include journals like "data in brief", "Data", and "Nature Scientific Data". Guidelines for what is required in the content of the paper, and the required information along with the data (meta data, data format, etc.) vary by journal.

# What belongs in the minimal structure for researchers

There are 8 pieces of content to consider for data sharing:

1. README: A Human readable description of the data
2. Codebook: Human readable dictionary of data contents
3. License: How to use and share the data
4. Citation: How you want your data to be cited
5. Machine readable meta data
6. Raw data: The original/first data provided
7. Scripts: To clean raw data ready for analysis
8. Analysis ready data: Final data used in analysis

```
proj-name/
         |
         | - README.md
         | - Codebook.csv
         | - LICENSE.md
         | - metadata.json
         | - raw-data/
                     | - raw-data.csv
                     | - 01-read-tidy-raw-data.csv
         | - data/
                 | - final-data.csv
```

_Figure XX. Example directory layout and structure for a data repository._

Out of these sections, the minimal viable format are:

1. README
2. Codebook
3. License
4. Citation
5. Analysis ready data

What goes into each of these sections is now described.

# README: A Human readable description of the data

The README is often the first place people will go to learn more about data. It is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":

- **Who** collected it
- **What** is the data
- **When** was it collected
- **Where** was it collected
- **How** is was collected

The README should be placed in the top level of the project. It should be brief, and provide links to the other aforementioned sections. Saving a README with the extension `.md` file gives the author the formatting benefits of `markdown`, making it simple to insert links, tables, and make lists. In systems like GitHub, a README file is detected and rendered in nice HTML on the repository by default.

# Codebook: Human readable dictionary of data contents

Codebooks provide human readable description of the variables in the data and contain information on:

- **variable names**
- **variable labels**
- **variable codes**, and
- special values for **missing data**.

**Variable names** are short, descriptive names with no spaces or special characters. For example, "job_position", "faculty_level", and "years_at_company". **Variable labels** are longer descriptions of variables. For example "University Job Position", "University Faculty Position Level", and  "Number of Years Worked at University" [@McGill-codebook]. **Variable codes** apply to categorical (factor) variables, and are the values for their contents. For example, 0 = no, 1 = yes, and 0 = male and 1 = female. These should be consistent across similar variables to avoid problems where `0 = yes` for one variable, and `1 =yes` in another. Date variables should have consistent formatting. For example, all date information could be in format "YYYY-MM-DD", and this should not be mixed with "YYYY-DD-MM". **Missing data** are values that should have been observed, but were not. The code for missingness should be documented in the codebook, and should nominally be `NA`. If the reason for missingness is known it should be recorded. For example, censored data, patient drop out, or measurement error can have different values, such as "unknown" or -99 [@White2013; @Broman2017].

Below is an example data dictionary table taken from the [tidy tuesday repository on incarceration trends](https://github.com/rfordatascience/tidytuesday). This includes information on the variable the class (type) of the variable, and a longer description of the variable.

__Table 1. The prisoner summary data dictionary, with columns on the variable, it's class, and a short description of the contents of the variable.__

|variable        |class     | Description |
|:---------------|:---------| :---------- |
|year            |integer (date)   | Year |
|urbanicity      |character | County-type (urban, suburban, small/mid, rural) |
|pop_category    |character | Category for population - either race, gender, or Total |
|rate_per_100000 |double    | Rate within a category for prison population per 100,000 people |

Data dictionary tables should be placed in the README and presented as a table. Every data dictionary should also be provided in its raw form in the repository, for example, saved as a CSV (see Figure XX.).

# License: How to use and share the data

Data with a license clearly establishes rules on how everyone can modify, use, and share data. Without a license, these rules are unclear, and can lead to problems with attribution and citation.

It can be overwhelming to try and the right license for a use case. Two licenses that are well suited for data sharing are:

1. Creative Commons Attribution 4.0 International Public License (CC BY), and
2. Creative Commons CC0 1.0 Universal (CC0)

**CC BY**

The CC BY enforces attribution and due credit by default, but gives a lot of freedom for its use. Data can be shared and adapted, even for commercial use, with the following conditions:

* You must provide appropriate credit to the source. This means listing the names of the creators.
* Link back to the CC BY license, and
* Clearly show if changes were made.
* Data cannot be sub-licensed, that is - a change to the existing license
* There is also no warranty, so the person or people who obtained the data cannot be held liable.

The journal PLOS Comp Bio requires that data submitted cannot be more restrictive than CC BY [@plos-comp-bio-data]. For a brief overview of the CC BY, suitable to include in a README, see @ccby-short. For the full license, see @ccby-long.

**CC0**

The CC0 is a “public domain” license. Data with a CC0 license means the data owners waive all their rights to the work, and it now "owned" by the public. The data can be freely shared, this means it can be copied, modified, and distributed, even for commercial purposes _without asking permission_. When using data with CC0, it is good practice to cite the original paper, but it is not required. If you wish to use the CC0, see https://creativecommons.org/choose/zero/.  For a brief overview of the CC0, see @cc0-short, and for the full license, see @cc0-long.

Other licenses or notices to be aware of are **copyrighted data**, and **data embargos**. If you are working with data **already copyrighted**, (for example under CC BY or CC0), you must give follow appropriate guidelines for giving credit. Data may also be under **embargo**. This means data cannot be shared more widely until a specific release time. If sharing data under an embargo, include detailed information on the embargo requirements in: the README, and in separate correspondence with those who receive the data.

# Citation: How you want your data to be cited

A DOI is a prerequisite for citation. When citing data, it only makes sense to cite datasets that have been deposited into a DataCite compliant repo. If, for example, the data are deposited in dryad or Zenodo, the best practice would be to copy the citation created by these repositories. Under the hood, DataCite provides the DOI. If a DOI is unavailable, a citation will be meaningless, as it cannot be tracked by any means.

# Machine readable metadata

The data dictionary provides _human readable_ information on the data. To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata. An excellent standard for metadata is [Table Schema](https://frictionlessdata.io/specs/table-schema/). Say for a dataset called "demographics" with three variables:

| age| height|nationality |
|---:|------:|:-----------|
|  12|  161.5|Australian  |
|  21|  181.2|American    |
|  37|  178.3|New Zealand |

_Table XX Example demographics table of age, height, and nationality._

```
{
  "name": "demographics",
  # here we list the data files in this dataset
  "resources": [
    {
      "path": "demographics.csv",
      "schema": {
        "fields": [
          {
            "name": "age",
            "type": "integer"
          },
          {
            "name": "height",
            "type": "number"
          },
          {
            "name": "nationality",
            "type": "string"
          }
        ]
      }
    }
  ]
}
```

_Figure XX Example snippet of some Table Schema data for a dataset with three variables. This provides a description of each field, and the type of field, and it's description._

This contains fields such as path, and a schema with subfields name and type, for each variable. It also provides information for, licensing and features such as line breaks, and delimiters. It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format. Table schema is baked into formats such as [csvy](http://csvy.org/), an extended csv format, which has additional front matter in a YAML format using Table Schema.

# Raw data: The original/first data provided

Raw data is usually the first format of the data provided before any tidying or cleaning of the data. If the raw data is a practical size to share, it should be shared in a folder called `raw-data`. The raw data should be in the form that was first received, even if it is in binary or some proprietary format.  If possible, data dictionaries of the raw data should be provided in this folder as well.

# Scripts: To clean raw data ready for analysis

Any code used to clean and tidy the raw data should be provided in the `raw-data` directory. Ideally this would involve only scripted languages, but if other practical steps were taken to clean up the data, these should be recorded in a plain text or markdown file.

# Analysis ready data: Final data used in analysis

The data used in the data analysis should be provided in a folder called `data`.
Ideally, the data should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, and observations in rows. Contrasting `raw data`, `tidy data`/`analysis data` should be in an easily readable plain-text format, such as CSV, tab separated, or semicolon separated. Binary or proprietary formats are discouraged in favour of interoperability.

## Tooling for packaging data

NOTE: not sure if this quite fits here, tooling is important, but at the moment we are not sure if we should describe all tooling, as I will have a lot of examples for R, and not for Python/Julia

Tooling for producing these this information speeds up the process of sharing data. To help create codebooks, there are R packages such as `dataMeta`, `memisc`. and `codebook`. Codebooks are implemented in other software such as STATA, which provides a "codebook" command. Data can be packaged up in a "data package" with [DataPackageR](https://github.com/RGLab/DataPackageR), which provides tools to wrap up data into an R package, while prioviding helpers with MD5sum checks that track versioning. Note that is different to [Frictionlessdata's tabular data package spec](https://frictionlessdata.io/docs/tabular-data-package/).

# Example datasets

### Dataset 1: Forest Census Plot on Barro Colorado Island

The ["Forest Ceneus Plot on Barro Colorado Island"](http://ctfs.si.edu/webatlas/datasets/bci/) provides an example of data that is time consuming to document, but also robust. It contains information on trees from a 50-hectare tree plot. Here they state:

> Censuses have been carried out in 1981-1983, 1985, 1990, 1995, 2000, 2005, 2010, and 2015. In each census, all free-standing woody stems at least 10 mm diameter at breast height were identified, tagged, and mapped. Over 350,000 individual trees have been censused over 35 years.

This is an enourmous, ongoing research effort. Brief context of the data is provided at the link, along with a link to the census plot data from 2012 at [this paper](https://repository.si.edu/handle/10088/20925). This provides an abstract and description of the data, along with co-authors, and a recommended acknowledgement.

This dataset falls in the top right of the graph, at high effort to prepare, and high ease of understanding. It is high effort to prepare because it is a large research effort that not only involves expensive research in terms of time. It is high ease of understanding because

The impact of this kind of research is huge, as it provides information on biodiversity ...,  ..., ... .

### Dataset 2: Sensor data

Datasets obtained from sensors such as meteorological data are typically easy to prepare and understand.  This is because the sensors have to be very specific about what they are measuring, so a lot of the description of the type of data collected happens upstream at the instrument-design level. This flows down to when data is collected. The telescope data from the Sloan Digital Sky Survey (SDSS) is a great example of this. It contains:

> ... the most detailed three-dimensional maps of the Universe ever made, with deep multi-color images of one third of the sky, and spectra for more than three million astronomical objects. --(https://www.sdss.org/)

The telescope data is very high quality, with the following features:

- Good metadata in SCHEMA, a machine readable standard
- Data in multiple formats:
- Released data in archives
- Data is available in a database
- Entire dataset available for download
- Interface to the data covers all skills
- The interface provides tutorials from school children to professional astronomers on how to read it
- Raw data before processed into database is also available

The SDSS provides different degrees of understanding, for different understanding. It is a huge project involving many people, and so would have provided enourmous effort to prepare, but has a high degree of understanding. It would be in the top right hand corner of the DARECO.

### Dataset 3: Most other datasets

Many datasets are "data dumped" into repositories with a paper. Phrases such as the following might be familiar:

> Researcher 1: "What do these columns mean?"
> Researcher 2: "Sorry, I created the data 14 years ago and I don't remember".

** Need to get an example of this.

## Summary

TODO: Needs a succint summary here

# Ten simple rules for publishing data

1. **Decide whether publishing is appropriate for your data** It is important for everyone to publish their data in a research data archive to ensure long-term availability, curated metadata, and that they are preserved in the future. However this step might not be necessary depending on the stage and purpose of your project. In cases where this step is cumbersome, ensuring that the data are readily available by other means.
2. **Include a README** file with your data archive. README files have a long history dating back to the 1970s in the context of free software and are a requirement of the GNU standard. In a more generic sense, README files are a form of simple documentation that describe the contents of a directory, along with brief descriptions, the type of license used and information on how to cite the collection.
3. **Provide a codebook** or glossary for each variable. Codebooks provide a free-form way to document the data, their structure and underlying details of how they were collected (https://www.emgo.nl/kc/codebook/). More importantly they provide a way for future consumers of the data to understand how the variables were named, how missing data were documented, along with any additional information that might impact how the data are interpreted.

4. **Provide a machine readable format** for your data that describes what each column contains. This can be a json schema file, which also helps websites like Google Dataset Search automatically discover your data.
5. **Provide the data in its rawest form** in a folder called "data-raw". This 
 ensures provenance of your data by keeping a master copy from which all
other tidied up data can be created.
6. **Provide [open source?] scripts** used to clean data from rawest form into tidy/analysis ready format. These describe the steps taken to prepare the data,
which helps explain and document any decisions taken during the cleaning phase.
These should ideally operate on some raw data stored in the "data-raw" folder (rule 5).
7. **Keep additional copies in more accessible locations**: Even if you archive the data somewhere long-term, keep a second (or additional copies) in locations that are more readily and easily accessible for data analysis workflows. These include services like GitHub, GitHub LFS and others where fast CDNs make access and reuse practical, but fall back to Zenodo or similar in case these services were to go away.
8. **Use a hash function like MD5 checksum** to ensure that the data you deposited are the same that are being made available to others. Hash functions not only remove redundancy by removing the need to download the same data again, but also help ensure that the data are valid.
9. **Only add a citation if your data has a DOI**. A citation only makes sense when your data has a DataCtie compliant DOI, which is automatically provided when data is published in repositories like Zenodo and Dryad. Even then a citation may not accrue references, but without a DOI, it is gauraunteed not to.
10. Store a copy of your data in simple formats such as csv: CSV files follow a flat and extremely simple schema and can be opened by any text editor. These files can also be easily parsed by any data science programming language. Modern software packages allow for fast reads without consuming too much memory and these tools also allow for easy modification of the data.

# Conclusions

TBD

# Literature Cited

### 2019-05-09 paper.Rmd
Typo fixes

@@
 
-```{r fig-effort-underrstanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to udnerstand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
+```{r fig-effort-underrstanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
 library(ggplot2)
@@
 This serves the dual purpose of exposing methods and data together, making it extremely easy for a researcher to simply install and load the library at the top of their script.
-CRAN's distributed network (along with the global content distribution network maintained by Rstudio) ensure that the package is quickly accessible to anyone in the R ecosystem.
+CRAN's distributed network (along with the global content distribution network maintained by RStudio) ensure that the package is quickly accessible to anyone in the R ecosystem.
 A second critical advantage in this approach is that one could also benefit from R's package documentation syntax to generate useful metadata for fast reference.
@@
 This is why we recommend, as a rule, that researchers also archive data in a long-term data repository.
-These include domain specific repositories (see SECTION) or more general purpose ones such as Zenodo or figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.
+These include domain specific repositories (see SECTION) or more general purpose ones such as Zenodo or Figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.
 
@@
 
-1. Database access [@data-access] for services like MySQL, Hive, and ODBC, similar to R’s DBI organisation [@R-DBI]
+1. Database access [@data-access] for services like MySQL, Hive, and ODBC, similar to R’s DBI organization [@R-DBI]
 2. Accessing data online via APIs, such as accessing the weather source NOAA [@NOAA-jl]
@@
 For example, in astronomy the data is hosted at the SDSS (SDSS), Genetic data can be hosted at GenBank (@GenBank), and so on.
-Some other places to put data include the open science framework: https://osf.io/, dataverse, https://dataverse.org/, and data.world, https://data.world/.
+Some other places to put data include the open science framework: https://osf.io/, Dataverse, https://dataverse.org/, and data.world, https://data.world/.
 Other datasets might be produced
@@
 It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format.
-Table schema is baked into formats such as [csvy](http://csvy.org/), an extended csv format, which has additional front matter in a YAML format using Table Schema.
+Table schema is baked into formats such as [csvy](http://csvy.org/), an extended `csv` format, which has additional front matter in a YAML format using Table Schema.
 
@@
 Contrasting `raw data`, `tidy data`/`analysis data` should be in an easily readable plain-text format, such as CSV, tab separated, or semicolon separated.
-Binary or proprietary formats are discouraged in favour of interoperability.
+Binary or proprietary formats are discouraged in favor of interoperability.
 
@@
 
-This is an enourmous, ongoing research effort.
+This is an enormous, ongoing research effort.
 Brief context of the data is provided at the link, along with a link to the census plot data from 2012 at [this paper](https://repository.si.edu/handle/10088/20925).
@@
 The SDSS provides different degrees of understanding, for different understanding.
-It is a huge project involving many people, and so would have provided enourmous effort to prepare, but has a high degree of understanding.
+It is a huge project involving many people, and so would have provided enormous effort to prepare, but has a high degree of understanding.
 It would be in the top right hand corner of the DARECO.

### 2019-05-09 paper.Rmd
remove curly quotes
use new preamble template

@@
 output:
-  bookdown::pdf_book:
-    base_format: rmarkdown::pdf_document
+  bookdown::pdf_document2:
+    toc: no
+    includes:
+      in_header: preamble.tex
 bibliography: references.bib
+header-includes: 
+  - \linespread{2} 
 ---
 
-```{r setup}
+```{r setup, echo = FALSE}
 knitr::opts_chunk$set(echo = FALSE,
@@
 
-> “Data! data! data!" he cried impatiently. "I can't make bricks without clay.” - Sherlock Holmes (PBS Series, 2010-2017)
+> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (PBS Series, 2010-2017)
 

### 2019-05-09 paper.Rmd
Fixed more typos

@@
 Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors (citation).
-Sharing datasets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work.
+Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work.
 But there is also a component of benefit to the broader scientific community.
@@
 The act of easy data sharing can also improves model training, as many different models can be tried and tested on latest data sources, closing the loop on research and application of statistical techniques.
-Existing datasets can be combined or linked with new or existing data, fostering the development and synthesis of new ideas and research areas.
+Existing data sets can be combined or linked with new or existing data, fostering the development and synthesis of new ideas and research areas.
 The biggest of these benefits is the overall increase in reproducibility.
@@
 A typical data scientist loads a dozen or two of these open source libraries at the top of a notebook and then rely on existing routines to rapidly read, transform, visualize, and model data.
-Each of these packages individually depend on a complex web of other packages, building upon existing work rather than reimplementing everything from scratch.
+Each of these packages individually depend on a complex web of other packages, building upon existing work rather than re-implementing everything from scratch.
 Working from script and a list of such dependencies, a data analyst can easily install all the necessary tools in any local or remote environment and reproduce the computation.
@@
 To frame discussion around the challenges of data documenting, we can think of two features: "Effort to prepare", and "Ease of understanding".
-We can plot these on two axes and place datasets along each to indicate the ease of understanding and preparation (see Figure \@ref(fig:fig-effort-underrstanding)).
+We can plot these on two axes and place datasets along each to indicate the ease of understanding and preparation (see Figure \@ref(fig:fig-effort-understanding)).
 
@@
 So it might not have a worthwhile tradeoff for ease of understanding and ease of preparation.
-This means effort put into preparing data documentation for a small survey to be made easy to use should be kept relatively simple and not overcomplicated.
+This means effort put into preparing data documentation for a small survey to be made easy to use should be kept relatively simple and not over complicated.
 
@@
 
-In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databaseses.
+In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databases.
 This approach **gets the data somewhere rather than nowhere**.
@@
 Codebooks are implemented in other software such as STATA, which provides a "codebook" command.
-Data can be packaged up in a "data package" with [DataPackageR](https://github.com/RGLab/DataPackageR), which provides tools to wrap up data into an R package, while prioviding helpers with MD5sum checks that track versioning.
+Data can be packaged up in a "data package" with [DataPackageR](https://github.com/RGLab/DataPackageR), which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that track versioning.
 Note that is different to [Frictionlessdata's tabular data package spec](https://frictionlessdata.io/docs/tabular-data-package/).
@@
 **Only add a citation if your data has a DOI**.
-A citation only makes sense when your data has a DataCtie compliant DOI, which is automatically provided when data is published in repositories like Zenodo and Dryad.
+A citation only makes sense when your data has a DataCite compliant DOI, which is automatically provided when data is published in repositories like Zenodo and Dryad.
 Even then a citation may not accrue references, but without a DOI, it is gauraunteed not to.

### 2019-05-09 paper.Rmd
Fixed more typos and text

@@
 Without access to good data, it becomes extremely difficult, if not impossible, to advance science.
-Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors (citation).
+Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors (CITATION).
 Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work.
@@
 **In this paper we focus on the practical side of sharing data for the purpose of reproducibility.** Our goal is to describe various methods in which an analyst can share their data with minimal friction.
-We steer clear of theoretical ideas such as the FAIR data principles [@Wilkinson2016] since they still do not help a researcher share their data.
+We steer clear of idealitstic ideas such as the FAIR data principles [@Wilkinson2016] since they still do not help a researcher share their data.
 We also skip the discussion around citation and credit because data citations are still poorly tracked and there is no standardized metric or a h-index equivalent for data as of this writing.
@@
 
-The first two of these challenges have largely been solved.
+The first two of these challenges have largely been solved (REFERENCE).
 
@@
 The first is contained in a folder with dozens of data files and several papers on the topic.
-The second contains only a single csv file.
+The second contains only a single CSV file.
 The first is too much; understanding it requires a high time cost.
@@
 The R ecosystem has many noteworthy examples of data-only packages.
-One exemplar.
 One exemplar is the `nycflights13` package by Hadley Wickham (https://cran.r-project.org/web/packages/nycflights13/index.html).
@@
 (766). Binary files comprise a bulk of the data shipped in the `data`
-folder (68.06%) with other plain text formats such as txt, csv, dat,
-json comprising less than one percent of data shipped in packages.
+folder (68.06%) with other plain text formats such as txt, CSV, dat,
+`json` comprising less than one percent of data shipped in packages.
 
@@
 
-Another common situation that researchers face is in dealing with data that fall somewhere between small (tabular data such as csvs that are in the order of bytes to megabytes to gigabytes that can be easily compressed) to large (what falls under the umbrella of `big data`).
+Another common situation that researchers face is in dealing with data that fall somewhere between small (tabular data such as CSV that are in the order of bytes to megabytes to gigabytes that can be easily compressed) to large (what falls under the umbrella of `big data`).
 The happy medium is generally data that are too big to fit on memory of most standard machines, but can successfully fit on disk (https://ropensci.github.io/arkdb/articles/articles/noapi.html).
@@
 These databases provide disk based storage and using language agnostic interfaces, a analyst can easily query these data in manageable chunks that don't overrun memory limitations.
-Using software packages such as arkdb (https://ropensci.github.io/arkdb/index.html) one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
+Using software packages such as `arkdb` (https://ropensci.github.io/arkdb/index.html) one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
 
@@
 For example, in astronomy the data is hosted at the SDSS (SDSS), Genetic data can be hosted at GenBank (@GenBank), and so on.
-Some other places to put data include the open science framework: https://osf.io/, Dataverse, https://dataverse.org/, and data.world, https://data.world/.
+Some other places to put data include the open science framework: https://osf.io/, Dataverse, https://dataverse.org/, and Data World, https://data.world/.
 Other datasets might be produced
@@
 Data for research has a different domain compared to data collected by governments, or by journalists.
-Many governments or civil organisations are now making their own data available through a government website interface.
-Media and journalism groups are also making their data available either through places like GitHub (https://github.com/the-pudding/data), or organisations such as data.world, or may self host their data (https://data.fivethirtyeight.com/).
+Many governments or civil organizations are now making their own data available through a government website interface.
+Media and journalism groups are also making their data available either through places like GitHub (https://github.com/the-pudding/data), or organizations such as Data World, or may self host their data (https://data.fivethirtyeight.com/).
 
@@
 
-Below is an example data dictionary table taken from the [tidy tuesday repository on incarceration trends](https://github.com/rfordatascience/tidytuesday).
+Below is an example data dictionary table taken from the [Tidy Tuesday repository on incarceration trends](https://github.com/rfordatascience/tidytuesday).
 This includes information on the variable the class (type) of the variable, and a longer description of the variable.
@@
 
-Other licenses or notices to be aware of are **copyrighted data**, and **data embargos**.
+Other licenses or notices to be aware of are **copyrighted data**, and **data embargoes**.
 If you are working with data **already copyrighted**, (for example under CC BY or CC0), you must give follow appropriate guidelines for giving credit.
@@
 
-TODO: Needs a succint summary here
+TODO: Needs a succinct summary here
 
@@
 A citation only makes sense when your data has a DataCite compliant DOI, which is automatically provided when data is published in repositories like Zenodo and Dryad.
-Even then a citation may not accrue references, but without a DOI, it is gauraunteed not to.
+Even then a citation may not accrue references, but without a DOI, it is guaranteed not to.
 10.
-Store a copy of your data in simple formats such as csv: CSV files follow a flat and extremely simple schema and can be opened by any text editor.
+Store a copy of your data in simple formats such as CSV: Plain text CSV files follow a flat and extremely simple schema and can be opened by any text editor.
 These files can also be easily parsed by any data science programming language.

### 2019-05-10 paper.Rmd
adds SDSS description

@@
   - \linespread{2} 
+  - \usepackage[utf8]{inputenc}
 ---
@@
 Datasets obtained from sensors such as meteorological data are typically easy to prepare and understand.
-This is because the sensors have to be very specific about what they are measuring, so a lot of the description of the type of data collected happens upstream at the instrument-design level.
+This is typically because the sensors have to be very specific about what they are measuring, so a lot of the description of the type of data collected happens upstream at the instrument-design level.
 This flows down to when data is collected.
 The telescope data from the Sloan Digital Sky Survey (SDSS) is a great example of this.
-It contains:
+It contains the most detailed, three dimensionsal maps of the universe ever made, and has been operated since 2000, now producing over 200Gb of data every day [@sdss-one-paper; @sdss-four-paper; @sdss-website].
+It is not possible to summarise the contents of the data, but briefly, the SDSS data includes photometric and spectroscopic information obtained from a 2.5m telescope at Apache Point Observatory in New Mexico.
 
-> ...
-the most detailed three-dimensional maps of the Universe ever made, with deep multi-color images of one third of the sky, and spectra for more than three million astronomical objects.
---(https://www.sdss.org/)
-
+The data are are free to use and publicly accessible, with the interface to the data being designed to cover all skills, from shcool children to professional astronomers.
 The telescope data is very high quality, with the following features:
@@
 - Good metadata in SCHEMA, a machine readable standard
-- Data in multiple formats:
+- Data in multiple formats
 - Released data in archives
 - Data is available in a database
-- Entire dataset available for download
-- Interface to the data covers all skills
-- The interface provides tutorials from school children to professional astronomers on how to read it
+- Entire datasets available for download
 - Raw data before processed into database is also available
 
-The SDSS provides different degrees of understanding, for different understanding.
-It is a huge project involving many people, and so would have provided enormous effort to prepare, but has a high degree of understanding.
-It would be in the top right hand corner of the DARECO.
+The SDSS provides different degrees of understanding.
+It is a huge project involving many people, and requiring a staggering and enormous effort to prepare.
+Despite this, it is still very easy to understand.
+This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-underrstanding)).
+The impact of this data is high, having changed the way we understand the past, present and future of the universe itself.
+This impact is surely due to the care and thought put into making the data accessible and public.
+It is a statement of what is possible with good data sharing practices.
 

### 2019-05-10 paper.Rmd
Added draft of BCI

@@
 
-The ["Forest Ceneus Plot on Barro Colorado Island"](http://ctfs.si.edu/webatlas/datasets/bci/) provides an example of data that is time consuming to document, but also robust.
-It contains information on trees from a 50-hectare tree plot.
-Here they state:
+Long-term field surveys, where the same type of data is repeatedly measured and collected over a long period of time, are an example of data collection where the time and financial investment would necessitate meticulously curated and easily accessible metadata.
+Both from the fact that the same protocol is being followed year after year, and that field data collection efforts are quite expensive, these data need to be well documented, with documentation available alongside the data.
 
-> Censuses have been carried out in 1981-1983, 1985, 1990, 1995, 2000, 2005, 2010, and 2015.
-In each census, all free-standing woody stems at least 10 mm diameter at breast height were identified, tagged, and mapped.
-Over 350,000 individual trees have been censused over 35 years.
-
-This is an enormous, ongoing research effort.
-Brief context of the data is provided at the link, along with a link to the census plot data from 2012 at [this paper](https://repository.si.edu/handle/10088/20925).
-This provides an abstract and description of the data, along with co-authors, and a recommended acknowledgement.
-
-This dataset falls in the top right of the graph, at high effort to prepare, and high ease of understanding.
-It is high effort to prepare because it is a large research effort that not only involves expensive research in terms of time.
-It is high ease of understanding because
-
-The impact of this kind of research is huge, as it provides information on biodiversity ...,  ..., ... .
+One example of a very laborious, long-term study is the 50-hectare plot on Barro Colorado Island in Panama.
+Researchers at the field station have censused every single tree in a pre-defined plot 7 times over the past 30 years.
+More than a million trees have been counted.
+The data and metadata however and hard to reach.
+To obtain the data, one must fill out a form and agree to terms and conditions which require approval and authorship prior to publication.
+The biggest challenge with this study is that the data are stored on a personal FTP server of one of the authors.
+While the data are available in CSV and binary Rdata formats, the data storage servers do not have any metadata, README files or codebooks.
+A separate, public archive hosts the metadata (https://repository.si.edu/handle/10088/20925) in a PDF file (https://repository.si.edu/bitstream/handle/10088/20925/RoutputFull.pdf?sequence=1&isAllowed=y) that describe the fields.
 

### 2019-05-11 paper.Rmd
replace some URLs with bib references; fix some references to figures

@@
 Each of these languages is supported by thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
-As of this writing, Python, Julia, and R have  167k packages (https://pypi.org/), ~ 14k packages (https://cran.r-project.org/) and ~2k  packages (https://pkg.julialang.org/) respectively.
+As of this writing, Python, Julia, and R have  167k packages [@pypi], ~ 14k packages [@cran] and ~2k  packages [@julia-pkgman] respectively.
 These packages form the building blocks of a data scientists daily workflow.
@@
 
-```{r fig-effort-underrstanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
+```{r fig-effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
 library(ggplot2)
@@
 The R ecosystem has many noteworthy examples of data-only packages.
-One exemplar is the `nycflights13` package by Hadley Wickham (https://cran.r-project.org/web/packages/nycflights13/index.html).
+One exemplar is the `nycflights13` package by Hadley Wickham [@nycflights].
 This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes.
@@
 
-Including data in a python package: https://python-packaging.readthedocs.io/en/latest/non-code-files.html
+Including data in a python package: [@include-data-py-pkg]
 A user can typically include data (or any other arbitrary files) inside a python package by declaring it in the manifest.
@@
 These databases provide disk based storage and using language agnostic interfaces, a analyst can easily query these data in manageable chunks that don't overrun memory limitations.
-Using software packages such as `arkdb` (https://ropensci.github.io/arkdb/index.html) one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
+Using software packages such as `arkdb` [@arkdb] one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
 
@@
 Many governments or civil organizations are now making their own data available through a government website interface.
-Media and journalism groups are also making their data available either through places like GitHub (https://github.com/the-pudding/data), or organizations such as Data World, or may self host their data (https://data.fivethirtyeight.com/).
+Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or organizations such as Data World, or may self host their data, such as five thirty eight [@data-538].
 
@@
 
-Below is an example data dictionary table taken from the [Tidy Tuesday repository on incarceration trends](https://github.com/rfordatascience/tidytuesday).
+Below is an example data dictionary table taken from the [Tidy Tuesday repository on incarceration trends](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-01-22) [@tidy-tuesday-incarcerate].
 This includes information on the variable the class (type) of the variable, and a longer description of the variable.
@@
 It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format.
-Table schema is baked into formats such as [csvy](http://csvy.org/), an extended `csv` format, which has additional front matter in a YAML format using Table Schema.
+Table schema is baked into formats such as csvy [@csvy], an extended `csv` format, which has additional front matter in a YAML format using Table Schema.
 
@@
 Despite this, it is still very easy to understand.
-This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-underrstanding)).
+This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-understanding)).
 The impact of this data is high, having changed the way we understand the past, present and future of the universe itself.

### 2019-05-15 paper.Rmd
replace most URLs with .bib citations.

@@
 For example, in astronomy the data is hosted at the SDSS (SDSS), Genetic data can be hosted at GenBank (@GenBank), and so on.
-Some other places to put data include the open science framework: https://osf.io/, Dataverse, https://dataverse.org/, and Data World, https://data.world/.
+Some other places to put data include the open science framework [@OSF], Dataverse, [@dataverse], and Data World, [@dataworld].
 Other datasets might be produced
@@
 When using data with CC0, it is good practice to cite the original paper, but it is not required.
-If you wish to use the CC0, see https://creativecommons.org/choose/zero/.
+If you wish to use the CC0, see @choose-cc0.
 For a brief overview of the CC0, see @cc0-short, and for the full license, see @cc0-long.
@@
 To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
-An excellent standard for metadata is [Table Schema](https://frictionlessdata.io/specs/table-schema/).
+An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
 Say for a dataset called "demographics" with three variables:
@@
 Codebooks are implemented in other software such as STATA, which provides a "codebook" command.
-Data can be packaged up in a "data package" with [DataPackageR](https://github.com/RGLab/DataPackageR), which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that track versioning.
-Note that is different to [Frictionlessdata's tabular data package spec](https://frictionlessdata.io/docs/tabular-data-package/).
+Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that track versioning.
+Note that is different to [Frictionlessdata's tabular data package spec [@frictionlessdata-data-pkg].
 

### 2019-05-15 paper.Rmd
remove hyperlink to tidy tuesday repo, as it now has a proper bib reference.

@@
 
-Below is an example data dictionary table taken from the [Tidy Tuesday repository on incarceration trends](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-01-22) [@tidy-tuesday-incarcerate].
+Below is an example data dictionary table taken from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate].
 This includes information on the variable the class (type) of the variable, and a longer description of the variable.

### 2019-05-15 paper.Rmd
cites Kirk2019 article in introduction, close #75

@@
 Without access to good data, it becomes extremely difficult, if not impossible, to advance science.
-Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors (CITATION).
-Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work.
+Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors [@Rowhani-Farid2016; @Stodden2018].
+Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [@Kirk2019].
 But there is also a component of benefit to the broader scientific community.

### 2019-05-15 paper.Rmd
add reference to introduction; restructure heading levels for section 9 on the minimal structure for researchers

@@
 Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors [@Rowhani-Farid2016; @Stodden2018].
-Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [@Kirk2019].
+Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [Popkin2019; @Kirk2019].
 But there is also a component of benefit to the broader scientific community.
@@
 
-# README: A Human readable description of the data
+## README: A Human readable description of the data
 
@@
 
-# Codebook: Human readable dictionary of data contents
+## Codebook: Human readable dictionary of data contents
 
@@
 
-# License: How to use and share the data
+## License: How to use and share the data
 
@@
 
-**CC BY**
+### CC BY
 
@@
 
-**CC0**
+### CC0
 
@@
 
-# Citation: How you want your data to be cited
+## Citation: How you want your data to be cited
 
@@
 
-# Machine readable metadata
+## Machine readable metadata
 
@@
 
-# Raw data: The original/first data provided
+## Raw data: The original/first data provided
 
@@
 
-# Scripts: To clean raw data ready for analysis
+## Scripts: To clean raw data ready for analysis
 
@@
 
-# Analysis ready data: Final data used in analysis
+## Analysis ready data: Final data used in analysis
 
@@
 
-## Tooling for packaging data
+# Tooling for packaging data
 
-NOTE: not sure if this quite fits here, tooling is important, but at the moment we are not sure if we should describe all tooling, as I will have a lot of examples for R, and not for Python/Julia
+<!-- NOTE: not sure if this quite fits here, tooling is important, but at the moment we are not sure if we should describe all tooling, as I will have a lot of examples for R, and not for Python/Julia -->
 
@@
 
-### Dataset 1: Forest Census Plot on Barro Colorado Island
+## Dataset 1: Forest Census Plot on Barro Colorado Island
 
@@
 
-### Dataset 2: Sensor data
+## Dataset 2: Sensor data
 
@@
 
-### Dataset 3: Most other datasets
+## Dataset 3: Most other datasets
 

### 2019-05-15 paper.Rmd
create proper tables and reference them using dynamic references

@@
                       out.width = "100%")
+
+library(kableExtra)
 ```
@@
 
-Below is an example data dictionary table taken from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate].
+Table \@ref(tab:table-data-incarcerate) shows an example data dictionary table taken from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate].
 This includes information on the variable the class (type) of the variable, and a longer description of the variable.
 
-__Table 1. The prisoner summary data dictionary, with columns on the variable, it's class, and a short description of the contents of the variable.__
+```{r table-data-incarcerate}
+dictionary_incarcerate <- data.frame(
+  Variable = c("year", "urbanicity", "pop_category", "rate_per_100000"),
+  Class  = c("integer (date)",
+             "character",
+             "character",
+             "double"),
+  description = c("Year",
+                  "County-type (urban, suburban, small/mid, rural)",
+                  "Category for population - either race, gender, or Total",
+                  "Rate within a category for prison population per 100,000 people")
+)
 
-|variable        |class     | Description |
-|:---------------|:---------| :---------- |
-|year            |integer (date)   | Year |
-|urbanicity      |character | County-type (urban, suburban, small/mid, rural) |
-|pop_category    |character | Category for population - either race, gender, or Total |
-|rate_per_100000 |double    | Rate within a category for prison population per 100,000 people |
+knitr::kable(dictionary_incarcerate,
+             caption = "The prisoner summary data dictionary, with columns on the variable, it's class, and a short description of the contents of the variable.",
+             booktabs = TRUE) %>%
+  kable_styling(latex_options = c("hold_position"))
+```
 
@@
 An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
-Say for a dataset called "demographics" with three variables:
+Say for a dataset called "demographics" with three variables (Table \@ref(tab:table-machine-readable))
 
-| age| height|nationality |
-|---:|------:|:-----------|
-|  12|  161.5|Australian  |
-|  21|  181.2|American    |
-|  37|  178.3|New Zealand |
+```{r table-machine-readable}
+table_machine_readable <- data.frame(
+  age = c(12L, 21L, 37L),
+  height = c(161.5, 181.2, 178.3),
+  nationality = c("Australian", "American", "New Zealand")
+)
 
-_Table XX Example demographics table of age, height, and nationality._
+knitr::kable(table_machine_readable,
+             caption = "Example demographics table of age, height, and nationality.",
+             booktabs = TRUE) %>%
+  kable_styling(latex_options = c("hold_position"))
+```
 

### 2019-05-15 paper.Rmd
reword small parts of challenges

@@
 Providing a well documented dataset alongside research requires meeting several challenges.
-This includes understanding all the pieces and features of documentation such as metadata, licenses, and raw data; where to literally share the data; how to ensure long term provenance of the data.
+This includes understanding all the pieces and features of documentation such as metadata, licenses, and raw data; knowing where to share the data to ensure long term storage.
 
-One challenge is that better documentation requires more time, and that higher impact data needs better documentation.
-
-Sharing data with colleagues: binary data is quick to share, but not as good as something like a simpler format like CSV (ask Laura DeCicco re this)
-
-Providing the right amount of documentation for your data is challenging.
-Too much is overwhelming; too little and you don't know anything.
-To illustrate this, imagine two datasets.
-The first is contained in a folder with dozens of data files and several papers on the topic.
-The second contains only a single CSV file.
-The first is too much; understanding it requires a high time cost.
-The second is too little; we don't know anything about what the data is or why it was collected.
-The challenge is to provide the right level of detail, somewhere between these two extremes.
+Better documentation requires more time. Higher impact data needs better documentation.
 
@@
 Data with high potential for impact should be made easier to understand and share.
-This can be seen in the research created from the 50 hectare plot.
+This can be seen in the research created from the 50 hectare plot, discussed in section ...
+- this dataset was challenging to understand without special domain expertise.
+
+Simpler format data like CSV is more reliable than binary serialised/compressed data (ask Laura DeCicco re this).
+
+Providing the right amount of documentation for your data is challenging.
+Too much is overwhelming; too little and you don't know anything.
+For example, compare a folder with dozens of data files and several papers, to a single CSV file.
+The first is too much, and understanding it requires a high time cost.
+The second is too little; we don't know anything about what the data is or why it was collected.
+The challenge is to provide the right level of detail, somewhere between these two extremes.
 

### 2019-05-15 paper.Rmd
condense data sharing in julia and python section

@@
 
-## Challenges
-
-**Challenges: higher potential impact means more time**
+## Higher potential impact means more time**
 
@@
 
-Sharing data in a programming language happens to some extent in other languages. We discuss how Julia and Python share and distribute data in the following section.
-
-# Julia
+Sharing data in a programming language happens to some extent in other languages, such as Julia and Python.
 
 Julia was established in 2012, and is a high level language with great performance.
-Searching METADATA.jl [@METADATA-jl] for packages mentioning data revealed three main categories of data packages in Julia:
-
-1. Database access [@data-access] for services like MySQL, Hive, and ODBC, similar to R’s DBI organization [@R-DBI]
-2. Accessing data online via APIs, such as accessing the weather source NOAA [@NOAA-jl]
-3.
-Data is provided directly.
-For example, accessing Econometric datasets [@econ-data-jl], face datasets [@face-data-jl], Market data [@market-data-jl], and even R datasets [@r-data-jl].
-It is telling that the datasets provided in R are useful, as there is even a Julia package, RDatasets that provides over 700 datasets from R packages.
-There are also packages for Accessing datasets used in vega [@vega-data-jl], for accessing common machine learning datasets [@ml-data-jl].
-There are also packages for generating fake data [@fake-data-jl], for which there are at least two R packages that do this, charlatan [@charlatan-data-r], and [@wakefield-data-r].
-
-# Python
+Similar to R, Julia provides packages for database access, API access, and data packages [@METADATA-jl].
+This includes access to MySQL databases, NOAA [@NOAA-jl], and actual datasets from Econometrics [@econ-data-jl], faces, [@face-data-jl], Market data [@market-data-jl], and even R datasets [@r-data-jl].
 
 Python is a popular general purpose language, used in engineering, statistics, and computer science. 
-One language agnostic solution is to use the Open Knowledge Foundations Frictionless data spec.
-This is implemented in both R and Python.
-
-Including data in a python package: [@include-data-py-pkg]
-A user can typically include data (or any other arbitrary files) inside a python package by declaring it in the manifest.
+Including data in a python package is straightforward; a user can include data (or any other arbitrary files) inside a python package by declaring it in the manifest [@include-data-py-pkg].
 
@@
 
+# Morgue
+
+One language agnostic solution is to use the Open Knowledge Foundations Frictionless data spec, implemented in both R and Python.
+
 # Literature Cited

### 2019-06-10 paper.Rmd
first pass at adding more about the SDSS data #80

--- 2019-05-15 paper.Rmd
@@
                       out.width = "100%")
-
-library(kableExtra)
 ```
@@
 Without access to good data, it becomes extremely difficult, if not impossible, to advance science.
-Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors [@Rowhani-Farid2016; @Stodden2018].
-Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [Popkin2019; @Kirk2019].
+Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors (CITATION).
+Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work.
 But there is also a component of benefit to the broader scientific community.
@@
 Each of these languages is supported by thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
-As of this writing, Python, Julia, and R have  167k packages [@pypi], ~ 14k packages [@cran] and ~2k  packages [@julia-pkgman] respectively.
+As of this writing, Python, Julia, and R have  167k packages (https://pypi.org/), ~ 14k packages (https://cran.r-project.org/) and ~2k  packages (https://pkg.julialang.org/) respectively.
 These packages form the building blocks of a data scientists daily workflow.
@@
 
-```{r fig-effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
+```{r fig-effort-underrstanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
 library(ggplot2)
@@
 The R ecosystem has many noteworthy examples of data-only packages.
-One exemplar is the `nycflights13` package by Hadley Wickham [@nycflights].
+One exemplar is the `nycflights13` package by Hadley Wickham (https://cran.r-project.org/web/packages/nycflights13/index.html).
 This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes.
@@
 
-Including data in a python package: [@include-data-py-pkg]
+Including data in a python package: https://python-packaging.readthedocs.io/en/latest/non-code-files.html
 A user can typically include data (or any other arbitrary files) inside a python package by declaring it in the manifest.
@@
 These databases provide disk based storage and using language agnostic interfaces, a analyst can easily query these data in manageable chunks that don't overrun memory limitations.
-Using software packages such as `arkdb` [@arkdb] one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
+Using software packages such as `arkdb` (https://ropensci.github.io/arkdb/index.html) one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
 
@@
 For example, in astronomy the data is hosted at the SDSS (SDSS), Genetic data can be hosted at GenBank (@GenBank), and so on.
-Some other places to put data include the open science framework [@OSF], Dataverse, [@dataverse], and Data World, [@dataworld].
+Some other places to put data include the open science framework: https://osf.io/, Dataverse, https://dataverse.org/, and Data World, https://data.world/.
 Other datasets might be produced
@@
 Many governments or civil organizations are now making their own data available through a government website interface.
-Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or organizations such as Data World, or may self host their data, such as five thirty eight [@data-538].
+Media and journalism groups are also making their data available either through places like GitHub (https://github.com/the-pudding/data), or organizations such as Data World, or may self host their data (https://data.fivethirtyeight.com/).
 
@@
 
-## README: A Human readable description of the data
+# README: A Human readable description of the data
 
@@
 
-## Codebook: Human readable dictionary of data contents
+# Codebook: Human readable dictionary of data contents
 
@@
 
-Table \@ref(tab:table-data-incarcerate) shows an example data dictionary table taken from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate].
+Below is an example data dictionary table taken from the [Tidy Tuesday repository on incarceration trends](https://github.com/rfordatascience/tidytuesday).
 This includes information on the variable the class (type) of the variable, and a longer description of the variable.
 
-```{r table-data-incarcerate}
-dictionary_incarcerate <- data.frame(
-  Variable = c("year", "urbanicity", "pop_category", "rate_per_100000"),
-  Class  = c("integer (date)",
-             "character",
-             "character",
-             "double"),
-  description = c("Year",
-                  "County-type (urban, suburban, small/mid, rural)",
-                  "Category for population - either race, gender, or Total",
-                  "Rate within a category for prison population per 100,000 people")
-)
+__Table 1. The prisoner summary data dictionary, with columns on the variable, it's class, and a short description of the contents of the variable.__
 
-knitr::kable(dictionary_incarcerate,
-             caption = "The prisoner summary data dictionary, with columns on the variable, it's class, and a short description of the contents of the variable.",
-             booktabs = TRUE) %>%
-  kable_styling(latex_options = c("hold_position"))
-```
+|variable        |class     | Description |
+|:---------------|:---------| :---------- |
+|year            |integer (date)   | Year |
+|urbanicity      |character | County-type (urban, suburban, small/mid, rural) |
+|pop_category    |character | Category for population - either race, gender, or Total |
+|rate_per_100000 |double    | Rate within a category for prison population per 100,000 people |
 
@@
 
-## License: How to use and share the data
+# License: How to use and share the data
 
@@
 
-### CC BY
+**CC BY**
 
@@
 
-### CC0
+**CC0**
 
@@
 When using data with CC0, it is good practice to cite the original paper, but it is not required.
-If you wish to use the CC0, see @choose-cc0.
+If you wish to use the CC0, see https://creativecommons.org/choose/zero/.
 For a brief overview of the CC0, see @cc0-short, and for the full license, see @cc0-long.
@@
 
-## Citation: How you want your data to be cited
+# Citation: How you want your data to be cited
 
@@
 
-## Machine readable metadata
+# Machine readable metadata
 
@@
 To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
-An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
-Say for a dataset called "demographics" with three variables (Table \@ref(tab:table-machine-readable))
+An excellent standard for metadata is [Table Schema](https://frictionlessdata.io/specs/table-schema/).
+Say for a dataset called "demographics" with three variables:
 
-```{r table-machine-readable}
-table_machine_readable <- data.frame(
-  age = c(12L, 21L, 37L),
-  height = c(161.5, 181.2, 178.3),
-  nationality = c("Australian", "American", "New Zealand")
-)
+| age| height|nationality |
+|---:|------:|:-----------|
+|  12|  161.5|Australian  |
+|  21|  181.2|American    |
+|  37|  178.3|New Zealand |
 
-knitr::kable(table_machine_readable,
-             caption = "Example demographics table of age, height, and nationality.",
-             booktabs = TRUE) %>%
-  kable_styling(latex_options = c("hold_position"))
-```
+_Table XX Example demographics table of age, height, and nationality._
 
@@
 It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format.
-Table schema is baked into formats such as csvy [@csvy], an extended `csv` format, which has additional front matter in a YAML format using Table Schema.
+Table schema is baked into formats such as [csvy](http://csvy.org/), an extended `csv` format, which has additional front matter in a YAML format using Table Schema.
 
-## Raw data: The original/first data provided
+# Raw data: The original/first data provided
 
@@
 
-## Scripts: To clean raw data ready for analysis
+# Scripts: To clean raw data ready for analysis
 
@@
 
-## Analysis ready data: Final data used in analysis
+# Analysis ready data: Final data used in analysis
 
@@
 
-# Tooling for packaging data
+## Tooling for packaging data
 
-<!-- NOTE: not sure if this quite fits here, tooling is important, but at the moment we are not sure if we should describe all tooling, as I will have a lot of examples for R, and not for Python/Julia -->
+NOTE: not sure if this quite fits here, tooling is important, but at the moment we are not sure if we should describe all tooling, as I will have a lot of examples for R, and not for Python/Julia
 
@@
 Codebooks are implemented in other software such as STATA, which provides a "codebook" command.
-Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that track versioning.
-Note that is different to [Frictionlessdata's tabular data package spec [@frictionlessdata-data-pkg].
+Data can be packaged up in a "data package" with [DataPackageR](https://github.com/RGLab/DataPackageR), which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that track versioning.
+Note that is different to [Frictionlessdata's tabular data package spec](https://frictionlessdata.io/docs/tabular-data-package/).
 
@@
 
-## Dataset 1: Forest Census Plot on Barro Colorado Island
+### Dataset 1: Forest Census Plot on Barro Colorado Island
 
-Long-term field surveys, where the same type of data is repeatedly measured and collected over a long period of time, are an example of data collection where the time and financial investment would necessitate meticulously curated and easily accessible metadata.
-Both from the fact that the same protocol is being followed year after year, and that field data collection efforts are quite expensive, these data need to be well documented, with documentation available alongside the data.
+The ["Forest Ceneus Plot on Barro Colorado Island"](http://ctfs.si.edu/webatlas/datasets/bci/) provides an example of data that is time consuming to document, but also robust.
+It contains information on trees from a 50-hectare tree plot.
+Here they state:
 
-One example of a very laborious, long-term study is the 50-hectare plot on Barro Colorado Island in Panama.
-Researchers at the field station have censused every single tree in a pre-defined plot 7 times over the past 30 years.
-More than a million trees have been counted.
-The data and metadata however and hard to reach.
-To obtain the data, one must fill out a form and agree to terms and conditions which require approval and authorship prior to publication.
-The biggest challenge with this study is that the data are stored on a personal FTP server of one of the authors.
-While the data are available in CSV and binary Rdata formats, the data storage servers do not have any metadata, README files or codebooks.
-A separate, public archive hosts the metadata (https://repository.si.edu/handle/10088/20925) in a PDF file (https://repository.si.edu/bitstream/handle/10088/20925/RoutputFull.pdf?sequence=1&isAllowed=y) that describe the fields.
+> Censuses have been carried out in 1981-1983, 1985, 1990, 1995, 2000, 2005, 2010, and 2015.
+In each census, all free-standing woody stems at least 10 mm diameter at breast height were identified, tagged, and mapped.
+Over 350,000 individual trees have been censused over 35 years.
 
-## Dataset 2: Sensor data
+This is an enormous, ongoing research effort.
+Brief context of the data is provided at the link, along with a link to the census plot data from 2012 at [this paper](https://repository.si.edu/handle/10088/20925).
+This provides an abstract and description of the data, along with co-authors, and a recommended acknowledgement.
+
+This dataset falls in the top right of the graph, at high effort to prepare, and high ease of understanding.
+It is high effort to prepare because it is a large research effort that not only involves expensive research in terms of time.
+It is high ease of understanding because
+
+The impact of this kind of research is huge, as it provides information on biodiversity ...,  ..., ... .
+
+### Dataset 2: Sensor data
 
@@
 The SDSS provides different degrees of understanding.
-It is a huge project involving many people, and requiring a staggering and enormous effort to prepare.
+It is a huge project involving many people, and a staggering effort to prepare.
 Despite this, it is still very easy to understand.
-This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-understanding)).
+This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-underrstanding)).
 The impact of this data is high, having changed the way we understand the past, present and future of the universe itself.
@@
 
-## Dataset 3: Most other datasets
+**include some references to the SDSS project and some of the key/impactful findings**
+
+As an example, the optical and spectral data are explained here: https://www.sdss.org/dr15/spectro/spectro_basics/ and can be downloaded here in special FITS dataformat, or even as a plain CSV: http://skyserver.sdss.org/dr15/en/tools/quicklook/summary.aspx?
+
+An example plot in Figure \@ref(fig:sdss-spectra)A shows the Wavelength against Flux. Here, flux is ..., and wave is ...
+
+```{r prepare-spectra-plot, include = FALSE}
+
+sdss_image_path <- here::here("figures", "sdss-spectra-example-plate-1678.png")
+library(magick)
+library(ggplot2)
+library(cowplot)
+
+spectra <- readr::read_csv(
+  here::here("data",
+             "plateid-1678_mjd-53433_fiberid-425_reduction2d-v5_7_0.csv")
+)
+
+gg_spectra <- ggplot(spectra,
+                     aes(x = Wavelength,
+                         y = Flux)) +
+  geom_line() + 
+  theme_minimal() + 
+  theme(aspect.ratio = 1)
+
+plot_spectra <- ggdraw() + 
+  draw_image(sdss_image_path)
+
+```
+
+```{r sdss-spectra, fig.cap = "Spectra image of wavelength against flux from SDSS. Part A shows the example image taken from SDSS website. Part B shows the replicated figure using data provided"}
+
+plot_grid(plot_spectra, gg_spectra, labels = "AUTO")
+```
+
+A great litmus test of reproducibility for this is the fact that from downloading the CSV from the website, a similar data visualisation can be produced, seen in Figure \@ref(fig:sdss-spectra)B:
+
+The data was downloaded from: http://dr15.sdss.org/optical/spectrum/view/data/format=csv?plateid=1678&mjd=53433&fiberid=425&reduction2d=v5_7_0
+
+This URL contains details of the data, specifically:
+
+* plateid = 1678
+* mjd = 53433
+* fiberid = 425
+* reduction2d = v5.7.0
+
+These are unique identifying features of the data, which tell us:
+
+* plateid is the plate that they use to explore a given section of night sky. It uniquely identifies a "plug plate" used to eollcet the data
+* mjd is an integer that marks the "modified julian date" of the night when observations was carried out.
+* fiberid is an integer that marks the fiber number used.
+* The version marked as `v5_7_0` is the pipeline version used for BOSS in DR12
+
+### Dataset 3: Most other datasets
 

### 2019-06-11 paper.Rmd
finish up sdss section, close #80

@@
 
+To demonstrate the variety of datasets and documentation practices, we now discuss three datasets of varying quality.
+
 ### Dataset 1: Forest Census Plot on Barro Colorado Island
@@
 
-Datasets obtained from sensors such as meteorological data are typically easy to prepare and understand.
-This is typically because the sensors have to be very specific about what they are measuring, so a lot of the description of the type of data collected happens upstream at the instrument-design level.
-This flows down to when data is collected.
+Datasets obtained from sensors such as meteorological data are typically easy to prepare and understand, since sensors measure specific features.
+This means the description of data type happens upstream at the instrument-design level, and flows down to data collection.
 The telescope data from the Sloan Digital Sky Survey (SDSS) is a great example of this.
-It contains the most detailed, three dimensionsal maps of the universe ever made, and has been operated since 2000, now producing over 200Gb of data every day [@sdss-one-paper; @sdss-four-paper; @sdss-website].
-It is not possible to summarise the contents of the data, but briefly, the SDSS data includes photometric and spectroscopic information obtained from a 2.5m telescope at Apache Point Observatory in New Mexico.
 
-The data are are free to use and publicly accessible, with the interface to the data being designed to cover all skills, from shcool children to professional astronomers.
+It is not possible to summarise the contents of the data, but briefly, the SDSS data includes photometric and spectroscopic information obtained from a 2.5m telescope at Apache Point Observatory in New Mexico, operating since 2000,s producing over 200Gb of data every day [@sdss-one-paper; @sdss-four-paper; @sdss-website].
+.
+This data has been used to produce the most detailed, three dimensionsal maps of the universe ever made.
+
+The data are are free to use and publicly accessible, with the interface to the data being designed to cover a wide range of skills.
+For example, the marvin service to streamline access and visualisation of MaNGA data [@marvin-service], through to raw data access (see Figure \@ref(fig:sdss-raw-data-screenshot)).
+
+```{r sdss-raw-data-screenshot, fig.cap = "Screenshot of Raw data avaialbe through DR15 FITS."}
+knitr::include_graphics(here::here("figures",
+                                   "sdss-raw-data-screenshot.png"))
+```
+
 The telescope data is very high quality, with the following features:
 
-- Good metadata in SCHEMA, a machine readable standard
-- Data in multiple formats
-- Released data in archives
+- Metadata in the machine readable standard, SCHEMA
+- Data in multiple formats (e.g., .csv and .fits)
+- Previously released data available in archives
 - Data is available in a database
@@
 
-The SDSS provides different degrees of understanding.
-It is a huge project involving many people, and a staggering effort to prepare.
-Despite this, it is still very easy to understand.
-This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-underrstanding)).
-The impact of this data is high, having changed the way we understand the past, present and future of the universe itself.
-This impact is surely due to the care and thought put into making the data accessible and public.
-It is a statement of what is possible with good data sharing practices.
+As an example, the optical and spectral data can be accessed in FITS data format, or even as a plain CSV, the first few rows shown in Table \@ref(tab:spectra-tab).
+An example plot in Figure \@ref(fig:sdss-spectra)A shows the Wavelength against Flux as produced on the SDSS website, while Figure \@ref(fig:sdss-spectra)B.
+The fact that figure Figure \@ref(fig:sdss-spectra)A is reproduced in Figure \@ref(fig:sdss-spectra)B demonstrates great reproducibility.
 
-**include some references to the SDSS project and some of the key/impactful findings**
-
-As an example, the optical and spectral data are explained here: https://www.sdss.org/dr15/spectro/spectro_basics/ and can be downloaded here in special FITS dataformat, or even as a plain CSV: http://skyserver.sdss.org/dr15/en/tools/quicklook/summary.aspx?
-
-An example plot in Figure \@ref(fig:sdss-spectra)A shows the Wavelength against Flux. Here, flux is ..., and wave is ...
-
-```{r prepare-spectra-plot, include = FALSE}
+```{r prepare-spectra-data, include = FALSE}
 
@@
 
-```{r sdss-spectra, fig.cap = "Spectra image of wavelength against flux from SDSS. Part A shows the example image taken from SDSS website. Part B shows the replicated figure using data provided"}
+```{r spectra-tab, echo = FALSE}
+library(kableExtra)
+head(spectra) %>%
+  knitr::kable(caption = "Example spectra data from SDSS, showing values for Wavelength, Flux, BestFit, and SkyFlux.",
+               booktabs = TRUE) %>%
+  kable_styling(latex_options = c("striped",
+                                  "hold_position"))
+
+```
+
+```{r sdss-spectra, fig.cap = "Spectra image of wavelength against flux from SDSS from two sources.
+Part A shows the example image taken from SDSS website.
+Part B shows the replicated figure using data provided as CSV, demonstrating great reproducibility."}
 
@@
 
-A great litmus test of reproducibility for this is the fact that from downloading the CSV from the website, a similar data visualisation can be produced, seen in Figure \@ref(fig:sdss-spectra)B:
+Metadata from the data cannot be contained in the CSV file, but is available in the URL the csv was downloaded from:
 
-The data was downloaded from: http://dr15.sdss.org/optical/spectrum/view/data/format=csv?plateid=1678&mjd=53433&fiberid=425&reduction2d=v5_7_0
+```
+http://dr15.sdss.org/optical/spectrum/view/data/format=csv?
+plateid=1678&mjd=53433&fiberid=425&reduction2d=v5_7_0
+```
 
-This URL contains details of the data, specifically:
+Which contains details of the data, specifically:
 
@@
 
-* plateid is the plate that they use to explore a given section of night sky. It uniquely identifies a "plug plate" used to eollcet the data
-* mjd is an integer that marks the "modified julian date" of the night when observations was carried out.
-* fiberid is an integer that marks the fiber number used.
-* The version marked as `v5_7_0` is the pipeline version used for BOSS in DR12
+* **plateid** is the plate that they use to explore a given section of night sky. It uniquely identifies a "plug plate" used to eollcet the data
+* **mjd** is an integer that marks the "modified julian date" of the night when observations was carried out.
+* **fiberid** is an integer that marks the fiber number used (the fiber optic cable used to transmit light to the sensor).
+* The **version** marked as `v5_7_0` is the pipeline version used for BOSS in DR12.
+
+As a researcher who does not work in astrophysics, this information on the data was quite easy to retrieve from the SDSS website section called "Understanding SDSS spectroscopic data": https://www.sdss.org/dr12/spectro/spectro_basics/.
+
+The SDSS provides data at different levels of complexity for different degrees of understanding.
+It is a huge project involving many people, and a staggering effort to prepare.
+Despite this, it is still very easy to understand.
+This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-understanding)).
+The impact of this data is high, having changed the way we understand the past, present and future of the universe itself.
+This impact is surely due to the care and thought put into making the data accessible and public.
+It is a statement of what is possible with good data sharing practices.
 

### 2019-06-11 paper.Rmd

--- 2019-05-15 paper.Rmd
@@
 
+To demonstrate the variety of datasets and documentation practices, we now discuss three datasets of varying quality.
+
 ## Dataset 1: Forest Census Plot on Barro Colorado Island
@@
 
-The SDSS provides different degrees of understanding.
-It is a huge project involving many people, and requiring a staggering and enormous effort to prepare.
+As an example, the optical and spectral data can be accessed in FITS data format, or even as a plain CSV, the first few rows shown in Table \@ref(tab:spectra-tab).
+An example plot in Figure \@ref(fig:sdss-spectra)A shows the Wavelength against Flux as produced on the SDSS website, while Figure \@ref(fig:sdss-spectra)B.
+The fact that figure Figure \@ref(fig:sdss-spectra)A is reproduced in Figure \@ref(fig:sdss-spectra)B demonstrates great reproducibility.
+
+```{r prepare-spectra-data, include = FALSE}
+
+sdss_image_path <- here::here("figures", "sdss-spectra-example-plate-1678.png")
+library(magick)
+library(ggplot2)
+library(cowplot)
+
+spectra <- readr::read_csv(
+  here::here("data",
+             "plateid-1678_mjd-53433_fiberid-425_reduction2d-v5_7_0.csv")
+)
+
+gg_spectra <- ggplot(spectra,
+                     aes(x = Wavelength,
+                         y = Flux)) +
+  geom_line() + 
+  theme_minimal() + 
+  theme(aspect.ratio = 1)
+
+plot_spectra <- ggdraw() + 
+  draw_image(sdss_image_path)
+
+```
+
+```{r spectra-tab, echo = FALSE}
+library(kableExtra)
+head(spectra) %>%
+  knitr::kable(caption = "Example spectra data from SDSS, showing values for Wavelength, Flux, BestFit, and SkyFlux.",
+               booktabs = TRUE) %>%
+  kable_styling(latex_options = c("striped",
+                                  "hold_position"))
+
+```
+
+```{r sdss-spectra, fig.cap = "Spectra image of wavelength against flux from SDSS from two sources.
+Part A shows the example image taken from SDSS website.
+Part B shows the replicated figure using data provided as CSV, demonstrating great reproducibility."}
+
+plot_grid(plot_spectra, gg_spectra, labels = "AUTO")
+```
+
+Metadata from the data cannot be contained in the CSV file, but is available in the URL the csv was downloaded from:
+
+```
+http://dr15.sdss.org/optical/spectrum/view/data/format=csv?
+plateid=1678&mjd=53433&fiberid=425&reduction2d=v5_7_0
+```
+
+Which contains details of the data, specifically:
+
+* plateid = 1678
+* mjd = 53433
+* fiberid = 425
+* reduction2d = v5.7.0
+
+These are unique identifying features of the data, which tell us:
+
+* **plateid** is the plate that they use to explore a given section of night sky. It uniquely identifies a "plug plate" used to eollcet the data
+* **mjd** is an integer that marks the "modified julian date" of the night when observations was carried out.
+* **fiberid** is an integer that marks the fiber number used (the fiber optic cable used to transmit light to the sensor).
+* The **version** marked as `v5_7_0` is the pipeline version used for BOSS in DR12.
+
+As a researcher who does not work in astrophysics, this information on the data was quite easy to retrieve from the SDSS website section called "Understanding SDSS spectroscopic data": https://www.sdss.org/dr12/spectro/spectro_basics/.
+
+The SDSS provides data at different levels of complexity for different degrees of understanding.
+It is a huge project involving many people, and a staggering effort to prepare.
 Despite this, it is still very easy to understand.
@@
 It is a statement of what is possible with good data sharing practices.
-
 ## Dataset 3: Most other datasets

### 2019-06-11 paper.Rmd
add slugs to headings so they can be referenced as sections in text. Fixes minor reference bug and moves something to morgue

@@
 
-# Introduction
+# Introduction {#intro}
 
@@
 Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors [@Rowhani-Farid2016; @Stodden2018].
-Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [Popkin2019; @Kirk2019].
+Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [@ Popkin2019; @Kirk2019].
 But there is also a component of benefit to the broader scientific community.
@@
 
-# Challenges in documenting your dataset
+# Challenges in documenting your dataset {#doco-challenge}
 
@@
 
-# Releasing data in R
+# Releasing data in R {#data-release-r}
 
@@
 
-# Data Sharing in Julia and Python
+# Data Sharing in Julia and Python {#julia-python}
 
@@
 
-# Dealing with medium to large data
+# Dealing with medium to large data {#med-large}
 
@@
 
-# Publishing and repositories
+# Publishing and repositories {#publish-repos}
 
@@
 
-# What belongs in the minimal structure for researchers
+# What belongs in the minimal structure for researchers {#minimal}
 
@@
 
-## README: A Human readable description of the data
+## README: A Human readable description of the data {#readme}
 
@@
 
-## Codebook: Human readable dictionary of data contents
+## Codebook: Human readable dictionary of data contents {#codebook}
 
@@
 
-## License: How to use and share the data
+## License: How to use and share the data {#license}
 
@@
 
-## Citation: How you want your data to be cited
+## Citation: How you want your data to be cited {#cite}
 
@@
 
-## Machine readable metadata
+## Machine readable metadata {#machine-metadata}
 
@@
 
-## Raw data: The original/first data provided
+## Raw data: The original or first data provided {#raw-data}
 
@@
 
-## Scripts: To clean raw data ready for analysis
+## Scripts: To clean raw data ready for analysis {#scripts}
 
@@
 
-## Analysis ready data: Final data used in analysis
+## Analysis ready data: Final data used in analysis {#final-data}
 
@@
 
-# Tooling for packaging data
+# Tooling for packaging data {#tooling}
 
@@
 
-# Example datasets
+# Example datasets {#example-data}
 
@@
 
-## Dataset 1: Forest Census Plot on Barro Colorado Island
+## Dataset 1: Forest Census Plot on Barro Colorado Island {#ex-data-1}
 
@@
 
-## Dataset 2: Sensor data
+## Dataset 2: Sensor data {#sdss-data}
 
@@
 It is a statement of what is possible with good data sharing practices.
-## Dataset 3: Most other datasets
+
+## Dataset 3: Most other datasets {#ex-data-3}
 
@@
 
-## Summary
+## Summary {#ex-data-summary}
 
@@
 
-# Ten simple rules for publishing data
+# Ten simple rules for publishing data {#10-rules}
 
@@
 
-# Conclusions
+# Conclusions {#conclusions}
 
@@
 
-# Morgue
-
-One language agnostic solution is to use the Open Knowledge Foundations Frictionless data spec, implemented in both R and Python.
-
 # Literature Cited

### 2019-06-13 paper.Rmd
Expanded on 10 simple rules

@@
 
-1.
-**Decide whether publishing is appropriate for your data** It is important for everyone to publish their data in a research data archive to ensure long-term availability, curated metadata, and that they are preserved in the future.
-However this step might not be necessary depending on the stage and purpose of your project.
-In cases where this step is cumbersome, ensuring that the data are readily available by other means.
+1. **Decide whether publishing is appropriate for your data** 
+It is critical to publish your data if they are the basis for a peer-reviewed research publication.
+To be broadly useful, your data must be deposited in a permanent archive (to ensure that it does not disappear from a ephemeral location such as your university website), along with useful metadata so that someone outside the project team can understand what the data mean.
+However, this level of effort is not always critical for all data science efforts.
+For small-scale projects such as ones where one might generate simulated datasets, this level of curation is highly unnecessary.
+Here making data available in a transient location like GitHub or a website is sufficient.
+
 2.
 **Include a README** file with your data archive.
-README files have a long history dating back to the 1970s in the context of free software and are a requirement of the GNU standard.
-In a more generic sense, README files are a form of simple documentation that describe the contents of a directory, along with brief descriptions, the type of license used and information on how to cite the collection.
+README files have a long history in software (https://medium.com/@NSomar/readme-md-history-and-components-a365aff07f10) and are named so that the ASCII systems capital letters filename would show this file first, making it the obvious place to put relevant information.
+In the context of datasets, READMEs are particularly useful when there are no reliable standards.
+The best practice here is to have one readme per dataset, with names that easily associate with the corresponding data files.
+In addition to metadata about the data such as variable names, units of measurement and code books, the readme should also contain information on licenses, authors/copyright, a title, and dates and locations where the data were collected.
+Additionally keywords, author identifiers such as ORCID, and funder information would be useful.
+
 3.
@@
 4.
-**Provide a machine readable format** for your data that describes what each column contains.
-This can be a json schema file, which also helps websites like Google Dataset Search automatically discover your data.
-5. **Provide the data in its rawest form** in a folder called "data-raw". This
- ensures provenance of your data by keeping a master copy from which all
-other tidied up data can be created.
-6. **Provide [open source?] scripts** used to clean data from rawest form into tidy/analysis ready format. These describe the steps taken to prepare the data,
-which helps explain and document any decisions taken during the cleaning phase.
+**Provide a machine readable format** for your data.
+When possible, provide machine readable metadata that map on to the open standards of [schema.org](https://schema.org/) and [JSON-LD](https://en.wikipedia.org/wiki/JSON-LD).
+These metadata provide all of the information described in the readme best practices (rule 2), but in a machine readable way and include much of the same information such as name, description, spatial/temporal coverage etc.
+One way to test if your metadata are machine readable is to use Google's structured testing data (https://search.google.com/structured-data/testing-tool/u/0/) to verify the integrity of the metadata.
+
+5.
+**Provide the data in its rawest form** in a folder called "data-raw".
+Keeping your data raw (also sometimes referred to as the sushi principle) is the safest way to ensure that your analysis can be reproduced from scratch.
+This approach not only lets you trace the provenance of any analysis but it also ensures further use of the data that a derived dataset may prevent.
+
+6.
+**Provide [open source?] scripts** used to clean data from rawest form into tidy/analysis ready format.
+raw data is often unusable without further processing (data munging or data cleaning), which are the steps necessary to detect and clean inaccurate, incomplete records, and missing values from a dataset.
+While datasets can be cleaned interactively, this approach is often very difficult to reproduce.
+It is a better practice to use scripts to batch process data cleaning.
+Scripts can be verified, version controlled and rerun without much overhead when mistakes are uncovered.
+These scripts describe the steps taken to prepare the data, which helps explain and document any decisions taken during the cleaning phase.
 These should ideally operate on some raw data stored in the "data-raw" folder (rule 5).
+
 7.
 **Keep additional copies in more accessible locations**: Even if you archive the data somewhere long-term, keep a second (or additional copies) in locations that are more readily and easily accessible for data analysis workflows.
-These include services like GitHub, GitHub LFS and others where fast CDNs make access and reuse practical, but fall back to Zenodo or similar in case these services were to go away.
+These include services like GitHub, GitHub LFS and others where fast content delivery networks (CDNs) make access and reuse practical.
+In the event that these services shut down or become unavailable, a long-term archival copy can still be accessed from a permanent repository such as Zenodo or Dryad and populated into a new data hosting service or technology.
+
 8.
 **Use a hash function like MD5 checksum** to ensure that the data you deposited are the same that are being made available to others.
-Hash functions not only remove redundancy by removing the need to download the same data again, but also help ensure that the data are valid.
+Hash values such as MD5 are short numeric values that can serve as digital signatures for large amounts of data.
+By sharing the hash in a README, authors can ensure that the data, particularly the version of data being used by the reader is the same.
+
 9.
@@
 Even then a citation may not accrue references, but without a DOI, it is guaranteed not to.
+
 10.
-Store a copy of your data in simple formats such as CSV: Plain text CSV files follow a flat and extremely simple schema and can be opened by any text editor.
-These files can also be easily parsed by any data science programming language.
-Modern software packages allow for fast reads without consuming too much memory and these tools also allow for easy modification of the data.
+**Stick with simple data formats** to ensure long-term usefulness of datasets in a way that is not tied to transient technologies (such as ever changing spreadsheet standards), store the data as plain text where possible.
+This can take the form of comma/tab separated files in most cases.
+This will ensure transparency and future proof your data.
 

### 2019-06-13 paper.Rmd
update codebook section to include Arslan article on codebook package, and also add dataspice

--- 2019-06-11 paper.Rmd
@@
 
-Codebooks provide human readable description of the variables in the data and contain information on:
+> A high-level summary, ideally combined with text explaining the structure and nature of the data set, also helps to explain an unfamiliar data set to researchers who want to reproduce analyses or reuse the data.
+
+Codebooks provide human readable description of the data, providing context on the nature and structure of the data.
+This helps someone not familiar with the data understand, and use the data.
+At a minimum they should contain:
 
@@
 **Variable labels** are longer descriptions of variables.
-For example "University Job Position", "University Faculty Position Level", and  "Number of Years Worked at University" [@McGill-codebook].
+For example "University Job Position", "University Faculty Position Level", and  "Number of Years Worked at University" [@McGill-codebook; @Arslan2019].
 **Variable codes** apply to categorical (factor) variables, and are the values for their contents.
@@
 Tooling for producing these this information speeds up the process of sharing data.
-To help create codebooks, there are R packages such as `dataMeta`, `memisc`.
-and `codebook`.
+Machine-readable metadata that can be indexed by google is created using the `dataspice` package [@dataspice].
+To help create codebooks, we recommend the `codebook` package in R [@Arslan2019], which also generates machine readable metadata.
 Codebooks are implemented in other software such as STATA, which provides a "codebook" command.

### 2019-06-13 paper.Rmd
Pushed draft of conclusions

--- 2019-06-13 paper.Rmd
@@
 
-TBD
+# Conclusions
+
+The open science literature has been advocating the importance of data sharing for a long time.
+However many of these articles appeal to the broader benefits of data sharing but rarely talk about the practical considerations of making data available alongside code.
+Trivial reproducibility has always relied up on the idea that as long as code exists on a platform such as GitHub and clearly lists open source dependencies, one should be able to run the code at a future time.
+Containerization efforts such as Docker have made it even easier to capture system dependencies and rapidly launch containers.
+But true reproducibility requires not just code but also data.
+
+Over the years researchers have made data available in a multitude of unreliable ways, including offering downloads from university websites, FTP repositories, and password protected servers, all of which are prone to bit rot.
+In rare cases where the data are deposited in a permanent archive, insufficient metadata and missing data processing scripts have made it harder and time intensive to map these to code.
+Our goal here is to describe a variety of ways in which a researcher can ship data, analogous to how easy it has become to push code to services like GitHub.
+The choice of technology and method all depends on the nature of the data, size, and complexity.
+Once the issue of access is solved, we also discuss how much metadata to include in order to make the data useful.
+Ultimately there is a tradeoff to consider on a case by case basis.
+Long-term studies with complex protocols require structured metadata which more standardized datasets such as those coming from sensors require far less effort.
+
+The key message for the reader is that accessing data for reproducibility should come with minimal barriers.
+Having users jump through data usage forms or other barriers is enough of a roadblock to dissuade users, which over time will make it hard to track the provenance of critical studies.
+For data with broad impact, one should further invest effort in documenting the metadata (either unstructured or structured depending on the complexity) and also focus on ensuring that at least one archival copy exists in a permanent archive.
+Following this continuum can ensure that more and more computational work becomes readily reproducible without unreasonable effort.
 

### 2019-06-13 paper.Rmd
Removed duplicate headings
added code images and suppress library messages

@@
 ```{r setup, echo = FALSE}
+
 knitr::opts_chunk$set(echo = FALSE,
@@
 
-library(kableExtra)
+suppressMessages(library(ggrepel))
+suppressMessages(library(tibble))
+suppressMessages(library(kableExtra))
+suppressMessages(library(magick))
+suppressMessages(library(ggplot2))
+suppressMessages(library(cowplot))
 ```
@@
 ```{r fig-effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
-library(ggplot2)
 
@@
 #   Y axis: <- Difficulty (easy - hard)
-
-library(tibble)
 
@@
                             ))
-
-library(ggrepel)
 
@@
 
-## Higher potential impact means more time**
+## Higher potential impact means more time
 
@@
 
+One basic suggested directory layout is given below in Figure \@ref(fig:code-directory).
+
+```{r code-directory, out.width = "75%", fig.align = "center", fig.cap = "Example directory layout and structure for a data repository."}
+knitr::include_graphics(here::here("figures",
+                                   "directory-structure.png"))
 ```
-proj-name/
-         |
-         | - README.md
-         | - Codebook.csv
-         | - LICENSE.md
-         | - metadata.json
-         | - raw-data/
-                     | - raw-data.csv
-                     | - 01-read-tidy-raw-data.csv
-         | - data/
-                 | - final-data.csv
-```
-
-_Figure XX. Example directory layout and structure for a data repository._
 
@@
 An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
-Say for a dataset called "demographics" with three variables (Table \@ref(tab:table-machine-readable))
+Say for a dataset called "demographics" with three variables (Table \@ref(tab:table-machine-readable)) and JSON equivalent in Figure \@ref(fig:json-code).
 
@@
 
+```{r json-code, fig.align = "center", out.width = "75%", fig.cap = "Example snippet of some Table Schema data for a dataset with three variables.
+This provides a description of each field, and the type of field, and it's description."}
+knitr::include_graphics(here::here("figures",
+                                   "json-code.png"))
 ```
-{
-  "name": "demographics",
-  # here we list the data files in this dataset
-  "resources": [
-    {
-      "path": "demographics.csv",
-      "schema": {
-        "fields": [
-          {
-            "name": "age",
-            "type": "integer"
-          },
-          {
-            "name": "height",
-            "type": "number"
-          },
-          {
-            "name": "nationality",
-            "type": "string"
-          }
-        ]
-      }
-    }
-  ]
-}
-```
-
-_Figure XX Example snippet of some Table Schema data for a dataset with three variables. This provides a description of each field, and the type of field, and it's description._
 
@@
 sdss_image_path <- here::here("figures", "sdss-spectra-example-plate-1678.png")
-library(magick)
-library(ggplot2)
-library(cowplot)
 
@@
 ```{r spectra-tab, echo = FALSE}
-library(kableExtra)
+
 head(spectra) %>%
@@
 
-# Ten simple rules for publishing data {#10-rules}
+# Ten simple rules for publishing data {#ten-rules}
 

### 2019-06-17 paper.Rmd
misc changes to section 2 on challanges/tradeoffs

@@
 Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors [@Rowhani-Farid2016; @Stodden2018].
-Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [@ Popkin2019; @Kirk2019].
+Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [@Popkin2019; @Kirk2019].
 But there is also a component of benefit to the broader scientific community.
@@
 
-A typical data scientist loads a dozen or two of these open source libraries at the top of a notebook and then rely on existing routines to rapidly read, transform, visualize, and model data.
-Each of these packages individually depend on a complex web of other packages, building upon existing work rather than re-implementing everything from scratch.
+A typical data analysis loads a dozen or two of these open source libraries at the top of a notebook and then relies on existing routines to rapidly read, transform, visualize, and model data.
+Each package depends on a complex web of other packages, building upon existing work rather than re-implementing everything from scratch.
 Working from script and a list of such dependencies, a data analyst can easily install all the necessary tools in any local or remote environment and reproduce the computation.
@@
 
-Although there are numerous public and private data repositories, none  of them function as package managers.
-Datasets are also far more diverse in complexity, size, and formats, making them particularly challenging to standardize or easily "install" where the code is running.
+Datasets are often far more diverse than code in terms of complexity, size, and formats.
+This makes them particularly challenging to standardize or easily "install" where the code is running.
+While there are numerous public and private data repositories, none of them function as package managers, which is what provides so much robustness to code sharing.
 As a result, data used in an analysis is often read from various locations (local, network, or cloud), various formats, varying levels of tidiness [@Wickham2014].
@@
 
+# Challenges in documenting your dataset {#doco-challenge}
+
+There are many features to consider when documenting, sharing, and publishing data.
+What format to store the data in?
+How do you describe the variables, observations, and collection methods?
+Do you include raw data?
+What about data used in analysis?
+What about metadata for the variables?
+And then there's linceses, and long term storage.
+It can be hard to know where to start.
+This section discusses challenges that come with these features of data documenting, sharing, and publication, setting up the discussion for the rest of the paper.
+
+## How far to go with data documenting
+
+To frame discussion around the challenges of data documenting, we can think of how an individual dataset fits on two features: "Effort to prepare", and "Ease of understanding", see Figure \@ref(fig:fig-effort-understanding).
+
 ```{r fig-data-sharing-workflow, fig.cap = "The steps required compared to effort required for data sharing for reproducibility."}
@@
 ```
-
-# Challenges in documenting your dataset {#doco-challenge}
-
-Providing a well documented dataset alongside research requires meeting several challenges.
-This includes understanding all the pieces and features of documentation such as metadata, licenses, and raw data; knowing where to share the data to ensure long term storage.
-
-Better documentation requires more time. Higher impact data needs better documentation.
-
-It takes more effort to prepare data to make it easy to understand and share.
-The challenge here is understanding how much time to spend to make it easier to understand before the gains in understanding are outweighed by the time taken to improve them.
-
-There is also a challenge in understanding the impact of data.
-Data with high potential for impact should be made easier to understand and share.
-This can be seen in the research created from the 50 hectare plot, discussed in section ...
-- this dataset was challenging to understand without special domain expertise.
-
-Simpler format data like CSV is more reliable than binary serialised/compressed data (ask Laura DeCicco re this).
-
-Providing the right amount of documentation for your data is challenging.
-Too much is overwhelming; too little and you don't know anything.
-For example, compare a folder with dozens of data files and several papers, to a single CSV file.
-The first is too much, and understanding it requires a high time cost.
-The second is too little; we don't know anything about what the data is or why it was collected.
-The challenge is to provide the right level of detail, somewhere between these two extremes.
-
-There are many features to consider when documenting data, such as a README file with human-readable descriptions of data and variables, data licenses, whether to include raw data, data cleaning scripts, and machine readable metadata.
-Together these all help make the data easier to understand.
-These features are not clearly described in the scientific literature in the context of including them alongside research.
-
-To frame discussion around the challenges of data documenting, we can think of two features: "Effort to prepare", and "Ease of understanding".
-We can plot these on two axes and place datasets along each to indicate the ease of understanding and preparation (see Figure \@ref(fig:fig-effort-understanding)).
 
@@
 
-## Higher potential impact means more time
+There are many features to include in data documentation.
+Not all of these are needed for every dataset.
+But, deciding which to include should be based upon the potential impact of the data.
+The more impactful the data, the more features to include.
+This is one of the first challanges (tradeoffs) to discuss: how to decide what to include with your data.
+
+Together these all help make the data easier to understand. These features are not clearly described in the scientific literature in the context of including them alongside research.
+
+## Data format
+
+Data that is easy to use should avoid being in a proprietary format that requires special software to read.
+For example, a CSV should be preferred over proprietary data formats, such as SPSS or SAS custom data types.
+
+There is also the challenge for format.
+A simpler format data like CSV (Comma Separated Value) is more reliable than binary serialised or compressed data, but can sometimes be slower to read in and out, and harder to share.
 
@@
 This means effort put into preparing data documentation for a small survey to be made easy to use should be kept relatively simple and not over complicated.
+Likewise, data that can be created via simulation from open source software could arguably not be archived - since it can be generated from scratch with code; a reproducible process with likely very low cost in person time to create.
 
@@
 
-Data that is easy to use should avoid being in a proprietary format that requires special software to read.
-For example, a CSV should be preferred over proprietary data formats, such as SPSS or SAS custom data types.
+High impact data should be easy to understand and use, but how do you assess its impact?
 
-## Understanding tradeoffs
+## Describing the variables and metadata
 
-To help explain these descriptive qualities of the data sharing, we discuss three datasets and their relationship of **Effort to prepare**, and **Ease of understanding** in section 5.
+## Including raw data
 
-One of the easiest solutions to making data available alongside an analysis is to include it in the same version controlled repository as the analysis scripts.
-This works well with the data file sizes are in the order of bytes to megabytes.
-When a reader clones the repository, the scripts and notebooks will be able to read all data files from relative paths with no further modifications to the code itself.
-Assuming the data would be useful beyond the context of the analysis, it would be beneficial to also archive it in a long-term repository.
+## Data used in analysis
+
+## Licenses
+
+## Long term storage
+
+## Just how much to include
+
+Another challenge is in the gross total amount of documentation.
+Too much documentation can be overwhelming; too little and you don't know anything.
+For example, compare a folder with dozens of data files and several papers, to a single CSV file.
+The first is too much, and understanding it requires a high time cost.
+The second is too little; we don't know anything about what the data is or why it was collected.
+The challenge is to provide the right level of detail, somewhere between these two extremes.
+This changes depending on the impact of the data, and the task at hand.
+
+Creating good documentation has the same challenge as doing good writing: it takes time, and it can be hard to know when you are done.
 
@@
 
-# Conclusions
-
 The open science literature has been advocating the importance of data sharing for a long time.
@@
 
+# Morgue
+
+## Challenges -  Understanding tradeoffs
+
+The research created from the 50 hectare plot, discussed in section \@ref(ex-data-1) - this dataset was challenging to understand without special domain expertise.
+
+Ultimately, creating good documentation has the same challenge as good writing: it takes time, and it can be hard to know when you are done.
+
+It can be hard to find the right amount of time to spend. Ideally, you want the gains in understanding not to be outweighed by the time taken to improve them.
+
+Impact is a difficult quantity to measure.
+
+To help explain these descriptive qualities of the data sharing, we discuss three datasets and their relationship of **Effort to prepare**, and **Ease of understanding** in section 5.
+
+One of the easiest solutions to making data available alongside an analysis is to include it in the same version controlled repository as the analysis scripts.
+This works well with the data file sizes are in the order of bytes to megabytes.
+When a reader clones the repository, the scripts and notebooks will be able to read all data files from relative paths with no further modifications to the code itself.
+Assuming the data would be useful beyond the context of the analysis, it would be beneficial to also archive it in a long-term repository.
+
 # Literature Cited

### 2019-06-17 paper.Rmd
more tinkering with the paper

@@
 
-There are many features to consider when documenting, sharing, and publishing data.
-What format to store the data in?
-How do you describe the variables, observations, and collection methods?
-Do you include raw data?
-What about data used in analysis?
-What about metadata for the variables?
-And then there's linceses, and long term storage.
-It can be hard to know where to start.
-This section discusses challenges that come with these features of data documenting, sharing, and publication, setting up the discussion for the rest of the paper.
+There are many features to include with data alongside publications.
+However, not all of these are needed for every dataset.
+This section discusses how to think about the degree of documentation to include.
+To frame discussion around the challenges of data documenting, we can think of how an individual dataset falls on two features: "Effort to prepare", and "Ease of understanding" in Figure \@ref(fig:fig-effort-understanding).
 
-## How far to go with data documenting
-
-To frame discussion around the challenges of data documenting, we can think of how an individual dataset fits on two features: "Effort to prepare", and "Ease of understanding", see Figure \@ref(fig:fig-effort-understanding).
-
-```{r fig-data-sharing-workflow, fig.cap = "The steps required compared to effort required for data sharing for reproducibility."}
-knitr::include_graphics("figures/data_sharing_workflow.png")
-```
-
-```{r fig-effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
+```{r fig-effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT).", out.width = "90%"}
 
@@
 
-dat_prep <- tibble(prep = c(15,20,25,95, 96),
-               understanding = c(15,20,95,95,96),
+dat_prep <- tibble(prep = c(15,20,25,90, 96, 98),
+               understanding = c(15,20,95,90,96, 80),
                   label = c("Data Dump",
                             "Most studies",
-                            "Sensor",
+                            "Sensor Data",
                             "50 hectare",
-                            "RCT"
+                            "RCT",
+                            "Simulation Study"
                             ))
@@
     segment.size = 0.2,
-    segment.color = "grey50") +
-  theme_minimal() +
+    segment.color = "grey50"
+    ) +
+  # theme_void() + 
+  theme(axis.line.x = element_line(arrow = arrow(length = unit(3, 'mm'))), 
+        axis.line.y = element_line(arrow = arrow(length = unit(3, 'mm'))),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank(),
+        aspect.ratio = 1
+  ) + 
   labs(x = "Effort to Prepare",
@@
 
-There are many features to include in data documentation.
-Not all of these are needed for every dataset.
-But, deciding which to include should be based upon the potential impact of the data.
-The more impactful the data, the more features to include.
-This is one of the first challanges (tradeoffs) to discuss: how to decide what to include with your data.
+```{r data-sharing-workflow, fig.cap = "The steps required compared to effort required for data sharing for reproducibility.", out.width = "90%"}
+knitr::include_graphics("figures/data_sharing_workflow.png")
+```
 
-Together these all help make the data easier to understand. These features are not clearly described in the scientific literature in the context of including them alongside research.
-
-## Data format
-
-Data that is easy to use should avoid being in a proprietary format that requires special software to read.
-For example, a CSV should be preferred over proprietary data formats, such as SPSS or SAS custom data types.
-
-There is also the challenge for format.
-A simpler format data like CSV (Comma Separated Value) is more reliable than binary serialised or compressed data, but can sometimes be slower to read in and out, and harder to share.
+To decide the extent of data documentation to include should be based on the impact of the data. The more impactful the data, the more features to include.
 
@@
 Therefore data with high potential impact requires more time and effort to prepare.
+
+High impact data should be easy to understand and use. Impact is hard to measure as a general measure, and may vary from field to field. An example is given below.
+
 For example, data from medical randomized control trials (RCTs) on cancer treatment can have high impact, so requires a lot of time and effort document.
@@
 
+Figure \@ref(fig:data-sharing-workflow) shows the different degrees or steps of this.
+
 Tradeoffs may change over time as the cost of making them easy to prepare and understand change.
@@
 
-High impact data should be easy to understand and use, but how do you assess its impact?
-
-## Describing the variables and metadata
-
-## Including raw data
-
-## Data used in analysis
-
-## Licenses
-
-## Long term storage
-
-## Just how much to include
-
-Another challenge is in the gross total amount of documentation.
-Too much documentation can be overwhelming; too little and you don't know anything.
-For example, compare a folder with dozens of data files and several papers, to a single CSV file.
-The first is too much, and understanding it requires a high time cost.
-The second is too little; we don't know anything about what the data is or why it was collected.
-The challenge is to provide the right level of detail, somewhere between these two extremes.
-This changes depending on the impact of the data, and the task at hand.
-
 Creating good documentation has the same challenge as doing good writing: it takes time, and it can be hard to know when you are done.
+Thinking about the potential impact of should help guide the researcher in how much they should document their data.
 
@@
 
+## Challenges - Data format
+
+Data that is easy to use should avoid being in a proprietary format that requires special software to read.
+For example, a more basic format such as CSV (Comma Separated Value) should be preferred over proprietary data formats, such as SPSS or SAS custom data types, or binary formats, as it is more reliable, although it can sometimes be slower to read in and out, and harder to share due to size.
+
 ## Challenges -  Understanding tradeoffs
+
+Another challenge is in the gross total amount of documentation.
+Too much documentation can be overwhelming; too little and you don't know anything.
+For example, compare a folder with dozens of data files and several papers, to a single CSV file.
+The first is too much, and understanding it requires a high time cost.
+The second is too little; we don't know anything about what the data is or why it was collected.
+The challenge is to provide the right level of detail, somewhere between these two extremes.
+This changes depending on the impact of the data, and the task at hand.
+
+There are many features to consider when documenting, sharing, and publishing data.
+What format to store the data in?
+How do you describe the variables, observations, and collection methods?
+Do you include raw data?
+What about data used in analysis?
+What about metadata for the variables?
+And then there's lincenses, and long term storage.
+It can be hard to know where to start, but together they all help make the data easier to understand.
+These features are not clearly described in the scientific literature in the context of including them alongside research.
 

### 2019-06-17 paper.Rmd
finished draft of data documenting :tada:

@@
 **In this paper we focus on the practical side of sharing data for the purpose of reproducibility.** Our goal is to describe various methods in which an analyst can share their data with minimal friction.
-We steer clear of idealitstic ideas such as the FAIR data principles [@Wilkinson2016] since they still do not help a researcher share their data.
+We steer clear of idealistic ideas such as the FAIR data principles [@Wilkinson2016] since they still do not help a researcher share their data.
 We also skip the discussion around citation and credit because data citations are still poorly tracked and there is no standardized metric or a h-index equivalent for data as of this writing.
@@
 However, not all of these are needed for every dataset.
-This section discusses how to think about the degree of documentation to include.
-To frame discussion around the challenges of data documenting, we can think of how an individual dataset falls on two features: "Effort to prepare", and "Ease of understanding" in Figure \@ref(fig:fig-effort-understanding).
+Working out which are needed is a challenge that is not discussed in the literature.
+This section discusses a framework to use decide how much they should document their data.
+To frame discussion around the challenges of data documenting, we can think of how an individual dataset falls on two features: "Effort to prepare", and "Ease of understanding" in Figure \@ref(fig:effort-understanding).
+The ideal space to be in the graph would be the top left hand corner.
+But what we notice is that taking more effort to prepare data means that the data is easier to understand.
 
-```{r fig-effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT).", out.width = "90%"}
+```{r effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT).", out.width = "90%"}
 
@@
 
-dat_prep <- tibble(prep = c(15,20,25,90, 96, 98),
-               understanding = c(15,20,95,90,96, 80),
+dat_prep <- tibble(prep = c(15,20,55,90, 96, 40),
+               understanding = c(15,20,95,90,96, 91),
                   label = c("Data Dump",
@@
 
+Data with higher potential impact and reuse should be made as easy to understand as possible; but it also requires more time and effort to prepare.
+Impact is hard to measure, and varies from field to field, but as an example, take some data from medical randomized control trials (RCTs) on cancer treatment.
+These can have high impact, so requires a lot of time and effort document.
+Comparatively, a small survey on a few demographics can have low impact.
+Placing these on the graph above, we see they might not have a worthwhile tradeoff for ease of understanding and ease of preparation.
+This means the effort put into preparing data documentation for a small survey should be kept relatively simple, not over complicated.
+Likewise, data that can be created via simulation from open source software could arguably not be shared since it can be generated from scratch with code; a reproducible process that requires computer time, not person time to create.
+
+Deciding how much data documentation to include should be based on the data's impact.
+The more impactful the data, the more documentation features to include.
+Figure \@ref(fig:data-sharing-workflow) shows the practical types of steps that can be taken against the effort required.
+
 ```{r data-sharing-workflow, fig.cap = "The steps required compared to effort required for data sharing for reproducibility.", out.width = "90%"}
-knitr::include_graphics("figures/data_sharing_workflow.png")
+knitr::include_graphics(here::here("figures",
+                                   "data_sharing_workflow.png"))
 ```
 
-To decide the extent of data documentation to include should be based on the impact of the data. The more impactful the data, the more features to include.
-
-To improve data impact it needs more time to prepare so it is easier to understand.
-Accordingly, data with higher potential for impact and reuse should be made as easy to understand as possible.
-Therefore data with high potential impact requires more time and effort to prepare.
-
-High impact data should be easy to understand and use. Impact is hard to measure as a general measure, and may vary from field to field. An example is given below.
-
-For example, data from medical randomized control trials (RCTs) on cancer treatment can have high impact, so requires a lot of time and effort document.
-Comparatively, a small survey on simple demographics can have low impact.
-So it might not have a worthwhile tradeoff for ease of understanding and ease of preparation.
-This means effort put into preparing data documentation for a small survey to be made easy to use should be kept relatively simple and not over complicated.
-Likewise, data that can be created via simulation from open source software could arguably not be archived - since it can be generated from scratch with code; a reproducible process with likely very low cost in person time to create.
-
-Figure \@ref(fig:data-sharing-workflow) shows the different degrees or steps of this.
-
-Tradeoffs may change over time as the cost of making them easy to prepare and understand change.
-If it becomes technologically cheaper and easier to document data more rigorously in the same time as it takes to do poorly now, then there is no reason why more people cannot do this, even when the benefits are not immediately apparent.
+Documentation challenges may evolve over time as the cost of making them easy to prepare and understand change.
+For example, new technologically may automate rigorous data documentation, so thorough documentation can take the same time as it takes to do poorly now.
+In this case, there would be no reason why more people cannot do this, even when the benefits are not immediately apparent.
 So if something might appear to have low impact, it should still be made easy to understand, if it doesn't take long to prepare that documentation.
 
-Creating good documentation has the same challenge as doing good writing: it takes time, and it can be hard to know when you are done.
-Thinking about the potential impact of should help guide the researcher in how much they should document their data.
+Creating good documentation has similar challenges to good writing: it takes time, and it can be hard to know when you are done.
+Thinking about two features: 1) the impact of data, and 2) the current effort to do each step, provides guidelines for the researcher to decide how much documentation they should provide for their data.
 
@@
 .
-This data has been used to produce the most detailed, three dimensionsal maps of the universe ever made.
+This data has been used to produce the most detailed, three dimensional maps of the universe ever made.
 
@@
 
+To improve data impact it needs more time to prepare so it is easier to understand.
+
 Another challenge is in the gross total amount of documentation.
@@
 What about metadata for the variables?
-And then there's lincenses, and long term storage.
+And then there's licenses, and long term storage.
 It can be hard to know where to start, but together they all help make the data easier to understand.

### 2019-06-18 paper.Rmd
remove an extra .
minor edits up to section 7

@@
 
-One low cost and easy way to distribute data alongside compute would be to package the datasets as a data only package or as part of something larger where the methods are being implemented.
+One low cost and easy way to distribute data alongside compute is to package the datasets as a data only package or as part of something larger where the methods are being implemented.
 The R ecosystem has many noteworthy examples of data-only packages.
-One exemplar is the `nycflights13` package by Hadley Wickham [@nycflights].
-This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes.
+One exemplar is the `nycflights13` package by Hadley Wickham [@nycflights], This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes.
 The package not only provides ready to use binary data but also ships raw data (in a folder called `data-raw`) along with scripts used to clean them.
@@
 To do so, one can package raw data alongside scripts.
-Long form documentation such as Vignettes can provide further detailed discussion on the process.
-Advanced users can simply skip the raw data and scripts and proceed directly to the fast binary data which can hold a lot of data when heavily compressed.
+Long form documentation such as vignettes can provide further detailed discussion on the process.
+Users can also skip the raw data and scripts and proceed directly to the fast binary data, which can hold a lot of data when heavily compressed.
+
 2.
@@
 For cases when this size is hard to maintain, CRAN recommends data-only packages that will be rarely updated.
-For a detailed discussion on this issue and alternative approaches, see [@Brooke_Anderson2017]
+For a detailed discussion on this issue and alternative approaches, see @Brooke_Anderson2017.
 
@@
 This is why we recommend, as a rule, that researchers also archive data in a long-term data repository.
-These include domain specific repositories (see SECTION) or more general purpose ones such as Zenodo or Figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.
+These include domain specific repositories (see Section \@ref(publish-repos)) or more general purpose ones such as Zenodo or Figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.
 
@@
 
-Sharing data in a programming language happens to some extent in other languages, such as Julia and Python.
-
-Julia was established in 2012, and is a high level language with great performance.
+Sharing data in a programming language happens to some extent in other languages, such as Julia and Python.  Julia was established in 2012, and is a high level language with great performance.
 Similar to R, Julia provides packages for database access, API access, and data packages [@METADATA-jl].
-This includes access to MySQL databases, NOAA [@NOAA-jl], and actual datasets from Econometrics [@econ-data-jl], faces, [@face-data-jl], Market data [@market-data-jl], and even R datasets [@r-data-jl].
+This includes access to MySQL databases, NOAA, and actual datasets from Econometrics, faces, Market data, and even R datasets  [@NOAA-jl; @econ-data-jl; @face-data-jl; @market-data-jl; @r-data-jl].
 
@@
 
-Another common situation that researchers face is in dealing with data that fall somewhere between small (tabular data such as CSV that are in the order of bytes to megabytes to gigabytes that can be easily compressed) to large (what falls under the umbrella of `big data`).
+Another common situation that researchers face is in dealing with data that fall somewhere between small and large.
+For example, small data could be tabular, as a CSV, in the order of bytes to megabytes to gigabytes that can be easily compressed, and large could be what falls under the umbrella of **big data**.
 The happy medium is generally data that are too big to fit on memory of most standard machines, but can successfully fit on disk (https://ropensci.github.io/arkdb/articles/articles/noapi.html).
@@
 
-As for making these files available alongside compute, one ingenious but short-term solution is to use the GitHub release feature to attach such large database dumps.
+To make these files available alongside compute, one ingenious but short-term solution is to use the GitHub release feature to attach such large database dumps.
 GitHub releases are designed to serve software releases and provide the opportunity to attach binary files.
 This mechanism can also be used to attach arbitrary files such as large data files, binary data, and database dumps as long as each file does not exceed 2gb.
-The R package `piggyback` allows for uploading and downloading such files to GitHub releases, which would make it easy for anyone to access data files wherever the script is being run.
-It's worth emphasizing again that this is a short-term solution that is dependent on GitHub maintaining this service.
+The R package `piggyback` allows for uploading and downloading such files to GitHub releases, making it easy for anyone to access data files wherever the script is being run.
+We emphasize again that this is a short-term solution that is dependent on GitHub maintaining this service.
 
@@
 One can share data in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify.
-Data can be shared in numerous ways including by placing it in a repository, packaging it with methods, or by using various free tiers of commercial services.
+Data can be shared in numerous ways, including placing it in a repository, packaging it with methods, or using various free tiers of commercial services.
 However, one must publish data when appropriate.
@@
 This approach **gets the data somewhere rather than nowhere**.
-It's minimal features means it is simple to maintain.
+Its minimal features means it is simple to maintain.
 A downsides is that it does not scale to larger data.
@@
 In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data.
-For example, in astronomy the data is hosted at the SDSS (SDSS), Genetic data can be hosted at GenBank (@GenBank), and so on.
+For example, in astronomy the data is hosted at the SDSS (SDSS) or Genetic data can be hosted at GenBank (@GenBank), and so on.
 Some other places to put data include the open science framework [@OSF], Dataverse, [@dataverse], and Data World, [@dataworld].
@@
 
-The data should contain information on metadata, data dictionaries, the README, and data used in analysis (ref section 2).
-No matter where data is submitted, there should ideally be a canonical data repo in one long term archive that links to others with a DOI
+The data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)).
+No matter where data is submitted, there should ideally be a canonical data repo in one long term archive that links to others with a DOI.
 
@@
 It is not possible to summarise the contents of the data, but briefly, the SDSS data includes photometric and spectroscopic information obtained from a 2.5m telescope at Apache Point Observatory in New Mexico, operating since 2000,s producing over 200Gb of data every day [@sdss-one-paper; @sdss-four-paper; @sdss-website].
-.
 This data has been used to produce the most detailed, three dimensional maps of the universe ever made.

### 2019-06-18 paper.Rmd

@@
 
-For a piece of computational research to be minimally reproducible, it requires three distinct elements:
-
-1. Code  
-2. Computing environment  
-3. Data
-
+For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) Code;  2) Computing environment, and 3) Data.
 The first two of these challenges have largely been solved (REFERENCE).
@@
 - **Where** was it collected
+- **Why** it was collected
 - **How** is was collected
@@
 ## Codebook: Human readable dictionary of data contents {#codebook}
-
-> A high-level summary, ideally combined with text explaining the structure and nature of the data set, also helps to explain an unfamiliar data set to researchers who want to reproduce analyses or reuse the data.
 
@@
 **Variable codes** apply to categorical (factor) variables, and are the values for their contents.
-For example, 0 = no, 1 = yes, and 0 = male and 1 = female.
-These should be consistent across similar variables to avoid problems where `0 = yes` for one variable, and `1 =yes` in another.
+For example, `0 = no, 1 = yes` for one variable, and `0 = inside` and `1 = outside`.
+These should be consistent across similar variables to avoid problems where `0 = yes` for one variable, and `1 = yes` in another.
 Date variables should have consistent formatting.
@@
 
-Data dictionary tables should be placed in the README and presented as a table.
-Every data dictionary should also be provided in its raw form in the repository, for example, saved as a CSV (see Figure XX.).
+Data dictionary tables should be placed in the README and presented as a table. Every data dictionary should also be provided in its raw form in the repository, for example, saved as a CSV.
 
@@
 
-Data with a license clearly establishes rules on how everyone can modify, use, and share data. Without a license, these rules are unclear, and can lead to problems with attribution and citation.
-
-It can be overwhelming to try and the right license for a use case. Two licenses that are well suited for data sharing are:
+Data with a license clearly establishes rules on how everyone can modify, use, and share data.
+Without a license, these rules are unclear, and can lead to problems with attribution and citation.
+It can be overwhelming to try and the right license for a use case.
+Two licenses that are well suited for data sharing are:
 
@@
 
-The data dictionary provides _human readable_ information on the data.
+The README, and data dictionary provides _human readable_ information on the data.
 To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
 An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
-Say for a dataset called "demographics" with three variables (Table \@ref(tab:table-machine-readable)) and JSON equivalent in Figure \@ref(fig:json-code).
+For example, a dataset called "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), which has the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).
 
@@
 This contains fields such as path, and a schema with subfields name and type, for each variable.
-It also provides information for, licensing and features such as line breaks, and delimiters.
+It also provides information for licensing and features such as line breaks, and delimiters.
 It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format.
@@
 Ideally, the data should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, and observations in rows.
-Contrasting `raw data`, `tidy data`/`analysis data` should be in an easily readable plain-text format, such as CSV, tab separated, or semicolon separated.
-Binary or proprietary formats are discouraged in favor of interoperability.
+Contrasting "raw data", "analysis ready data" should be in an easily readable plain-text format, such as CSV, tab separated, or semicolon separated.
+Binary or proprietary formats are discouraged in favor of interoperability, as it requires special software to read, even if it is sometimes slower to read in and out, and harder to share due to size.
 
@@
 
-Tooling for producing these this information speeds up the process of sharing data.
+Tooling for producing data documentation information speeds up the process of sharing data.
 Machine-readable metadata that can be indexed by google is created using the `dataspice` package [@dataspice].
@@
 Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that track versioning.
-Note that is different to [Frictionlessdata's tabular data package spec [@frictionlessdata-data-pkg].
+Note that is different to Frictionlessdata's tabular data package spec [@frictionlessdata-data-pkg].
 
@@
 
-To demonstrate the variety of datasets and documentation practices, we now discuss three datasets of varying quality.
+We now explore the variety of documentation practices of a few selected datasets.
 
@@
 More than a million trees have been counted.
-The data and metadata however and hard to reach.
+The data and metadata however are hard to reach.
 To obtain the data, one must fill out a form and agree to terms and conditions which require approval and authorship prior to publication.
@@
 
-Datasets obtained from sensors such as meteorological data are typically easy to prepare and understand, since sensors measure specific features.
-This means the description of data type happens upstream at the instrument-design level, and flows down to data collection.
-The telescope data from the Sloan Digital Sky Survey (SDSS) is a great example of this.
+Datasets obtained from sensors such as meteorological data are typically easy to prepare and understand.
+This is because sensors measure specific features, so the description of data type happens upstream at the instrument-design level, and flows down to data collection.
+The telescope data from the Sloan Digital Sky Survey (SDSS) is a great example of sensor data.
 
-It is not possible to summarise the contents of the data, but briefly, the SDSS data includes photometric and spectroscopic information obtained from a 2.5m telescope at Apache Point Observatory in New Mexico, operating since 2000,s producing over 200Gb of data every day [@sdss-one-paper; @sdss-four-paper; @sdss-website].
+The SDSS data includes photometric and spectroscopic information obtained from a 2.5m telescope at Apache Point Observatory in New Mexico, operating since 2000, producing over 200Gb of data every day [@sdss-one-paper; @sdss-four-paper; @sdss-website].
 This data has been used to produce the most detailed, three dimensional maps of the universe ever made.
@@
 
-```{r sdss-raw-data-screenshot, fig.cap = "Screenshot of Raw data avaialbe through DR15 FITS."}
+```{r sdss-raw-data-screenshot, fig.cap = "Screenshot of Raw data avaialbe through DR15 FITS.", out.width = "60%", fig.align = "center"}
 knitr::include_graphics(here::here("figures",
@@
 - Metadata in the machine readable standard, SCHEMA
-- Data in multiple formats (e.g., .csv and .fits)
+- Data in multiple formats (e.g., .csv and .fits, database)
 - Previously released data available in archives
-- Data is available in a database
 - Entire datasets available for download
@@
 
-As an example, the optical and spectral data can be accessed in FITS data format, or even as a plain CSV, the first few rows shown in Table \@ref(tab:spectra-tab).
-An example plot in Figure \@ref(fig:sdss-spectra)A shows the Wavelength against Flux as produced on the SDSS website, while Figure \@ref(fig:sdss-spectra)B.
-The fact that figure Figure \@ref(fig:sdss-spectra)A is reproduced in Figure \@ref(fig:sdss-spectra)B demonstrates great reproducibility.
+For example, the optical and spectral data can be accessed in FITS data format, or even as a plain CSV, the first few rows shown in Table \@ref(tab:spectra-tab).
+Wavelength plotted against Flux is shown in Figure \@ref(fig:sdss-spectra), with \@ref(fig:sdss-spectra)A showing the output from the SDSS website, and Figure \@ref(fig:sdss-spectra)B showing the output from the CSV.
+The fact that figure Figure \@ref(fig:sdss-spectra)A is virtually replicated in Figure \@ref(fig:sdss-spectra)B demonstrates great reproducibility.
 
@@
 Part A shows the example image taken from SDSS website.
-Part B shows the replicated figure using data provided as CSV, demonstrating great reproducibility."}
+Part B shows the replicated figure using data provided as CSV, demonstrating great reproducibility.", out.width = "90%", fig.align = "center"}
 
@@
 
-Metadata from the data cannot be contained in the CSV file, but is available in the URL the csv was downloaded from:
-
-```
-http://dr15.sdss.org/optical/spectrum/view/data/format=csv?
-plateid=1678&mjd=53433&fiberid=425&reduction2d=v5_7_0
-```
+Metadata from the data cannot be contained in the CSV file, but is available in the URL the csv was downloaded from: `http://dr15.sdss.org/optical/spectrum/view/data/format=csv?`
+`plateid=1678&mjd=53433&fiberid=425&reduction2d=v5_7_0`
 
@@
 
-* **plateid** is the plate that they use to explore a given section of night sky. It uniquely identifies a "plug plate" used to eollcet the data
-* **mjd** is an integer that marks the "modified julian date" of the night when observations was carried out.
-* **fiberid** is an integer that marks the fiber number used (the fiber optic cable used to transmit light to the sensor).
+* **plateid:** the unique plate used to explore a given section of night sky. 
+* **mjd:** an integer that marks the "modified julian date" of the night of observation.
+* **fiberid:** an integer of the fiber optic cable used to transmit light to the sensor.
 * The **version** marked as `v5_7_0` is the pipeline version used for BOSS in DR12.
@@
 Despite this, it is still very easy to understand.
-This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:fig-effort-understanding)).
+This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:effort-understanding)).
 The impact of this data is high, having changed the way we understand the past, present and future of the universe itself.
@@
 It is critical to publish your data if they are the basis for a peer-reviewed research publication.
-To be broadly useful, your data must be deposited in a permanent archive (to ensure that it does not disappear from a ephemeral location such as your university website), along with useful metadata so that someone outside the project team can understand what the data mean.
+To be broadly useful, your data must be deposited in a permanent archive, to ensure that it does not disappear from a ephemeral location such as your university website.
+It shoudl also contain useful metadata so that someone outside the project team can understand what the data mean.
 However, this level of effort is not always critical for all data science efforts.
@@
 In the context of datasets, READMEs are particularly useful when there are no reliable standards.
-The best practice here is to have one readme per dataset, with names that easily associate with the corresponding data files.
-In addition to metadata about the data such as variable names, units of measurement and code books, the readme should also contain information on licenses, authors/copyright, a title, and dates and locations where the data were collected.
+The best practice here is to have one README per dataset, with names that easily associate with the corresponding data files.
+In addition to metadata about the data such as variable names, units of measurement and code books, the README should also contain information on licenses, authors/copyright, a title, and dates and locations where the data were collected.
 Additionally keywords, author identifiers such as ORCID, and funder information would be useful.
@@
 When possible, provide machine readable metadata that map on to the open standards of [schema.org](https://schema.org/) and [JSON-LD](https://en.wikipedia.org/wiki/JSON-LD).
-These metadata provide all of the information described in the readme best practices (rule 2), but in a machine readable way and include much of the same information such as name, description, spatial/temporal coverage etc.
+These metadata provide all of the information described in the README best practices (rule 2), but in a machine readable way and include much of the same information such as name, description, spatial/temporal coverage etc.
 One way to test if your metadata are machine readable is to use Google's structured testing data (https://search.google.com/structured-data/testing-tool/u/0/) to verify the integrity of the metadata.
@@
 **Provide [open source?] scripts** used to clean data from rawest form into tidy/analysis ready format.
-raw data is often unusable without further processing (data munging or data cleaning), which are the steps necessary to detect and clean inaccurate, incomplete records, and missing values from a dataset.
+Raw data is often unusable without further processing (data munging or data cleaning), which are the steps necessary to detect and clean inaccurate, incomplete records, and missing values from a dataset.
 While datasets can be cleaned interactively, this approach is often very difficult to reproduce.
@@
 A citation only makes sense when your data has a DataCite compliant DOI, which is automatically provided when data is published in repositories like Zenodo and Dryad.
-Even then a citation may not accrue references, but without a DOI, it is guaranteed not to.
+Although a citation may not accrue references, without a DOI it is guaranteed not to.
 
 10.
-**Stick with simple data formats** to ensure long-term usefulness of datasets in a way that is not tied to transient technologies (such as ever changing spreadsheet standards), store the data as plain text where possible.
+**Stick with simple data formats** to ensure long-term usefulness of datasets in a way that is not tied to transient technologies (such as ever changing spreadsheet standards), store the data as plain text (e.g., CSV) where possible.
 This can take the form of comma/tab separated files in most cases.
@@
 The open science literature has been advocating the importance of data sharing for a long time.
-However many of these articles appeal to the broader benefits of data sharing but rarely talk about the practical considerations of making data available alongside code.
-Trivial reproducibility has always relied up on the idea that as long as code exists on a platform such as GitHub and clearly lists open source dependencies, one should be able to run the code at a future time.
-Containerization efforts such as Docker have made it even easier to capture system dependencies and rapidly launch containers.
-But true reproducibility requires not just code but also data.
+Many of these articles appeal to the broader benefits of data sharing, but rarely talk about the practical considerations of making data available alongside code.
+Trivial reproducibility has always relied upon the idea that as long as code exists on a platform such as GitHub, and clearly lists open source dependencies, one should be able to run the code at a future time.
+Containerization efforts such as Docker have made it even easier to capture system dependencies, and rapidly launch containers.
+But true reproducibility requires not just code, but also data.
 
@@
 Our goal here is to describe a variety of ways in which a researcher can ship data, analogous to how easy it has become to push code to services like GitHub.
-The choice of technology and method all depends on the nature of the data, size, and complexity.
+The choice of technology and method all depends on the nature of the data, and its size and complexity.
 Once the issue of access is solved, we also discuss how much metadata to include in order to make the data useful.
 Ultimately there is a tradeoff to consider on a case by case basis.
-Long-term studies with complex protocols require structured metadata which more standardized datasets such as those coming from sensors require far less effort.
+Long-term studies with complex protocols require structured metadata, while more standardized datasets such as those coming from sensors require far less effort.
 
@@
 
-# Morgue
-
-## Challenges - Data format
-
-Data that is easy to use should avoid being in a proprietary format that requires special software to read.
-For example, a more basic format such as CSV (Comma Separated Value) should be preferred over proprietary data formats, such as SPSS or SAS custom data types, or binary formats, as it is more reliable, although it can sometimes be slower to read in and out, and harder to share due to size.
-
-## Challenges -  Understanding tradeoffs
-
-To improve data impact it needs more time to prepare so it is easier to understand.
-
-Another challenge is in the gross total amount of documentation.
-Too much documentation can be overwhelming; too little and you don't know anything.
-For example, compare a folder with dozens of data files and several papers, to a single CSV file.
-The first is too much, and understanding it requires a high time cost.
-The second is too little; we don't know anything about what the data is or why it was collected.
-The challenge is to provide the right level of detail, somewhere between these two extremes.
-This changes depending on the impact of the data, and the task at hand.
-
-There are many features to consider when documenting, sharing, and publishing data.
-What format to store the data in?
-How do you describe the variables, observations, and collection methods?
-Do you include raw data?
-What about data used in analysis?
-What about metadata for the variables?
-And then there's licenses, and long term storage.
-It can be hard to know where to start, but together they all help make the data easier to understand.
-These features are not clearly described in the scientific literature in the context of including them alongside research.
-
-The research created from the 50 hectare plot, discussed in section \@ref(ex-data-1) - this dataset was challenging to understand without special domain expertise.
-
-Ultimately, creating good documentation has the same challenge as good writing: it takes time, and it can be hard to know when you are done.
-
-It can be hard to find the right amount of time to spend. Ideally, you want the gains in understanding not to be outweighed by the time taken to improve them.
-
-Impact is a difficult quantity to measure.
-
-To help explain these descriptive qualities of the data sharing, we discuss three datasets and their relationship of **Effort to prepare**, and **Ease of understanding** in section 5.
-
-One of the easiest solutions to making data available alongside an analysis is to include it in the same version controlled repository as the analysis scripts.
-This works well with the data file sizes are in the order of bytes to megabytes.
-When a reader clones the repository, the scripts and notebooks will be able to read all data files from relative paths with no further modifications to the code itself.
-Assuming the data would be useful beyond the context of the analysis, it would be beneficial to also archive it in a long-term repository.
-
 # Literature Cited

### 2019-06-21 paper.Rmd
Took a pass through and added more refs and edits

@@
 This primarily comes in the form of potential for reuse in other contexts along use for training and teaching (@McKiernan2016).
-Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing doesn't put the authors at a competitive disadvantage, sharing data will always have a net positive benefit.
+Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing doesn't put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
 First and foremost, sharing data along with other artifacts can make it easier for a reader to independently reproduce the results, thereby increasing transparency and trust in the work.
@@
 For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) Code;  2) Computing environment, and 3) Data.
-The first two of these challenges have largely been solved (REFERENCE).
+The first two of these challenges have largely been solved [@Poisot2019].
 
@@
 These packages form the building blocks of a data scientists daily workflow.
+In a modern computational workflow it is trivial for a user to `pip install` a package in Python, or use `install.packages` in R to install all software dependencies for a project.
+By relying on the idea of having research compendia [@gentleman2007statistical] or something as simple as a requirements file, one can easily install all the necessary scaffolding.
+Data on the other hard are rarely accessible that easily.
 
@@
 We discuss how to share and/or publish data and cover various tradeoffs when deciding how much to do.
-We skip detailed discussions of the minutiae of data preparation (covered elsewhere in various articles), licenses (there are not many options when it comes to data), or citation (as of this writing, data citation is still in its infancy).
+We skip detailed discussions of the minutiae of data preparation (TODO: We need to cite these articles), licenses (there are not many options when it comes to data), or citation (as of this writing, data citation is still in its infancy).
 We also analyze the state of data contained inside software packages, shared and made available as part of modern data analysis.
@@
 
-Sharing data in a programming language happens to some extent in other languages, such as Julia and Python.  Julia was established in 2012, and is a high level language with great performance.
+Sharing data in a programming language happens to some extent in other languages, such as Julia and Python.
+Julia was established in 2012, and is a high level language with great performance [@bezanson2012julia].
 Similar to R, Julia provides packages for database access, API access, and data packages [@METADATA-jl].
@@
 This mechanism can also be used to attach arbitrary files such as large data files, binary data, and database dumps as long as each file does not exceed 2gb.
-The R package `piggyback` allows for uploading and downloading such files to GitHub releases, making it easy for anyone to access data files wherever the script is being run.
+The R package `piggyback` (`https://docs.ropensci.org/piggyback/`) allows for uploading and downloading such files to GitHub releases, making it easy for anyone to access data files wherever the script is being run.
 We emphasize again that this is a short-term solution that is dependent on GitHub maintaining this service.

### 2019-06-21 paper.Rmd
Added Ellis and Woo references

@@
 We discuss how to share and/or publish data and cover various tradeoffs when deciding how much to do.
-We skip detailed discussions of the minutiae of data preparation (TODO: We need to cite these articles), licenses (there are not many options when it comes to data), or citation (as of this writing, data citation is still in its infancy).
+We skip detailed discussions of the minutiae of data preparation [@Broman2017; @Ellis2017], or citation (as of this writing, data citation is still in its infancy).
 We also analyze the state of data contained inside software packages, shared and made available as part of modern data analysis.

### 2019-06-26 paper.Rmd
add bioconductor description, close #17

@@
 For a detailed discussion on this issue and alternative approaches, see @Brooke_Anderson2017.
+
+Unlike CRAN, bioconductor, does not have a 5 megabyte limit for package size.
+They have a strict specification for how to organise genomic data in a package, so they can be used for data analysis out of the box.
+This is a robust solution, but only works for homogeneous data as found with bioinformatics.
+For example, data in their ExperimentData section (bioconductor.org/packages/3.9/data/experiment) can be used in a reliable way.
+However, such a strict standard would be impossible to enforce or scale on a repository as heterogeneous as CRAN.
+
+However, these packages are still generally used for demonstration or testing purposes, over generating new knowledge for papers.
 

### 2019-06-28 paper.Rmd
Added Binder reference

@@
 The computing environment is also easily captured with modern tools such as Docker [@Boettiger2015; @Jupyter2018].
-Modern tools such as Binder (citation) can parse Docker files and dependency trees to provide on demand, live notebooks in R and Python that a reader can immediately execute in the browser without dealing with the challenges of local installation.
+Modern tools such as Binder [@Jupyter2018] can parse Docker files and dependency trees to provide on demand, live notebooks in R and Python that a reader can immediately execute in the browser without dealing with the challenges of local installation.
 This makes it simple to load a specific environment to run any analysis.

### 2019-08-19 paper.Rmd
Nick's comments addressing Anna's feedback.

@@
 Working out which are needed is a challenge that is not discussed in the literature.
-This section discusses a framework to use decide how much they should document their data.
+This section discusses a framework for users, to help them decide how much they should document their data.
 To frame discussion around the challenges of data documenting, we can think of how an individual dataset falls on two features: "Effort to prepare", and "Ease of understanding" in Figure \@ref(fig:effort-understanding).
@@
 Impact is hard to measure, and varies from field to field, but as an example, take some data from medical randomized control trials (RCTs) on cancer treatment.
-These can have high impact, so requires a lot of time and effort document.
+These can have high impact, so requires a lot of time and effort to document.
 Comparatively, a small survey on a few demographics can have low impact.
@@
 
+Creating good documentation has similar challenges to good writing: it takes time, and it can be hard to know when you are done.
+Thinking about two features: 1) the impact of data, and 2) the current effort to do each step, provides guidelines for the researcher to decide how much documentation they should provide for their data.
+
 Documentation challenges may evolve over time as the cost of making them easy to prepare and understand change.
@@
 So if something might appear to have low impact, it should still be made easy to understand, if it doesn't take long to prepare that documentation.
-
-Creating good documentation has similar challenges to good writing: it takes time, and it can be hard to know when you are done.
-Thinking about two features: 1) the impact of data, and 2) the current effort to do each step, provides guidelines for the researcher to decide how much documentation they should provide for their data.
 
@@
 One can share data in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify.
-Data can be shared in numerous ways, including placing it in a repository, packaging it with methods, or using various free tiers of commercial services.
+Data can be shared in numerous ways, including placing it in a repository, packaging it with methods, or using various free tiers of commercial services (such as dropbox or google drive).
 However, one must publish data when appropriate.
@@
 Its minimal features means it is simple to maintain.
+In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organise.
 A downsides is that it does not scale to larger data.
-Self hosting the data is an option, but we discourage this, as it may succumb to bit rot.
+Self hosting the data is an option, but we discourage this, as it may succumb to bit rot -.
 
 In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data.
-For example, in astronomy the data is hosted at the SDSS (SDSS) or Genetic data can be hosted at GenBank (@GenBank), and so on.
+For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank), and so on.
 Some other places to put data include the open science framework [@OSF], Dataverse, [@dataverse], and Data World, [@dataworld].
@@
 Zenodo is able to accommodate larger file sizes upon request.
-
-**Principles for all three**
-
-The data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)).
-No matter where data is submitted, there should ideally be a canonical data repo in one long term archive that links to others with a DOI.
 
@@
 
-# What belongs in the minimal structure for researchers {#minimal}
+# What is the minimal structure to share data for analysis {#minimal}
+
+**Principles for all three**
+
+To share data analysis data, there should be some minimal set of requirements.
+For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)).
+No matter where data is submitted, there should ideally be a canonical data repo in one long term archive that links to others with a DOI.
 
@@
 
-The README is often the first place people will go to learn more about data. It is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":
+The README is often the first place people will go to learn more about data, and helps users navigate the repository.
+It is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":
 
@@
 
-**Variable names** are short, descriptive names with no spaces or special characters.
-For example, "job_position", "faculty_level", and "years_at_company".
+**Variable names** should be short, descriptive names with no spaces or special characters.
+In the most common tabular data case, these correspond to column names, for example, "job_position", "faculty_level", and "years_at_company".
 **Variable labels** are longer descriptions of variables.
@@
 **Variable codes** apply to categorical (factor) variables, and are the values for their contents.
-For example, `0 = no, 1 = yes` for one variable, and `0 = inside` and `1 = outside`.
+For example, a variable could contain answers to a question, with values and codes like so: `0 = no, 1 = yes`, or for another variable indicating indoors or outdoors: `0 = inside` and `1 = outside`.
 These should be consistent across similar variables to avoid problems where `0 = yes` for one variable, and `1 = yes` in another.
@@
 For example, all date information could be in format "YYYY-MM-DD", and this should not be mixed with "YYYY-DD-MM".
+Date formatting should be described in the variable labels.
 **Missing data** are values that should have been observed, but were not.
@@
 If the reason for missingness is known it should be recorded.
-For example, censored data, patient drop out, or measurement error can have different values, such as "unknown" or -99 [@White2013; @Broman2017].
+For example, censored data, patient drop out, or measurement error can have different values, such as "unknown", -99, or other value codes [@White2013; @Broman2017].
 
@@
 
-A DOI is a prerequisite for citation.
-When citing data, it only makes sense to cite datasets that have been deposited into a DataCite compliant repo.
-If, for example, the data are deposited in dryad or Zenodo, the best practice would be to copy the citation created by these repositories.
-Under the hood, DataCite provides the DOI.
+A Digital Object Identifier (DOI), is a unique identifier for a digital object such as a paper, poster, or software, and is a prerequisite for citation.
+For example, if the data are deposited in dryad or Zenodo, the best practice would be to copy the citation created by these repositories.
+Under the hood, the DOI is "minted" by DataCite, who provide DOIs for research outputs and data.
+This means that when citing data, it only makes sense to cite datasets that have been deposited into a DataCite compliant repository.
 If a DOI is unavailable, a citation will be meaningless, as it cannot be tracked by any means.
@@
 To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
+Machine readable metadata also provide a structure that means the data can be indexed and searched.
 An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
@@
 ```{r json-code, fig.align = "center", out.width = "75%", fig.cap = "Example snippet of some Table Schema data for a dataset with three variables.
-This provides a description of each field, and the type of field, and it's description."}
+This provides a description of each field, and the type of field."}
 knitr::include_graphics(here::here("figures",
@@
 
-This contains fields such as path, and a schema with subfields name and type, for each variable.
-It also provides information for licensing and features such as line breaks, and delimiters.
+This contains fields such as "path", which describe the file path to the data, and a schema with subfields name and type, for each variable.
+It also provides information for licensing and features such as line breaks, and delimiters, useful to ensure the data is correctly imported into a machine.
 It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format.
-Table schema is baked into formats such as csvy [@csvy], an extended `csv` format, which has additional front matter in a YAML format using Table Schema.
+Table schema is baked into formats such as csvy [@csvy], an extended `csv` format, which has additional front matter in a lightweight markup format (YAML) using Table Schema.
+
+Expand on:
+
+- Explain formats like XML, as people might come into formats that wany XML.
+- The power lies in the contrast to a csv codebook, as there is a defined, nested structure.
+- Ontologies and domain specific metadata standards such as EML might also be worth mentioning.
+- introduce the holy grail of the semantic web: JSON-LD
+- readers just been advised to make a:
+  - a csv codebook,
+  - put that in the README
+  - then make a JSON file. 
+Is there a way to discuss strategies on harmonising all these formats?
 
@@
 
-Tooling for producing data documentation information speeds up the process of sharing data.
+Tooling for producing data documentation information speeds up and simplifies the process of sharing data.
 Machine-readable metadata that can be indexed by google is created using the `dataspice` package [@dataspice].

### 2019-08-20 paper.Rmd

@@
 - introduce the holy grail of the semantic web: JSON-LD
+
 - readers just been advised to make a:
@@
 Is there a way to discuss strategies on harmonising all these formats?
+
+- we could map out how these things link together (e.g., you can turn JSON into xml with package ...)
+- make a table that describes how to do something in R, or Python, or Julia
+-
 

### 2019-08-20 paper.Rmd
Use data dictionary not codebook, close #96

@@
 1. README: A Human readable description of the data
-2. Codebook: Human readable dictionary of data contents
+2. Data dictionary: Human readable dictionary of data contents
 3. License: How to use and share the data
 4. Citation: How you want your data to be cited
-5. Machine readable meta data
+5. Machine readable meta data: Make your data searchable
 6. Raw data: The original/first data provided
@@
 1. README
-2. Codebook
+2. Data dictionary
 3. License
@@
 
-## Codebook: Human readable dictionary of data contents {#codebook}
+## Data dictionary: Human readable dictionary of data contents {#data-dictionary}
 
-Codebooks provide human readable description of the data, providing context on the nature and structure of the data.
+A data dictionary provide human readable description of the data, providing context on the nature and structure of the data.
 This helps someone not familiar with the data understand, and use the data.
@@
 **Missing data** are values that should have been observed, but were not.
-The code for missingness should be documented in the codebook, and should nominally be `NA`.
+The code for missingness should be documented in the data dictionary, and should nominally be `NA`.
 If the reason for missingness is known it should be recorded.
@@
 - Explain formats like XML, as people might come into formats that wany XML.
-- The power lies in the contrast to a csv codebook, as there is a defined, nested structure.
+- The power lies in the contrast to a csv data dictionary, as there is a defined, nested structure.
 - Ontologies and domain specific metadata standards such as EML might also be worth mentioning.
@@
 - readers just been advised to make a:
-  - a csv codebook,
+  - a csv data dictionary,
   - put that in the README
@@
 Machine-readable metadata that can be indexed by google is created using the `dataspice` package [@dataspice].
-To help create codebooks, we recommend the `codebook` package in R [@Arslan2019], which also generates machine readable metadata.
-Codebooks are implemented in other software such as STATA, which provides a "codebook" command.
+To help create a data dictionary, we recommend the `codebook` package in R [@Arslan2019], which also generates machine readable metadata.
+Data dictionaries are implemented in other software such as STATA, which provides a "codebook" command.
 Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that track versioning.
@@
 The biggest challenge with this study is that the data are stored on a personal FTP server of one of the authors.
-While the data are available in CSV and binary Rdata formats, the data storage servers do not have any metadata, README files or codebooks.
+While the data are available in CSV and binary Rdata formats, the data storage servers do not have any metadata, README files or a data dictionary.
 A separate, public archive hosts the metadata (https://repository.si.edu/handle/10088/20925) in a PDF file (https://repository.si.edu/bitstream/handle/10088/20925/RoutputFull.pdf?sequence=1&isAllowed=y) that describe the fields.
@@
 3.
-**Provide a codebook** or glossary for each variable.
-Codebooks provide a free-form way to document the data, their structure and underlying details of how they were collected (https://www.emgo.nl/kc/codebook/).
+**Provide a data dictionary** or glossary for each variable.
+Data dictionaries provide a free-form way to document the data, their structure and underlying details of how they were collected (https://www.emgo.nl/kc/codebook/).
 More importantly they provide a way for future consumers of the data to understand how the variables were named, how missing data were documented, along with any additional information that might impact how the data are interpreted.

### 2019-09-02 paper.Rmd
update order of sections - resolves #97
remove Julia and Python section

@@
 How does data format and size impact availability?
-
-# Challenges in documenting your dataset {#doco-challenge}
-
-There are many features to include with data alongside publications.
-However, not all of these are needed for every dataset.
-Working out which are needed is a challenge that is not discussed in the literature.
-This section discusses a framework for users, to help them decide how much they should document their data.
-To frame discussion around the challenges of data documenting, we can think of how an individual dataset falls on two features: "Effort to prepare", and "Ease of understanding" in Figure \@ref(fig:effort-understanding).
-The ideal space to be in the graph would be the top left hand corner.
-But what we notice is that taking more effort to prepare data means that the data is easier to understand.
-
-```{r effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT).", out.width = "90%"}
-
-# There are two Axis:
-#   X axis: <- Robustness (brittle - robust)
-#   Y axis: <- Difficulty (easy - hard)
-
-dat_prep <- tibble(prep = c(15,20,55,90, 96, 40),
-               understanding = c(15,20,95,90,96, 91),
-                  label = c("Data Dump",
-                            "Most studies",
-                            "Sensor Data",
-                            "50 hectare",
-                            "RCT",
-                            "Simulation Study"
-                            ))
-
-ggplot(dat_prep,
-       aes(x = prep,
-           y = understanding,
-           label = label)) +
-  geom_point() +
-  geom_label_repel(
-    segment.size = 0.2,
-    segment.color = "grey50"
-    ) +
-  # theme_void() + 
-  theme(axis.line.x = element_line(arrow = arrow(length = unit(3, 'mm'))), 
-        axis.line.y = element_line(arrow = arrow(length = unit(3, 'mm'))),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank(),
-        aspect.ratio = 1
-  ) + 
-  labs(x = "Effort to Prepare",
-       y = "Ease of Understanding") +
-  xlim(0,100) +
-  ylim(0, 100)
-
-```
-
-Data with higher potential impact and reuse should be made as easy to understand as possible; but it also requires more time and effort to prepare.
-Impact is hard to measure, and varies from field to field, but as an example, take some data from medical randomized control trials (RCTs) on cancer treatment.
-These can have high impact, so requires a lot of time and effort to document.
-Comparatively, a small survey on a few demographics can have low impact.
-Placing these on the graph above, we see they might not have a worthwhile tradeoff for ease of understanding and ease of preparation.
-This means the effort put into preparing data documentation for a small survey should be kept relatively simple, not over complicated.
-Likewise, data that can be created via simulation from open source software could arguably not be shared since it can be generated from scratch with code; a reproducible process that requires computer time, not person time to create.
-
-Deciding how much data documentation to include should be based on the data's impact.
-The more impactful the data, the more documentation features to include.
-Figure \@ref(fig:data-sharing-workflow) shows the practical types of steps that can be taken against the effort required.
-
-```{r data-sharing-workflow, fig.cap = "The steps required compared to effort required for data sharing for reproducibility.", out.width = "90%"}
-knitr::include_graphics(here::here("figures",
-                                   "data_sharing_workflow.png"))
-```
-
-Creating good documentation has similar challenges to good writing: it takes time, and it can be hard to know when you are done.
-Thinking about two features: 1) the impact of data, and 2) the current effort to do each step, provides guidelines for the researcher to decide how much documentation they should provide for their data.
-
-Documentation challenges may evolve over time as the cost of making them easy to prepare and understand change.
-For example, new technologically may automate rigorous data documentation, so thorough documentation can take the same time as it takes to do poorly now.
-In this case, there would be no reason why more people cannot do this, even when the benefits are not immediately apparent.
-So if something might appear to have low impact, it should still be made easy to understand, if it doesn't take long to prepare that documentation.
-
-# Releasing data in R {#data-release-r}
-
-One low cost and easy way to distribute data alongside compute is to package the datasets as a data only package or as part of something larger where the methods are being implemented.
-The R ecosystem has many noteworthy examples of data-only packages.
-One exemplar is the `nycflights13` package by Hadley Wickham [@nycflights], This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes.
-The package not only provides ready to use binary data but also ships raw data (in a folder called `data-raw`) along with scripts used to clean them.
-The package was originally created as a way to provide example data to teach tidy data principles and serves as a great model for how one can publish a data package in R.
-
-A more common use case is to include data as part of a regular package where analytical methods are being implemented and shared.
-This serves the dual purpose of exposing methods and data together, making it extremely easy for a researcher to simply install and load the library at the top of their script.
-CRAN's distributed network (along with the global content distribution network maintained by RStudio) ensure that the package is quickly accessible to anyone in the R ecosystem.
-A second critical advantage in this approach is that one could also benefit from R's package documentation syntax to generate useful metadata for fast reference.
-This approach can also easily be adapted to other languages.
-Python for example, is far more flexible with respect to including arbitrary files as part of a package.
-
-**Other benefits of packaging data in R**
-
-1.
-Packaging data can be a very powerful pedagogical tool to help researchers understand how to transform data and prepare it for further analysis.
-To do so, one can package raw data alongside scripts.
-Long form documentation such as vignettes can provide further detailed discussion on the process.
-Users can also skip the raw data and scripts and proceed directly to the fast binary data, which can hold a lot of data when heavily compressed.
-
-2.
-When the primary motivation for shipping data is to illustrate visualization techniques or to run examples, one can skip the raw data and processing scripts and only include the binary data.
-As long as the total package size doesn't exceed 5 megabytes, it would be acceptable as part of CRAN.
-For cases when this size is hard to maintain, CRAN recommends data-only packages that will be rarely updated.
-For a detailed discussion on this issue and alternative approaches, see @Brooke_Anderson2017.
-
-Unlike CRAN, bioconductor, does not have a 5 megabyte limit for package size.
-They have a strict specification for how to organise genomic data in a package, so they can be used for data analysis out of the box.
-This is a robust solution, but only works for homogeneous data as found with bioinformatics.
-For example, data in their ExperimentData section (bioconductor.org/packages/3.9/data/experiment) can be used in a reliable way.
-However, such a strict standard would be impossible to enforce or scale on a repository as heterogeneous as CRAN.
-
-However, these packages are still generally used for demonstration or testing purposes, over generating new knowledge for papers.
-
-One major disadvantage of packaging data inside R is that it makes the data availability very language centric.
-Non R users are unlikely to download and export data out of a package.
-This is why we recommend, as a rule, that researchers also archive data in a long-term data repository.
-These include domain specific repositories (see Section \@ref(publish-repos)) or more general purpose ones such as Zenodo or Figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.
-
-Of the 15539 packages on Central R Archive Network (CRAN), 6278 contain
-datasets either as binary data (5903 packages) or as external datasets
-(766). Binary files comprise a bulk of the data shipped in the `data`
-folder (68.06%) with other plain text formats such as txt, CSV, dat,
-`json` comprising less than one percent of data shipped in packages.
-
-# Data Sharing in Julia and Python {#julia-python}
-
-Sharing data in a programming language happens to some extent in other languages, such as Julia and Python.
-Julia was established in 2012, and is a high level language with great performance [@bezanson2012julia].
-Similar to R, Julia provides packages for database access, API access, and data packages [@METADATA-jl].
-This includes access to MySQL databases, NOAA, and actual datasets from Econometrics, faces, Market data, and even R datasets  [@NOAA-jl; @econ-data-jl; @face-data-jl; @market-data-jl; @r-data-jl].
-
-Python is a popular general purpose language, used in engineering, statistics, and computer science. 
-Including data in a python package is straightforward; a user can include data (or any other arbitrary files) inside a python package by declaring it in the manifest [@include-data-py-pkg].
-
-# Dealing with medium to large data {#med-large}
-
-Another common situation that researchers face is in dealing with data that fall somewhere between small and large.
-For example, small data could be tabular, as a CSV, in the order of bytes to megabytes to gigabytes that can be easily compressed, and large could be what falls under the umbrella of **big data**.
-The happy medium is generally data that are too big to fit on memory of most standard machines, but can successfully fit on disk (https://ropensci.github.io/arkdb/articles/articles/noapi.html).
-In this case, users who do not have the support to maintain resource intensive database servers can instead rely on light serverless databases such as MonetDB or SQLite.
-These databases provide disk based storage and using language agnostic interfaces, a analyst can easily query these data in manageable chunks that don't overrun memory limitations.
-Using software packages such as `arkdb` [@arkdb] one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
-
-To make these files available alongside compute, one ingenious but short-term solution is to use the GitHub release feature to attach such large database dumps.
-GitHub releases are designed to serve software releases and provide the opportunity to attach binary files.
-This mechanism can also be used to attach arbitrary files such as large data files, binary data, and database dumps as long as each file does not exceed 2gb.
-The R package `piggyback` (`https://docs.ropensci.org/piggyback/`) allows for uploading and downloading such files to GitHub releases, making it easy for anyone to access data files wherever the script is being run.
-We emphasize again that this is a short-term solution that is dependent on GitHub maintaining this service.
-
-# Publishing and repositories {#publish-repos}
-
-It is worth distinguishing between sharing data and publishing data.
-One can share data in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify.
-Data can be shared in numerous ways, including placing it in a repository, packaging it with methods, or using various free tiers of commercial services (such as dropbox or google drive).
-However, one must publish data when appropriate.
-
-There are a three common options for publishing data in research:
-
-1. **Least Moving Parts**
-2. **Domain Specific Venue**
-3. **Long Term Archive**
-
-We discuss each of these options and provide recommendations on how to publish data in these areas.
-
-In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databases.
-This approach **gets the data somewhere rather than nowhere**.
-Its minimal features means it is simple to maintain.
-In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organise.
-A downsides is that it does not scale to larger data.
-Self hosting the data is an option, but we discourage this, as it may succumb to bit rot -.
-
-In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data.
-For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank), and so on.
-Some other places to put data include the open science framework [@OSF], Dataverse, [@dataverse], and Data World, [@dataworld].
-Other datasets might be produced
-
-The purpose, use, and origin of the data is an important component to consider.
-Data for research has a different domain compared to data collected by governments, or by journalists.
-Many governments or civil organizations are now making their own data available through a government website interface.
-Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or organizations such as Data World, or may self host their data, such as five thirty eight [@data-538].
-
-This is a good option when the data is appropriate for the domain.
-This is often decided by community standards.
-We recommend adhering to community standards for a location to publish your data, as this will likely improve data reuse.
-The guidelines suggested in this paper for sharing the data should be included.
-
-A **Long Term Archive** is the best option to share the data.
-Long term archives provide information such as DOI (Digital Object Identifier) that make it easier to cite.
-Data can be placed in a long term archive and a DOI can be minted.
-This DOI can then be used in other venues, such as domain specific or even self hosted, and will ensure that the projects refer back appropriately.
-
-If the dataset you are shipping has a research application, the most relevant home for it would be Zenodo, a research data repository.
-Launched in 2013 in a joint collaboration between openAIRE and CERN, Zenodo provides a free, archival location for any researcher to depositor their datasets.
-The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases.
-Zenodo is able to accommodate larger file sizes upon request.
-
-## Publishing data in a data repository
-
-## Publishing data through a data journal
-
-Data used in publications are often shared in the supplementary materials of articles, or served on repositories such as the Dryad Digital Repository (@dryad).
-Dryad makes data from scientific publications discoverable, reusable, and citable.
-It is well funded through grants from the NSF, European Commission.
-
-To provide better context around the data used in research and better expose data for reuse, journals are now adding "data papers".
-These are specifically designed for publishing articles about the data, and sharing it.
-This benefits both researchers and readers.
-Researchers receive credit for data they have collected or produced.
-Readers get more context about the data.
-
-Data papers are similar to research articles, they have titles, authors, affiliations, abstract, and references.
-They generally require an explanation of why the data is useful to others, a direct link to the data, description of the design, materials, and methods.
-Other information on the subject area, data type, format, and related articles are usually required.
-
-Whilst useful, these requirements do not tell the author how to actually structure the data and folders.
-Instead providing ideas on what they should include.
-This is a useful step towards improving data reuse, but it _lacks some minimal structure that allows a researcher to have a predictable way to access and interpret the data_.
-
-Other journals operating in this space include journals like "data in brief", "Data", and "Nature Scientific Data".
-Guidelines for what is required in the content of the paper, and the required information along with the data (meta data, data format, etc.) vary by journal.
 
@@
 Binary or proprietary formats are discouraged in favor of interoperability, as it requires special software to read, even if it is sometimes slower to read in and out, and harder to share due to size.
+
+# Releasing data in R {#data-release-r}
+
+One low cost and easy way to distribute data alongside compute is to package the datasets as a data only package or as part of something larger where the methods are being implemented.
+The R ecosystem has many noteworthy examples of data-only packages.
+One exemplar is the `nycflights13` package by Hadley Wickham [@nycflights], This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes.
+The package not only provides ready to use binary data but also ships raw data (in a folder called `data-raw`) along with scripts used to clean them.
+The package was originally created as a way to provide example data to teach tidy data principles and serves as a great model for how one can publish a data package in R.
+
+A more common use case is to include data as part of a regular package where analytical methods are being implemented and shared.
+This serves the dual purpose of exposing methods and data together, making it extremely easy for a researcher to simply install and load the library at the top of their script.
+CRAN's distributed network (along with the global content distribution network maintained by RStudio) ensure that the package is quickly accessible to anyone in the R ecosystem.
+A second critical advantage in this approach is that one could also benefit from R's package documentation syntax to generate useful metadata for fast reference.
+This approach can also easily be adapted to other languages.
+Python for example, is far more flexible with respect to including arbitrary files as part of a package.
+
+**Other benefits of packaging data in R**
+
+1.
+Packaging data can be a very powerful pedagogical tool to help researchers understand how to transform data and prepare it for further analysis.
+To do so, one can package raw data alongside scripts.
+Long form documentation such as vignettes can provide further detailed discussion on the process.
+Users can also skip the raw data and scripts and proceed directly to the fast binary data, which can hold a lot of data when heavily compressed.
+
+2.
+When the primary motivation for shipping data is to illustrate visualization techniques or to run examples, one can skip the raw data and processing scripts and only include the binary data.
+As long as the total package size doesn't exceed 5 megabytes, it would be acceptable as part of CRAN.
+For cases when this size is hard to maintain, CRAN recommends data-only packages that will be rarely updated.
+For a detailed discussion on this issue and alternative approaches, see @Brooke_Anderson2017.
+
+Unlike CRAN, bioconductor, does not have a 5 megabyte limit for package size.
+They have a strict specification for how to organise genomic data in a package, so they can be used for data analysis out of the box.
+This is a robust solution, but only works for homogeneous data as found with bioinformatics.
+For example, data in their ExperimentData section (bioconductor.org/packages/3.9/data/experiment) can be used in a reliable way.
+However, such a strict standard would be impossible to enforce or scale on a repository as heterogeneous as CRAN.
+
+However, these packages are still generally used for demonstration or testing purposes, over generating new knowledge for papers.
+
+One major disadvantage of packaging data inside R is that it makes the data availability very language centric.
+Non R users are unlikely to download and export data out of a package.
+This is why we recommend, as a rule, that researchers also archive data in a long-term data repository.
+These include domain specific repositories (see Section \@ref(publish-repos)) or more general purpose ones such as Zenodo or Figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.
+
+Of the 15539 packages on Central R Archive Network (CRAN), 6278 contain
+datasets either as binary data (5903 packages) or as external datasets
+(766). Binary files comprise a bulk of the data shipped in the `data`
+folder (68.06%) with other plain text formats such as txt, CSV, dat,
+`json` comprising less than one percent of data shipped in packages.
+
+# Dealing with medium to large data {#med-large}
+
+Another common situation that researchers face is in dealing with data that fall somewhere between small and large.
+For example, small data could be tabular, as a CSV, in the order of bytes to megabytes to gigabytes that can be easily compressed, and large could be what falls under the umbrella of **big data**.
+The happy medium is generally data that are too big to fit on memory of most standard machines, but can successfully fit on disk (https://ropensci.github.io/arkdb/articles/articles/noapi.html).
+In this case, users who do not have the support to maintain resource intensive database servers can instead rely on light serverless databases such as MonetDB or SQLite.
+These databases provide disk based storage and using language agnostic interfaces, a analyst can easily query these data in manageable chunks that don't overrun memory limitations.
+Using software packages such as `arkdb` [@arkdb] one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
+
+To make these files available alongside compute, one ingenious but short-term solution is to use the GitHub release feature to attach such large database dumps.
+GitHub releases are designed to serve software releases and provide the opportunity to attach binary files.
+This mechanism can also be used to attach arbitrary files such as large data files, binary data, and database dumps as long as each file does not exceed 2gb.
+The R package `piggyback` (`https://docs.ropensci.org/piggyback/`) allows for uploading and downloading such files to GitHub releases, making it easy for anyone to access data files wherever the script is being run.
+We emphasize again that this is a short-term solution that is dependent on GitHub maintaining this service.
+
+# Challenges in documenting your dataset {#doco-challenge}
+
+There are many features to include with data alongside publications.
+However, not all of these are needed for every dataset.
+Working out which are needed is a challenge that is not discussed in the literature.
+This section discusses a framework for users, to help them decide how much they should document their data.
+To frame discussion around the challenges of data documenting, we can think of how an individual dataset falls on two features: "Effort to prepare", and "Ease of understanding" in Figure \@ref(fig:effort-understanding).
+The ideal space to be in the graph would be the top left hand corner.
+But what we notice is that taking more effort to prepare data means that the data is easier to understand.
+
+```{r effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT).", out.width = "90%"}
+
+# There are two Axis:
+#   X axis: <- Robustness (brittle - robust)
+#   Y axis: <- Difficulty (easy - hard)
+
+dat_prep <- tibble(prep = c(15,20,55,90, 96, 40),
+               understanding = c(15,20,95,90,96, 91),
+                  label = c("Data Dump",
+                            "Most studies",
+                            "Sensor Data",
+                            "50 hectare",
+                            "RCT",
+                            "Simulation Study"
+                            ))
+
+ggplot(dat_prep,
+       aes(x = prep,
+           y = understanding,
+           label = label)) +
+  geom_point() +
+  geom_label_repel(
+    segment.size = 0.2,
+    segment.color = "grey50"
+    ) +
+  # theme_void() + 
+  theme(axis.line.x = element_line(arrow = arrow(length = unit(3, 'mm'))), 
+        axis.line.y = element_line(arrow = arrow(length = unit(3, 'mm'))),
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank(),
+        aspect.ratio = 1
+  ) + 
+  labs(x = "Effort to Prepare",
+       y = "Ease of Understanding") +
+  xlim(0,100) +
+  ylim(0, 100)
+
+```
+
+Data with higher potential impact and reuse should be made as easy to understand as possible; but it also requires more time and effort to prepare.
+Impact is hard to measure, and varies from field to field, but as an example, take some data from medical randomized control trials (RCTs) on cancer treatment.
+These can have high impact, so requires a lot of time and effort to document.
+Comparatively, a small survey on a few demographics can have low impact.
+Placing these on the graph above, we see they might not have a worthwhile tradeoff for ease of understanding and ease of preparation.
+This means the effort put into preparing data documentation for a small survey should be kept relatively simple, not over complicated.
+Likewise, data that can be created via simulation from open source software could arguably not be shared since it can be generated from scratch with code; a reproducible process that requires computer time, not person time to create.
+
+Deciding how much data documentation to include should be based on the data's impact.
+The more impactful the data, the more documentation features to include.
+Figure \@ref(fig:data-sharing-workflow) shows the practical types of steps that can be taken against the effort required.
+
+```{r data-sharing-workflow, fig.cap = "The steps required compared to effort required for data sharing for reproducibility.", out.width = "90%"}
+knitr::include_graphics(here::here("figures",
+                                   "data_sharing_workflow.png"))
+```
+
+Creating good documentation has similar challenges to good writing: it takes time, and it can be hard to know when you are done.
+Thinking about two features: 1) the impact of data, and 2) the current effort to do each step, provides guidelines for the researcher to decide how much documentation they should provide for their data.
+
+Documentation challenges may evolve over time as the cost of making them easy to prepare and understand change.
+For example, new technologically may automate rigorous data documentation, so thorough documentation can take the same time as it takes to do poorly now.
+In this case, there would be no reason why more people cannot do this, even when the benefits are not immediately apparent.
+So if something might appear to have low impact, it should still be made easy to understand, if it doesn't take long to prepare that documentation.
+
+# Publishing and repositories {#publish-repos}
+
+It is worth distinguishing between sharing data and publishing data.
+One can share data in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify.
+Data can be shared in numerous ways, including placing it in a repository, packaging it with methods, or using various free tiers of commercial services (such as dropbox or google drive).
+However, one must publish data when appropriate.
+
+There are a three common options for publishing data in research:
+
+1. **Least Moving Parts**
+2. **Domain Specific Venue**
+3. **Long Term Archive**
+
+We discuss each of these options and provide recommendations on how to publish data in these areas.
+
+In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databases.
+This approach **gets the data somewhere rather than nowhere**.
+Its minimal features means it is simple to maintain.
+In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organise.
+A downsides is that it does not scale to larger data.
+Self hosting the data is an option, but we discourage this, as it may succumb to bit rot -.
+
+In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data.
+For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank), and so on.
+Some other places to put data include the open science framework [@OSF], Dataverse, [@dataverse], and Data World, [@dataworld].
+Other datasets might be produced
+
+The purpose, use, and origin of the data is an important component to consider.
+Data for research has a different domain compared to data collected by governments, or by journalists.
+Many governments or civil organizations are now making their own data available through a government website interface.
+Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or organizations such as Data World, or may self host their data, such as five thirty eight [@data-538].
+
+This is a good option when the data is appropriate for the domain.
+This is often decided by community standards.
+We recommend adhering to community standards for a location to publish your data, as this will likely improve data reuse.
+The guidelines suggested in this paper for sharing the data should be included.
+
+A **Long Term Archive** is the best option to share the data.
+Long term archives provide information such as DOI (Digital Object Identifier) that make it easier to cite.
+Data can be placed in a long term archive and a DOI can be minted.
+This DOI can then be used in other venues, such as domain specific or even self hosted, and will ensure that the projects refer back appropriately.
+
+If the dataset you are shipping has a research application, the most relevant home for it would be Zenodo, a research data repository.
+Launched in 2013 in a joint collaboration between openAIRE and CERN, Zenodo provides a free, archival location for any researcher to depositor their datasets.
+The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases.
+Zenodo is able to accommodate larger file sizes upon request.
+
+## Publishing data in a data repository
+
+## Publishing data through a data journal
+
+Data used in publications are often shared in the supplementary materials of articles, or served on repositories such as the Dryad Digital Repository (@dryad).
+Dryad makes data from scientific publications discoverable, reusable, and citable.
+It is well funded through grants from the NSF, European Commission.
+
+To provide better context around the data used in research and better expose data for reuse, journals are now adding "data papers".
+These are specifically designed for publishing articles about the data, and sharing it.
+This benefits both researchers and readers.
+Researchers receive credit for data they have collected or produced.
+Readers get more context about the data.
+
+Data papers are similar to research articles, they have titles, authors, affiliations, abstract, and references.
+They generally require an explanation of why the data is useful to others, a direct link to the data, description of the design, materials, and methods.
+Other information on the subject area, data type, format, and related articles are usually required.
+
+Whilst useful, these requirements do not tell the author how to actually structure the data and folders.
+Instead providing ideas on what they should include.
+This is a useful step towards improving data reuse, but it _lacks some minimal structure that allows a researcher to have a predictable way to access and interpret the data_.
+
+Other journals operating in this space include journals like "data in brief", "Data", and "Nature Scientific Data".
+Guidelines for what is required in the content of the paper, and the required information along with the data (meta data, data format, etc.) vary by journal.
 

### 2019-09-09 paper.Rmd
add some more detail on analysis vs raw data, remove some details on SDSS data

@@
 5. Analysis ready data
+6. Scripts to tidy analysis raw data into ready data
 
@@
 The data used in the data analysis should be provided in a folder called `data`.
-Ideally, the data should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, and observations in rows.
-Contrasting "raw data", "analysis ready data" should be in an easily readable plain-text format, such as CSV, tab separated, or semicolon separated.
+Ideally, the data should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, observations in rows, and only one value per cell.
+
+In contrast to "raw data", "analysis ready data" should be in an easily readable plain-text format, such as comma-, tab- or semicolon-separated.
 Binary or proprietary formats are discouraged in favor of interoperability, as it requires special software to read, even if it is sometimes slower to read in and out, and harder to share due to size.
+There are times where Tidy Data format can inflate the size of data due to repetition, which make it difficult to share - in these cases it might be worthwhile to consider multiple smaller, connected tables or other easily compressible formats such as apache Arrow.
 
@@
 
-Metadata from the data cannot be contained in the CSV file, but is available in the URL the csv was downloaded from: `http://dr15.sdss.org/optical/spectrum/view/data/format=csv?`
-`plateid=1678&mjd=53433&fiberid=425&reduction2d=v5_7_0`
-
-Which contains details of the data, specifically:
-
-* plateid = 1678
-* mjd = 53433
-* fiberid = 425
-* reduction2d = v5.7.0
-
-These are unique identifying features of the data, which tell us:
-
-* **plateid:** the unique plate used to explore a given section of night sky. 
-* **mjd:** an integer that marks the "modified julian date" of the night of observation.
-* **fiberid:** an integer of the fiber optic cable used to transmit light to the sensor.
-* The **version** marked as `v5_7_0` is the pipeline version used for BOSS in DR12.
-
-As a researcher who does not work in astrophysics, this information on the data was quite easy to retrieve from the SDSS website section called "Understanding SDSS spectroscopic data": https://www.sdss.org/dr12/spectro/spectro_basics/.
-
 The SDSS provides data at different levels of complexity for different degrees of understanding.

### 2019-10-14 paper.Rmd
fixes #79 and a few other issues

@@
 Data dictionaries are implemented in other software such as STATA, which provides a "codebook" command.
-Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that track versioning.
-Note that is different to Frictionlessdata's tabular data package spec [@frictionlessdata-data-pkg].
+Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that allow  checksum file comparisons for versioning.
+Note that is different to Frictionless Data's tabular data package spec [@frictionlessdata-data-pkg].
 
@@
 More than a million trees have been counted.
-The data and metadata however are hard to reach.
+The data and metadata however are unnecessarily difficult to reach.
 To obtain the data, one must fill out a form and agree to terms and conditions which require approval and authorship prior to publication.
 The biggest challenge with this study is that the data are stored on a personal FTP server of one of the authors.
-While the data are available in CSV and binary Rdata formats, the data storage servers do not have any metadata, README files or a data dictionary.
+While the data are available in CSV and binary `Rdata` formats, the data storage servers do not have any metadata, README files or a data dictionary.
 A separate, public archive hosts the metadata (https://repository.si.edu/handle/10088/20925) in a PDF file (https://repository.si.edu/bitstream/handle/10088/20925/RoutputFull.pdf?sequence=1&isAllowed=y) that describe the fields.
@@
 
-Many datasets are "data dumped" into repositories with a paper. Phrases such as the following might be familiar:
-
-> Researcher 1: "What do these columns mean?"
-> Researcher 2: "Sorry, I created the data 14 years ago and I don't remember".
-
-** Need to get an example of this.
-
-## Summary {#ex-data-summary}
-
-TODO: Needs a succinct summary here
+The vast majority of datasets are handled in an ad-hoc manner [@Michener2001] and dumped in transient data stores with little usable metadata.
+The value of these datasets decline over time until they are useless.
+We recommend that researchers consider the tradeoff in long-term value of the dataset and effort required to document and archive.
 

### 2019-12-08 paper.Rmd
Fixes #51
add author affiliations and resolve #101

@@
 ---
-title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility
+title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.
+author: Nicholas J Tierney^1,2,*^, Karthik Ram^3,*^
 output:
@@
 ---
+
+1 = Monash University, Department of Econometrics and Business Statistics
+2 = Australia Centre of Excellence for Mathematical and Statistical Frontiers
+3 = University of California, Berkeley
+* = Both authors contributed equally to the work
 
@@
 
-In contrast to "raw data", "analysis ready data" should be in an easily readable plain-text format, such as comma-, tab- or semicolon-separated.
+In contrast to "raw data", "analysis ready data" should be in an easily readable plain-text format, such as comma-, tab- or semicolon-separated files.
+These typically have file extensions like `.csv`, `.tsv` or `.txt`.
 Binary or proprietary formats are discouraged in favor of interoperability, as it requires special software to read, even if it is sometimes slower to read in and out, and harder to share due to size.
-There are times where Tidy Data format can inflate the size of data due to repetition, which make it difficult to share - in these cases it might be worthwhile to consider multiple smaller, connected tables or other easily compressible formats such as apache Arrow.
+For example, R data formats like `.rds`, `.rda`, or SPSS, STATA, or SAS data formats, `.sav`, `.dta`, or `.sas7bdat`.
 
@@
 
-Unlike CRAN, bioconductor, does not have a 5 megabyte limit for package size.
+Unlike CRAN, bioconductor, does not have a 5 megabyte limit for package size and offers a more liberal data inclusion policy.
 They have a strict specification for how to organise genomic data in a package, so they can be used for data analysis out of the box.

### 2019-12-09 paper.Rmd
working through dotpoints and add acknowledgements outline

@@
 
+In contrasst to a CSV data dictionary, JSON-LD has a defined, nested structure, which means more can be communicated efficiently, while still maintaining readability, and avoiding extra writing and repetition that comes with a plain CSV where everything is defined from scratch.
+
+Another rich format is XML, the eXtensible Markup Language.
+This was an early iteration on the idea of a plain text format that was both human and machine readable.
+It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable.
+However, one example of its usage is in EML, the Ecological Metadata Language.
+XML and JSON can be converted into each others respective formats
+
 Expand on:
 
-- Explain formats like XML, as people might come into formats that wany XML.
-- The power lies in the contrast to a csv data dictionary, as there is a defined, nested structure.
-- Ontologies and domain specific metadata standards such as EML might also be worth mentioning.
-- introduce the holy grail of the semantic web: JSON-LD
+- The holy grail of the semantic web: JSON-LD
 
@@
 - make a table that describes how to do something in R, or Python, or Julia
--
 
@@
 It is worth distinguishing between sharing data and publishing data.
-One can share data in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify.
+Data can be shared in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify.
 Data can be shared in numerous ways, including placing it in a repository, packaging it with methods, or using various free tiers of commercial services (such as dropbox or google drive).
@@
 A downsides is that it does not scale to larger data.
-Self hosting the data is an option, but we discourage this, as it may succumb to bit rot -.
+Self hosting the data is an option, but we discourage this, as it may succumb to bit rot - where data is corrupted, degraded, or servers turn off.
 
@@
 For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank), and so on.
-Some other places to put data include the open science framework [@OSF], Dataverse, [@dataverse], and Data World, [@dataworld].
-Other datasets might be produced
 
@@
 Many governments or civil organizations are now making their own data available through a government website interface.
-Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or organizations such as Data World, or may self host their data, such as five thirty eight [@data-538].
+Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or may self host their data, such as five thirty eight [@data-538].
 
@@
 
-If the dataset you are shipping has a research application, the most relevant home for it would be Zenodo, a research data repository.
-Launched in 2013 in a joint collaboration between openAIRE and CERN, Zenodo provides a free, archival location for any researcher to depositor their datasets.
+If the dataset you are shipping has a research application, the most relevant home for it would be the research data repositories **zenodo** or **dryad**.
+Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to depositor their datasets.
 The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases.
 Zenodo is able to accommodate larger file sizes upon request.
+Dryad performs data curation
+
+```{r create-draft-table, }
+tibble::tribble(
+  ~Organisation,              ~Pros, ~Cons,
+  "FigShare",                 "",     "",
+  "Dryad",                    "",     "",
+  "Zenodo",                   "",     "",
+  "Institutional Repository", "",     "",
+) %>% 
+  knitr::kable(caption = "Table Comparing common data submission platforms")
+```
 
@@
 
+# Acknowledgements
+
+- Mile McBain
+- Anna Krystalli
+- Daniella Lowenberg
+
 # Literature Cited

### 2019-12-09 paper.Rmd
tidy up authorship addresses; add sentence on open data commons open database license; expand on JSON vs XML; expand on dryad

@@
 title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.
-author: Nicholas J Tierney^1,2,*^, Karthik Ram^3,*^
+author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)
 output:
@@
 1 = Monash University, Department of Econometrics and Business Statistics
+
 2 = Australia Centre of Excellence for Mathematical and Statistical Frontiers
+
 3 = University of California, Berkeley
-* = Both authors contributed equally to the work
+
+\* = Both authors contributed equally to the work
 
@@
 
-> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (PBS Series, 2010-2017)
+> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (The Adventure of the Copper Beeches by Sir Arthur Conan Doyle)
 
@@
 If sharing data under an embargo, include detailed information on the embargo requirements in: the README, and in separate correspondence with those who receive the data.
+Databases may also be licensed with the Open Data Commons Open Database License: https://opendatacommons.org/licenses/odbl/, which provides provisions for sharing, creating, and adapting, provided that work is attributed, shared, and kept open.
 
@@
 
-In contrasst to a CSV data dictionary, JSON-LD has a defined, nested structure, which means more can be communicated efficiently, while still maintaining readability, and avoiding extra writing and repetition that comes with a plain CSV where everything is defined from scratch.
+In contrast to a CSV data dictionary, JSON-LD has a defined, nested structure, which means more can be communicated efficiently, while still maintaining readability, and avoiding extra writing and repetition that comes with a plain CSV where everything is defined from scratch.
 
-Another rich format is XML, the eXtensible Markup Language.
+Another rich format is XML, the e**X**tensible **M**arkup **L**anguage.
 This was an early iteration on the idea of a plain text format that was both human and machine readable.
 It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable.
-However, one example of its usage is in EML, the Ecological Metadata Language.
-XML and JSON can be converted into each others respective formats
-
-Expand on:
-
-- The holy grail of the semantic web: JSON-LD
-
-- readers just been advised to make a:
-  - a csv data dictionary,
-  - put that in the README
-  - then make a JSON file. 
-Is there a way to discuss strategies on harmonising all these formats?
-
-- we could map out how these things link together (e.g., you can turn JSON into xml with package ...)
-- make a table that describes how to do something in R, or Python, or Julia
+XML is used is in EML, the Ecological Metadata Language.
+XML differs to JSON in that JSON is shorter, and quicker and easier to read and write.
+JSON can also be parsed with standard JavaScript, whereas XML must be parsed by a special XML parser.
+JSON has key-value pairs of information, whereas XML represents data in a tree structure.
+XML and JSON can be converted into each others respective formats, for example with the `json` or `xml2` R packages.
 
@@
 Zenodo is able to accommodate larger file sizes upon request.
-Dryad performs data curation
+Dryad will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, but requires a data publishing fee.
 
-```{r create-draft-table, }
+```{r create-draft-table}
 tibble::tribble(
@@
 ) %>% 
-  knitr::kable(caption = "Table Comparing common data submission platforms")
+  knitr::kable(caption = "Comparison of common data submission platforms.")
 ```
@@
 - Daniella Lowenberg
+- ACEMS International Mobility Programme
 

### 2019-12-09 paper.Rmd
add in text reference for ellis and broman

@@
 This is the question we tackle in this paper.
-While there are several papers that serve as best-practice guides for formatting data and getting them ready for sharing, the aims of this paper are a bit different.
+While there are several papers that serve as best-practice guides for formatting data [@Broman2017] and getting them ready for sharing [@Ellis2017], the aims of this paper are a bit different.
 Our aim to address the issue of data in the context of reproducibility in data science workflows.

### 2019-12-09 paper.Rmd
final changes before review by Karthik

@@
 For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)).
-No matter where data is submitted, there should ideally be a canonical data repo in one long term archive that links to others with a DOI.
+No matter where data is submitted, there should ideally be a canonical data repository in one long term archive that links to others with a DOI.
 
@@
 This was an early iteration on the idea of a plain text format that was both human and machine readable.
-It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable.
-XML is used is in EML, the Ecological Metadata Language.
-XML differs to JSON in that JSON is shorter, and quicker and easier to read and write.
-JSON can also be parsed with standard JavaScript, whereas XML must be parsed by a special XML parser.
-JSON has key-value pairs of information, whereas XML represents data in a tree structure.
+It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable as JSON.
+JSON is shorter, and quicker and easier to read and write than XML, and can also be parsed with standard JavaScript, whereas XML must be parsed by a special XML parser.
 XML and JSON can be converted into each others respective formats, for example with the `json` or `xml2` R packages.
+XML is still used as metadata for data storage, for example in EML, the Ecological Metadata Language [@EML-about].
 
@@
 
-If the dataset you are shipping has a research application, the most relevant home for it would be the research data repositories **zenodo** or **dryad**.
+If the dataset you are shipping has a research application, the most relevant home would be the research data repositories **zenodo** or **dryad**.
 Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to depositor their datasets.
@@
 Zenodo is able to accommodate larger file sizes upon request.
-Dryad will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, but requires a data publishing fee.
+The Dryad Digital Repository (@dryad) will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, in exchange for a data publishing fee.
 
@@
 To be broadly useful, your data must be deposited in a permanent archive, to ensure that it does not disappear from a ephemeral location such as your university website.
-It shoudl also contain useful metadata so that someone outside the project team can understand what the data mean.
+It should also contain useful metadata so that someone outside the project team can understand what the data mean.
 However, this level of effort is not always critical for all data science efforts.

### 2019-12-19 paper.Rmd
tidy up up to section 2

@@
 Among the different areas with a computational bent, the bioinformatics community in particular has a strong culture around open source [@Gentleman2004], and has made releasing code and associated software a recognized mainstream activity.
-Many journals in these fields have also pushed authors to submit code (model implementations) alongside their papers, with a few journals going as far as providing a "reproducibility review" [@Peng2011].
+Many journals in these fields have also pushed authors to submit code (in the form of model implementations) alongside their papers, with a few journals going as far as providing a "reproducibility review" [@Peng2011].
 
@@
 This rapid growth in data science has largely been fueled by easy access to open source software tools.
-Programming languages such as Python, R and Julia help scientists implement and share new methods to work with data.
+Programming languages such as Python, R and Julia help scientists implement and share new methods to work with data [@rcore; @python; @julia].
 Each of these languages is supported by thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
@@
 This makes it simple to load a specific environment to run any analysis.
-Code is handled by version control with tools like Git and GitHub [@git; @github], paired with archiving such as Zenodo provide access to code (particularly model implementations)[@zenodo].
+Code is handled by version control with tools like Git and GitHub [@git; @github], paired with archiving frameworks such as Zenodo provide access to code (particularly model implementations)[@zenodo].
 All the necessary software is available from various package managers (and their numerous geographic mirrors and archives) making it easy to install any version of a software package.
@@
 While there are several papers that serve as best-practice guides for formatting data [@Broman2017] and getting them ready for sharing [@Ellis2017], the aims of this paper are a bit different.
-Our aim to address the issue of data in the context of reproducibility in data science workflows.
-In particular we discuss the various tradeoffs one has to consider when documenting and sharing data, when it is worth publishing and how this would change depending on the use case, perceived impact, and potential audience.
+Our aim is to address the issue of data in the context of reproducibility in data science workflows.
+In particular we discuss the various tradeoffs one has to consider when documenting and sharing data, when it is worth publishing, and how this would change depending on the use case, perceived impact, and potential audience.
 
 We discuss how to share and/or publish data and cover various tradeoffs when deciding how much to do.
-We skip detailed discussions of the minutiae of data preparation [@Broman2017; @Ellis2017], or citation (as of this writing, data citation is still in its infancy).
+We skip detailed discussions of data preparation [@Broman2017; @Ellis2017], or citation (as of this writing, data citation is still in its infancy).
 We also analyze the state of data contained inside software packages, shared and made available as part of modern data analysis.
@@
 
-# What is the minimal structure to share data for analysis {#minimal}
-
-**Principles for all three**
+# A minimal structure to share data for analysis {#minimal}
 
@@
 For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)).
-No matter where data is submitted, there should ideally be a canonical data repository in one long term archive that links to others with a DOI.
+No matter where data is submitted, there should ideally be a canonical data repository in one long term archive that links to others.
+It should also have an accession number, or Digital Object Identification (DOI) number, which allows for it to be cited (discussed further in \@ref(cite).
 
@@
 
-Out of these sections, the minimal viable format are:
+Out of these sections, the minimal viable format are the following provided in one main directory folder:
 
@@
 
-What goes into each of these sections is now described.
+These sections are now described.
 

### 2019-12-19 paper.Rmd
relatively minor updates, including adding a new data codebook

@@
 
-Out of these sections, the minimal viable format are the following provided in one main directory folder:
+Out of these sections, the minimal files to provide, in order of most importance should be provided in one main directory folder:
 
@@
 
-The README is often the first place people will go to learn more about data, and helps users navigate the repository.
+In the context of datasets, READMEs are particularly useful when there are no reliable standards. The best practice here is to have one README per dataset.
+
+A README is often the first place people will go to learn more about anything in a folder - they are very common in softwar, and historically were includd so that the uppecase lettrs of README meant ti would be at the top of a directory.
+In the context of data, a README when there are no reliable standards.
 It is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":
@@
 
-The README should be placed in the top level of the project.
-It should be brief, and provide links to the other aforementioned sections.
+The README should be placed in the top level of the project, and with one README per dataset.
+It should be brief, and provide links to the other aforementioned sections, and be in one directory.
 Saving a README with the extension `.md` file gives the author the formatting benefits of `markdown`, making it simple to insert links, tables, and make lists.
@@
 This helps someone not familiar with the data understand, and use the data.
-At a minimum they should contain:
+At a minimum they should contain the following pieces of information about the data:
 
@@
 
+Figure \@ref(fig:variables) shows an example of the data and data dictionary details.
 **Variable names** should be short, descriptive names with no spaces or special characters.
-In the most common tabular data case, these correspond to column names, for example, "job_position", "faculty_level", and "years_at_company".
+In the most common tabular data case, these correspond to column names, for example, "job_position", "faculty_level", and "years_at_uni".
 **Variable labels** are longer descriptions of variables.
@@
 **Variable codes** apply to categorical (factor) variables, and are the values for their contents.
-For example, a variable could contain answers to a question, with values and codes like so: `0 = no, 1 = yes`, or for another variable indicating indoors or outdoors: `0 = inside` and `1 = outside`.
+For example, a variable could contain answers to a question, with values and codes like so: `0 = no, 1 = yes`, to indicate if someone is in statistics, for example.
 These should be consistent across similar variables to avoid problems where `0 = yes` for one variable, and `1 = yes` in another.
@@
 For example, censored data, patient drop out, or measurement error can have different values, such as "unknown", -99, or other value codes [@White2013; @Broman2017].
+
+```{r variables, fig.cap = "Four tables - the data, and it's variable names, variable codes, and the meaning of missing values", fig.height = 8}
+library(tibble)
+
+dat <- tibble::tribble(
+  ~job_position, ~outside, ~statistics, ~date_started,
+   "Lecturer",        1,        1,      "2019-01-01",
+   "Professor",       0,        0,      NA,
+   "Senior Lecturer", 1,        1,      "unknown"
+)
+
+variable_labels <- tibble::tribble(
+  ~variable, ~description,
+  "job_position", "University Job Position",
+  "outside",      "Did this person study outside this university?",
+  "statistics",   "Is this perosn in the statistics faculty",
+  "date_started", "What date did this perosn start their position?"
+)
+
+variable_codes <- tibble::tribble(
+  ~variable,        ~code,
+   "job_position",   "text description of job position",
+   "outside",       "0 = not studied outside, \n1 = has studied outside",
+   "statistics",    "0 = is not in statistics dept., \n1 = is currently in statistics dept.",
+   "date_started",   "date variable in format YYYY-MM-DD, \ne.g., 2001-01-01 is 1st January, 2001"
+)
+
+missing_values <- tibble::tribble(
+  ~value, ~meaning,
+  NA,     "Missing value",
+  "unknown", "Missing value"
+)
+
+library(gridExtra)
+grid.arrange(
+  tableGrob(dat),
+  tableGrob(variable_labels),
+  tableGrob(variable_codes),
+  tableGrob(missing_values),
+  ncol = 1
+)
+
+```
 
@@
              "double"),
-  description = c("Year",
+  Description = c("Year",
                   "County-type (urban, suburban, small/mid, rural)",
@@
 knitr::kable(dictionary_incarcerate,
-             caption = "The prisoner summary data dictionary, with columns on the variable, it's class, and a short description of the contents of the variable.",
+             caption = "The prisoner summary data dictionary, with columns for the variable, its class, and a short description of the contents of the variable.",
              booktabs = TRUE) %>%
@@
 
-Data dictionary tables should be placed in the README and presented as a table. Every data dictionary should also be provided in its raw form in the repository, for example, saved as a CSV.
+Data dictionary tables should be placed in the README and presented as a table.
+Every data dictionary should also be provided in its raw form in the repository, for example, saved as a CSV, so that they don't become trapped in a README.
 
@@
 Data with a CC0 license means the data owners waive all their rights to the work, and it now "owned" by the public.
-The data can be freely shared, this means it can be copied, modified, and distributed, even for commercial purposes _without asking permission_.
+The data can be freely shared, copied, modified, and distributed, even for commercial purposes _without asking permission_.
 When using data with CC0, it is good practice to cite the original paper, but it is not required.
@@
 A Digital Object Identifier (DOI), is a unique identifier for a digital object such as a paper, poster, or software, and is a prerequisite for citation.
-For example, if the data are deposited in dryad or Zenodo, the best practice would be to copy the citation created by these repositories.
+For example, if the data are deposited in repositories like Dryad or Zenodo, the best practice would be to copy the citation created by these repositories.
 Under the hood, the DOI is "minted" by DataCite, who provide DOIs for research outputs and data.
@@
 
-The README, and data dictionary provides _human readable_ information on the data.
+The README and data dictionary provides _human readable_ information on the data.
 To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
-Machine readable metadata also provide a structure that means the data can be indexed and searched.
+Machine readable metadata also provide a structure that allow the data to be indexed and searched online, through services scuh as google datasets search.
 An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
@@
 This was an early iteration on the idea of a plain text format that was both human and machine readable.
-It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable as JSON.
+It is powerful and extensible (it powers the entire Microsoft Office suite), but for metadata purposes it is not as human readable as JSON.
 JSON is shorter, and quicker and easier to read and write than XML, and can also be parsed with standard JavaScript, whereas XML must be parsed by a special XML parser.
@@
 
-Unlike CRAN, bioconductor, does not have a 5 megabyte limit for package size and offers a more liberal data inclusion policy.
+Unlike CRAN, bioconductor does not have a 5 megabyte limit for package size and offers a more liberal data inclusion policy.
 They have a strict specification for how to organise genomic data in a package, so they can be used for data analysis out of the box.
@@
 However, such a strict standard would be impossible to enforce or scale on a repository as heterogeneous as CRAN.
-
 However, these packages are still generally used for demonstration or testing purposes, over generating new knowledge for papers.
@@
 
-There are a three common options for publishing data in research:
+There are three common options for publishing data in research:
 
@@
 In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organise.
-A downsides is that it does not scale to larger data.
+A downside is that it does not scale to larger data.
 Self hosting the data is an option, but we discourage this, as it may succumb to bit rot - where data is corrupted, degraded, or servers turn off.
@@
 Dryad makes data from scientific publications discoverable, reusable, and citable.
-It is well funded through grants from the NSF, European Commission.
+It is well funded through grants from the National Science Foundation and European Commission.
 
@@
 This benefits both researchers and readers.
-Researchers receive credit for data they have collected or produced.
-Readers get more context about the data.
+Researchers receive credit for data they have collected or produced, and readers get more context about the data.
 
@@
 
-Whilst useful, these requirements do not tell the author how to actually structure the data and folders.
-Instead providing ideas on what they should include.
+Whilst useful, these requirements do not tell the author how to actually structure the data and folders for reuse.
+Instead, they provide ideas on what they should include.
 This is a useful step towards improving data reuse, but it _lacks some minimal structure that allows a researcher to have a predictable way to access and interpret the data_.
@@
 
-The data are are free to use and publicly accessible, with the interface to the data being designed to cover a wide range of skills.
+The data are free to use and publicly accessible, with the interface to the data being designed to cover a wide range of skills.
 For example, the marvin service to streamline access and visualisation of MaNGA data [@marvin-service], through to raw data access (see Figure \@ref(fig:sdss-raw-data-screenshot)).
@@
 - Metadata in the machine readable standard, SCHEMA
-- Data in multiple formats (e.g., .csv and .fits, database)
+- Data in multiple formats (e.g., .csv, .fits, and database)
 - Previously released data available in archives
@@
 To be broadly useful, your data must be deposited in a permanent archive, to ensure that it does not disappear from a ephemeral location such as your university website.
-It should also contain useful metadata so that someone outside the project team can understand what the data mean.
+It should also contain useful metadata so that someone outside the project team can understand what the data, and variables, mean.
 However, this level of effort is not always critical for all data science efforts.
@@
 **Stick with simple data formats** to ensure long-term usefulness of datasets in a way that is not tied to transient technologies (such as ever changing spreadsheet standards), store the data as plain text (e.g., CSV) where possible.
-This can take the form of comma/tab separated files in most cases.
-This will ensure transparency and future proof your data.
+This can take the form of comma, or tab separated files in most cases.
+This will ensure transparency, and future proof your data.
 
@@
 Over the years researchers have made data available in a multitude of unreliable ways, including offering downloads from university websites, FTP repositories, and password protected servers, all of which are prone to bit rot.
-In rare cases where the data are deposited in a permanent archive, insufficient metadata and missing data processing scripts have made it harder and time intensive to map these to code.
-Our goal here is to describe a variety of ways in which a researcher can ship data, analogous to how easy it has become to push code to services like GitHub.
+In rare cases where the data are deposited in a permanent archive, insufficient metadata and missing data processing scripts have made it harder and more time intensive to map these to code.
+Our goal here is to describe a variety of ways a researcher can ship data, analogous to how easy it has become to push code to services like GitHub.
 The choice of technology and method all depends on the nature of the data, and its size and complexity.

### 2019-12-19 paper.Rmd
reference google data search

@@
 To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
-Machine readable metadata also provide a structure that allow the data to be indexed and searched online, through services scuh as google datasets search.
+Machine readable metadata also provide a structure that allow the data to be indexed and searched online, through services scuh as google datasets search [@google-data-search].
 An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].

### 2020-01-10 paper.Rmd
adds abstract (#91) and various phrasing

--- 2019-12-19 paper.Rmd
@@
 
-1 = Monash University, Department of Econometrics and Business Statistics
+1 = Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia
 
-2 = Australia Centre of Excellence for Mathematical and Statistical Frontiers
+2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)
 
-3 = University of California, Berkeley
+3 = University of California, Berkeley, USA
 
 \* = Both authors contributed equally to the work
+
+**Abstract**
+
+Data makes science possible.
+Sharing data improves visibility, and makes the research process transparent.
+This increases trust in the work, and allows for independent reproduction of results.
+However, a large proportion of data from published research is often only available to the original authors.
+Despite the obvious benefits of sharing data, and scientists' advocating for the importance of sharing data, most advice on sharing data discusses its broader benefits, rather than the practical considerations of sharing.
+This paper provides practical, actionable advice on how to actually share data alongside research.
+The key message is sharing data falls on a continuum, and entering it should come with minimal barriers.
 
@@
 
+# Introduction {#intro}
+
 > "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (The Adventure of the Copper Beeches by Sir Arthur Conan Doyle)
-
-# Introduction {#intro}
 
@@
 No matter where data is submitted, there should ideally be a canonical data repository in one long term archive that links to others.
-It should also have an accession number, or Digital Object Identification (DOI) number, which allows for it to be cited (discussed further in \@ref(cite).
+It should also have an accession number, or Digital Object Identification (DOI) number, which allows for it to be cited (discussed further in \@ref(cite)).
 
@@
 
-Out of these sections, the minimal files to provide, in order of most importance should be provided in one main directory folder:
+From these sections, the minimal files to provide in order of most importance are:
 
@@
 
-In the context of datasets, READMEs are particularly useful when there are no reliable standards. The best practice here is to have one README per dataset.
-
+In the context of datasets, READMEs are particularly useful when there are no reliable standards.
+The best practice here is to have one README per dataset.
 A README is often the first place people will go to learn more about anything in a folder - they are very common in softwar, and historically were includd so that the uppecase lettrs of README meant ti would be at the top of a directory.
@@
 
-Table \@ref(tab:table-data-incarcerate) shows an example data dictionary table taken from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate].
-This includes information on the variable the class (type) of the variable, and a longer description of the variable.
+Table \@ref(tab:table-data-incarcerate) shows an example data dictionary table from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate].
+This includes information on the variable, its class (type), and a longer description.
 
@@
 
-Data dictionary tables should be placed in the README and presented as a table.
-Every data dictionary should also be provided in its raw form in the repository, for example, saved as a CSV, so that they don't become trapped in a README.
+Data dictionaries should be placed in the README and presented as a table.
+Every data dictionary should also be provided in its raw form (e.g., a CSV) in the repository, so they aren't "trapped" in the README.
 
@@
 
-The CC0 is a “public domain” license.
+The CC0 is a "public domain" license.
 Data with a CC0 license means the data owners waive all their rights to the work, and it now "owned" by the public.
@@
 To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
-Machine readable metadata also provide a structure that allow the data to be indexed and searched online, through services scuh as google datasets search.
+This provides a structure allowing the data to be indexed and searched online, through services scuh as google datasets search [@google-data-search].
 An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
-For example, a dataset called "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), which has the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).
+For example, a dataset "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), and the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).
 
@@
 
-This contains fields such as "path", which describe the file path to the data, and a schema with subfields name and type, for each variable.
+This contains fields such as "path", describing the file path to the data, and a schema with subfields name and type, for each variable.
 It also provides information for licensing and features such as line breaks, and delimiters, useful to ensure the data is correctly imported into a machine.
-It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format.
+It is built on JSON, a lightweight, human-friendly, machine readable data-interchange format.
 Table schema is baked into formats such as csvy [@csvy], an extended `csv` format, which has additional front matter in a lightweight markup format (YAML) using Table Schema.
@@
 The data used in the data analysis should be provided in a folder called `data`.
-Ideally, the data should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, observations in rows, and only one value per cell.
+Ideally, the data "analysis ready data" should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, observations in rows, and only one value per cell.
 
@@
 
-## Publishing data in a data repository
+## Publishing data in a data repository (need some help here)
 

### 2020-01-10 paper.Rmd
minor wording changes

@@
 They have a strict specification for how to organise genomic data in a package, so they can be used for data analysis out of the box.
-This is a robust solution, but only works for homogeneous data as found with bioinformatics.
-For example, data in their ExperimentData section (bioconductor.org/packages/3.9/data/experiment) can be used in a reliable way.
-However, such a strict standard would be impossible to enforce or scale on a repository as heterogeneous as CRAN.
-However, these packages are still generally used for demonstration or testing purposes, over generating new knowledge for papers.
+While a robust solution, it only works for homogeneous data as found with bioinformatics.
+For example, data in their ExperimentData section (bioconductor.org/packages/3.9/data/experiment) can be used reliably.
+Such a strict standard would be impossible to enforce or scale on a repository as heterogeneous as CRAN.
+However, CRAN packages are still generally used for demonstration or testing purposes, over generating new knowledge for papers.
 
@@
 Documentation challenges may evolve over time as the cost of making them easy to prepare and understand change.
-For example, new technologically may automate rigorous data documentation, so thorough documentation can take the same time as it takes to do poorly now.
+For example, if new technologically automates rigorous data documentation, thorough documentation can take the same time it takes to do poorly now.
 In this case, there would be no reason why more people cannot do this, even when the benefits are not immediately apparent.
@@
 A citation only makes sense when your data has a DataCite compliant DOI, which is automatically provided when data is published in repositories like Zenodo and Dryad.
-Although a citation may not accrue references, without a DOI it is guaranteed not to.
+While a citation may not accrue references, without a DOI it is guaranteed not to.
 
@@
 Many of these articles appeal to the broader benefits of data sharing, but rarely talk about the practical considerations of making data available alongside code.
-Trivial reproducibility has always relied upon the idea that as long as code exists on a platform such as GitHub, and clearly lists open source dependencies, one should be able to run the code at a future time.
+Trivial reproducibility has always relied upon the idea that as long as code exists on a platform (e.g., GitHub), and clearly lists open source dependencies, one should be able to run the code at a future time.
 Containerization efforts such as Docker have made it even easier to capture system dependencies, and rapidly launch containers.

### 2020-01-13 paper.Rmd
Added some small changes to affliation and funding

@@
 
-3 = University of California, Berkeley, USA
+3 = Berkeley Institute for Data Science, University of California, Berkeley, USA
 
@@
 - ACEMS International Mobility Programme
+- Helmsley Charitable Trust, Gordon and Betty Moore Foundation, Sloan Foundation.
 

### 2020-01-16 paper.Rmd

@@
 
-## Publishing data in a data repository (need some help here)
+## Publishing data in a data repository
+
+Scientists have numerous venues to deposit their data, many of which are ephemeral, despite the convenience they offer.
+For data that are not narrow in scope with limited potential for reuse, we recommend publishing data in an established data repository such as Dryad.
+In addition to providing a long-term home for data, these repositories also include data curation by a professional data manager, compliance with open source standards, and metrics.
 

### 2020-01-20 paper.Rmd
update makefile to address #109

--- 2020-01-13 paper.Rmd
@@
 suppressMessages(library(cowplot))
+suppressMessages(library(gridExtra))
 ```
@@
 It should be brief, and provide links to the other aforementioned sections, and be in one directory.
+It should also contain any other guidance for the user on how to read and interpret the whole directory.
+For example, explaining where tidy and raw data are stored, and other scripts for tidying.
+
 Saving a README with the extension `.md` file gives the author the formatting benefits of `markdown`, making it simple to insert links, tables, and make lists.
@@
 ```{r variables, fig.cap = "Four tables - the data, and it's variable names, variable codes, and the meaning of missing values", fig.height = 8}
-library(tibble)
 
-dat <- tibble::tribble(
+dat <- tribble(
   ~job_position, ~outside, ~statistics, ~date_started,
@@
 
-variable_labels <- tibble::tribble(
+variable_labels <- tribble(
   ~variable, ~description,
@@
 
-variable_codes <- tibble::tribble(
+variable_codes <- tribble(
   ~variable,        ~code,
@@
 
-missing_values <- tibble::tribble(
+missing_values <- tribble(
   ~value, ~meaning,
@@
 
-library(gridExtra)
 grid.arrange(
@@
 Without a license, these rules are unclear, and can lead to problems with attribution and citation.
-It can be overwhelming to try and the right license for a use case.
+It can be overwhelming to try and find the right license for a use case.
 Two licenses that are well suited for data sharing are:
@@
 2. Creative Commons CC0 1.0 Universal (CC0)
+
+Other licenses or notices to be aware of are **copyrighted data**, and **data embargoes**.
+If you are working with data **already copyrighted**, (for example under CC BY or CC0), you must give follow appropriate guidelines for giving credit.
+Data may also be under **embargo**.
+This means data cannot be shared more widely until a specific release time.
+If sharing data under an embargo, include detailed information on the embargo requirements in: the README, and in separate correspondence with those who receive the data.
+Databases may also be licensed with the Open Data Commons Open Database License: https://opendatacommons.org/licenses/odbl/, which provides provisions for sharing, creating, and adapting, provided that work is attributed, shared, and kept open.
+
+Once you decide on a license, you should provide a LICENSE or LICENSE.md file that contains the entire license in the top level of the directory.
 
@@
 
-Other licenses or notices to be aware of are **copyrighted data**, and **data embargoes**.
-If you are working with data **already copyrighted**, (for example under CC BY or CC0), you must give follow appropriate guidelines for giving credit.
-Data may also be under **embargo**.
-This means data cannot be shared more widely until a specific release time.
-If sharing data under an embargo, include detailed information on the embargo requirements in: the README, and in separate correspondence with those who receive the data.
-Databases may also be licensed with the Open Data Commons Open Database License: https://opendatacommons.org/licenses/odbl/, which provides provisions for sharing, creating, and adapting, provided that work is attributed, shared, and kept open.
-
 ## Citation: How you want your data to be cited {#cite}
@@
 If a DOI is unavailable, a citation will be meaningless, as it cannot be tracked by any means.
+A file named `citation` should be placed in the directory, at the same level as the `README`.
+It should contain a DOI, and could be in `.bibtex` format.
 
@@
 
+To create appropriate metadata, we recommend metadata generators such as `dataspice` or `codebook` [@dataspice; @Arslan2019].
+These help automatically create the strucutred JSON data.
+This metadata should be provided in a folder called "metadata", which should be provided for every dataset.
+
 ## Raw data: The original or first data provided {#raw-data}
@@
 Raw data is usually the first format of the data provided before any tidying or cleaning of the data.
-If the raw data is a practical size to share, it should be shared in a folder called `raw-data`.
+If the raw data is a practical size to share, it should be shared in a folder called `data-raw`.
 The raw data should be in the form that was first received, even if it is in binary or some proprietary format.
@@
 
-Any code used to clean and tidy the raw data should be provided in the `raw-data` directory.
+Any code used to clean and tidy the raw data should be provided in the `data-raw` directory.
 Ideally this would involve only scripted languages, but if other practical steps were taken to clean up the data, these should be recorded in a plain text or markdown file.
@@
 Using software packages such as `arkdb` [@arkdb] one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
+Another option is to break large text data files into chunks with named sequences (e.g., `teaching-1.csv`, `teaching-2.csv`, etc.).
 
@@
 If the dataset you are shipping has a research application, the most relevant home would be the research data repositories **zenodo** or **dryad**.
-Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to depositor their datasets.
+Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to deposit their datasets.
 The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases.
@@
 
-sdss_image_path <- here::here("figures", "sdss-spectra-example-plate-1678.png")
+sdss_image_path <- here::here("figures",
+                              "sdss-spectra-example-plate-1678.png")
 

### 2020-01-22 paper.Rmd
use material design for docs/index.html
fixes typos and updates docs site

@@
 
-**Abstract**
+# Abstract {-}
 
@@
 This primarily comes in the form of potential for reuse in other contexts along use for training and teaching (@McKiernan2016).
-Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing doesn't put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
+Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing does not put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
 First and foremost, sharing data along with other artifacts can make it easier for a reader to independently reproduce the results, thereby increasing transparency and trust in the work.
@@
 By relying on the idea of having research compendia [@gentleman2007statistical] or something as simple as a requirements file, one can easily install all the necessary scaffolding.
-Data on the other hard are rarely accessible that easily.
+Data on the other hand are rarely accessible that easily.
 
@@
 In the context of datasets, READMEs are particularly useful when there are no reliable standards.
-The best practice here is to have one README per dataset.
-A README is often the first place people will go to learn more about anything in a folder - they are very common in softwar, and historically were includd so that the uppecase lettrs of README meant ti would be at the top of a directory.
-In the context of data, a README when there are no reliable standards.
-It is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":
+A README is often the first place people will go to learn more about anything in a folder - they are very common in software, and historically were included so the uppercase letters of README meant it would be at the top of a directory.
+The README is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":
 
@@
   "outside",      "Did this person study outside this university?",
-  "statistics",   "Is this perosn in the statistics faculty",
+  "statistics",   "Is this person in the statistics faculty",
   "date_started", "What date did this perosn start their position?"
@@
 To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
-This provides a structure allowing the data to be indexed and searched online, through services scuh as google datasets search [@google-data-search].
+This provides a structure allowing the data to be indexed and searched online, through services such as google datasets search [@google-data-search].
 An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
@@
 
-This contains fields such as "path", describing the file path to the data, and a schema with subfields name and type, for each variable.
+This contains fields such as "path", describing the file path to the data, and a schema with sub-fields name and type, for each variable.
 It also provides information for licensing and features such as line breaks, and delimiters, useful to ensure the data is correctly imported into a machine.
@@
 To create appropriate metadata, we recommend metadata generators such as `dataspice` or `codebook` [@dataspice; @Arslan2019].
-These help automatically create the strucutred JSON data.
+These help automatically create the structured JSON data.
 This metadata should be provided in a folder called "metadata", which should be provided for every dataset.
@@
 When the primary motivation for shipping data is to illustrate visualization techniques or to run examples, one can skip the raw data and processing scripts and only include the binary data.
-As long as the total package size doesn't exceed 5 megabytes, it would be acceptable as part of CRAN.
+As long as the total package size does not exceed 5 megabytes, it would be acceptable as part of CRAN.
 For cases when this size is hard to maintain, CRAN recommends data-only packages that will be rarely updated.
@@
 Unlike CRAN, bioconductor does not have a 5 megabyte limit for package size and offers a more liberal data inclusion policy.
-They have a strict specification for how to organise genomic data in a package, so they can be used for data analysis out of the box.
+They have a strict specification for how to organize genomic data in a package, so they can be used for data analysis out of the box.
 While a robust solution, it only works for homogeneous data as found with bioinformatics.
@@
 
-Deciding how much data documentation to include should be based on the data's impact.
+Deciding how much data documentation to include should be based on the impact of the data.
 The more impactful the data, the more documentation features to include.
@@
 Documentation challenges may evolve over time as the cost of making them easy to prepare and understand change.
-For example, if new technologically automates rigorous data documentation, thorough documentation can take the same time it takes to do poorly now.
+For example, if new technology automates rigorous data documentation, thorough documentation can take the same time it takes to do poorly now.
 In this case, there would be no reason why more people cannot do this, even when the benefits are not immediately apparent.
-So if something might appear to have low impact, it should still be made easy to understand, if it doesn't take long to prepare that documentation.
+So if something might appear to have low impact, it should still be made easy to understand, if it does not take long to prepare that documentation.
 
@@
 Its minimal features means it is simple to maintain.
-In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organise.
+In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organize.
 A downside is that it does not scale to larger data.
@@
 The Dryad Digital Repository (@dryad) will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, in exchange for a data publishing fee.
-
-```{r create-draft-table}
-tibble::tribble(
-  ~Organisation,              ~Pros, ~Cons,
-  "FigShare",                 "",     "",
-  "Dryad",                    "",     "",
-  "Zenodo",                   "",     "",
-  "Institutional Repository", "",     "",
-) %>% 
-  knitr::kable(caption = "Comparison of common data submission platforms.")
-```
 
@@
 
-# Literature Cited
+# Literature Cited {-}

### 2020-02-05 paper.Rmd
Update paper.Rmd
Typo
mention OSF - closes #112

@@
 A Digital Object Identifier (DOI), is a unique identifier for a digital object such as a paper, poster, or software, and is a prerequisite for citation.
-For example, if the data are deposited in repositories like Dryad or Zenodo, the best practice would be to copy the citation created by these repositories.
+For example, if the data are deposited in repositories like Dryad, Zenodo, or the Open Science Framework, the best practice would be to copy the citation created by these repositories.
 Under the hood, the DOI is "minted" by DataCite, who provide DOIs for research outputs and data.
@@
 
-If the dataset you are shipping has a research application, the most relevant home would be the research data repositories **zenodo** or **dryad**.
+If the dataset you are shipping has a research application, the most relevant home would be the research data repositories **zenodo**, **dryad**, or **Open Science Framework**.
 Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to deposit their datasets.
@@
 The Dryad Digital Repository (@dryad) will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, in exchange for a data publishing fee.
+The OSF is a tool that captures the research process online.
+This ranges from conceiving research ideas, study design, writing papers, to storing research data.
+The entire history of the project is recorded, so it promotes centralized, and transparent workflows.
+Although not specifically designed for data, like Dryad or Zenodo, OSF does provide a DOI minting service, and a more wholistic approach, which might be appealing to users [@OSF].
 

### 2020-02-05 paper.Rmd
comment on package installation - closes #111

@@
 These packages form the building blocks of a data scientists daily workflow.
-In a modern computational workflow it is trivial for a user to `pip install` a package in Python, or use `install.packages` in R to install all software dependencies for a project.
-By relying on the idea of having research compendia [@gentleman2007statistical] or something as simple as a requirements file, one can easily install all the necessary scaffolding.
+
+In a modern computational workflow a user can `pip install` a package in Python, or use `install.packages` in R to install all software dependencies for a project.
+By relying on the idea of having research compendia [@gentleman2007statistical] or something as simple as a requirements file, it is relatively straightforward to install all the necessary scaffolding.
 Data on the other hand are rarely accessible that easily.

### 2020-02-05 paper.Rmd
implements suggestions from #109. Closes #109

@@
 \* = Both authors contributed equally to the work
+
+Correspondence to:
+
+Nicholas Tierney (nicholas.tierney@gmail.com) & Karthik Ram (k@karthik.io)
 
@@
 The README and data dictionary provides _human readable_ information on the data.
-To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata.
+To ensure data types are preserved - for example, dates are dates, names are characters - there needs to be some form of _machine readable_ metadata.
 This provides a structure allowing the data to be indexed and searched online, through services such as google datasets search [@google-data-search].
-An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema].
+An excellent standard for metadata is Table Schema written by frictionless data [@Fowler2017].
 For example, a dataset "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), and the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).
@@
 
-Any code used to clean and tidy the raw data should be provided in the `data-raw` directory.
+Any code used to clean and tidy the raw data, should be provided in the `data-raw` directory.
+This cleaned up data, analysis ready data should be placed in the `data/` directory.
 Ideally this would involve only scripted languages, but if other practical steps were taken to clean up the data, these should be recorded in a plain text or markdown file.
@@
 
-In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databases.
+In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or in a serverless database.
 This approach **gets the data somewhere rather than nowhere**.
@@
 In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data.
-For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank), and so on.
+For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank).
 
@@
 Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that allow  checksum file comparisons for versioning.
-Note that is different to Frictionless Data's tabular data package spec [@frictionlessdata-data-pkg].
+Note that is different to Frictionless Data's tabular data package spec [@Fowler2017].
 
@@
 
-```{r sdss-spectra, fig.cap = "Spectra image of wavelength against flux from SDSS from two sources.
+```{r sdss-spectra, fig.cap = "Visualisations from SDSS can be effecticely replicated.
+The spectra image of wavelength against flux from SDSS from two sources.
 Part A shows the example image taken from SDSS website.
-Part B shows the replicated figure using data provided as CSV, demonstrating great reproducibility.", out.width = "90%", fig.align = "center"}
+Part B shows the replicated figure using data provided as CSV - although there is some data cleaning or filtering since the numbers are not identical, it still demonstrates the great potential of reproducibility.", out.width = "90%", fig.align = "center"}
 

### 2020-02-06 paper.Rmd
keep_tex: true

@@
   bookdown::pdf_document2:
+    keep_tex: true
     toc: no

### 2020-02-06 paper.tex
keep_tex: true

--- 2020-02-05 paper.Rmd
@@
----
-title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.
-author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)
-output:
-  bookdown::pdf_document2:
-    toc: no
-    includes:
-      in_header: preamble.tex
-bibliography: references.bib
-header-includes: 
-  - \linespread{2} 
-  - \usepackage[utf8]{inputenc}
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.},
+  pdfauthor={Nicholas J Tierney* (1,2) \& Karthik Ram* (3)},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+%% Any special functions or other packages can be loaded here.
+
+%% FIGURE FLOATING taken from
+%% https://stackoverflow.com/questions/16626462/figure-position-in-markdown-when-converting-to-pdf-with-knitr-and-pandoc
+\usepackage{float}
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+    \expandafter\origfigure\expandafter[H]
+} {
+    \endorigfigure
+}
+
+%% MONASH STUFF TAKEN FROM ROB
+
+%% SPACING
+\RequirePackage{setspace}
+\spacing{1.5}
+
+%% HEADERS AND FOOTERS
+\RequirePackage{fancyhdr}
+\pagestyle{fancy}
+\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
+\makeatletter
+\lhead{\textsf{\expandafter{\@title}}}
+\makeatother
+\rhead{}
+\cfoot{}
+\setlength{\headheight}{15pt}
+\renewcommand{\headrulewidth}{0.4pt}
+\renewcommand{\footrulewidth}{0.4pt}
+\fancypagestyle{plain}{%
+\fancyhf{} % clear all header and footer fields
+\fancyfoot[C]{\sffamily\thepage} % except the center
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrulewidth}{0pt}}
+
+%% LINE AND PAGE BREAKING
+\sloppy
+\clubpenalty = 10000
+\widowpenalty = 10000
+\brokenpenalty = 10000
+\RequirePackage{microtype}
+
+%% PARAGRAPH BREAKS
+\setlength{\parskip}{1.4ex}
+\setlength{\parindent}{0em}
+
+%% CAPTIONS
+\RequirePackage{caption}
+\DeclareCaptionStyle{italic}[justification=centering]
+ {labelfont={bf},textfont={it},labelsep=colon}
+\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
+\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
+\linespread{2}
+\usepackage[utf8]{inputenc}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
+
+\title{A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.}
+\author{Nicholas J Tierney* (1,2) \& Karthik Ram* (3)}
+\date{}
+
+\begin{document}
+\maketitle
 
@@
 
-\* = Both authors contributed equally to the work
+* = Both authors contributed equally to the work
 
@@
 
-Nicholas Tierney (nicholas.tierney@gmail.com) & Karthik Ram (k@karthik.io)
+Nicholas Tierney (\href{mailto:nicholas.tierney@gmail.com}{\nolinkurl{nicholas.tierney@gmail.com}}) \& Karthik Ram (\href{mailto:k@karthik.io}{\nolinkurl{k@karthik.io}})
 
-# Abstract {-}
+\hypertarget{abstract}{%
+\section*{Abstract}\label{abstract}}
+\addcontentsline{toc}{section}{Abstract}
 
@@
 
-```{r setup, echo = FALSE}
+\hypertarget{intro}{%
+\section{Introduction}\label{intro}}
 
-knitr::opts_chunk$set(echo = FALSE,
-                      out.width = "100%")
-
-suppressMessages(library(ggrepel))
-suppressMessages(library(tibble))
-suppressMessages(library(kableExtra))
-suppressMessages(library(magick))
-suppressMessages(library(ggplot2))
-suppressMessages(library(cowplot))
-suppressMessages(library(gridExtra))
-```
-
-# Introduction {#intro}
-
-> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (The Adventure of the Copper Beeches by Sir Arthur Conan Doyle)
+\begin{quote}
+``Data! data! data!'' he cried impatiently. ``I can't make bricks without clay.'' - Sherlock Holmes (The Adventure of the Copper Beeches by Sir Arthur Conan Doyle)
+\end{quote}
 
@@
 Without access to good data, it becomes extremely difficult, if not impossible, to advance science.
-Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors [@Rowhani-Farid2016; @Stodden2018].
-Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work [@Popkin2019; @Kirk2019].
+Yet, a large majority of data on which published research papers are based rarely see the light of day and are only visible to the original authors (Rowhani-Farid and Barnett 2016; Stodden, Seiler, and Ma 2018).
+Sharing data sets upon publication of a research paper has benefits to individual researchers, often through increased visibility of the work (Popkin 2019; Kirk and Norton 2019).
 But there is also a component of benefit to the broader scientific community.
-This primarily comes in the form of potential for reuse in other contexts along use for training and teaching (@McKiernan2016).
+This primarily comes in the form of potential for reuse in other contexts along use for training and teaching (McKiernan et al.
+(2016)).
 Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing does not put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
@@
 To meet these minimal standards, there must be easy access to the data, models, and code.
-Among the different areas with a computational bent, the bioinformatics community in particular has a strong culture around open source [@Gentleman2004], and has made releasing code and associated software a recognized mainstream activity.
-Many journals in these fields have also pushed authors to submit code (in the form of model implementations) alongside their papers, with a few journals going as far as providing a "reproducibility review" [@Peng2011].
+Among the different areas with a computational bent, the bioinformatics community in particular has a strong culture around open source (Gentleman et al.
+2004), and has made releasing code and associated software a recognized mainstream activity.
+Many journals in these fields have also pushed authors to submit code (in the form of model implementations) alongside their papers, with a few journals going as far as providing a ``reproducibility review'' (Peng 2011).
 
-**In this paper we focus on the practical side of sharing data for the purpose of reproducibility.** Our goal is to describe various methods in which an analyst can share their data with minimal friction.
-We steer clear of idealistic ideas such as the FAIR data principles [@Wilkinson2016] since they still do not help a researcher share their data.
+\textbf{In this paper we focus on the practical side of sharing data for the purpose of reproducibility.} Our goal is to describe various methods in which an analyst can share their data with minimal friction.
+We steer clear of idealistic ideas such as the FAIR data principles (Wilkinson et al.
+2016) since they still do not help a researcher share their data.
 We also skip the discussion around citation and credit because data citations are still poorly tracked and there is no standardized metric or a h-index equivalent for data as of this writing.
@@
 For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) Code; 2) Computing environment, and 3) Data.
-The first two of these challenges have largely been solved [@Poisot2019].
+The first two of these challenges have largely been solved (Poisot et al.
+2019).
 
-Although code sharing in science had a rocky start [@Barnes2010], more and more code writing by scientists is being shared, partly due to the rapid increase in training opportunities made available by organizations like The Carpentries, combined with the rapid adoption of Github by scientists [@Ram2013].
-The bigger driver for this may also be connected to the rise in popularity of data science as a discipline distinct from statistics [@Donoho2017].
+Although code sharing in science had a rocky start (Barnes 2010), more and more code writing by scientists is being shared, partly due to the rapid increase in training opportunities made available by organizations like The Carpentries, combined with the rapid adoption of Github by scientists (Ram 2013).
+The bigger driver for this may also be connected to the rise in popularity of data science as a discipline distinct from statistics (Donoho 2017).
 This rapid growth in data science has largely been fueled by easy access to open source software tools.
-Programming languages such as Python, R and Julia help scientists implement and share new methods to work with data [@rcore; @python; @julia].
+Programming languages such as Python, R and Julia help scientists implement and share new methods to work with data (R Core Team 2019; ``Python,'' n.d.; Bezanson et al.
+2017).
 Each of these languages is supported by thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
-As of this writing, Python, Julia, and R have  167k packages [@pypi], ~ 14k packages [@cran] and ~2k  packages [@julia-pkgman] respectively.
+As of this writing, Python, Julia, and R have 167k packages (``Pypi,'' n.d.), \textasciitilde{} 14k packages (``Cran,'' n.d.) and \textasciitilde2k packages (``Julia-Pkgman,'' n.d.) respectively.
 These packages form the building blocks of a data scientists daily workflow.
 
-In a modern computational workflow a user can `pip install` a package in Python, or use `install.packages` in R to install all software dependencies for a project.
-By relying on the idea of having research compendia [@gentleman2007statistical] or something as simple as a requirements file, it is relatively straightforward to install all the necessary scaffolding.
+In a modern computational workflow a user can \texttt{pip\ install} a package in Python, or use \texttt{install.packages} in R to install all software dependencies for a project.
+By relying on the idea of having research compendia (Gentleman and Temple Lang 2007) or something as simple as a requirements file, it is relatively straightforward to install all the necessary scaffolding.
 Data on the other hand are rarely accessible that easily.
@@
 
-The computing environment is also easily captured with modern tools such as Docker [@Boettiger2015; @Jupyter2018].
-Modern tools such as Binder [@Jupyter2018] can parse Docker files and dependency trees to provide on demand, live notebooks in R and Python that a reader can immediately execute in the browser without dealing with the challenges of local installation.
+The computing environment is also easily captured with modern tools such as Docker (Boettiger 2015; Jupyter 2018).
+Modern tools such as Binder (Jupyter 2018) can parse Docker files and dependency trees to provide on demand, live notebooks in R and Python that a reader can immediately execute in the browser without dealing with the challenges of local installation.
 This makes it simple to load a specific environment to run any analysis.
-Code is handled by version control with tools like Git and GitHub [@git; @github], paired with archiving frameworks such as Zenodo provide access to code (particularly model implementations)[@zenodo].
+Code is handled by version control with tools like Git and GitHub (``Git,'' n.d.; ``Github,'' n.d.), paired with archiving frameworks such as Zenodo provide access to code (particularly model implementations)(Zenodo 2016).
 All the necessary software is available from various package managers (and their numerous geographic mirrors and archives) making it easy to install any version of a software package.
@@
 Datasets are often far more diverse than code in terms of complexity, size, and formats.
-This makes them particularly challenging to standardize or easily "install" where the code is running.
+This makes them particularly challenging to standardize or easily ``install'' where the code is running.
 While there are numerous public and private data repositories, none of them function as package managers, which is what provides so much robustness to code sharing.
-As a result, data used in an analysis is often read from various locations (local, network, or cloud), various formats, varying levels of tidiness [@Wickham2014].
+As a result, data used in an analysis is often read from various locations (local, network, or cloud), various formats, varying levels of tidiness (Wickham 2014).
 There is also the overhead associated with data publishing, the act of archiving data in a repository that also mints permanent identifiers, that are not required of all projects due to the effort and time involved.
@@
 This is the question we tackle in this paper.
-While there are several papers that serve as best-practice guides for formatting data [@Broman2017] and getting them ready for sharing [@Ellis2017], the aims of this paper are a bit different.
+While there are several papers that serve as best-practice guides for formatting data (Broman and Woo 2017) and getting them ready for sharing (Ellis and Leek 2017), the aims of this paper are a bit different.
 Our aim is to address the issue of data in the context of reproducibility in data science workflows.
@@
 We discuss how to share and/or publish data and cover various tradeoffs when deciding how much to do.
-We skip detailed discussions of data preparation [@Broman2017; @Ellis2017], or citation (as of this writing, data citation is still in its infancy).
+We skip detailed discussions of data preparation (Broman and Woo 2017; Ellis and Leek 2017), or citation (as of this writing, data citation is still in its infancy).
 We also analyze the state of data contained inside software packages, shared and made available as part of modern data analysis.
@@
 
-# A minimal structure to share data for analysis {#minimal}
+\hypertarget{minimal}{%
+\section{A minimal structure to share data for analysis}\label{minimal}}
 
 To share data analysis data, there should be some minimal set of requirements.
-For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)).
+For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \ref{minimal}).
 No matter where data is submitted, there should ideally be a canonical data repository in one long term archive that links to others.
-It should also have an accession number, or Digital Object Identification (DOI) number, which allows for it to be cited (discussed further in \@ref(cite)).
+It should also have an accession number, or Digital Object Identification (DOI) number, which allows for it to be cited (discussed further in \ref{cite}).
 
@@
 
-1. README: A Human readable description of the data
-2. Data dictionary: Human readable dictionary of data contents
-3. License: How to use and share the data
-4. Citation: How you want your data to be cited
-5. Machine readable meta data: Make your data searchable
-6. Raw data: The original/first data provided
-7. Scripts: To clean raw data ready for analysis
-8. Analysis ready data: Final data used in analysis
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  README: A Human readable description of the data
+\item
+  Data dictionary: Human readable dictionary of data contents
+\item
+  License: How to use and share the data
+\item
+  Citation: How you want your data to be cited
+\item
+  Machine readable meta data: Make your data searchable
+\item
+  Raw data: The original/first data provided
+\item
+  Scripts: To clean raw data ready for analysis
+\item
+  Analysis ready data: Final data used in analysis
+\end{enumerate}
 
-One basic suggested directory layout is given below in Figure \@ref(fig:code-directory).
+One basic suggested directory layout is given below in Figure \ref{fig:code-directory}.
 
-```{r code-directory, out.width = "75%", fig.align = "center", fig.cap = "Example directory layout and structure for a data repository."}
-knitr::include_graphics(here::here("figures",
-                                   "directory-structure.png"))
-```
+\begin{figure}
+
+{\centering \includegraphics[width=0.75\linewidth]{/Users/ntie0001/github/njtierney/ddd/figures/directory-structure}
+
+}
+
+\caption{Example directory layout and structure for a data repository.}\label{fig:code-directory}
+\end{figure}
 
@@
 
-1. README
-2. Data dictionary
-3. License
-4. Citation
-5. Analysis ready data
-6. Scripts to tidy analysis raw data into ready data
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  README
+\item
+  Data dictionary
+\item
+  License
+\item
+  Citation
+\item
+  Analysis ready data
+\item
+  Scripts to tidy analysis raw data into ready data
+\end{enumerate}
 
@@
 
-## README: A Human readable description of the data {#readme}
+\hypertarget{readme}{%
+\subsection{README: A Human readable description of the data}\label{readme}}
 
@@
 A README is often the first place people will go to learn more about anything in a folder - they are very common in software, and historically were included so the uppercase letters of README meant it would be at the top of a directory.
-The README is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":
+The README is meant for someone to read and understand more about the data and contains the ``who, what, when, where, why, \& how'':
 
-- **Who** collected it
-- **What** is the data
-- **When** was it collected
-- **Where** was it collected
-- **Why** it was collected
-- **How** is was collected
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Who} collected it
+\item
+  \textbf{What} is the data
+\item
+  \textbf{When} was it collected
+\item
+  \textbf{Where} was it collected
+\item
+  \textbf{Why} it was collected
+\item
+  \textbf{How} is was collected
+\end{itemize}
 
@@
 
-Saving a README with the extension `.md` file gives the author the formatting benefits of `markdown`, making it simple to insert links, tables, and make lists.
+Saving a README with the extension \texttt{.md} file gives the author the formatting benefits of \texttt{markdown}, making it simple to insert links, tables, and make lists.
 In systems like GitHub, a README file is detected and rendered in nice HTML on the repository by default.
 
-## Data dictionary: Human readable dictionary of data contents {#data-dictionary}
+\hypertarget{data-dictionary}{%
+\subsection{Data dictionary: Human readable dictionary of data contents}\label{data-dictionary}}
 
@@
 
-- **variable names**
-- **variable labels**
-- **variable codes**, and
-- special values for **missing data**.
+\begin{itemize}
+\tightlist
+\item
+  \textbf{variable names}
+\item
+  \textbf{variable labels}
+\item
+  \textbf{variable codes}, and
+\item
+  special values for \textbf{missing data}.
+\end{itemize}
 
-Figure \@ref(fig:variables) shows an example of the data and data dictionary details.
-**Variable names** should be short, descriptive names with no spaces or special characters.
-In the most common tabular data case, these correspond to column names, for example, "job_position", "faculty_level", and "years_at_uni".
-**Variable labels** are longer descriptions of variables.
-For example "University Job Position", "University Faculty Position Level", and  "Number of Years Worked at University" [@McGill-codebook; @Arslan2019].
-**Variable codes** apply to categorical (factor) variables, and are the values for their contents.
-For example, a variable could contain answers to a question, with values and codes like so: `0 = no, 1 = yes`, to indicate if someone is in statistics, for example.
-These should be consistent across similar variables to avoid problems where `0 = yes` for one variable, and `1 = yes` in another.
+Figure \ref{fig:variables} shows an example of the data and data dictionary details.
+\textbf{Variable names} should be short, descriptive names with no spaces or special characters.
+In the most common tabular data case, these correspond to column names, for example, ``job\_position'', ``faculty\_level'', and ``years\_at\_uni''.
+\textbf{Variable labels} are longer descriptions of variables.
+For example ``University Job Position'', ``University Faculty Position Level'', and ``Number of Years Worked at University'' (``Codebook Cookbook: A Guide to Writing a Good Codebook for Data Analysis Projects in Medicine'' n.d.; Arslan 2019).
+\textbf{Variable codes} apply to categorical (factor) variables, and are the values for their contents.
+For example, a variable could contain answers to a question, with values and codes like so: \texttt{0\ =\ no,\ 1\ =\ yes}, to indicate if someone is in statistics, for example.
+These should be consistent across similar variables to avoid problems where \texttt{0\ =\ yes} for one variable, and \texttt{1\ =\ yes} in another.
 Date variables should have consistent formatting.
-For example, all date information could be in format "YYYY-MM-DD", and this should not be mixed with "YYYY-DD-MM".
+For example, all date information could be in format ``YYYY-MM-DD'', and this should not be mixed with ``YYYY-DD-MM''.
 Date formatting should be described in the variable labels.
-**Missing data** are values that should have been observed, but were not.
-The code for missingness should be documented in the data dictionary, and should nominally be `NA`.
+\textbf{Missing data} are values that should have been observed, but were not.
+The code for missingness should be documented in the data dictionary, and should nominally be \texttt{NA}.
 If the reason for missingness is known it should be recorded.
-For example, censored data, patient drop out, or measurement error can have different values, such as "unknown", -99, or other value codes [@White2013; @Broman2017].
+For example, censored data, patient drop out, or measurement error can have different values, such as ``unknown'', -99, or other value codes (White et al.
+2013; Broman and Woo 2017).
 
-```{r variables, fig.cap = "Four tables - the data, and it's variable names, variable codes, and the meaning of missing values", fig.height = 8}
+\begin{figure}
+\includegraphics[width=1\linewidth]{paper_files/figure-latex/variables-1} \caption{Four tables - the data, and it's variable names, variable codes, and the meaning of missing values}\label{fig:variables}
+\end{figure}
 
-dat <- tribble(
-  ~job_position, ~outside, ~statistics, ~date_started,
-   "Lecturer",        1,        1,      "2019-01-01",
-   "Professor",       0,        0,      NA,
-   "Senior Lecturer", 1,        1,      "unknown"
-)
-
-variable_labels <- tribble(
-  ~variable, ~description,
-  "job_position", "University Job Position",
-  "outside",      "Did this person study outside this university?",
-  "statistics",   "Is this person in the statistics faculty",
-  "date_started", "What date did this perosn start their position?"
-)
-
-variable_codes <- tribble(
-  ~variable,        ~code,
-   "job_position",   "text description of job position",
-   "outside",       "0 = not studied outside, \n1 = has studied outside",
-   "statistics",    "0 = is not in statistics dept., \n1 = is currently in statistics dept.",
-   "date_started",   "date variable in format YYYY-MM-DD, \ne.g., 2001-01-01 is 1st January, 2001"
-)
-
-missing_values <- tribble(
-  ~value, ~meaning,
-  NA,     "Missing value",
-  "unknown", "Missing value"
-)
-
-grid.arrange(
-  tableGrob(dat),
-  tableGrob(variable_labels),
-  tableGrob(variable_codes),
-  tableGrob(missing_values),
-  ncol = 1
-)
-
-```
-
-Table \@ref(tab:table-data-incarcerate) shows an example data dictionary table from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate].
+Table \ref{tab:table-data-incarcerate} shows an example data dictionary table from the Tidy Tuesday repository on incarceration trends (``Tidy-Tuesday-Incarcerate,'' n.d.).
 This includes information on the variable, its class (type), and a longer description.
 
-```{r table-data-incarcerate}
-dictionary_incarcerate <- data.frame(
-  Variable = c("year", "urbanicity", "pop_category", "rate_per_100000"),
-  Class  = c("integer (date)",
-             "character",
-             "character",
-             "double"),
-  Description = c("Year",
-                  "County-type (urban, suburban, small/mid, rural)",
-                  "Category for population - either race, gender, or Total",
-                  "Rate within a category for prison population per 100,000 people")
-)
+\begin{table}[!h]
 
-knitr::kable(dictionary_incarcerate,
-             caption = "The prisoner summary data dictionary, with columns for the variable, its class, and a short description of the contents of the variable.",
-             booktabs = TRUE) %>%
-  kable_styling(latex_options = c("hold_position"))
-```
+\caption{\label{tab:table-data-incarcerate}The prisoner summary data dictionary, with columns for the variable, its class, and a short description of the contents of the variable.}
+\centering
+\begin{tabular}[t]{lll}
+\toprule
+Variable & Class & Description\\
+\midrule
+year & integer (date) & Year\\
+urbanicity & character & County-type (urban, suburban, small/mid, rural)\\
+pop\_category & character & Category for population - either race, gender, or Total\\
+rate\_per\_100000 & double & Rate within a category for prison population per 100,000 people\\
+\bottomrule
+\end{tabular}
+\end{table}
 
 Data dictionaries should be placed in the README and presented as a table.
-Every data dictionary should also be provided in its raw form (e.g., a CSV) in the repository, so they aren't "trapped" in the README.
+Every data dictionary should also be provided in its raw form (e.g., a CSV) in the repository, so they aren't ``trapped'' in the README.
 
-## License: How to use and share the data {#license}
+\hypertarget{license}{%
+\subsection{License: How to use and share the data}\label{license}}
 
@@
 
-1. Creative Commons Attribution 4.0 International Public License (CC BY), and
-2. Creative Commons CC0 1.0 Universal (CC0)
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Creative Commons Attribution 4.0 International Public License (CC BY), and
+\item
+  Creative Commons CC0 1.0 Universal (CC0)
+\end{enumerate}
 
-Other licenses or notices to be aware of are **copyrighted data**, and **data embargoes**.
-If you are working with data **already copyrighted**, (for example under CC BY or CC0), you must give follow appropriate guidelines for giving credit.
-Data may also be under **embargo**.
+Other licenses or notices to be aware of are \textbf{copyrighted data}, and \textbf{data embargoes}.
+If you are working with data \textbf{already copyrighted}, (for example under CC BY or CC0), you must give follow appropriate guidelines for giving credit.
+Data may also be under \textbf{embargo}.
 This means data cannot be shared more widely until a specific release time.
 If sharing data under an embargo, include detailed information on the embargo requirements in: the README, and in separate correspondence with those who receive the data.
-Databases may also be licensed with the Open Data Commons Open Database License: https://opendatacommons.org/licenses/odbl/, which provides provisions for sharing, creating, and adapting, provided that work is attributed, shared, and kept open.
+Databases may also be licensed with the Open Data Commons Open Database License: \url{https://opendatacommons.org/licenses/odbl/}, which provides provisions for sharing, creating, and adapting, provided that work is attributed, shared, and kept open.
 
@@
 
-### CC BY
+\hypertarget{cc-by}{%
+\subsubsection{CC BY}\label{cc-by}}
 
@@
 
-* You must provide appropriate credit to the source. This means listing the names of the creators.
-* Link back to the CC BY license, and
-* Clearly show if changes were made.
-* Data cannot be sub-licensed, that is - a change to the existing license
-* There is also no warranty, so the person or people who obtained the data cannot be held liable.
+\begin{itemize}
+\tightlist
+\item
+  You must provide appropriate credit to the source. This means listing the names of the creators.
+\item
+  Link back to the CC BY license, and
+\item
+  Clearly show if changes were made.
+\item
+  Data cannot be sub-licensed, that is - a change to the existing license
+\item
+  There is also no warranty, so the person or people who obtained the data cannot be held liable.
+\end{itemize}
 
-The journal PLOS Comp Bio requires that data submitted cannot be more restrictive than CC BY [@plos-comp-bio-data].
-For a brief overview of the CC BY, suitable to include in a README, see @ccby-short.
-For the full license, see @ccby-long.
+The journal PLOS Comp Bio requires that data submitted cannot be more restrictive than CC BY (``PLOS Computational Biology,'' n.d.).
+For a brief overview of the CC BY, suitable to include in a README, see (``CCBY Short Guide,'' n.d.).
+For the full license, see (``CCBY Full License,'' n.d.).
 
-### CC0
+\hypertarget{cc0}{%
+\subsubsection{CC0}\label{cc0}}
 
-The CC0 is a "public domain" license.
-Data with a CC0 license means the data owners waive all their rights to the work, and it now "owned" by the public.
-The data can be freely shared, copied, modified, and distributed, even for commercial purposes _without asking permission_.
+The CC0 is a ``public domain'' license.
+Data with a CC0 license means the data owners waive all their rights to the work, and it now ``owned'' by the public.
+The data can be freely shared, copied, modified, and distributed, even for commercial purposes \emph{without asking permission}.
 When using data with CC0, it is good practice to cite the original paper, but it is not required.
-If you wish to use the CC0, see @choose-cc0.
-For a brief overview of the CC0, see @cc0-short, and for the full license, see @cc0-long.
+If you wish to use the CC0, see (``Choose-Cc0,'' n.d.).
+For a brief overview of the CC0, see (``CC0 Short Guide,'' n.d.), and for the full license, see (``CC0 Full License,'' n.d.).
 
-## Citation: How you want your data to be cited {#cite}
+\hypertarget{cite}{%
+\subsection{Citation: How you want your data to be cited}\label{cite}}
 
@@
 For example, if the data are deposited in repositories like Dryad, Zenodo, or the Open Science Framework, the best practice would be to copy the citation created by these repositories.
-Under the hood, the DOI is "minted" by DataCite, who provide DOIs for research outputs and data.
+Under the hood, the DOI is ``minted'' by DataCite, who provide DOIs for research outputs and data.
 This means that when citing data, it only makes sense to cite datasets that have been deposited into a DataCite compliant repository.
 If a DOI is unavailable, a citation will be meaningless, as it cannot be tracked by any means.
-A file named `citation` should be placed in the directory, at the same level as the `README`.
-It should contain a DOI, and could be in `.bibtex` format.
+A file named \texttt{citation} should be placed in the directory, at the same level as the \texttt{README}.
+It should contain a DOI, and could be in \texttt{.bibtex} format.
 
-## Machine readable metadata {#machine-metadata}
+\hypertarget{machine-metadata}{%
+\subsection{Machine readable metadata}\label{machine-metadata}}
 
-The README and data dictionary provides _human readable_ information on the data.
-To ensure data types are preserved - for example, dates are dates, names are characters - there needs to be some form of _machine readable_ metadata.
-This provides a structure allowing the data to be indexed and searched online, through services such as google datasets search [@google-data-search].
-An excellent standard for metadata is Table Schema written by frictionless data [@Fowler2017].
-For example, a dataset "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), and the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).
+The README and data dictionary provides \emph{human readable} information on the data.
+To ensure data types are preserved - for example, dates are dates, names are characters - there needs to be some form of \emph{machine readable} metadata.
+This provides a structure allowing the data to be indexed and searched online, through services such as google datasets search (Castelvecchi 2018).
+An excellent standard for metadata is Table Schema written by frictionless data (Fowler, Barratt, and Walsh 2017).
+For example, a dataset ``demographics'' with three variables is shown in Table \ref{tab:table-machine-readable}, and the Java Script Object Notation (JSON) equivalent in Figure \ref{fig:json-code}.
 
-```{r table-machine-readable}
-table_machine_readable <- data.frame(
-  age = c(12L, 21L, 37L),
-  height = c(161.5, 181.2, 178.3),
-  nationality = c("Australian", "American", "New Zealand")
-)
+\begin{table}[!h]
 
-knitr::kable(table_machine_readable,
-             caption = "Example demographics table of age, height, and nationality.",
-             booktabs = TRUE) %>%
-  kable_styling(latex_options = c("hold_position"))
-```
+\caption{\label{tab:table-machine-readable}Example demographics table of age, height, and nationality.}
+\centering
+\begin{tabular}[t]{rrl}
+\toprule
+age & height & nationality\\
+\midrule
+12 & 161.5 & Australian\\
+21 & 181.2 & American\\
+37 & 178.3 & New Zealand\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-```{r json-code, fig.align = "center", out.width = "75%", fig.cap = "Example snippet of some Table Schema data for a dataset with three variables.
-This provides a description of each field, and the type of field."}
-knitr::include_graphics(here::here("figures",
-                                   "json-code.png"))
-```
+\begin{figure}
 
-This contains fields such as "path", describing the file path to the data, and a schema with sub-fields name and type, for each variable.
+{\centering \includegraphics[width=0.75\linewidth]{/Users/ntie0001/github/njtierney/ddd/figures/json-code}
+
+}
+
+\caption{Example snippet of some Table Schema data for a dataset with three variables. This provides a description of each field, and the type of field.}\label{fig:json-code}
+\end{figure}
+
+This contains fields such as ``path'', describing the file path to the data, and a schema with sub-fields name and type, for each variable.
 It also provides information for licensing and features such as line breaks, and delimiters, useful to ensure the data is correctly imported into a machine.
 It is built on JSON, a lightweight, human-friendly, machine readable data-interchange format.
-Table schema is baked into formats such as csvy [@csvy], an extended `csv` format, which has additional front matter in a lightweight markup format (YAML) using Table Schema.
+Table schema is baked into formats such as csvy (``Csvy,'' n.d.), an extended \texttt{csv} format, which has additional front matter in a lightweight markup format (YAML) using Table Schema.
 
@@
 
-Another rich format is XML, the e**X**tensible **M**arkup **L**anguage.
+Another rich format is XML, the e\textbf{X}tensible \textbf{M}arkup \textbf{L}anguage.
 This was an early iteration on the idea of a plain text format that was both human and machine readable.
@@
 JSON is shorter, and quicker and easier to read and write than XML, and can also be parsed with standard JavaScript, whereas XML must be parsed by a special XML parser.
-XML and JSON can be converted into each others respective formats, for example with the `json` or `xml2` R packages.
-XML is still used as metadata for data storage, for example in EML, the Ecological Metadata Language [@EML-about].
+XML and JSON can be converted into each others respective formats, for example with the \texttt{json} or \texttt{xml2} R packages.
+XML is still used as metadata for data storage, for example in EML, the Ecological Metadata Language (n.d.).
 
-To create appropriate metadata, we recommend metadata generators such as `dataspice` or `codebook` [@dataspice; @Arslan2019].
+To create appropriate metadata, we recommend metadata generators such as \texttt{dataspice} or \texttt{codebook} (\emph{Dataspice: Create Lightweight Schema.org Descriptions of Dataset} 2018; Arslan 2019).
 These help automatically create the structured JSON data.
-This metadata should be provided in a folder called "metadata", which should be provided for every dataset.
+This metadata should be provided in a folder called ``metadata'', which should be provided for every dataset.
 
-## Raw data: The original or first data provided {#raw-data}
+\hypertarget{raw-data}{%
+\subsection{Raw data: The original or first data provided}\label{raw-data}}
 
 Raw data is usually the first format of the data provided before any tidying or cleaning of the data.
-If the raw data is a practical size to share, it should be shared in a folder called `data-raw`.
+If the raw data is a practical size to share, it should be shared in a folder called \texttt{data-raw}.
 The raw data should be in the form that was first received, even if it is in binary or some proprietary format.
@@
 
-## Scripts: To clean raw data ready for analysis {#scripts}
+\hypertarget{scripts}{%
+\subsection{Scripts: To clean raw data ready for analysis}\label{scripts}}
 
-Any code used to clean and tidy the raw data, should be provided in the `data-raw` directory.
-This cleaned up data, analysis ready data should be placed in the `data/` directory.
+Any code used to clean and tidy the raw data, should be provided in the \texttt{data-raw} directory.
+This cleaned up data, analysis ready data should be placed in the \texttt{data/} directory.
 Ideally this would involve only scripted languages, but if other practical steps were taken to clean up the data, these should be recorded in a plain text or markdown file.
 
-## Analysis ready data: Final data used in analysis {#final-data}
+\hypertarget{final-data}{%
+\subsection{Analysis ready data: Final data used in analysis}\label{final-data}}
 
-The data used in the data analysis should be provided in a folder called `data`.
-Ideally, the data "analysis ready data" should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, observations in rows, and only one value per cell.
+The data used in the data analysis should be provided in a folder called \texttt{data}.
+Ideally, the data ``analysis ready data'' should be in ``Tidy Data'' format (Wickham 2014), where tidy data contains variables in columns, observations in rows, and only one value per cell.
 
-In contrast to "raw data", "analysis ready data" should be in an easily readable plain-text format, such as comma-, tab- or semicolon-separated files.
-These typically have file extensions like `.csv`, `.tsv` or `.txt`.
+In contrast to ``raw data'', ``analysis ready data'' should be in an easily readable plain-text format, such as comma-, tab- or semicolon-separated files.
+These typically have file extensions like \texttt{.csv}, \texttt{.tsv} or \texttt{.txt}.
 Binary or proprietary formats are discouraged in favor of interoperability, as it requires special software to read, even if it is sometimes slower to read in and out, and harder to share due to size.
-For example, R data formats like `.rds`, `.rda`, or SPSS, STATA, or SAS data formats, `.sav`, `.dta`, or `.sas7bdat`.
+For example, R data formats like \texttt{.rds}, \texttt{.rda}, or SPSS, STATA, or SAS data formats, \texttt{.sav}, \texttt{.dta}, or \texttt{.sas7bdat}.
 
-# Releasing data in R {#data-release-r}
+\hypertarget{data-release-r}{%
+\section{Releasing data in R}\label{data-release-r}}
 
@@
 The R ecosystem has many noteworthy examples of data-only packages.
-One exemplar is the `nycflights13` package by Hadley Wickham [@nycflights], This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes.
-The package not only provides ready to use binary data but also ships raw data (in a folder called `data-raw`) along with scripts used to clean them.
+One exemplar is the \texttt{nycflights13} package by Hadley Wickham (Wickham 2018), This package makes available airline data for all flights departing New York city in 2013 in a tidy format, with distinct tables for metadata on airlines, airports, weather, and planes.
+The package not only provides ready to use binary data but also ships raw data (in a folder called \texttt{data-raw}) along with scripts used to clean them.
 The package was originally created as a way to provide example data to teach tidy data principles and serves as a great model for how one can publish a data package in R.
@@
 
-**Other benefits of packaging data in R**
+\textbf{Other benefits of packaging data in R}
 
-1.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
   Packaging data can be a very powerful pedagogical tool to help researchers understand how to transform data and prepare it for further analysis.
@@
 Users can also skip the raw data and scripts and proceed directly to the fast binary data, which can hold a lot of data when heavily compressed.
-
-2.
+\item
   When the primary motivation for shipping data is to illustrate visualization techniques or to run examples, one can skip the raw data and processing scripts and only include the binary data.
@@
 For cases when this size is hard to maintain, CRAN recommends data-only packages that will be rarely updated.
-For a detailed discussion on this issue and alternative approaches, see @Brooke_Anderson2017.
+For a detailed discussion on this issue and alternative approaches, see Brooke Anderson and Eddelbuettel (2017).
+\end{enumerate}
 
@@
 This is why we recommend, as a rule, that researchers also archive data in a long-term data repository.
-These include domain specific repositories (see Section \@ref(publish-repos)) or more general purpose ones such as Zenodo or Figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.
+These include domain specific repositories (see Section \ref{publish-repos}) or more general purpose ones such as Zenodo or Figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.
 
@@
 datasets either as binary data (5903 packages) or as external datasets
-(766). Binary files comprise a bulk of the data shipped in the `data`
-folder (68.06%) with other plain text formats such as txt, CSV, dat,
-`json` comprising less than one percent of data shipped in packages.
+(766). Binary files comprise a bulk of the data shipped in the \texttt{data}
+folder (68.06\%) with other plain text formats such as txt, CSV, dat,
+\texttt{json} comprising less than one percent of data shipped in packages.
 
-# Dealing with medium to large data {#med-large}
+\hypertarget{med-large}{%
+\section{Dealing with medium to large data}\label{med-large}}
 
 Another common situation that researchers face is in dealing with data that fall somewhere between small and large.
-For example, small data could be tabular, as a CSV, in the order of bytes to megabytes to gigabytes that can be easily compressed, and large could be what falls under the umbrella of **big data**.
-The happy medium is generally data that are too big to fit on memory of most standard machines, but can successfully fit on disk (https://ropensci.github.io/arkdb/articles/articles/noapi.html).
+For example, small data could be tabular, as a CSV, in the order of bytes to megabytes to gigabytes that can be easily compressed, and large could be what falls under the umbrella of \textbf{big data}.
+The happy medium is generally data that are too big to fit on memory of most standard machines, but can successfully fit on disk (\url{https://ropensci.github.io/arkdb/articles/articles/noapi.html}).
 In this case, users who do not have the support to maintain resource intensive database servers can instead rely on light serverless databases such as MonetDB or SQLite.
 These databases provide disk based storage and using language agnostic interfaces, a analyst can easily query these data in manageable chunks that don't overrun memory limitations.
-Using software packages such as `arkdb` [@arkdb] one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
-Another option is to break large text data files into chunks with named sequences (e.g., `teaching-1.csv`, `teaching-2.csv`, etc.).
+Using software packages such as \texttt{arkdb} (Boettiger 2018) one could easily chunk large data from flat text files to these lite databases without running into memory limitations.
+Another option is to break large text data files into chunks with named sequences (e.g., \texttt{teaching-1.csv}, \texttt{teaching-2.csv}, etc.).
 
@@
 This mechanism can also be used to attach arbitrary files such as large data files, binary data, and database dumps as long as each file does not exceed 2gb.
-The R package `piggyback` (`https://docs.ropensci.org/piggyback/`) allows for uploading and downloading such files to GitHub releases, making it easy for anyone to access data files wherever the script is being run.
+The R package \texttt{piggyback} (\texttt{https://docs.ropensci.org/piggyback/}) allows for uploading and downloading such files to GitHub releases, making it easy for anyone to access data files wherever the script is being run.
 We emphasize again that this is a short-term solution that is dependent on GitHub maintaining this service.
 
-# Challenges in documenting your dataset {#doco-challenge}
+\hypertarget{doco-challenge}{%
+\section{Challenges in documenting your dataset}\label{doco-challenge}}
 
@@
 This section discusses a framework for users, to help them decide how much they should document their data.
-To frame discussion around the challenges of data documenting, we can think of how an individual dataset falls on two features: "Effort to prepare", and "Ease of understanding" in Figure \@ref(fig:effort-understanding).
+To frame discussion around the challenges of data documenting, we can think of how an individual dataset falls on two features: ``Effort to prepare'', and ``Ease of understanding'' in Figure \ref{fig:effort-understanding}.
 The ideal space to be in the graph would be the top left hand corner.
@@
 
-```{r effort-understanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT).", out.width = "90%"}
-
-# There are two Axis:
-#   X axis: <- Robustness (brittle - robust)
-#   Y axis: <- Difficulty (easy - hard)
-
-dat_prep <- tibble(prep = c(15,20,55,90, 96, 40),
-               understanding = c(15,20,95,90,96, 91),
-                  label = c("Data Dump",
-                            "Most studies",
-                            "Sensor Data",
-                            "50 hectare",
-                            "RCT",
-                            "Simulation Study"
-                            ))
-
-ggplot(dat_prep,
-       aes(x = prep,
-           y = understanding,
-           label = label)) +
-  geom_point() +
-  geom_label_repel(
-    segment.size = 0.2,
-    segment.color = "grey50"
-    ) +
-  # theme_void() + 
-  theme(axis.line.x = element_line(arrow = arrow(length = unit(3, 'mm'))), 
-        axis.line.y = element_line(arrow = arrow(length = unit(3, 'mm'))),
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank(),
-        aspect.ratio = 1
-  ) + 
-  labs(x = "Effort to Prepare",
-       y = "Ease of Understanding") +
-  xlim(0,100) +
-  ylim(0, 100)
-
-```
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/effort-understanding-1} \caption{There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT).}\label{fig:effort-understanding}
+\end{figure}
 
@@
 The more impactful the data, the more documentation features to include.
-Figure \@ref(fig:data-sharing-workflow) shows the practical types of steps that can be taken against the effort required.
+Figure \ref{fig:data-sharing-workflow} shows the practical types of steps that can be taken against the effort required.
 
-```{r data-sharing-workflow, fig.cap = "The steps required compared to effort required for data sharing for reproducibility.", out.width = "90%"}
-knitr::include_graphics(here::here("figures",
-                                   "data_sharing_workflow.png"))
-```
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{/Users/ntie0001/github/njtierney/ddd/figures/data_sharing_workflow} \caption{The steps required compared to effort required for data sharing for reproducibility.}\label{fig:data-sharing-workflow}
+\end{figure}
 
@@
 
-# Publishing and repositories {#publish-repos}
+\hypertarget{publish-repos}{%
+\section{Publishing and repositories}\label{publish-repos}}
 
@@
 
-1. **Least Moving Parts**
-2. **Domain Specific Venue**
-3. **Long Term Archive**
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \textbf{Least Moving Parts}
+\item
+  \textbf{Domain Specific Venue}
+\item
+  \textbf{Long Term Archive}
+\end{enumerate}
 
@@
 
-In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or in a serverless database.
-This approach **gets the data somewhere rather than nowhere**.
+In \textbf{Least Moving Parts}, the data might be published with an R package, or as part of a GitHub release using piggyback (``Piggyback,'' n.d.), or in a serverless database.
+This approach \textbf{gets the data somewhere rather than nowhere}.
 Its minimal features means it is simple to maintain.
@@
 
-In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data.
-For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank).
+In \textbf{Domain Specific Venue}, data can be published in journal data papers, or venues specific to the type of data.
+For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (Benson et al.
+(2005)).
 
@@
 Many governments or civil organizations are now making their own data available through a government website interface.
-Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or may self host their data, such as five thirty eight [@data-538].
+Media and journalism groups are also making their data available either through places like GitHub (``Pudding-Data,'' n.d.), or may self host their data, such as five thirty eight (``Data-538,'' n.d.).
 
@@
 
-A **Long Term Archive** is the best option to share the data.
+A \textbf{Long Term Archive} is the best option to share the data.
 Long term archives provide information such as DOI (Digital Object Identifier) that make it easier to cite.
@@
 
-If the dataset you are shipping has a research application, the most relevant home would be the research data repositories **zenodo**, **dryad**, or **Open Science Framework**.
+If the dataset you are shipping has a research application, the most relevant home would be the research data repositories \textbf{zenodo}, \textbf{dryad}, or \textbf{Open Science Framework}.
 Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to deposit their datasets.
@@
 Zenodo is able to accommodate larger file sizes upon request.
-The Dryad Digital Repository (@dryad) will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, in exchange for a data publishing fee.
+The Dryad Digital Repository ((``Dryad,'' n.d.)) will take data from any field of research, and perform human quality control and assistance of the data, with the ability to link data with a journal publication, in exchange for a data publishing fee.
 The OSF is a tool that captures the research process online.
@@
 The entire history of the project is recorded, so it promotes centralized, and transparent workflows.
-Although not specifically designed for data, like Dryad or Zenodo, OSF does provide a DOI minting service, and a more wholistic approach, which might be appealing to users [@OSF].
+Although not specifically designed for data, like Dryad or Zenodo, OSF does provide a DOI minting service, and a more wholistic approach, which might be appealing to users (Spies et al.
+2012).
 
-## Publishing data in a data repository
+\hypertarget{publishing-data-in-a-data-repository}{%
+\subsection{Publishing data in a data repository}\label{publishing-data-in-a-data-repository}}
 
@@
 
-## Publishing data through a data journal
+\hypertarget{publishing-data-through-a-data-journal}{%
+\subsection{Publishing data through a data journal}\label{publishing-data-through-a-data-journal}}
 
-Data used in publications are often shared in the supplementary materials of articles, or served on repositories such as the Dryad Digital Repository (@dryad).
+Data used in publications are often shared in the supplementary materials of articles, or served on repositories such as the Dryad Digital Repository ((``Dryad,'' n.d.)).
 Dryad makes data from scientific publications discoverable, reusable, and citable.
@@
 
-To provide better context around the data used in research and better expose data for reuse, journals are now adding "data papers".
+To provide better context around the data used in research and better expose data for reuse, journals are now adding ``data papers''.
 These are specifically designed for publishing articles about the data, and sharing it.
@@
 Instead, they provide ideas on what they should include.
-This is a useful step towards improving data reuse, but it _lacks some minimal structure that allows a researcher to have a predictable way to access and interpret the data_.
+This is a useful step towards improving data reuse, but it \emph{lacks some minimal structure that allows a researcher to have a predictable way to access and interpret the data}.
 
-Other journals operating in this space include journals like "data in brief", "Data", and "Nature Scientific Data".
+Other journals operating in this space include journals like ``data in brief'', ``Data'', and ``Nature Scientific Data''.
 Guidelines for what is required in the content of the paper, and the required information along with the data (meta data, data format, etc.) vary by journal.
 
-# Tooling for packaging data {#tooling}
-
-<!-- NOTE: not sure if this quite fits here, tooling is important, but at the moment we are not sure if we should describe all tooling, as I will have a lot of examples for R, and not for Python/Julia -->
+\hypertarget{tooling}{%
+\section{Tooling for packaging data}\label{tooling}}
 
 Tooling for producing data documentation information speeds up and simplifies the process of sharing data.
-Machine-readable metadata that can be indexed by google is created using the `dataspice` package [@dataspice].
-To help create a data dictionary, we recommend the `codebook` package in R [@Arslan2019], which also generates machine readable metadata.
-Data dictionaries are implemented in other software such as STATA, which provides a "codebook" command.
-Data can be packaged up in a "data package" with DataPackageR [@DataPackageR], which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that allow  checksum file comparisons for versioning.
-Note that is different to Frictionless Data's tabular data package spec [@Fowler2017].
+Machine-readable metadata that can be indexed by google is created using the \texttt{dataspice} package (\emph{Dataspice: Create Lightweight Schema.org Descriptions of Dataset} 2018).
+To help create a data dictionary, we recommend the \texttt{codebook} package in R (Arslan 2019), which also generates machine readable metadata.
+Data dictionaries are implemented in other software such as STATA, which provides a ``codebook'' command.
+Data can be packaged up in a ``data package'' with DataPackageR (Greg Finak 2019), which provides tools to wrap up data into an R package, while providing helpers with MD5sum checks that allow checksum file comparisons for versioning.
+Note that is different to Frictionless Data's tabular data package spec (Fowler, Barratt, and Walsh 2017).
 
-# Example datasets {#example-data}
+\hypertarget{example-data}{%
+\section{Example datasets}\label{example-data}}
 
@@
 
-## Dataset 1: Forest Census Plot on Barro Colorado Island {#ex-data-1}
+\hypertarget{ex-data-1}{%
+\subsection{Dataset 1: Forest Census Plot on Barro Colorado Island}\label{ex-data-1}}
 
@@
 The biggest challenge with this study is that the data are stored on a personal FTP server of one of the authors.
-While the data are available in CSV and binary `Rdata` formats, the data storage servers do not have any metadata, README files or a data dictionary.
-A separate, public archive hosts the metadata (https://repository.si.edu/handle/10088/20925) in a PDF file (https://repository.si.edu/bitstream/handle/10088/20925/RoutputFull.pdf?sequence=1&isAllowed=y) that describe the fields.
+While the data are available in CSV and binary \texttt{Rdata} formats, the data storage servers do not have any metadata, README files or a data dictionary.
+A separate, public archive hosts the metadata (\url{https://repository.si.edu/handle/10088/20925}) in a PDF file (\url{https://repository.si.edu/bitstream/handle/10088/20925/RoutputFull.pdf?sequence=1\&isAllowed=y}) that describe the fields.
 
-## Dataset 2: Sensor data {#sdss-data}
+\hypertarget{sdss-data}{%
+\subsection{Dataset 2: Sensor data}\label{sdss-data}}
 
@@
 
-The SDSS data includes photometric and spectroscopic information obtained from a 2.5m telescope at Apache Point Observatory in New Mexico, operating since 2000, producing over 200Gb of data every day [@sdss-one-paper; @sdss-four-paper; @sdss-website].
+The SDSS data includes photometric and spectroscopic information obtained from a 2.5m telescope at Apache Point Observatory in New Mexico, operating since 2000, producing over 200Gb of data every day (York 2000; Blanton et al.
+2017; ``Sdss-Website,'' n.d.).
 This data has been used to produce the most detailed, three dimensional maps of the universe ever made.
@@
 The data are free to use and publicly accessible, with the interface to the data being designed to cover a wide range of skills.
-For example, the marvin service to streamline access and visualisation of MaNGA data [@marvin-service], through to raw data access (see Figure \@ref(fig:sdss-raw-data-screenshot)).
+For example, the marvin service to streamline access and visualisation of MaNGA data (Cherinka et al.
+2018), through to raw data access (see Figure \ref{fig:sdss-raw-data-screenshot}).
 
-```{r sdss-raw-data-screenshot, fig.cap = "Screenshot of Raw data avaialbe through DR15 FITS.", out.width = "60%", fig.align = "center"}
-knitr::include_graphics(here::here("figures",
-                                   "sdss-raw-data-screenshot.png"))
-```
+\begin{figure}
+
+{\centering \includegraphics[width=0.6\linewidth]{/Users/ntie0001/github/njtierney/ddd/figures/sdss-raw-data-screenshot}
+
+}
+
+\caption{Screenshot of Raw data avaialbe through DR15 FITS.}\label{fig:sdss-raw-data-screenshot}
+\end{figure}
 
@@
 
-- Metadata in the machine readable standard, SCHEMA
-- Data in multiple formats (e.g., .csv, .fits, and database)
-- Previously released data available in archives
-- Entire datasets available for download
-- Raw data before processed into database is also available
+\begin{itemize}
+\tightlist
+\item
+  Metadata in the machine readable standard, SCHEMA
+\item
+  Data in multiple formats (e.g., .csv, .fits, and database)
+\item
+  Previously released data available in archives
+\item
+  Entire datasets available for download
+\item
+  Raw data before processed into database is also available
+\end{itemize}
 
-For example, the optical and spectral data can be accessed in FITS data format, or even as a plain CSV, the first few rows shown in Table \@ref(tab:spectra-tab).
-Wavelength plotted against Flux is shown in Figure \@ref(fig:sdss-spectra), with \@ref(fig:sdss-spectra)A showing the output from the SDSS website, and Figure \@ref(fig:sdss-spectra)B showing the output from the CSV.
-The fact that figure Figure \@ref(fig:sdss-spectra)A is virtually replicated in Figure \@ref(fig:sdss-spectra)B demonstrates great reproducibility.
+For example, the optical and spectral data can be accessed in FITS data format, or even as a plain CSV, the first few rows shown in Table \ref{tab:spectra-tab}.
+Wavelength plotted against Flux is shown in Figure \ref{fig:sdss-spectra}, with \ref{fig:sdss-spectra}A showing the output from the SDSS website, and Figure \ref{fig:sdss-spectra}B showing the output from the CSV.
+The fact that figure Figure \ref{fig:sdss-spectra}A is virtually replicated in Figure \ref{fig:sdss-spectra}B demonstrates great reproducibility.
 
-```{r prepare-spectra-data, include = FALSE}
+\begin{table}[!h]
 
-sdss_image_path <- here::here("figures",
-                              "sdss-spectra-example-plate-1678.png")
+\caption{\label{tab:spectra-tab}Example spectra data from SDSS, showing values for Wavelength, Flux, BestFit, and SkyFlux.}
+\centering
+\begin{tabular}[t]{rrrr}
+\toprule
+Wavelength & Flux & BestFit & SkyFlux\\
+\midrule
+\rowcolor{gray!6}  3801.893 & 48.513 & 44.655 & 14.002\\
+3802.770 & 54.516 & 42.890 & 14.113\\
+\rowcolor{gray!6}  3803.645 & 52.393 & 47.813 & 14.142\\
+3804.522 & 45.273 & 42.612 & 14.186\\
+\rowcolor{gray!6}  3805.397 & 51.529 & 39.659 & 14.221\\
+\addlinespace
+3806.273 & 44.530 & 44.183 & 14.349\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-spectra <- readr::read_csv(
-  here::here("data",
-             "plateid-1678_mjd-53433_fiberid-425_reduction2d-v5_7_0.csv")
-)
+\begin{figure}
 
-gg_spectra <- ggplot(spectra,
-                     aes(x = Wavelength,
-                         y = Flux)) +
-  geom_line() + 
-  theme_minimal() + 
-  theme(aspect.ratio = 1)
+{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/sdss-spectra-1}
 
-plot_spectra <- ggdraw() + 
-  draw_image(sdss_image_path)
+}
 
-```
-
-```{r spectra-tab, echo = FALSE}
-
-head(spectra) %>%
-  knitr::kable(caption = "Example spectra data from SDSS, showing values for Wavelength, Flux, BestFit, and SkyFlux.",
-               booktabs = TRUE) %>%
-  kable_styling(latex_options = c("striped",
-                                  "hold_position"))
-
-```
-
-```{r sdss-spectra, fig.cap = "Visualisations from SDSS can be effecticely replicated.
+\caption{Visualisations from SDSS can be effecticely replicated.
 The spectra image of wavelength against flux from SDSS from two sources.
 Part A shows the example image taken from SDSS website.
-Part B shows the replicated figure using data provided as CSV - although there is some data cleaning or filtering since the numbers are not identical, it still demonstrates the great potential of reproducibility.", out.width = "90%", fig.align = "center"}
-
-plot_grid(plot_spectra, gg_spectra, labels = "AUTO")
-```
+Part B shows the replicated figure using data provided as CSV - although there is some data cleaning or filtering since the numbers are not identical, it still demonstrates the great potential of reproducibility.}\label{fig:sdss-spectra}
+\end{figure}
 
@@
 Despite this, it is still very easy to understand.
-This is a further reflection on the idea that high effort  can create highly understandable data, (see Figure \@ref(fig:effort-understanding)).
+This is a further reflection on the idea that high effort can create highly understandable data, (see Figure \ref{fig:effort-understanding}).
 The impact of this data is high, having changed the way we understand the past, present and future of the universe itself.
@@
 
-## Dataset 3: Most other datasets {#ex-data-3}
+\hypertarget{ex-data-3}{%
+\subsection{Dataset 3: Most other datasets}\label{ex-data-3}}
 
-The vast majority of datasets are handled in an ad-hoc manner [@Michener2001] and dumped in transient data stores with little usable metadata.
+The vast majority of datasets are handled in an ad-hoc manner (Michener et al.
+2001) and dumped in transient data stores with little usable metadata.
 The value of these datasets decline over time until they are useless.
@@
 
-# Ten simple rules for publishing data {#ten-rules}
+\hypertarget{ten-rules}{%
+\section{Ten simple rules for publishing data}\label{ten-rules}}
 
-1. **Decide whether publishing is appropriate for your data** 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Decide whether publishing is appropriate for your data}
   It is critical to publish your data if they are the basis for a peer-reviewed research publication.
@@
 Here making data available in a transient location like GitHub or a website is sufficient.
-
-2.
-**Include a README** file with your data archive.
-README files have a long history in software (https://medium.com/@NSomar/readme-md-history-and-components-a365aff07f10) and are named so that the ASCII systems capital letters filename would show this file first, making it the obvious place to put relevant information.
+\item
+  \textbf{Include a README} file with your data archive.
+README files have a long history in software (\url{https://medium.com/@NSomar/readme-md-history-and-components-a365aff07f10}) and are named so that the ASCII systems capital letters filename would show this file first, making it the obvious place to put relevant information.
 In the context of datasets, READMEs are particularly useful when there are no reliable standards.
@@
 Additionally keywords, author identifiers such as ORCID, and funder information would be useful.
-
-3.
-**Provide a data dictionary** or glossary for each variable.
-Data dictionaries provide a free-form way to document the data, their structure and underlying details of how they were collected (https://www.emgo.nl/kc/codebook/).
+\item
+  \textbf{Provide a data dictionary} or glossary for each variable.
+Data dictionaries provide a free-form way to document the data, their structure and underlying details of how they were collected (\url{https://www.emgo.nl/kc/codebook/}).
 More importantly they provide a way for future consumers of the data to understand how the variables were named, how missing data were documented, along with any additional information that might impact how the data are interpreted.
-
-4.
-**Provide a machine readable format** for your data.
-When possible, provide machine readable metadata that map on to the open standards of [schema.org](https://schema.org/) and [JSON-LD](https://en.wikipedia.org/wiki/JSON-LD).
+\item
+  \textbf{Provide a machine readable format} for your data.
+When possible, provide machine readable metadata that map on to the open standards of \href{https://schema.org/}{schema.org} and \href{https://en.wikipedia.org/wiki/JSON-LD}{JSON-LD}.
 These metadata provide all of the information described in the README best practices (rule 2), but in a machine readable way and include much of the same information such as name, description, spatial/temporal coverage etc.
-One way to test if your metadata are machine readable is to use Google's structured testing data (https://search.google.com/structured-data/testing-tool/u/0/) to verify the integrity of the metadata.
-
-5.
-**Provide the data in its rawest form** in a folder called "data-raw".
+One way to test if your metadata are machine readable is to use Google's structured testing data (\url{https://search.google.com/structured-data/testing-tool/u/0/}) to verify the integrity of the metadata.
+\item
+  \textbf{Provide the data in its rawest form} in a folder called ``data-raw''.
 Keeping your data raw (also sometimes referred to as the sushi principle) is the safest way to ensure that your analysis can be reproduced from scratch.
 This approach not only lets you trace the provenance of any analysis but it also ensures further use of the data that a derived dataset may prevent.
-
-6.
-**Provide [open source?] scripts** used to clean data from rawest form into tidy/analysis ready format.
+\item
+  \textbf{Provide {[}open source?{]} scripts} used to clean data from rawest form into tidy/analysis ready format.
 Raw data is often unusable without further processing (data munging or data cleaning), which are the steps necessary to detect and clean inaccurate, incomplete records, and missing values from a dataset.
@@
 These scripts describe the steps taken to prepare the data, which helps explain and document any decisions taken during the cleaning phase.
-These should ideally operate on some raw data stored in the "data-raw" folder (rule 5).
-
-7.
-**Keep additional copies in more accessible locations**: Even if you archive the data somewhere long-term, keep a second (or additional copies) in locations that are more readily and easily accessible for data analysis workflows.
+These should ideally operate on some raw data stored in the ``data-raw'' folder (rule 5).
+\item
+  \textbf{Keep additional copies in more accessible locations}: Even if you archive the data somewhere long-term, keep a second (or additional copies) in locations that are more readily and easily accessible for data analysis workflows.
 These include services like GitHub, GitHub LFS and others where fast content delivery networks (CDNs) make access and reuse practical.
 In the event that these services shut down or become unavailable, a long-term archival copy can still be accessed from a permanent repository such as Zenodo or Dryad and populated into a new data hosting service or technology.
-
-8.
-**Use a hash function like MD5 checksum** to ensure that the data you deposited are the same that are being made available to others.
+\item
+  \textbf{Use a hash function like MD5 checksum} to ensure that the data you deposited are the same that are being made available to others.
 Hash values such as MD5 are short numeric values that can serve as digital signatures for large amounts of data.
 By sharing the hash in a README, authors can ensure that the data, particularly the version of data being used by the reader is the same.
-
-9.
-**Only add a citation if your data has a DOI**.
+\item
+  \textbf{Only add a citation if your data has a DOI}.
 A citation only makes sense when your data has a DataCite compliant DOI, which is automatically provided when data is published in repositories like Zenodo and Dryad.
 While a citation may not accrue references, without a DOI it is guaranteed not to.
-
-10.
-**Stick with simple data formats** to ensure long-term usefulness of datasets in a way that is not tied to transient technologies (such as ever changing spreadsheet standards), store the data as plain text (e.g., CSV) where possible.
+\item
+  \textbf{Stick with simple data formats} to ensure long-term usefulness of datasets in a way that is not tied to transient technologies (such as ever changing spreadsheet standards), store the data as plain text (e.g., CSV) where possible.
 This can take the form of comma, or tab separated files in most cases.
 This will ensure transparency, and future proof your data.
+\end{enumerate}
 
-# Conclusions {#conclusions}
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
 
@@
 
-# Acknowledgements
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
 
-- Mile McBain
-- Anna Krystalli
-- Daniella Lowenberg
-- ACEMS International Mobility Programme
-- Helmsley Charitable Trust, Gordon and Betty Moore Foundation, Sloan Foundation.
+\begin{itemize}
+\tightlist
+\item
+  Mile McBain
+\item
+  Anna Krystalli
+\item
+  Daniella Lowenberg
+\item
+  ACEMS International Mobility Programme
+\item
+  Helmsley Charitable Trust, Gordon and Betty Moore Foundation, Sloan Foundation.
+\end{itemize}
 
-# Literature Cited {-}
+\hypertarget{literature-cited}{%
+\section*{Literature Cited}\label{literature-cited}}
+\addcontentsline{toc}{section}{Literature Cited}
+
+\hypertarget{refs}{}
+\begin{cslreferences}
+\leavevmode\hypertarget{ref-Arslan2019}{}%
+Arslan, Ruben C. 2019. ``How to Automatically Document Data with the Codebook Package to Facilitate Data Reuse.'' \emph{Advances in Methods and Practices in Psychological Science} 2 (2): 169--87.
+
+\leavevmode\hypertarget{ref-Barnes2010}{}%
+Barnes, Nick. 2010. ``Publish Your Computer Code: It Is Good Enough.'' \emph{Nature} 467 (7317): 753--53. \url{https://doi.org/10.1038/467753a}.
+
+\leavevmode\hypertarget{ref-GenBank}{}%
+Benson, Dennis A, Ilene Karsch-Mizrachi, David J Lipman, James Ostell, and David L Wheeler. 2005. ``GenBank.'' \emph{Nucleic Acids Research} 33 (Database issue): D34--8.
+
+\leavevmode\hypertarget{ref-julia}{}%
+Bezanson, Jeff, Alan Edelman, Stefan Karpinski, and Viral B Shah. 2017. ``Julia: A Fresh Approach to Numerical Computing.'' \emph{SIAM Review} 59 (1): 65--98. \url{https://doi.org/10.1137/141000671}.
+
+\leavevmode\hypertarget{ref-sdss-four-paper}{}%
+Blanton, Michael R, Matthew A Bershady, Bela Abolfathi, Franco D Albareti, Carlos Allende Prieto, Andres Almeida, Javier Alonso-García, et al.
+2017.
+``Sloan Digital Sky Survey IV: Mapping the Milky Way, Nearby Galaxies, and the Distant Universe,'' February.
+\url{http://arxiv.org/abs/1703.00052}.
+
+\leavevmode\hypertarget{ref-arkdb}{}%
+Boettiger, Carl. 2018. \emph{Arkdb: Archive and Unarchive Databases Using Flat Files}. \url{https://CRAN.R-project.org/package=arkdb}.
+
+\leavevmode\hypertarget{ref-Boettiger2015}{}%
+---------. 2015. ``An Introduction to Docker for Reproducible Research.'' \emph{ACM SIGOPS Operating Systems Review} 49 (1): 71--79. \url{https://doi.org/10.1145/2723872.2723882}.
+
+\leavevmode\hypertarget{ref-Broman2017}{}%
+Broman, Karl W., and Kara H. Woo. 2017. ``Data Organization in Spreadsheets.'' \emph{The American Statistician} 72 (1): 2--10. \url{https://doi.org/10.1080/00031305.2017.1375989}.
+
+\leavevmode\hypertarget{ref-Brooke_Anderson2017}{}%
+Brooke Anderson, G, and Dirk Eddelbuettel. 2017. ``Hosting Data Packages via Drat: A Case Study with Hurricane Exposure Data.'' \emph{The R Journal} 9 (1): 486--97.
+
+\leavevmode\hypertarget{ref-google-data-search}{}%
+Castelvecchi, Davide. 2018. ``Google Unveils Search Engine for Open Data.'' \emph{Nature} 561 (7722): 161--62.
+
+\leavevmode\hypertarget{ref-cc0-long}{}%
+``CC0 Full License.'' n.d. https://creativecommons.org/publicdomain/zero/1.0/legalcode.
+
+\leavevmode\hypertarget{ref-cc0-short}{}%
+``CC0 Short Guide.'' n.d. https://creativecommons.org/publicdomain/zero/1.0/.
+
+\leavevmode\hypertarget{ref-ccby-long}{}%
+``CCBY Full License.'' n.d. https://creativecommons.org/licenses/by/4.0/legalcode.
+
+\leavevmode\hypertarget{ref-ccby-short}{}%
+``CCBY Short Guide.'' n.d. https://creativecommons.org/licenses/by/4.0/.
+
+\leavevmode\hypertarget{ref-marvin-service}{}%
+Cherinka, Brian, Brett H Andrews, José Sánchez-Gallego, Joel Brownstein, María Argudo-Fernández, Michael Blanton, Kevin Bundy, et al.
+2018.
+``Marvin: A Toolkit for Streamlined Access and Visualization of the SDSS-IV MaNGA Data Set,'' December.
+\url{http://arxiv.org/abs/1812.03833}.
+
+\leavevmode\hypertarget{ref-choose-cc0}{}%
+``Choose-Cc0.'' n.d. https://creativecommons.org/choose/zero/.
+
+\leavevmode\hypertarget{ref-McGill-codebook}{}%
+``Codebook Cookbook: A Guide to Writing a Good Codebook for Data Analysis Projects in Medicine.'' n.d.
+Accessed December 18, 2018.
+\url{http://www.medicine.mcgill.ca/epidemiology/joseph/pbelisle/CodebookCookbook/CodebookCookbook.pdf}.
+
+\leavevmode\hypertarget{ref-cran}{}%
+``Cran.'' n.d. https://cran.r-project.org/.
+
+\leavevmode\hypertarget{ref-csvy}{}%
+``Csvy.'' n.d. http://csvy.org/.
+
+\leavevmode\hypertarget{ref-data-538}{}%
+``Data-538.'' n.d. https://data.fivethirtyeight.com/.
+
+\leavevmode\hypertarget{ref-dataspice}{}%
+\emph{Dataspice: Create Lightweight Schema.org Descriptions of Dataset}. 2018. \url{https://github.com/ropenscilabs/dataspice}.
+
+\leavevmode\hypertarget{ref-Donoho2017}{}%
+Donoho, David.
+2017.
+``50 Years of Data Science.'' \emph{Journal of Computational and Graphical Statistics: A Joint Publication of American Statistical Association, Institute of Mathematical Statistics, Interface Foundation of North America} 26 (4): 745--66.
+
+\leavevmode\hypertarget{ref-dryad}{}%
+``Dryad.'' n.d. https://datadryad.org/.
+
+\leavevmode\hypertarget{ref-Ellis2017}{}%
+Ellis, Shannon E., and Jeffrey T. Leek. 2017. ``How to Share Data for Collaboration.'' \emph{The American Statistician} 72 (1): 53--57. \url{https://doi.org/10.1080/00031305.2017.1375987}.
+
+\leavevmode\hypertarget{ref-Fowler2017}{}%
+Fowler, Dan, Jo Barratt, and Paul Walsh. 2017. ``Frictionless Data: Making Research Data Quality Visible.'' \emph{IJDC} 12 (2): 274--85.
+
+\leavevmode\hypertarget{ref-Gentleman2004}{}%
+Gentleman, Robert C, Vincent J Carey, Douglas M Bates, Ben Bolstad, Marcel Dettling, Sandrine Dudoit, Byron Ellis, et al.
+2004.
+\emph{Genome Biology} 5 (10): R80.
+\url{https://doi.org/10.1186/gb-2004-5-10-r80}.
+
+\leavevmode\hypertarget{ref-gentleman2007statistical}{}%
+Gentleman, Robert, and Duncan Temple Lang. 2007. ``Statistical Analyses and Reproducible Research.'' \emph{Journal of Computational and Graphical Statistics} 16 (1): 1--23.
+
+\leavevmode\hypertarget{ref-git}{}%
+``Git.'' n.d. https://git-scm.com/about.
+
+\leavevmode\hypertarget{ref-github}{}%
+``Github.'' n.d. https://github.com/.
+
+\leavevmode\hypertarget{ref-DataPackageR}{}%
+Greg Finak. 2019. \emph{DataPackageR: Construct Reproducible Analytic Data Sets as R Packages}. \url{https://doi.org/10.5281/zenodo.2620378}.
+
+\leavevmode\hypertarget{ref-julia-pkgman}{}%
+``Julia-Pkgman.'' n.d. https://pkg.julialang.org/.
+
+\leavevmode\hypertarget{ref-Jupyter2018}{}%
+Jupyter.
+2018.
+``Binder 2.0 - Reproducible, Interactive, Sharable Environments for Science at Scale.'' \emph{Proceedings of the 17th Python in Science Conference}.
+\url{https://doi.org/10.25080/Majora-4af1f417-011}.
+
+\leavevmode\hypertarget{ref-Kirk2019}{}%
+Kirk, Rebecca, and Larry Norton. 2019. ``Supporting Data Sharing.'' \emph{NPJ Breast Cancer} 5 (February): 8.
+
+\leavevmode\hypertarget{ref-McKiernan2016}{}%
+McKiernan, Erin C, Philip E Bourne, C Titus Brown, Stuart Buck, Amye Kenall, Jennifer Lin, Damon McDougall, et al.
+2016.
+``How Open Science Helps Researchers Succeed.'' \emph{eLife} 5 (July).
+\url{https://doi.org/10.7554/elife.16800}.
+
+\leavevmode\hypertarget{ref-Michener2001}{}%
+Michener, William, James Brunt, Kristin Vanderbilt, and Sevilleta Program. 2001. ``Ecological Informatics: A Long-Term Ecological Research Perspective'' 51 (January).
+
+\leavevmode\hypertarget{ref-Peng2011}{}%
+Peng, R. D. 2011. ``Reproducible Research in Computational Science.'' \emph{Science} 334 (6060): 1226--7. \url{https://doi.org/10.1126/science.1213847}.
+
+\leavevmode\hypertarget{ref-piggyback}{}%
+``Piggyback.'' n.d. https://github.com/ropensci/piggyback.
+
+\leavevmode\hypertarget{ref-plos-comp-bio-data}{}%
+``PLOS Computational Biology.'' n.d. \emph{PLOS ONE}. Public Library of Science. \url{https://journals.plos.org/ploscompbiol/s/data-availability}.
+
+\leavevmode\hypertarget{ref-Poisot2019}{}%
+Poisot, Timothée, Anne Bruneau, Andrew Gonzalez, Dominique Gravel, and Pedro Peres-Neto.
+2019.
+``Ecological Data Should Not Be so Hard to Find and Reuse.'' \emph{Trends in Ecology \& Evolution} 34 (6): 494--96.
+\url{https://doi.org/10.1016/j.tree.2019.04.005}.
+
+\leavevmode\hypertarget{ref-Popkin2019}{}%
+Popkin, Gabriel. 2019. ``Data Sharing and How It Can Benefit Your Scientific Career.'' \emph{Nature} 569 (7756): 445--47.
+
+\leavevmode\hypertarget{ref-pudding-data}{}%
+``Pudding-Data.'' n.d. https://github.com/the-pudding/data.
+
+\leavevmode\hypertarget{ref-pypi}{}%
+``Pypi.'' n.d. https://pypi.org/.
+
+\leavevmode\hypertarget{ref-python}{}%
+``Python.'' n.d. https://www.python.org/.
+
+\leavevmode\hypertarget{ref-Ram2013}{}%
+Ram, Karthik.
+2013.
+``Git Can Facilitate Greater Reproducibility and Increased Transparency in Science.'' \emph{Source Code for Biology and Medicine} 8 (1).
+\url{https://doi.org/10.1186/1751-0473-8-7}.
+
+\leavevmode\hypertarget{ref-rcore}{}%
+R Core Team. 2019. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
+
+\leavevmode\hypertarget{ref-Rowhani-Farid2016}{}%
+Rowhani-Farid, Anisa, and Adrian G Barnett. 2016. ``Has Open Data Arrived at the British Medical Journal (BMJ)? An Observational Study.'' \emph{BMJ Open} 6 (10): e011784.
+
+\leavevmode\hypertarget{ref-sdss-website}{}%
+``Sdss-Website.'' n.d. https://www.sdss.org/.
+
+\leavevmode\hypertarget{ref-OSF}{}%
+Spies, Jeffrey R, Brian A Nosek, Sheila Miguez, Ben B Blohowiak, Michael Cohn, Elizabeth Bartmess, and L Simon. 2012. ``Open Science Framework.'' \emph{Open Science Collaboration}.
+
+\leavevmode\hypertarget{ref-Stodden2018}{}%
+Stodden, Victoria, Jennifer Seiler, and Zhaokun Ma.
+2018.
+``An Empirical Analysis of Journal Policy Effectiveness for Computational Reproducibility.'' \emph{Proceedings of the National Academy of Sciences of the United States of America} 115 (11): 2584--9.
+
+\leavevmode\hypertarget{ref-tidy-tuesday-incarcerate}{}%
+``Tidy-Tuesday-Incarcerate.'' n.d. https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-01-22.
+
+\leavevmode\hypertarget{ref-White2013}{}%
+White, Ethan P, Elita Baldridge, Zachary T Brym, Kenneth J Locey, Daniel J McGlinn, and Sarah R Supp.
+2013.
+``Nine Simple Ways to Make It Easier to (Re)use Your Data.'' \emph{Ideas in Ecology and Evolution} 6 (2).
+
+\leavevmode\hypertarget{ref-Wickham2014}{}%
+Wickham, Hadley. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10). \url{https://doi.org/10.18637/jss.v059.i10}.
+
+\leavevmode\hypertarget{ref-nycflights}{}%
+---------. 2018. \emph{Nycflights13: Flights That Departed Nyc in 2013}. \url{https://CRAN.R-project.org/package=nycflights13}.
+
+\leavevmode\hypertarget{ref-Wilkinson2016}{}%
+Wilkinson, Mark D., Michel Dumontier, IJsbrand Jan Aalbersberg, Gabrielle Appleton, Myles Axton, Arie Baak, Niklas Blomberg, et al.
+2016.
+``The FAIR Guiding Principles for Scientific Data Management and Stewardship.'' \emph{Scientific Data} 3 (March): 160018.
+\url{https://doi.org/10.1038/sdata.2016.18}.
+
+\leavevmode\hypertarget{ref-sdss-one-paper}{}%
+York, D G. 2000. ``The Sloan Digital Sky Survey: Technical Summary,'' June. \url{http://arxiv.org/abs/astro-ph/0006396}.
+
+\leavevmode\hypertarget{ref-zenodo}{}%
+Zenodo. 2016. ``Zenodo.''
+
+\leavevmode\hypertarget{ref-EML-about}{}%
+n.d. Accessed December 18, 2018. \url{https://knb.ecoinformatics.org/\#external//emlparser/docs/index.html}.
+\end{cslreferences}
+
+\end{document}

### 2020-02-06 paper.tex
Creates "submission.tar.gz", which is the right .tar.gz file for submission to arxiv

@@
 
-{\centering \includegraphics[width=0.75\linewidth]{/Users/ntie0001/github/njtierney/ddd/figures/directory-structure}
+{\centering \includegraphics[width=0.75\linewidth]{figures/directory-structure}
 
@@
 
-{\centering \includegraphics[width=0.75\linewidth]{/Users/ntie0001/github/njtierney/ddd/figures/json-code}
+{\centering \includegraphics[width=0.75\linewidth]{figures/json-code}
 
@@
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{/Users/ntie0001/github/njtierney/ddd/figures/data_sharing_workflow} \caption{The steps required compared to effort required for data sharing for reproducibility.}\label{fig:data-sharing-workflow}
+\includegraphics[width=0.9\linewidth]{figures/data_sharing_workflow} \caption{The steps required compared to effort required for data sharing for reproducibility.}\label{fig:data-sharing-workflow}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{/Users/ntie0001/github/njtierney/ddd/figures/sdss-raw-data-screenshot}
+{\centering \includegraphics[width=0.6\linewidth]{figures/sdss-raw-data-screenshot}
 

### 2020-05-20 paper.Rmd
push some minor changes

@@
 This primarily comes in the form of potential for reuse in other contexts along use for training and teaching (@McKiernan2016).
-Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing does not put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
+Assuming that the data have no privacy concerns (e.g., human participants, locations of critically endangered species), or that the act of sharing does not put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
 First and foremost, sharing data along with other artifacts can make it easier for a reader to independently reproduce the results, thereby increasing transparency and trust in the work.
-The act of easy data sharing can also improves model training, as many different models can be tried and tested on latest data sources, closing the loop on research and application of statistical techniques.
+The act of easy data sharing can also improve model training, as many different models can be tried and tested on latest data sources, closing the loop on research and application of statistical techniques.
 Existing data sets can be combined or linked with new or existing data, fostering the development and synthesis of new ideas and research areas.
@@
 
-For nearly two decades, researchers who work in areas related to computational science have pushed for better standards to verify scientific claims, especially in areas where a full replication of the study would be prohibitively expensive.
+For nearly two decades, researchers who work in areas related to computational science have pushed for better standards to verify scientific claims, especially in areas where a full replication of the study would be very expensive.
 To meet these minimal standards, there must be easy access to the data, models, and code.
@@
 **In this paper we focus on the practical side of sharing data for the purpose of reproducibility.** Our goal is to describe various methods in which an analyst can share their data with minimal friction.
-We steer clear of idealistic ideas such as the FAIR data principles [@Wilkinson2016] since they still do not help a researcher share their data.
+We steer clear of high-level ideas such as the FAIR data principles [@Wilkinson2016] since they still do not help a researcher share their data.
 We also skip the discussion around citation and credit because data citations are still poorly tracked and there is no standardized metric or a h-index equivalent for data as of this writing.
@@
 As a result, data used in an analysis is often read from various locations (local, network, or cloud), various formats, varying levels of tidiness [@Wickham2014].
-There is also the overhead associated with data publishing, the act of archiving data in a repository that also mints permanent identifiers, that are not required of all projects due to the effort and time involved.
+There is also the overhead associated with data publishing and the act of archiving data in a repository that also mints permanent identifiers.
+These are not required of most projects for a combination of reasons: an absence of a culture of sharing data, a lack of academic reward, and the effort and time involved.
 It is worth drawing the distinction between data sharing (making the data available with little effort) and data publishing (archiving the data in a long-term repository, with our without curated metadata).
@@
 While there are several papers that serve as best-practice guides for formatting data [@Broman2017] and getting them ready for sharing [@Ellis2017], the aims of this paper are a bit different.
-Our aim is to address the issue of data in the context of reproducibility in data science workflows.
+Our aim is to address the issue of data in the context of reproducibility in data science workflows, which might include a data analysis from data reading, cleaning, analysis, and communication.
 In particular we discuss the various tradeoffs one has to consider when documenting and sharing data, when it is worth publishing, and how this would change depending on the use case, perceived impact, and potential audience.
@@
 
-One basic suggested directory layout is given below in Figure \@ref(fig:code-directory).
+One tried and tested directory layout is given below in Figure \@ref(fig:code-directory).
 
@@
 **Missing data** are values that should have been observed, but were not.
+Although for some cases this might include values are are not-applicable, for example variables on "transplant_received" and "transplant_date", then "transplant_date" would be missing if "transplant_received" was "no".
 The code for missingness should be documented in the data dictionary, and should nominally be `NA`.
@@
 (766). Binary files comprise a bulk of the data shipped in the `data`
-folder (68.06%) with other plain text formats such as txt, CSV, dat,
+folder (68%) with other plain text formats such as txt, CSV, dat,
 `json` comprising less than one percent of data shipped in packages.
@@
 The more impactful the data, the more documentation features to include.
+It should also be noted that good documentation can help the author if they return to do more analysis or check a result.
 Figure \@ref(fig:data-sharing-workflow) shows the practical types of steps that can be taken against the effort required.
@@
 
-Other journals operating in this space include journals like "data in brief", "Data", and "Nature Scientific Data".
+Other journals operating in this space include journals like "data in brief", "Data", and "Scientific Data".
 Guidelines for what is required in the content of the paper, and the required information along with the data (meta data, data format, etc.) vary by journal.
@@
 Long-term studies with complex protocols require structured metadata, while more standardized datasets such as those coming from sensors require far less effort.
+It is also worth pointing out that making work reproducible is a skill, that can be improved with practice.
 
@@
 
-- Mile McBain
-- Anna Krystalli
-- Daniella Lowenberg
-- ACEMS International Mobility Programme
-- Helmsley Charitable Trust, Gordon and Betty Moore Foundation, Sloan Foundation.
+We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Miles McBain, Anna Krystalli, Daniella Lowenberg.
+We'd also like to thanks the ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
 

### 2020-05-20 paper.tex
push some minor changes

--- 2020-02-06 paper.tex
@@
 (2016)).
-Assuming that the data have no privacy concerns (e.g., human subjects, locations of critically endangered species), or that the act of sharing does not put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
+Assuming that the data have no privacy concerns (e.g., human participants, locations of critically endangered species), or that the act of sharing does not put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
 First and foremost, sharing data along with other artifacts can make it easier for a reader to independently reproduce the results, thereby increasing transparency and trust in the work.
-The act of easy data sharing can also improves model training, as many different models can be tried and tested on latest data sources, closing the loop on research and application of statistical techniques.
+The act of easy data sharing can also improve model training, as many different models can be tried and tested on latest data sources, closing the loop on research and application of statistical techniques.
 Existing data sets can be combined or linked with new or existing data, fostering the development and synthesis of new ideas and research areas.
@@
 \textbf{In this paper we focus on the practical side of sharing data for the purpose of reproducibility.} Our goal is to describe various methods in which an analyst can share their data with minimal friction.
-We steer clear of idealistic ideas such as the FAIR data principles (Wilkinson et al.
+We steer clear of high-level ideas such as the FAIR data principles (Wilkinson et al.
 2016) since they still do not help a researcher share their data.
@@
 As a result, data used in an analysis is often read from various locations (local, network, or cloud), various formats, varying levels of tidiness (Wickham 2014).
-There is also the overhead associated with data publishing, the act of archiving data in a repository that also mints permanent identifiers, that are not required of all projects due to the effort and time involved.
+There is also the overhead associated with data publishing and the act of archiving data in a repository that also mints permanent identifiers.
+These are not required of most projects for a combination of reasons: an absence of a culture of sharing data, a lack of academic reward, and the effort and time involved.
 It is worth drawing the distinction between data sharing (making the data available with little effort) and data publishing (archiving the data in a long-term repository, with our without curated metadata).
@@
 While there are several papers that serve as best-practice guides for formatting data (Broman and Woo 2017) and getting them ready for sharing (Ellis and Leek 2017), the aims of this paper are a bit different.
-Our aim is to address the issue of data in the context of reproducibility in data science workflows.
+Our aim is to address the issue of data in the context of reproducibility in data science workflows, which might include a data analysis from data reading, cleaning, analysis, and communication.
 In particular we discuss the various tradeoffs one has to consider when documenting and sharing data, when it is worth publishing, and how this would change depending on the use case, perceived impact, and potential audience.
@@
 
-One basic suggested directory layout is given below in Figure \ref{fig:code-directory}.
+One tried and tested directory layout is given below in Figure \ref{fig:code-directory}.
 
@@
 \textbf{Missing data} are values that should have been observed, but were not.
+Although for some cases this might include values are are not-applicable, for example variables on ``transplant\_received'' and ``transplant\_date'', then ``transplant\_date'' would be missing if ``transplant\_received'' was ``no''.
 The code for missingness should be documented in the data dictionary, and should nominally be \texttt{NA}.
@@
 (766). Binary files comprise a bulk of the data shipped in the \texttt{data}
-folder (68.06\%) with other plain text formats such as txt, CSV, dat,
+folder (68\%) with other plain text formats such as txt, CSV, dat,
 \texttt{json} comprising less than one percent of data shipped in packages.
@@
 The more impactful the data, the more documentation features to include.
+It should also be noted that good documentation can help the author if they return to do more analysis or check a result.
 Figure \ref{fig:data-sharing-workflow} shows the practical types of steps that can be taken against the effort required.
@@
 
-Other journals operating in this space include journals like ``data in brief'', ``Data'', and ``Nature Scientific Data''.
+Other journals operating in this space include journals like ``data in brief'', ``Data'', and ``Scientific Data''.
 Guidelines for what is required in the content of the paper, and the required information along with the data (meta data, data format, etc.) vary by journal.
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Mile McBain
-\item
-  Anna Krystalli
-\item
-  Daniella Lowenberg
-\item
-  ACEMS International Mobility Programme
-\item
-  Helmsley Charitable Trust, Gordon and Betty Moore Foundation, Sloan Foundation.
-\end{itemize}
+We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Miles McBain, Anna Krystalli, Daniella Lowenberg.
+We'd also like to thanks the ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
 

### 2020-05-29 paper.Rmd
update paper to submit to patterns

@@
 
-We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Miles McBain, Anna Krystalli, Daniella Lowenberg.
-We'd also like to thanks the ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
+We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Anna Krystalli, Daniella Lowenberg, and Miles McBain.
+We'd also like to thank those who provided feedback on the paper when it went up for presubmission: Adrian Barnett, Raymond B Huey Christian Thurn, , Christina Maimone, and github user, 'dschief001'.
+Finally, we thank the following organisations for their support ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
+
+# Bigger Picture Statement {-}
+
+Without data, there is no science.
+Science needs to be reproducible so we can trust the results, and progress as a field.
+But without data, we do not have true reproducibility.
+Currently, the role of data in reproducibility is not adequately discussed.
+We believe this is in part because discussions around data are centered around both data quality and data sharing.
+However, these are in fact two separate (but related) topics.
+We provide immediately usable guidelines to sharing data alongside your research, with minimal barriers, that a researcher could pick up and use today.
+
+In the future, we hope that the culture around data sharing will change, and sharing data will be rewarded in science.
+We want the new normal to be an expectation of data to be submitted with every research publication, and for these datasets to be easily discovered, shared, and extended in other analysis.
+
+# Bio {-}
+
+**Nicholas Tierney**
+
+Nicholas Tierney has an honours degree in Psychological Science, and a PhD in Statistics.
+He is a Lecturer at Monash University, teaching introduction to data analysis and researching how to improve data analysis workflow.
+This ranges from exploratory data analysis, to statistical modelling, calculating diagnostics, and how we draw inferences, and make decisions from data.
+Crucial to this work is producing high quality software to accompany each research idea.
+Nick is a keen outdoorsman, and loves hiking, trail running, rock climbing, photography, and making coffee.
+
+**Karthik Ram**
+
+Karthik is a senior data scientist at the Berkeley Institute of Data Science and a co-founder of the rOpenSci project.
+Karthik is also a senior PI at the Berkeley Initiative for Global Change Biology.
+Prior to joining Berkeley, he earned his PhD in Ecology &amp; Evolution from the University of California, Davis.
+Karthik's interests are focused on reproducible research, especially as it applies to global change.
+Much of his recent work focuses on building tools and services around open data and growing diverse data science communities.
 

### 2020-05-29 cover-letter/cover-letter.Rmd
submitted to patterns

---
from: Nicholas Tierney
fromqualifications: BPsySc (Hons), PhD (Statistics)
fromtitle: Lecturer
fromWWW: njtierney.com
fromemail: nicholas.tierney@gmail.com
fromphone: +61 3 9905 9352
fromsignature: sigfile.png
toname: Dr. Sarah Callaghan
toaddress: Patterns
opening: Dear Dr. Sarah Callaghan,
closing: Regards, \newline Dr. Nicholas Tierney \newline Lecturer
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider our article "A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility" for publication in the Patterns. This article describes an accessible approach to sharing data alongside research. The key message of the paper is that accessing data for reproducibility should come with minimal barriers. 
Along with the PDF is an accompanying .zip file, which contains all of the code, data, and rmarkdown document to reproduce the paper. This article is not under consideration elsewhere, although we have submitted a pre-print of this article to arxiv.

We hope that you find this paper of interest to the readers of the Patterns,

### 2020-08-12 arxiv/paper.tex
Reorged paper for new submission

--- 2020-05-20 paper.tex
@@
 
-We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Miles McBain, Anna Krystalli, Daniella Lowenberg.
-We'd also like to thanks the ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
+We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Anna Krystalli, Daniella Lowenberg, and Miles McBain.
+We'd also like to thank those who provided feedback on the paper when it went up for presubmission: Adrian Barnett, Raymond B Huey Christian Thurn, , Christina Maimone, and github user, `dschief001'.
+Finally, we thank the following organisations for their support ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
+
+\hypertarget{bigger-picture-statement}{%
+\section*{Bigger Picture Statement}\label{bigger-picture-statement}}
+\addcontentsline{toc}{section}{Bigger Picture Statement}
+
+Without data, there is no science.
+Science needs to be reproducible so we can trust the results, and progress as a field.
+But without data, we do not have true reproducibility.
+Currently, the role of data in reproducibility is not adequately discussed.
+We believe this is in part because discussions around data are centered around both data quality and data sharing.
+However, these are in fact two separate (but related) topics.
+We provide immediately usable guidelines to sharing data alongside your research, with minimal barriers, that a researcher could pick up and use today.
+
+In the future, we hope that the culture around data sharing will change, and sharing data will be rewarded in science.
+We want the new normal to be an expectation of data to be submitted with every research publication, and for these datasets to be easily discovered, shared, and extended in other analysis.
+
+\hypertarget{bio}{%
+\section*{Bio}\label{bio}}
+\addcontentsline{toc}{section}{Bio}
+
+\textbf{Nicholas Tierney}
+
+Nicholas Tierney has an honours degree in Psychological Science, and a PhD in Statistics.
+He is a Lecturer at Monash University, teaching introduction to data analysis and researching how to improve data analysis workflow.
+This ranges from exploratory data analysis, to statistical modelling, calculating diagnostics, and how we draw inferences, and make decisions from data.
+Crucial to this work is producing high quality software to accompany each research idea.
+Nick is a keen outdoorsman, and loves hiking, trail running, rock climbing, photography, and making coffee.
+
+\textbf{Karthik Ram}
+
+Karthik is a senior data scientist at the Berkeley Institute of Data Science and a co-founder of the rOpenSci project.
+Karthik is also a senior PI at the Berkeley Initiative for Global Change Biology.
+Prior to joining Berkeley, he earned his PhD in Ecology \& Evolution from the University of California, Davis.
+Karthik's interests are focused on reproducible research, especially as it applies to global change.
+Much of his recent work focuses on building tools and services around open data and growing diverse data science communities.
 

### 2020-08-19 arxiv/paper.Rmd
update new-paper and fix most of the citations

--- 2020-05-29 paper.Rmd
@@
 Nicholas Tierney has an honours degree in Psychological Science, and a PhD in Statistics.
-He is a Lecturer at Monash University, teaching introduction to data analysis and researching how to improve data analysis workflow.
-This ranges from exploratory data analysis, to statistical modelling, calculating diagnostics, and how we draw inferences, and make decisions from data.
+He is a Lecturer at Monash University, teaching introduction to data analysis.
+His ranges from exploratory data analysis, to statistical modelling, calculating diagnostics, and how we draw inferences, and make decisions from data.
 Crucial to this work is producing high quality software to accompany each research idea.

### 2020-08-19 new-paper.Rmd
update new-paper and fix most of the citations

---
title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.
author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
header-includes: 
  - \setstretch{1.45}
  - \usepackage[utf8]{inputenc}
---

```{=tex}
\begin{singlespace}

1 = Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia

2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)

3 = Berkeley Institute for Data Science, University of California, Berkeley, USA

* = Both authors contributed equally to the work

Correspondence to:

Nicholas Tierney (nicholas.tierney@gmail.com) and Karthik Ram (k@karthik.io)

\end{singlespace}
```
# Abstract {.unnumbered}

```{=tex}
\begin{singlespace}

Data makes science possible. Sharing data improves visibility, and makes the research process transparent. This increases trust in the work, and allows for independent reproduction of results. However, a large proportion of data from published research is often only available to the original authors. Despite the obvious benefits of sharing data, and scientists' advocating for the importance of sharing data, most advice on sharing data discusses its broader benefits, rather than the practical considerations of sharing. This paper provides practical, actionable advice on how to actually share data alongside research with minimal barriers.

\end{singlespace}
```
```{r setup, echo = FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      out.width = "60%",
                      fig.align = "center")

suppressMessages(library(ggrepel))
suppressMessages(library(tibble))
suppressMessages(library(kableExtra))
suppressMessages(library(magick))
suppressMessages(library(ggplot2))
suppressMessages(library(cowplot))
suppressMessages(library(gridExtra))
```

# Introduction {#intro}

> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (The Adventure of the Copper Beeches by Sir Arthur Conan Doyle)

The idea of open and transparent science dates back to the founding principles of The Royal Society of London in 1660. The movement, which has seen shifting support over the years, has enjoyed growing popularity over the past decade [@Nielsen2020; @Peng2011] mainly among early career researchers [@McKiernan2016]. A key motivation for practicing openness is the notion that sharing all research artifacts will allow others to reproduce and verify results, thereby improving trust and verification of those findings. Jon Clarebout famously wrote in 1992 that reproducibility in computationally oriented fields had become so trivial that non experts could do the task. Clarebout's optimistic vision was a little ahead of its time as the tools and technologies available in the early nineties for literate programming presented too high of a barrier for most researchers to adopt without incurring a significant loss in productivity [@Barnes2010].

Meaningful progress on computational reproducibility, the ability of an independent researcher to recreate figures, tables and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence. Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by organizations like The Carpentries, combined with the growing adoption of Github by scientists [@Ram2013; @Zagalsky2015]. The bigger driver for this is more likely driven by the growing popularity of data science as a discipline distinct from statistics [@Donoho2017]. This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools. Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data [@RCore; @python; @julia]. Each of these languages enjoys support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular. Modern scientists now have access to many interactive computational frameworks (e.g. Jupyter, RMarkdown, Stencila, Observable) that can interact with kernels from dozens of programming languages. Beyond notebooks to run the code, a reader will also require access to the same software dependencies. In a study of applied computational research, [@Collberg2014] and colleagues were only able to install and run less than 20% of the software described in publications. In addition to locating software, one still has to install the right versions. Although this problem is far from solved, technologies like Docker have alleviated some of these concerns. Researchers can share Dockerfiles to recreate computational environments [@Boettiger2015]. Services like Binder can then leverage Docker and make possible to share the computational environments in which these notebooks were run. The barriers to verification are now low enough that trivial reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].

Notebooks however, are an incomplete solution for computational reproducibility. For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) Code; 2) Computing environment, and 3) Data. There has been tremendous progress on the first two, despite some limitations of existing tools and practices [@Poisot2019]. While it took nearly two decades for Clarebout's vision to become reality, improvements in technology and motivations played a strong role in the culture change. Sharing code wasn't just easy at first [@Barnes2010], but became a social and normative part of academic culture. The next stage of this evolution, turning code into software, also became popular with increasing training opportunities [@Wilson2014; @Teal2015], tooling [@devtools], and recognition as a scholarly contribution [@Smith2018]. Getting researchers to share usable data has been more of a struggle.

```{r pyramid, fig.cap = "Strategy for culture and behavior change, taken from the Center for Open Science: https://www.cos.io/blog/strategy-for-culture-change . To reach the top, all the lower levels need to be adequately achieved."}
# pyramid image
knitr::include_graphics("figures/pyramid.png")
```

```{r incentable-table, fig.cap = "Table of the mechanism for behavior change, the incentive, and our assessment of where the elements of Data, Code, and Computational Environment rank in terms of completing these aspects. We note that Data is often required, but the preceding steps are not required. This is in contrast to Code having no policy"}
library(gt)
incentive_table <- tibble(
  Mechanism = c("Policy", "Incentives", "Communities", "UI/UX", "Infrastructure"),
  Incentive = c("Required", "Rewarding", "Normative", "Easy", "Possible"),
  Data = c(100, 0, 0, 0, 100),
  Code = c(0, rep(100, 4)),
  `Computational Environment` = c(0, 0, 0, 50, 100)
  )

incentive_table %>%
  gt() %>%
  data_color(
    columns = vars(Data, Code, `Computational Environment`),
    colors = scales::col_numeric(
      # Using a function from paletteer to generate a vector of colors
      # Note that you need to wrap paletteer_d outputs in as.character()
      palette = colorspace::sequential_hcl(n = 3, palette = "Greens2", rev = TRUE),
      # Domain is from scales::col_numeric
      # Domain = The possible values that can be mapped
      # We don't HAVE to set a range for this since
      # we're using it inside data_color()
      domain = NULL
    )
  ) %>%
  gtsave("figures/gt-incentive-table.png")

# knitr::include_graphics("figures/gt-incentive-table.png")
```

Nosek [-@Nosek2019] describes the strategies for culture change, especially when the status quo is too entrenched. Beyond motivation, skills, tools, and incentives are critical to enable such change. In a five step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, then improve the infrastructure to make it easy, then increase awareness to make it normative. Once there is a critical mass, find ways to reward those who put it into practice. After these steps, it would then be possible to require this of everyone Figure 1, table 1). This framework provides a useful lens through which to examine adoption of code, computational environments and data for reproducibility. With code, we have slowly progressed through these stages. While no journals require that authors share code, the venues to share, along with the necessary training have paved the way for incentives and credit. All of the necessary foundation to make code sharing required are in place. For computational environments, Docker makes it possible to share them in most common use cases [@Boettiger2015]. But running them still requires expertise and time on the part of reviewers. More recent open services like Binder [@Jupyter2018] make it possible to run notebooks atop Docker images, but the technology is still new and slowly gaining adoption. There is still a long way to go before sharing computational environments in academia becomes mainstream. Finally, for data, we have implemented these steps out of order. The FAIR data principles were introduced in 2016 with good intentions [@Wilkinson2016] and describe the desired attributes for research data. That they be findable (good metadata), accessible (deposited somewhere persistent), interoperable (easily combined with other datasets), and be reusable. We as a community started out by making data required before demonstrating how this would directly benefit those who share data. Researchers are asked to deposit their data in one of many institutional or commercial repositories. Many journals even require it as a condition for publication via policy. However, the cost of preparing data for reuse remains high, and the career incentives to do so are almost non-existent. Until we make data reuse easier and more rewarding, no level of policy is going to improve the state of data sharing.

Despite the increase in data being shared alongside publications [@Culina2020], the status quo of data sharing in science is not good. [@Rowhani-Farid2016] showed that 7 of 157 (4.5%) sampled papers in the British Medical Journal shared data, and only 14% of articles claiming open sharing of data actually shared their data. Similarly, [@Stodden2018] showed only 36% of papers in the journal *Science* shared data, despite the Journal's data sharing policy. [@Colavizza2020] reflects similar findings for PLOS and BMC journals. There are numerous examples of similar problems in other fields, such as psychology, biology, and bioinformatics. The reality is that while data sharing is increasing in many fields such as ecology, the amount of data that can be reused is likely lower [@Culina2020] . To make shared data useful, we need to realign incentives for researchers.

Tools to prepare datasets and metadata (*the infrastructure*) and best practices (guidelines) for reuse are sparse [@Arslan2019; @Roche2015]. Unlike code, datasets are far more diverse in terms of heterogeneity, size, and formats. This makes them particularly challenging to standardize or easily "install" where the code is running. There are some efforts working towards improving this process, such as gogetdata , the 'dat' project, and a private company, Quilt [@goget; @dat; @Quilt].

Depositing data with rich metadata is laborious making it a pro forma activity outside of speciality repositories like Dryad that employ curators. Lastly, the infrastructure to track and get credit for data is non-existent. Datasets currently do not accrue citation credit since they are not tracked.

## Improving the state of reusable data sharing

As academic data scientists, we spend considerable time cleaning messy data and believe that preparing data for easy reuse can be a catalyst to improve the quality of data and pave the way for other incentives such as community adoptions, which will then create the culture change needed to credit data similar to publications. In this note, we describe a set of practices that researchers can adopt to make data more useful in the context of reproducibility. It should not be a huge burden to share data in a format that is readily usable by others. And similarly, this burden should be reduced by those who want to use another's data.

To make data more reusable, we need to make it easier to generate metadata, standardise good practices within and across communities, and make reusable data sharing a recognized, normalised, part of research. Here we present a set of best practices to enable this change. These include documentation (\@ref(readme)), licensing (\@ref(licenses)), citation (\@ref(cite)), code (\@ref(pkgs)), raw and analysis ready data (\@ref(rawdata)). We also discuss considerations for sharing of data of various sizes \@ref(size).

# Documentation (READMEs) {.readme}

A README is often the first place users will look to learn more about a directory [@Prana2019]. The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/> and [@Prana2019]. For example, a README in a data repository might explain which directories contain data, tidying scripts, and also why data was collected. The README should be placed in the top level of the project, and there can be multiple README files in a project. In the context of datasets, READMEs are particularly useful when there are no reliable standards.

# Licences {.licenses}

A license for data clearly establishes the rules for its use, and whether it can be modified or shared. Without a license, or with a non-standard license, these rules are unclear, and lead to issues with use, attribution and citation. We recommend using a license from Creative Commons [@creativecommons], specifically, we recommend CCBY and CC0 as they are commonly used, and provide simple rules for attribution, warranty, and use. We recommend against non-standard licenses, such as those custom written by the authors, as some terms can render them unusable. Once a license is chosen, a LICENSE file should be provided, containing the entire license in the top level of the directory. The license type should be mentioned in the README, if used.

CC BY requires appropriate credit and attribution, but data can be shared and adapted, even for commercial use, provided changes are described. The CC0 is a "public domain" license, meaning data owners waive all rights, and the work is now "owned" by the public. CC0 data can be freely shared, copied, modified, and distributed, even for commercial purposes **without asking permission**. When using data with CC0, it is good practice to cite the original work, but is not required. Both CCBY and CC0 provide no warranty. Brief and full overviews of CCBY and CC0 can be found at [@ccby-short; @ccby-long; @cc0-short; @cc0-long].

Some data repositories and journals will automatically require a license when submitting data. For example, Dryad requires CC0 (<https://blog.datadryad.org/2011/10/05/why-does-dryad-use-cc0/> and <https://datadryad.org/stash/faq>) and the journal PLOS Comp Bio, requires licences for data that are not more restrictive than CC BY [@plos-comp-bio-data].

```{r dir-layout, fig.cap = "Example directory layout and structure for sharing data. Note ..."}
# insert carbon.sh figure showing file structure layout
knitr::include_graphics("figures/carbon-dir-layout.png")
```

# Metadata {.metadata}

There are three benefits to a dataset having metadata: (1) They provide context by describing variables to humans; (2) They avoid errors by describing and preserving variable types for machine readability; (3) They make data easier to find online by facilitating indexing [@google-data-search]. Three pieces of metadata should be considered: (1) **variable names**, (2) **variable descriptions**, and (3) **unit text**. These should be curated by a person. Metadata created should be clearly labelled in a folder or in files named METADATA.

**Variable names** are short, descriptive names with no spaces or special characters, such as "job\_position" or "date". **Variable description** describes the measured variables, e.g., "University Job Position" or "Date Started", and should also define *variable codes* and *missing data*. **Unit text** defines the type of variable recorded (e.g., string, date, numeric, integer), and avoids serious errors of misrepresented data (e.g., a column of a gene sequence being interpreted as a date). For more details on this we recommend reading [@Broman2017; @Ellis2017; @Arslan2019].

Metadata should provided in plain text files following a standard format. Such formats include the Ecological Metadata Language [@EML-about] (EML), which uses XML (The eXtensible Markup Language), and Table Schema, which uses JSON-LD (Java Script Object Notation - Linked Data) [@ref](fig:json-ld). One implementation of this is Frictionless Data [@Fowler2017], which provides a language agnostic way to document data - to learn more see [@frictionless-data-guide].

EML is widely used in Ecology (for example, <https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-cwt.3091.13>). JSON-LD provides advantages over EML, by having fields to facilitate data discovery with google dataset search [@EML-data]. We recommend JSON-LD over EML, as it is more human readable, shorter, and quicker and easier to read and write. For more information on tooling to assist generating metadata, see <https://knb.ecoinformatics.org/tools/morpho> and [@Arslan2019].

```{r json-ld, fig.cap = "Example snippet of some Table Schema data for a dataset with three variables. This provides a description of each field, and the type of field."}
knitr::include_graphics("figures/carbon-json-ld.png")
```

# Raw data vs analysis read data {.rawdata}

Raw data is the first format of data provided before any cleaning has taken place (e.g., data in some binary or proprietary format). It should be placed in its own folder. Data used in analysis should be provided, especially if computationally intense to derive. It should be in its own folder, and to maximise interoperability, recorded in a plain text (e.g., csv, tsv), not a binary format (e.g., .rds, .sav, .xlsx). Ideally data should be in "Tidy Data" form, where variables are in columns, observations in rows, and each cell has only one value [@Wickham2014].

We don't have particular recommendations for non tabular of data, like binary images, but given the ubiquity of tables in papers, these recommendations will be broadly useful. For other data formats we recommend using a copy of the most robust common format and avoid the latest tools that may be more brittle and likely to change in the future. For example, for an image, a TIFF image in addition to a custom photoshop file.

# Data cleaning scripts {.scripts}

Code used to transform data from its raw format into analysis ready form should be kept in the same place as raw data (Section \@ref(fig:dir-layout)). Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file.

# Citation: How you want your data to be cited {.cite}

Citation is important for provenance, and credit. Citing data clarifies the data version used (provenance), and provides authors a citation (credit). To ensure these are provided, data needs to have an accession number, such as a Digital Object Identifier (DOI). A DOI is a permanent, unique identifier for a digital object such as a paper, poster, or software, and is a prerequisite for citation. If a DOI or accession number is unavailable, a citation will be meaningless, as it cannot be tracked by any means. A file named CITATION should be placed in the directory at the top level. It should contain a DOI, and could be in `.bibtex` format - [\@ref(fig:dir-layout). Avenues that provide DOIs are discussed in \@ref(avenues-to-share-data)].

# Avenues to share data {.avenues}

There are different avenues to share data, and although the terms are sometimes used interchangeably, they are, in fact, different. These include: (1) Data sharing: Making data publicly available; (2) Data archiving: Depositing data somewhere permanent; (3) Data publishing: writing a paper to explain the archived data.

As discussed in \@ref(cite), we recommend data are archived using a platform that generates an accession number such as a DOI. For sharing data collected for research, the most relevant homes are the research data repositories **zenodo**, **dryad**, or **Open Science Framework** (OSF). In addition to minting DOIs, they also provide citation templates that can be used in \@ref(cite).

Zenodo can be contacted to arrange special cases for storage of large data. Special hosting can be arranged via specially contacting zenodo or your research institute. If using a research institute's library service to share data, ensure a DOI is provided so it can be accurately identified and cited.

There are many alternatives to these three, for example, an astronomy project may host its data at the Sloan Digital Sky Survey (SDSS), and genetic data can be hosted at GenBank [@GenBank]. Importantly, these all provide some form of accession number. We recommend adhering to community standards when deciding where to publish data, as this will likely improve data reuse. Large data may also be shared via domain specific venues (e.g., in genomics or astronomy), where appropriate.

We recommend against depositing data in local file storage that does not provide DOI or accession numbers. For example, [@50Hectares] is hosted on a university server and accessed via FTP. It is not possible to effectively provide effective provenance and credit.

Although it is fine to share data among collaborators using temporary cloud storage (like Dropbox, Google Drive), once the paper is public, the data needs to be hosted somewhere persistent. While this approach is useful while iterating on the paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc). Just archive the data.

Journals provide data papers, where they provide a formal publication explaining data that is archived. This is a familiar avenue for researchers to receive credit and citation. Writing a data paper can be very useful for researchers, however, data papers lack guidelines on how to structure data for effective reuse and sharing. We recommend partnering with [@dryad] to archive data and link it with research papers, as they provide a person to help curate your data, improving usability.

# Data as part of open source packages {.pkgs}

Data can be distributed alongside code using a format such as a package in R, Python, and Julia [@nycflights; @penguins; @python-data; @r-data-julia]. A package or module structure provides an advantage as it can be "installed" and loaded just like code. Despite the advantages of sharing data as a software package, there are a few major disadvantages. Firstly, packaging data inside a software package makes the data availability language centric. Users who work in other languages are unlikely to download and export data out of a package. Secondly, data size can be a limitation depending on the repository (CRAN, the R repository, for example, limits package sizes to 5mb although there are workarounds [@Brooke_Anderson2017]). Other fields, such as bioinformatics have a process of sharing data through repositories like Bioconductor, with larger data size. Thirdly, there is the cost of authors and data curators/custodians understanding intricacies of package development in order to share data.

Packaged data and archival on venues like Zenodo are not mutually exclusive. We recommend researchers first archive data in a long-term data repository (Section \@ref(avenues-to-share-data)), before including it in a package. This makes the data easily citable wherever data is referenced.

# Considerations for data size {.size}

How data is shared can change depending on its size. Defining data size will always be somewhat ill-defined, since storage capacity and norms constantly change. We consider data sizes in reference to a moderately priced laptop, where small data fits on computer memory (RAM), medium data does not fit on RAM, and big data does not fit on the hard disk. No matter the size, the data should be archived with an accession number or DOI.

Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(raw-data-vs-analysis-read-data)), or in software \@ref(data-as-part-of-open-source-packages). For medium data, consider a venue like zenodo for archiving. Sharing medium data might require it to be broken into separate chunks for reading (e.g., `teaching-1.csv`, `teaching-2.csv`, etc.), and scripts provided to read it. Large data will require liaising with services from \@ref(avenues-to-share-data), or local research institute library to make special arrangements, to ensure there is some way to gain access, and also maintain a DOI.

# Conclusion

Although data sharing is widely mandated by funders, journals, and academic societies, the amount of usable data remains appallingly low. The mere act of depositing data does not make it reusable or useful to a potential reader. Although data citations do not currently impact promotion and tenure packets, there is anecdotal evidence that papers sharing usable data accrue more citations [@Piwowar2007]. Over time, the more useful we make the data, the more likely incentive mechanisms will take hold.

The steps leading up to requiring data sharing are still vague and disincentivized. In comparison, although few journals require that code be submitted with publications, it is more likely to be adopted given the foundation we have put in place. We have taught researchers how to write modular code via popular training initiatives such as The Carpentries. Commercial tools like GitHub have made code sharing fun and productive. For more advanced researchers, data science languages have made it easier to package and disseminate code as software. If we now require mandatory code sharing, it is quite likely that more of it will be useful than the data currently being shared.

To make data sharing truly useful, we need to make the process easier, improve tools and training, then reward and incentivize researchers who engage in such practices. Only then can data truly be useful.

# Acknowledgements {.unnumbered}

We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Anna Krystalli, Daniella Lowenberg, and Miles McBain. We'd also like to thank those who provided feedback on the paper when it went up for presubmission: Adrian Barnett, Raymond B Huey, Christian Thurn, Christina Maimone, and github user, 'dschief001'. Finally, we thank the following organisations for their support ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.

# Literature Cited {.unnumbered}

::: {#refs}
:::

\newpage

# Bigger Picture Statement {.unnumbered}

Without data, there is no science. Science needs to be reproducible so we can trust the results, and progress as a field. But without data, we do not have true reproducibility. Currently, the role of data in reproducibility is not adequately discussed. We believe this is in part because discussions around data are centered around both data quality and data sharing. However, these are in fact two separate (but related) topics. We provide immediately usable guidelines to sharing data alongside your research, with minimal barriers, that a researcher could pick up and use today.

In the future, we hope that the culture around data sharing will change, and sharing data will be rewarded in science. We want the new normal to be an expectation of data to be submitted with every research publication, and for these datasets to be easily discovered, shared, and extended in other analysis.

# Bio {.unnumbered}

**Nicholas Tierney**

Nicholas Tierney has an honours degree in Psychological Science, and a PhD in Statistics. He is a Lecturer at Monash University, teaching introduction to data analysis. His ranges from exploratory data analysis, to statistical modelling, calculating diagnostics, and how we draw inferences, and make decisions from data. Crucial to this work is producing high quality software to accompany each research idea. Nick is a keen outdoorsman, and loves hiking, trail running, rock climbing, photography, and making coffee.

**Karthik Ram**

Karthik is a senior data scientist at the Berkeley Institute of Data Science and a co-founder of the rOpenSci project. Karthik is also a senior PI at the Berkeley Initiative for Global Change Biology. Prior to joining Berkeley, he earned his PhD in Ecology & Evolution from the University of California, Davis. Karthik's interests are focused on reproducible research, especially as it applies to global change. Much of his recent work focuses on building tools and services around open data and growing diverse data science communities.

### 2020-08-19 figures/carbon/create-carbon-dir-structure.R

@@ -3,6 +3,7 @@ egg <-
   'proj-name
 ├── README.md
 ├── LICENSE
+├── METADATA
 ├── raw-data
 │   ├── raw-data.csv
 │   └── 01-read-tidy-raw.R

### 2020-08-19 figures/carbon/create-carbon-json-ld.R

@@ -1,0 +1,37 @@
+library(carbonate)
+json_ld <- '{
+  "name": "demographics",
+  # here we list the data files in this dataset
+  "resources": [
+    {
+      "path": "demographics.csv",
+      "schema": {
+        "fields": [
+          {
+            "name": "age",
+            "type": "integer"
+          },
+          {
+            "name": "height",
+            "type": "number"
+          },
+          {
+            "name": "nationality",
+            "type": "string"
+          }
+          ]
+      }
+    }
+    ]
+}
+'
+
+json_ld
+
+x <- carbon$new(readr::read_lines(json_ld))
+x$language <- "JSON"
+x$template <- "solarized light"
+x$add_window_control <- FALSE
+x$add_drop_shadow <- FALSE
+x$add_line_number <- TRUE
+x$browse()

### 2020-08-19 new-paper.Rmd
adding some of the missing references

@@
 This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools.
-Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data [@RCore; @python; @julia].
+Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data [@rcore; @python; @julia].
 Each of these languages enjoys support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
@@
 This makes them particularly challenging to standardize or easily "install" where the code is running.
-There are some efforts working towards improving this process, such as gogetdata , the 'dat' project, and a private company, Quilt [@goget; @dat; @Quilt].
+There are some efforts working towards improving this process, such as gogetdata , the 'dat' project, and a private company, Quilt [@goget; @dat; @quilt].
 
@@
 Here we present a set of best practices to enable this change.
-These include documentation (\@ref(readme)), licensing (\@ref(licenses)), citation (\@ref(cite)), code (\@ref(pkgs)), raw and analysis ready data (\@ref(rawdata)).
-We also discuss considerations for sharing of data of various sizes \@ref(size).
+These include documentation (\@ref(documentation-readmes)), licensing (\@ref(licences)), citation (\@ref(citation-how-you-want-your-data-to-be-cited)), code (\@ref(data-as-part-of-open-source-packages)), raw and analysis ready data (\@ref(raw-data-vs-analysis-read-data)).
+We also discuss considerations for sharing of data of various sizes \@ref(considerations-for-data-size).
 
@@
 Without a license, or with a non-standard license, these rules are unclear, and lead to issues with use, attribution and citation.
-We recommend using a license from Creative Commons [@creativecommons], specifically, we recommend CCBY and CC0 as they are commonly used, and provide simple rules for attribution, warranty, and use.
+We recommend using a license from Creative Commons (<https://creativecommons.org/>), specifically, we recommend CCBY and CC0 as they are commonly used, and provide simple rules for attribution, warranty, and use.
 We recommend against non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
@@
 Metadata should provided in plain text files following a standard format.
-Such formats include the Ecological Metadata Language [@EML-about] (EML), which uses XML (The eXtensible Markup Language), and Table Schema, which uses JSON-LD (Java Script Object Notation - Linked Data) [@ref](fig:json-ld).
-One implementation of this is Frictionless Data [@Fowler2017], which provides a language agnostic way to document data - to learn more see [@frictionless-data-guide].
+Such formats include the Ecological Metadata Language [@EML-about] (EML), which uses XML (The eXtensible Markup Language), and Table Schema, which uses JSON-LD (Java Script Object Notation - Linked Data) \@ref(fig:json-ld).
+One implementation of this is Frictionless Data [@Fowler2017], which provides a language agnostic way to document data - to learn more see <https://frictionlessdata.io/guide/>.
 
-EML is widely used in Ecology (for example, <https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-cwt.3091.13>).
+EML is widely used in Ecology (for example [@Maerz2014]).
 JSON-LD provides advantages over EML, by having fields to facilitate data discovery with google dataset search [@EML-data].
@@
 A file named CITATION should be placed in the directory at the top level.
-It should contain a DOI, and could be in `.bibtex` format - [\@ref(fig:dir-layout).
+It should contain a DOI, and could be in `.bibtex` format - \@ref(fig:dir-layout).
 Avenues that provide DOIs are discussed in \@ref(avenues-to-share-data)].
@@
 
-As discussed in \@ref(cite), we recommend data are archived using a platform that generates an accession number such as a DOI.
+As discussed in \@ref(citation-how-you-want-your-data-to-be-cited), we recommend data are archived using a platform that generates an accession number such as a DOI.
 For sharing data collected for research, the most relevant homes are the research data repositories **zenodo**, **dryad**, or **Open Science Framework** (OSF).
-In addition to minting DOIs, they also provide citation templates that can be used in \@ref(cite).
+In addition to minting DOIs, they also provide citation templates that can be used in \@ref(citation-how-you-want-your-data-to-be-cited).
 
@@
 
-Data can be distributed alongside code using a format such as a package in R, Python, and Julia [@nycflights; @penguins; @python-data; @r-data-julia].
+Data can be distributed alongside code using a format such as a package in R, Python, and Julia [@nycflights; @horst-penguins; @data-python-pkg; @r-data-julia].
 A package or module structure provides an advantage as it can be "installed" and loaded just like code.

### 2020-08-19 new-paper.Rmd
tinker with stubs
remove "we recommend"s and provide concrete statements.

@@
 
-1 = Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia
-
-2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)
-
-3 = Berkeley Institute for Data Science, University of California, Berkeley, USA
-
-* = Both authors contributed equally to the work
-
-Correspondence to:
-
+1 = Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia\\
+2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)\\
+3 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
+* = Both authors contributed equally to the work\\
+Correspondence to:\\
 Nicholas Tierney (nicholas.tierney@gmail.com) and Karthik Ram (k@karthik.io)
@@
 Here we present a set of best practices to enable this change.
-These include documentation (\@ref(documentation-readmes)), licensing (\@ref(licences)), citation (\@ref(citation-how-you-want-your-data-to-be-cited)), code (\@ref(data-as-part-of-open-source-packages)), raw and analysis ready data (\@ref(raw-data-vs-analysis-read-data)).
-We also discuss considerations for sharing of data of various sizes \@ref(considerations-for-data-size).
+These include documentation (Section \@ref(readme)), licensing (Section \@ref(licenses)), citation (Section \@ref(cite)), raw and analysis ready data (Section \@ref(rawdata)), code (Section \@ref(scripts)), and sharing in packages (Section \@ref(pkgs)).
+We also discuss considerations for sharing of data of various sizes (Section \@ref(size)).
 
-# Documentation (READMEs) {.readme}
+# Documentation (READMEs) {#readme}
 
+Use a README file with your data project.
 A README is often the first place users will look to learn more about a directory [@Prana2019].
@@
 
-# Licences {.licenses}
+# Licences {#licenses}
 
+Provide a license with your data.
 A license for data clearly establishes the rules for its use, and whether it can be modified or shared.
 Without a license, or with a non-standard license, these rules are unclear, and lead to issues with use, attribution and citation.
-We recommend using a license from Creative Commons (<https://creativecommons.org/>), specifically, we recommend CCBY and CC0 as they are commonly used, and provide simple rules for attribution, warranty, and use.
-We recommend against non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
+Use a license from Creative Commons (<https://creativecommons.org/>), specifically, CCBY and CC0 as they are commonly used, and provide simple rules for attribution, warranty, and use.
+Do not use non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
 Once a license is chosen, a LICENSE file should be provided, containing the entire license in the top level of the directory.
@@
 
-# Metadata {.metadata}
+# Metadata {#metadata}
 
+Use metadata with your data.
 There are three benefits to a dataset having metadata: (1) They provide context by describing variables to humans; (2) They avoid errors by describing and preserving variable types for machine readability; (3) They make data easier to find online by facilitating indexing [@google-data-search].
@@
 **Unit text** defines the type of variable recorded (e.g., string, date, numeric, integer), and avoids serious errors of misrepresented data (e.g., a column of a gene sequence being interpreted as a date).
-For more details on this we recommend reading [@Broman2017; @Ellis2017; @Arslan2019].
+For more details, see [@Broman2017; @Ellis2017; @Arslan2019].
 
@@
 JSON-LD provides advantages over EML, by having fields to facilitate data discovery with google dataset search [@EML-data].
-We recommend JSON-LD over EML, as it is more human readable, shorter, and quicker and easier to read and write.
+Use JSON-LD over EML, as it is more human readable, shorter, and quicker and easier to read and write.
 For more information on tooling to assist generating metadata, see <https://knb.ecoinformatics.org/tools/morpho> and [@Arslan2019].
@@
 
-# Raw data vs analysis read data {.rawdata}
+# Raw data vs analysis ready data {#rawdata}
 
+Make raw and analysis ready data available.
 Raw data is the first format of data provided before any cleaning has taken place (e.g., data in some binary or proprietary format).
@@
 We don't have particular recommendations for non tabular of data, like binary images, but given the ubiquity of tables in papers, these recommendations will be broadly useful.
-For other data formats we recommend using a copy of the most robust common format and avoid the latest tools that may be more brittle and likely to change in the future.
+For other data formats use a copy of the most robust common format and avoid the latest tools that may be more brittle and likely to change in the future.
 For example, for an image, a TIFF image in addition to a custom photoshop file.
 
-# Data cleaning scripts {.scripts}
+# Data cleaning scripts {#scripts}
 
-Code used to transform data from its raw format into analysis ready form should be kept in the same place as raw data (Section \@ref(fig:dir-layout)).
+Share code used to transform data from its raw format into analysis ready form.
+This should be kept in the same place as raw data (Section \@ref(fig:dir-layout)).
 Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file.
 
-# Citation: How you want your data to be cited {.cite}
+# Citation: How you want your data to be cited {#cite}
 
+Describe how you want your data to be cited.
 Citation is important for provenance, and credit.
@@
 A file named CITATION should be placed in the directory at the top level.
-It should contain a DOI, and could be in `.bibtex` format - \@ref(fig:dir-layout).
-Avenues that provide DOIs are discussed in \@ref(avenues-to-share-data)].
+It should contain a DOI, and could be in `.bibtex` format (Figure \@ref(fig:dir-layout)).
+Avenues that provide DOIs are discussed in Section \@ref(avenues).
 
-# Avenues to share data {.avenues}
+# Avenues to share data {#avenues}
 
@@
 
-As discussed in \@ref(citation-how-you-want-your-data-to-be-cited), we recommend data are archived using a platform that generates an accession number such as a DOI.
+As discussed in \@ref(citation-how-you-want-your-data-to-be-cited),archive data using a platform generating an accession number such as a DOI.
 For sharing data collected for research, the most relevant homes are the research data repositories **zenodo**, **dryad**, or **Open Science Framework** (OSF).
@@
 Importantly, these all provide some form of accession number.
-We recommend adhering to community standards when deciding where to publish data, as this will likely improve data reuse.
+Adhere to community standards when deciding where to publish data, as this will likely improve data reuse.
 Large data may also be shared via domain specific venues (e.g., in genomics or astronomy), where appropriate.
 
-We recommend against depositing data in local file storage that does not provide DOI or accession numbers.
+Do not deposit data in local file storage that does not provide DOI or accession numbers.
 For example, [@50Hectares] is hosted on a university server and accessed via FTP.
@@
 Writing a data paper can be very useful for researchers, however, data papers lack guidelines on how to structure data for effective reuse and sharing.
-We recommend partnering with [@dryad] to archive data and link it with research papers, as they provide a person to help curate your data, improving usability.
+Partner with Dryad [@dryad] to link archived data with research papers.
+Dryad will also provide a person to help curate your data, improving usability.
 
-# Data as part of open source packages {.pkgs}
+# Data as part of open source packages {#pkgs}
 
@@
 Packaged data and archival on venues like Zenodo are not mutually exclusive.
-We recommend researchers first archive data in a long-term data repository (Section \@ref(avenues-to-share-data)), before including it in a package.
+First, archive data in a long-term data repository (Section \@ref(avenues)), before including it in a package.
 This makes the data easily citable wherever data is referenced.
 
-# Considerations for data size {.size}
+# Considerations for data size {#size}
 
@@
 
-Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(raw-data-vs-analysis-read-data)), or in software \@ref(data-as-part-of-open-source-packages).
+Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software \@ref(data-as-part-of-open-source-packages).
 For medium data, consider a venue like zenodo for archiving.
 Sharing medium data might require it to be broken into separate chunks for reading (e.g., `teaching-1.csv`, `teaching-2.csv`, etc.), and scripts provided to read it.
-Large data will require liaising with services from \@ref(avenues-to-share-data), or local research institute library to make special arrangements, to ensure there is some way to gain access, and also maintain a DOI.
+Large data will require liaising with services from \@ref(avenues), or local research institute library to make special arrangements, to ensure there is some way to gain access, and also maintain a DOI.
 

### 2020-08-19 figures/carbon/create-carbon-dir-structure.R
add carbon figures

@@ -3,7 +3,12 @@ egg <-
   'proj-name
 ├── README.md
 ├── LICENSE
-├── METADATA
+├── CITATION
+├── metadata
+│   ├── access.csv
+│   ├── attributed.csv
+│   ├── biblio.csv
+│   └── creators.csv
 ├── raw-data
 │   ├── raw-data.csv
 │   └── 01-read-tidy-raw.R

### 2020-08-25 new-paper.Rmd
read table image

@@
 This is in contrast to Code having no policy"}
-library(gt)
-incentive_table <- tibble(
-  Mechanism = c("Policy", "Incentives", "Communities", "UI/UX", "Infrastructure"),
-  Incentive = c("Required", "Rewarding", "Normative", "Easy", "Possible"),
-  Data = c(100, 0, 0, 0, 100),
-  Code = c(0, rep(100, 4)),
-  `Computational Environment` = c(0, 0, 0, 50, 100)
-  )
-
-incentive_table %>%
-  gt() %>%
-  data_color(
-    columns = vars(Data, Code, `Computational Environment`),
-    colors = scales::col_numeric(
-      # Using a function from paletteer to generate a vector of colors
-      # Note that you need to wrap paletteer_d outputs in as.character()
-      palette = colorspace::sequential_hcl(n = 3, palette = "Greens2", rev = TRUE),
-      # Domain is from scales::col_numeric
-      # Domain = The possible values that can be mapped
-      # We don't HAVE to set a range for this since
-      # we're using it inside data_color()
-      domain = NULL
-    )
-  ) %>%
-  gtsave("figures/gt-incentive-table.png")
-
-# knitr::include_graphics("figures/gt-incentive-table.png")
+knitr::include_graphics("figures/gt-incentive-table.png")
 ```

### 2020-08-25 figures/create-incentives-table.R

@@ -1,0 +1,27 @@
+library(gt)
+incentive_table <- tibble(
+  Mechanism = c("Policy", "Incentives", "Communities", "UI/UX", "Infrastructure"),
+  Incentive = c("Required", "Rewarding", "Normative", "Easy", "Possible"),
+  Data = c(100, 0, 0, 0, 100),
+  Code = c(0, rep(100, 4)),
+  `Computational Environment` = c(0, 0, 0, 50, 100)
+)
+
+incentive_table %>%
+  gt() %>%
+  data_color(
+    columns = vars(Data, Code, `Computational Environment`),
+    colors = scales::col_numeric(
+      # Using a function from paletteer to generate a vector of colors
+      # Note that you need to wrap paletteer_d outputs in as.character()
+      palette = colorspace::sequential_hcl(n = 3, palette = "Greens2", rev = TRUE),
+      # Domain is from scales::col_numeric
+      # Domain = The possible values that can be mapped
+      # We don't HAVE to set a range for this since
+      # we're using it inside data_color()
+      domain = NULL
+    )
+  ) %>%
+  gtsave("figures/gt-incentive-table.png")
+
+# knitr::include_graphics("figures/gt-incentive-table.png")

### 2020-08-25 new-paper.Rmd
update 50 acres

@@
 Do not deposit data in local file storage that does not provide DOI or accession numbers.
-For example, [@50Hectares] is hosted on a university server and accessed via FTP.
-It is not possible to effectively provide effective provenance and credit.
+For example, [@50Hectares] was hosted on a university server and accessed via FTP for over a decade.
+It was not possible to effectively provide effective provenance and credit.
+Only recently was the data deposited in Dryad [@condit-2019].
 

### 2020-08-25 new-paper.Rmd
clarify point on 50 hectares paper

--- 2020-08-25 new-paper.Rmd
@@
 Do not deposit data in local file storage that does not provide DOI or accession numbers.
-For example, [@50Hectares] is hosted on a university server and accessed via FTP.
-It is not possible to effectively provide effective provenance and credit.
+For example, [@50Hectares] was hosted on a university server and only accessible via a FTP server since 1982.
+It was not possible to easily access or cite the data without an explicit authorship agreement among other conditions.
+Only recently were the complete dataset deposited in Dryad without such conditions [@condit-2019].
 

### 2020-08-25 new-paper.Rmd
updates title, adds possible titles and references.

@@
 ---
-title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.
+title: Meaningful data sharing.
 author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)
@@
 \begin{singlespace}
+Other titles under consideration:
+
+- Data sharing practices to realign incentives and enable positive/meaningful culture change
+
+- Enabling culture change for meaningful data sharing
+
+- Enabling culture change for meaningful data sharing: Aligning incentives and best practices
+
+- Principles of sharing data
+
+- Elements of meaningful data sharing
+
+- Better practices in data sharing to enable culture change
 
@@
 Correspondence to:\\
-Nicholas Tierney (nicholas.tierney@gmail.com) and Karthik Ram (k@karthik.io)
+Nicholas Tierney (nicholas.tierney@gmail.com) and Karthik Ram (karthik.ram@berkeley.edu)
 
@@
 EML is widely used in Ecology (for example [@Maerz2014]).
-JSON-LD provides advantages over EML, by having fields to facilitate data discovery with google dataset search [@EML-data].
+JSON-LD provides advantages over EML, by having fields to facilitate data discovery with google dataset search [@google-data-search].
 Use JSON-LD over EML, as it is more human readable, shorter, and quicker and easier to read and write.
@@
 
-As discussed in \@ref(citation-how-you-want-your-data-to-be-cited),archive data using a platform generating an accession number such as a DOI.
+As discussed in \@ref(cite),archive data using a platform generating an accession number such as a DOI.
 For sharing data collected for research, the most relevant homes are the research data repositories **zenodo**, **dryad**, or **Open Science Framework** (OSF).
-In addition to minting DOIs, they also provide citation templates that can be used in \@ref(citation-how-you-want-your-data-to-be-cited).
+In addition to minting DOIs, they also provide citation templates that can be used in \@ref(cite).
 
 Zenodo can be contacted to arrange special cases for storage of large data.
-Special hosting can be arranged via specially contacting zenodo or your research institute.
+Special hosting can be arranged via specially contacting Zenodo or your research institute.
 If using a research institute's library service to share data, ensure a DOI is provided so it can be accurately identified and cited.
@@
 Do not deposit data in local file storage that does not provide DOI or accession numbers.
-For example, [@50Hectares] was hosted on a university server and only accessible via a FTP server since 1982.
+For example, [@Hubbell1999] was hosted on a university server and only accessible via a FTP server since 1982.
 It was not possible to easily access or cite the data without an explicit authorship agreement among other conditions.
@@
 Although it is fine to share data among collaborators using temporary cloud storage (like Dropbox, Google Drive), once the paper is public, the data needs to be hosted somewhere persistent.
-While this approach is useful while iterating on the paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc).
+While this approach is useful while iterating on the paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc.).
 Just archive the data.
@@
 Users who work in other languages are unlikely to download and export data out of a package.
-Secondly, data size can be a limitation depending on the repository (CRAN, the R repository, for example, limits package sizes to 5mb although there are workarounds [@Brooke_Anderson2017]).
+Secondly, data size can be a limitation depending on the repository (CRAN, the R repository, for example, limits package sizes to 5Mb although there are workarounds [@Brooke_Anderson2017]).
 Other fields, such as bioinformatics have a process of sharing data through repositories like Bioconductor, with larger data size.
@@
 Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software \@ref(data-as-part-of-open-source-packages).
-For medium data, consider a venue like zenodo for archiving.
+For medium data, consider a venue like Zenodo for archiving.
 Sharing medium data might require it to be broken into separate chunks for reading (e.g., `teaching-1.csv`, `teaching-2.csv`, etc.), and scripts provided to read it.

### 2020-08-25 new-paper.Rmd
remove . from title

@@
 ---
-title: Meaningful data sharing.
+title: Meaningful data sharing
 author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)

### 2020-08-25 new-paper.Rmd
added footnotes (not needed anymore, but there it is)

--- 2020-08-25 new-paper.Rmd
@@
 bibliography: references.bib
+csl: nature-communications.csl
 header-includes: 
@@
 This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools.
-Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data [@rcore; @python; @julia].
+Programming languages such as Python[^1], R and Julia have helped scientists implement and share new methods to work with data [@rcore; @julia].
 Each of these languages enjoys support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
@@
 The barriers to verification are now low enough that trivial reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].
+
+[^1]: <https://www.python.org/>
 
@@
 This is in contrast to Code having no policy"}
-knitr::include_graphics("figures/gt-incentive-table.png")
+library(gt)
+incentive_table <- tibble(
+  Mechanism = c("Policy", "Incentives", "Communities", "UI/UX", "Infrastructure"),
+  Incentive = c("Required", "Rewarding", "Normative", "Easy", "Possible"),
+  Data = c(100, 0, 0, 0, 100),
+  Code = c(0, rep(100, 4)),
+  `Computational Environment` = c(0, 0, 0, 50, 100)
+  )
+
+incentive_table %>%
+  gt() %>%
+  data_color(
+    columns = vars(Data, Code, `Computational Environment`),
+    colors = scales::col_numeric(
+      # Using a function from paletteer to generate a vector of colors
+      # Note that you need to wrap paletteer_d outputs in as.character()
+      palette = colorspace::sequential_hcl(n = 3, palette = "Greens2", rev = TRUE),
+      # Domain is from scales::col_numeric
+      # Domain = The possible values that can be mapped
+      # We don't HAVE to set a range for this since
+      # we're using it inside data_color()
+      domain = NULL
+    )
+  ) %>%
+  gtsave("figures/gt-incentive-table.png")
+
+# knitr::include_graphics("figures/gt-incentive-table.png")
 ```
@@
 This makes them particularly challenging to standardize or easily "install" where the code is running.
-There are some efforts working towards improving this process, such as gogetdata , the 'dat' project, and a private company, Quilt [@goget; @dat; @quilt].
+There are some efforts working towards improving this process, such as gogetdata[^2] , dat[^3], and Quilt[^4].
+
+[^2]: <https://gogetdata.github.io/>
+
+[^3]: <https://dat.foundation/>
+
+[^4]: <https://quiltdata.com/>
 
@@
 A README is often the first place users will look to learn more about a directory [@Prana2019].
-The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/> and [@Prana2019].
+The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how[^5] [@Prana2019].
 For example, a README in a data repository might explain which directories contain data, tidying scripts, and also why data was collected.
@@
 In the context of datasets, READMEs are particularly useful when there are no reliable standards.
+
+[^5]: <https://betta.io/blog/2017/02/07/developer-experience-github-readmes/>
 
@@
 Without a license, or with a non-standard license, these rules are unclear, and lead to issues with use, attribution and citation.
-Use a license from Creative Commons (<https://creativecommons.org/>), specifically, CCBY and CC0 as they are commonly used, and provide simple rules for attribution, warranty, and use.
+Use a license from Creative Commons[^6], specifically, CCBY and CC0 as they are commonly used, and provide simple rules for attribution, warranty, and use.
 Do not use non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
@@
 The license type should be mentioned in the README, if used.
+
+[^6]: <https://creativecommons.org/>
 
@@
 Some data repositories and journals will automatically require a license when submitting data.
-For example, Dryad requires CC0 (<https://blog.datadryad.org/2011/10/05/why-does-dryad-use-cc0/> and <https://datadryad.org/stash/faq>) and the journal PLOS Comp Bio, requires licences for data that are not more restrictive than CC BY [@plos-comp-bio-data].
+For example, Dryad requires CC0[^7][^8], and the journal PLOS Comp Bio, requires licences for data that are not more restrictive than CC BY[^9].
+
+[^7]: <https://blog.datadryad.org/2011/10/05/why-does-dryad-use-cc0/>
+
+[^8]: <https://datadryad.org/stash/faq>
+
+[^9]: <https://journals.plos.org/ploscompbiol/s/data-availability>
 
@@
 
-**Variable names** are short, descriptive names with no spaces or special characters, such as "job\_position" or "date".
+**Variable names** are short, descriptive names with no spaces or special characters, such as "job_position" or "date".
 **Variable description** describes the measured variables, e.g., "University Job Position" or "Date Started", and should also define *variable codes* and *missing data*.
@@
 Do not deposit data in local file storage that does not provide DOI or accession numbers.
-For example, [@50Hectares] was hosted on a university server and only accessible via a FTP server since 1982.
-It was not possible to easily access or cite the data without an explicit authorship agreement among other conditions.
-Only recently were the complete dataset deposited in Dryad without such conditions [@condit-2019].
+For example, [@50Hectares] is hosted on a university server and accessed via FTP.
+It is not possible to effectively provide effective provenance and credit.
 
@@
 Writing a data paper can be very useful for researchers, however, data papers lack guidelines on how to structure data for effective reuse and sharing.
-Partner with Dryad [@dryad] to link archived data with research papers.
+Partner with Dryad[^10] to link archived data with research papers.
 Dryad will also provide a person to help curate your data, improving usability.
+
+[^10]: <https://datadryad.org/>
 
@@
 
-Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software \@ref(data-as-part-of-open-source-packages).
+Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software \@ref(pkgs).
 For medium data, consider a venue like zenodo for archiving.

### 2020-09-18 new-paper.Rmd
notes from meeting

@@
 
+**note: other little bit to say**
+
+- "Here are some high level ideas on how to share data, that might obvious but are not often considered"
+
 # Documentation (READMEs) {#readme}
 
-Use a README file with your data project.
-A README is often the first place users will look to learn more about a directory [@Prana2019].
+Use a README file with your data project to communicate its purpose and contents.
+A README is often the first place users will look to learn more about a project [@Prana2019].
 The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/> and [@Prana2019].
@@
 In the context of datasets, READMEs are particularly useful when there are no reliable standards.
+For example, the data from Dryad, https://datadryad.org/stash/dataset/doi:10.5061/dryad.rn6k1 could be greatly enhanced by providing a README file that explains what the variables in the data collected are, and what the different sheets of the excel worksheet contain.
 
@@
 Without a license, or with a non-standard license, these rules are unclear, and lead to issues with use, attribution and citation.
-Use a license from Creative Commons (<https://creativecommons.org/>), specifically, CCBY and CC0 as they are commonly used, and provide simple rules for attribution, warranty, and use.
-Do not use non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
-Once a license is chosen, a LICENSE file should be provided, containing the entire license in the top level of the directory.
-The license type should be mentioned in the README, if used.
-
-CC BY requires appropriate credit and attribution, but data can be shared and adapted, even for commercial use, provided changes are described.
+Use the CC0 license, as it is commonly used, and provide simple rules for attribution, warranty, and use.
 The CC0 is a "public domain" license, meaning data owners waive all rights, and the work is now "owned" by the public.
@@
 When using data with CC0, it is good practice to cite the original work, but is not required.
-Both CCBY and CC0 provide no warranty.
-Brief and full overviews of CCBY and CC0 can be found at [@ccby-short; @ccby-long; @cc0-short; @cc0-long].
+For more information on the benefit of using CC0, and why CCBY is not always appropriate, see https://osc.universityofcalifornia.edu/2016/09/cc-by-and-data-not-always-a-good-fit/.
+Short and full overviews of CCBY and CC0 can be found at [@cc0-short; @cc0-long].
 
-Some data repositories and journals will automatically require a license when submitting data.
-For example, Dryad requires CC0 (<https://blog.datadryad.org/2011/10/05/why-does-dryad-use-cc0/> and <https://datadryad.org/stash/faq>) and the journal PLOS Comp Bio, requires licences for data that are not more restrictive than CC BY [@plos-comp-bio-data].
+Do not use non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
+Once a license is chosen, a LICENSE file should be provided, containing the entire license in the top level of the directory (Figure \@ref(fig:dir-layout)).
+The license type should be mentioned in the README, if used.
 
@@
 For more details, see [@Broman2017; @Ellis2017; @Arslan2019].
-
-Metadata should provided in plain text files following a standard format.
-Such formats include the Ecological Metadata Language [@EML-about] (EML), which uses XML (The eXtensible Markup Language), and Table Schema, which uses JSON-LD (Java Script Object Notation - Linked Data) \@ref(fig:json-ld).
-One implementation of this is Frictionless Data [@Fowler2017], which provides a language agnostic way to document data - to learn more see <https://frictionlessdata.io/guide/>.
-
-EML is widely used in Ecology (for example [@Maerz2014]).
-JSON-LD provides advantages over EML, by having fields to facilitate data discovery with google dataset search [@google-data-search].
-Use JSON-LD over EML, as it is more human readable, shorter, and quicker and easier to read and write.
-For more information on tooling to assist generating metadata, see <https://knb.ecoinformatics.org/tools/morpho> and [@Arslan2019].
-
-```{r json-ld, fig.cap = "Example snippet of some Table Schema data for a dataset with three variables. This provides a description of each field, and the type of field."}
-knitr::include_graphics("figures/carbon-json-ld.png")
-```
+Metadata can be encoded into plain text formats that follow a set standard, such as EML or SCHEMA-LD.
+A deep understanding of these formats isn't required to use them, for more details, and guidance on generating them see @dataspice and [@Arslan2019].
 
@@
 
+(**note: this bit needs some work**)
+A trusted repository (such as Dryad) will manage citation for you.
+However, a CITATION file allows you to mention specifically how you want your data to be cited - this could point to the paper that used or described the data and collection methods, for example.
+
 # Avenues to share data {#avenues}
 
-There are different avenues to share data, and although the terms are sometimes used interchangeably, they are, in fact, different.
-These include: (1) Data sharing: Making data publicly available; (2) Data archiving: Depositing data somewhere permanent; (3) Data publishing: writing a paper to explain the archived data.
+(**note: on what to say**)
+1.
+Data should go into Domain area - e.g., gbif for species data, or genbif or genbank for genetics data @GenBank.
+More information can be found at https://www.nature.com/sdata/policies/repositories.
+Importantly, domain areas should provide some form of accession number.
 
 As discussed in \@ref(cite),archive data using a platform generating an accession number such as a DOI.
-For sharing data collected for research, the most relevant homes are the research data repositories **zenodo**, **dryad**, or **Open Science Framework** (OSF).
+For sharing data collected for research, the most relevant homes for general data are the research data repositories **zenodo** and **dryad**.
 In addition to minting DOIs, they also provide citation templates that can be used in \@ref(cite).
 
-Zenodo can be contacted to arrange special cases for storage of large data.
-Special hosting can be arranged via specially contacting Zenodo or your research institute.
 If using a research institute's library service to share data, ensure a DOI is provided so it can be accurately identified and cited.
-
-There are many alternatives to these three, for example, an astronomy project may host its data at the Sloan Digital Sky Survey (SDSS), and genetic data can be hosted at GenBank [@GenBank].
-Importantly, these all provide some form of accession number.
-Adhere to community standards when deciding where to publish data, as this will likely improve data reuse.
-Large data may also be shared via domain specific venues (e.g., in genomics or astronomy), where appropriate.
 
@@
 
+**mention bit about**
+
+- https://www.the-scientist.com/news-opinion/top-spider-biologists-research-under-fire-67083
+- https://www.nature.com/articles/d41586-020-00287-y
+
 Journals provide data papers, where they provide a formal publication explaining data that is archived.
@@
 Dryad will also provide a person to help curate your data, improving usability.
+
+**bit removed from this section**
+
+There are different avenues to share data, and although the terms are sometimes used interchangeably, they are, in fact, different.
+These include: (1) Data sharing: Making data publicly available; (2) Data archiving: Depositing data somewhere permanent; (3) Data publishing: writing a paper to explain the archived data.
 

### 2020-09-22 new-paper.Rmd
minor wording/flow clean up of license, metadata, citation.

@@
 We note that Data is often required, but the preceding steps are not required.
-This is in contrast to Code having no policy"}
+This is in contrast to Code having no policy."}
 knitr::include_graphics("figures/gt-incentive-table.png")
@@
 Without a license, or with a non-standard license, these rules are unclear, and lead to issues with use, attribution and citation.
-Use the CC0 license, as it is commonly used, and provide simple rules for attribution, warranty, and use.
+Use the CC0 license, as it provides simple rules for attribution, warranty, and use, and is very common.
 The CC0 is a "public domain" license, meaning data owners waive all rights, and the work is now "owned" by the public.
@@
 
-Do not use non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
+Avoid using non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
 Once a license is chosen, a LICENSE file should be provided, containing the entire license in the top level of the directory (Figure \@ref(fig:dir-layout)).
@@
 Use metadata with your data.
-There are three benefits to a dataset having metadata: (1) They provide context by describing variables to humans; (2) They avoid errors by describing and preserving variable types for machine readability; (3) They make data easier to find online by facilitating indexing [@google-data-search].
+There are three benefits to a dataset having metadata: (1) Provide context by describing variables to humans; (2) Avoid errors by describing and preserving variable types for machine readability; (3) Make data easier to find online by facilitating indexing [@google-data-search].
 Three pieces of metadata should be considered: (1) **variable names**, (2) **variable descriptions**, and (3) **unit text**.
@@
 **Variable description** describes the measured variables, e.g., "University Job Position" or "Date Started", and should also define *variable codes* and *missing data*.
-**Unit text** defines the type of variable recorded (e.g., string, date, numeric, integer), and avoids serious errors of misrepresented data (e.g., a column of a gene sequence being interpreted as a date).
+**Unit text** defines the type of variable recorded (e.g., string, date, numeric, integer), this avoids serious errors of misrepresented data (e.g., a column of a gene sequence being interpreted as a date).
 For more details, see [@Broman2017; @Ellis2017; @Arslan2019].
 Metadata can be encoded into plain text formats that follow a set standard, such as EML or SCHEMA-LD.
-A deep understanding of these formats isn't required to use them, for more details, and guidance on generating them see @dataspice and [@Arslan2019].
+A deep understanding of these formats isn't required to use them, for more details, and guidance on generating them see @dataspice and @Arslan2019.
 
@@
 
-Describe how you want your data to be cited.
+Describe how you want your data to be cited in a CITATION file.
 Citation is important for provenance, and credit.
@@
 To ensure these are provided, data needs to have an accession number, such as a Digital Object Identifier (DOI).
-A DOI is a permanent, unique identifier for a digital object such as a paper, poster, or software, and is a prerequisite for citation.
-If a DOI or accession number is unavailable, a citation will be meaningless, as it cannot be tracked by any means.
-A file named CITATION should be placed in the directory at the top level.
-It should contain a DOI, and could be in `.bibtex` format (Figure \@ref(fig:dir-layout)).
+A DOI is a permanent, unique identifier for a digital object such as a paper, poster, or software.
+Access numbers are a prerequisite for citation - if they are unavailable, a citation will be meaningless, as it cannot be tracked by any means.
+A file named CITATION should be placed in the directory at the top level (Figure \@ref(fig:dir-layout)).
+It should contain an accession number, and could be in `.bibtex` format.
 Avenues that provide DOIs are discussed in Section \@ref(avenues).
 
-(**note: this bit needs some work**)
-A trusted repository (such as Dryad) will manage citation for you.
-However, a CITATION file allows you to mention specifically how you want your data to be cited - this could point to the paper that used or described the data and collection methods, for example.
+Repositories such as Dryad will manage citation.
+However it is worth noting that a CITATION file allows you to customise how you want your data to be cited.
+For example, to cite the paper that used or described the data and collection methods.
 
@@
 
-As discussed in \@ref(cite),archive data using a platform generating an accession number such as a DOI.
+As discussed in \@ref(cite), archive data using a platform generating an accession number such as a DOI.
 For sharing data collected for research, the most relevant homes for general data are the research data repositories **zenodo** and **dryad**.

### 2020-09-22 new-paper.Rmd
working on avenues

@@
 
-(**note: on what to say**)
-1.
-Data should go into Domain area - e.g., gbif for species data, or genbif or genbank for genetics data @GenBank.
-More information can be found at https://www.nature.com/sdata/policies/repositories.
-Importantly, domain areas should provide some form of accession number.
+(**note: still need to work on this**)
+
+Data should be deposited into a relevant domain repository.
+For example, gbif [@gbif] for species data, or genbif or genbank for genetics data @GenBank.
+Information on other domain specific repositories can be found at https://www.nature.com/sdata/policies/repositories.
 
@@
 For sharing data collected for research, the most relevant homes for general data are the research data repositories **zenodo** and **dryad**.
-In addition to minting DOIs, they also provide citation templates that can be used in \@ref(cite).
+In addition to minting DOIs, they also provide citation templates.
 

### 2020-09-23 new-paper.Rmd
add some comments from meeting, finish draft of avenues section

@@
 
+> Someone can mention links to a Dockerfile fro the README, or a text file that describes versions of software used, date when things last ran successfully etc.
+> Also include the equivalent of a requirements.txt
+> The date of use, and versions of software should be recorded in these scripts, so users can in the future can replicate similar computing environment or computing conditions.
+A more ideal approach would be to clearly describe the computing environment (e.g., a Dockerfile).
+
 # Licences {#licenses}
@@
 Data used in analysis should be provided, especially if computationally intense to derive.
-It should be in its own folder, and to maximise interoperability, recorded in a plain text (e.g., csv, tsv), not a binary format (e.g., .rds, .sav, .xlsx).
+It should be in its own folder, and to maximise interoperability, recorded in a plain text (e.g., csv, tsv) and if possible a binary format (e.g., .rds, .sav, .xlsx, Apache Parquet).
+The binary format allows for immediate reuse in the intended state, while the interopability of plain text versions remain stable over time.
+
 Ideally data should be in "Tidy Data" form, where variables are in columns, observations in rows, and each cell has only one value [@Wickham2014].
@@
 Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file.
+Authors should run the scripts in a clean environment to make sure everything runs and that the code doesn’t rely on other local artifacts.
 
@@
 
-(**note: still need to work on this**)
-
-Data should be deposited into a relevant domain repository.
-For example, gbif [@gbif] for species data, or genbif or genbank for genetics data @GenBank.
+Data should be deposited into a relevant domain repository that generates an accession number such as a DOI.
+For example, gbif [@gbif] for species data, or genbank for genetics data [@GenBank].
 Information on other domain specific repositories can be found at https://www.nature.com/sdata/policies/repositories.
-
-As discussed in \@ref(cite), archive data using a platform generating an accession number such as a DOI.
-For sharing data collected for research, the most relevant homes for general data are the research data repositories **zenodo** and **dryad**.
+Data not falling into domain repositories can be submitted to research data repositories **zenodo** and **dryad**.
 In addition to minting DOIs, they also provide citation templates.
@@
 If using a research institute's library service to share data, ensure a DOI is provided so it can be accurately identified and cited.
-
 Do not deposit data in local file storage that does not provide DOI or accession numbers.
@@
 Just archive the data.
-
-**mention bit about**
-
-- https://www.the-scientist.com/news-opinion/top-spider-biologists-research-under-fire-67083
-- https://www.nature.com/articles/d41586-020-00287-y
+Providing back up data is particularly relevant in light of the case in [@Schleunes2020; @Viglione2020], where data was suspected of being tampered.
+The use of raw data allowed for issues to be identified, the absence of raw data not being available in other publications created difficulty in verifying the validity of the data.
 
@@
 Dryad will also provide a person to help curate your data, improving usability.
-
-**bit removed from this section**
-
-There are different avenues to share data, and although the terms are sometimes used interchangeably, they are, in fact, different.
-These include: (1) Data sharing: Making data publicly available; (2) Data archiving: Depositing data somewhere permanent; (3) Data publishing: writing a paper to explain the archived data.
 

### 2020-09-23 new-paper.Rmd
remove carbon sh figure for dir layout and other unused figures

@@
 Avoid using non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
-Once a license is chosen, a LICENSE file should be provided, containing the entire license in the top level of the directory (Figure \@ref(fig:dir-layout)).
+Once a license is chosen, a LICENSE file should be provided, containing the entire license in the top level of the directory.
 The license type should be mentioned in the README, if used.
-
-```{r dir-layout, fig.cap = "Example directory layout and structure for sharing data. Note ..."}
-# insert carbon.sh figure showing file structure layout
-knitr::include_graphics("figures/carbon-dir-layout.png")
-```
 
@@
 Share code used to transform data from its raw format into analysis ready form.
-This should be kept in the same place as raw data (Section \@ref(fig:dir-layout)).
+This should be kept in the same place as raw data.
 Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file.
@@
 Access numbers are a prerequisite for citation - if they are unavailable, a citation will be meaningless, as it cannot be tracked by any means.
-A file named CITATION should be placed in the directory at the top level (Figure \@ref(fig:dir-layout)).
+A file named CITATION should be placed in the directory at the top level.
 It should contain an accession number, and could be in `.bibtex` format.

### 2020-09-23 new-paper.Rmd
best practices --> minimally acceptable practices
other minor spellchecks

@@
 
-- Enabling culture change for meaningful data sharing: Aligning incentives and best practices
+- Enabling culture change for meaningful data sharing: Aligning incentives and minimally acceptable practices
 
@@
 The movement, which has seen shifting support over the years, has enjoyed growing popularity over the past decade [@Nielsen2020; @Peng2011] mainly among early career researchers [@McKiernan2016].
-A key motivation for practicing openness is the notion that sharing all research artifacts will allow others to reproduce and verify results, thereby improving trust and verification of those findings.
+A key motivation for practicing openness is the notion that sharing all research artefacts will allow others to reproduce and verify results, thereby improving trust and verification of those findings.
 Jon Clarebout famously wrote in 1992 that reproducibility in computationally oriented fields had become so trivial that non experts could do the task.
@@
 Meaningful progress on computational reproducibility, the ability of an independent researcher to recreate figures, tables and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence.
-Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by organizations like The Carpentries, combined with the growing adoption of Github by scientists [@Ram2013; @Zagalsky2015].
+Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by organizations like The Carpentries, combined with the growing adoption of GitHub by scientists [@Ram2013; @Zagalsky2015].
 The bigger driver for this is more likely driven by the growing popularity of data science as a discipline distinct from statistics [@Donoho2017].
@@
 
-Tools to prepare datasets and metadata (*the infrastructure*) and best practices (guidelines) for reuse are sparse [@Arslan2019; @Roche2015].
+Tools to prepare datasets and metadata (*the infrastructure*) and minimally acceptable practices (guidelines) for reuse are sparse [@Arslan2019; @Roche2015].
 Unlike code, datasets are far more diverse in terms of heterogeneity, size, and formats.
@@
 To make data more reusable, we need to make it easier to generate metadata, standardise good practices within and across communities, and make reusable data sharing a recognized, normalised, part of research.
-Here we present a set of best practices to enable this change.
+Here we present a set of minimally acceptable practices to enable this change.
 These include documentation (Section \@ref(readme)), licensing (Section \@ref(licenses)), citation (Section \@ref(cite)), raw and analysis ready data (Section \@ref(rawdata)), code (Section \@ref(scripts)), and sharing in packages (Section \@ref(pkgs)).
@@
 
-> Someone can mention links to a Dockerfile fro the README, or a text file that describes versions of software used, date when things last ran successfully etc.
-> Also include the equivalent of a requirements.txt
-> The date of use, and versions of software should be recorded in these scripts, so users can in the future can replicate similar computing environment or computing conditions.
-A more ideal approach would be to clearly describe the computing environment (e.g., a Dockerfile).
+The date of use, and versions of software should be recorded in the README, so users can in the future can replicate similar computing environment or computing conditions.
+A more ideal approach would be to clearly describe the computing environment (e.g., a Dockerfile) and packages used (e.g., requirements.txt).
 
@@
 It should be in its own folder, and to maximise interoperability, recorded in a plain text (e.g., csv, tsv) and if possible a binary format (e.g., .rds, .sav, .xlsx, Apache Parquet).
-The binary format allows for immediate reuse in the intended state, while the interopability of plain text versions remain stable over time.
-
+The binary format allows for immediate reuse in the intended state, while the interoperability of plain text versions remain stable over time.
 Ideally data should be in "Tidy Data" form, where variables are in columns, observations in rows, and each cell has only one value [@Wickham2014].
@@
 Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file.
-Authors should run the scripts in a clean environment to make sure everything runs and that the code doesn’t rely on other local artifacts.
+Authors should run the scripts in a clean environment to ensure everything runs and that the code doesn't rely on other local artefacts, such as their own folder layout not available to others.
 
@@
 We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Anna Krystalli, Daniella Lowenberg, and Miles McBain.
-We'd also like to thank those who provided feedback on the paper when it went up for presubmission: Adrian Barnett, Raymond B Huey, Christian Thurn, Christina Maimone, and github user, 'dschief001'.
+We'd also like to thank those who provided feedback on the paper when it went up for presubmission: Adrian Barnett, Raymond B Huey, Christian Thurn, Christina Maimone, and GitHub user, 'dschief001'.
 Finally, we thank the following organisations for their support ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.

### 2020-09-23 new-paper.Rmd

@@
 
-Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software \@ref(data-as-part-of-open-source-packages).
+Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software \@ref(pkgs).
 For medium data, consider a venue like Zenodo for archiving.

### 2020-09-23 new-paper.Rmd
Finished rest of Karthik's tasks

@@
 The movement, which has seen shifting support over the years, has enjoyed growing popularity over the past decade [@Nielsen2020; @Peng2011] mainly among early career researchers [@McKiernan2016].
-A key motivation for practicing openness is the notion that sharing all research artefacts will allow others to reproduce and verify results, thereby improving trust and verification of those findings.
+A key motivation for practicing openness is the notion that sharing all research artifacts will allow others to reproduce and verify results, thereby improving trust and verification of those findings.
 Jon Clarebout famously wrote in 1992 that reproducibility in computationally oriented fields had become so trivial that non experts could do the task.
@@
 
-```{r pyramid, fig.cap = "Strategy for culture and behavior change, taken from the Center for Open Science: https://www.cos.io/blog/strategy-for-culture-change .
-To reach the top, all the lower levels need to be adequately achieved."}
-# pyramid image
-knitr::include_graphics("figures/pyramid.png")
-```
-
 ```{r incentable-table, fig.cap = "Table of the mechanism for behavior change, the incentive, and our assessment of where the elements of Data, Code, and Computational Environment rank in terms of completing these aspects.
@@
 This is in contrast to Code having no policy."}
-knitr::include_graphics("figures/gt-incentive-table.png")
+knitr::include_graphics("figures/incentives_pyramid.png")
 ```
@@
 In the context of datasets, READMEs are particularly useful when there are no reliable standards.
-For example, the data from Dryad, https://datadryad.org/stash/dataset/doi:10.5061/dryad.rn6k1 could be greatly enhanced by providing a README file that explains what the variables in the data collected are, and what the different sheets of the excel worksheet contain.
+Although data repositories contain a large number of datasets associated with peer reviewed papers, not all are usable (example: https://datadryad.org/stash/dataset/doi:10.5061/dryad.rn6k1).
+Some of the more recent submissions contain usage notes and a README, which considerably improve reuse potential (e.g.
+https://datadryad.org/stash/dataset/doi:10.5061/dryad.3xsj3txbz).
+\
 
@@
 Provide a license with your data.
-A license for data clearly establishes the rules for its use, and whether it can be modified or shared.
-Without a license, or with a non-standard license, these rules are unclear, and lead to issues with use, attribution and citation.
-Use the CC0 license, as it provides simple rules for attribution, warranty, and use, and is very common.
-The CC0 is a "public domain" license, meaning data owners waive all rights, and the work is now "owned" by the public.
-CC0 data can be freely shared, copied, modified, and distributed, even for commercial purposes **without asking permission**.
-When using data with CC0, it is good practice to cite the original work, but is not required.
+A license for data establishes the rules for its use, and whether it can be modified or shared.
+Without a license, or with a non-standard one, these rules are unclear, and can lead to issues with use, attribution and citation.
+While there are a handful of licenses for data [@data_world_inc_common_2019], we recommend the CC0 license, as it provides simple rules for attribution, warranty, and use, and is very common.
+The CC0 allows data owners to waive all rights and data licensed as such can be freely shared, copied, modified, and distributed, even for commercial purposes without permission.
 For more information on the benefit of using CC0, and why CCBY is not always appropriate, see https://osc.universityofcalifornia.edu/2016/09/cc-by-and-data-not-always-a-good-fit/.
 Short and full overviews of CCBY and CC0 can be found at [@cc0-short; @cc0-long].
-
-Avoid using non-standard licenses, such as those custom written by the authors, as some terms can render them unusable.
-Once a license is chosen, a LICENSE file should be provided, containing the entire license in the top level of the directory.
-The license type should be mentioned in the README, if used.
 
@@
 Data used in analysis should be provided, especially if computationally intense to derive.
-It should be in its own folder, and to maximise interoperability, recorded in a plain text (e.g., csv, tsv) and if possible a binary format (e.g., .rds, .sav, .xlsx, Apache Parquet).
+It should be in its own folder, and to maximize interoperability, recorded in a plain text (e.g., csv, tsv) and if possible a binary format (e.g., .rds, .sav, .xlsx, Apache Parquet).
 The binary format allows for immediate reuse in the intended state, while the interoperability of plain text versions remain stable over time.
@@
 Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file.
-Authors should run the scripts in a clean environment to ensure everything runs and that the code doesn't rely on other local artefacts, such as their own folder layout not available to others.
+Authors should run the scripts in a clean environment to ensure everything runs and that the code doesn't rely on other local artifacts, such as their own folder layout not available to others.
 
@@
 
-Describe how you want your data to be cited in a CITATION file.
-Citation is important for provenance, and credit.
-Citing data clarifies the data version used (provenance), and provides authors a citation (credit).
-To ensure these are provided, data needs to have an accession number, such as a Digital Object Identifier (DOI).
-A DOI is a permanent, unique identifier for a digital object such as a paper, poster, or software.
-Access numbers are a prerequisite for citation - if they are unavailable, a citation will be meaningless, as it cannot be tracked by any means.
-A file named CITATION should be placed in the directory at the top level.
-It should contain an accession number, and could be in `.bibtex` format.
-Avenues that provide DOIs are discussed in Section \@ref(avenues).
-
-Repositories such as Dryad will manage citation.
-However it is worth noting that a CITATION file allows you to customise how you want your data to be cited.
-For example, to cite the paper that used or described the data and collection methods.
+One critical advantage of depositing data in a trusted repository \@ref(avenues) is that it comes with many benefits such as preservation, indexing,  discoverability and most importantly, citation with persistent identifier.
+Authors can take an extra step by including their own CITATION file to give credit to a paper where the data and associated methods are described in detail.
 
@@
 We'd also like to thank those who provided feedback on the paper when it went up for presubmission: Adrian Barnett, Raymond B Huey, Christian Thurn, Christina Maimone, and GitHub user, 'dschief001'.
-Finally, we thank the following organisations for their support ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
+Finally, we thank the following organizations for their support ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
 
@@
 Currently, the role of data in reproducibility is not adequately discussed.
-We believe this is in part because discussions around data are centered around both data quality and data sharing.
+We believe this is in part because discussions around data are centered on both data quality and data sharing.
 However, these are in fact two separate (but related) topics.
@@
 Prior to joining Berkeley, he earned his PhD in Ecology & Evolution from the University of California, Davis.
-Karthik's interests are focused on reproducible research, especially as it applies to global change.
+Karthik's interests are focused on reproducible research, as it applies to global change.
 Much of his recent work focuses on building tools and services around open data and growing diverse data science communities.

### 2020-09-23 new-paper.Rmd
minor changes - removed hanging text on "other bit to say" and minor adjustment to license section

@@
 \begin{singlespace}
+
 Other titles under consideration:
 
-- Data sharing practices to realign incentives and enable positive/meaningful culture change
-
-- Enabling culture change for meaningful data sharing
-
-- Enabling culture change for meaningful data sharing: Aligning incentives and minimally acceptable practices
-
-- Principles of sharing data
-
-- Elements of meaningful data sharing
-
-- Better practices in data sharing to enable culture change
+- Data sharing practices to realign incentives and enable positive/meaningful culture change\\
+- Enabling culture change for meaningful data sharing\\
+- Enabling culture change for meaningful data sharing: Aligning incentives and minimally acceptable practices\\
+- Principles of sharing data\\
+- Elements of meaningful data sharing\\
+- Minimally acceptable practices in data sharing to enable culture change\\
 
@@
 
-**note: other little bit to say**
-
-- "Here are some high level ideas on how to share data, that might obvious but are not often considered"
-
 # Documentation (READMEs) {#readme}
@@
 
-Provide a license with your data.
+Provide a license file with your data.
 A license for data establishes the rules for its use, and whether it can be modified or shared.
@@
 While there are a handful of licenses for data [@data_world_inc_common_2019], we recommend the CC0 license, as it provides simple rules for attribution, warranty, and use, and is very common.
-The CC0 allows data owners to waive all rights and data licensed as such can be freely shared, copied, modified, and distributed, even for commercial purposes without permission.
+The CC0 means data owners waive all rights and data licensed as such can be freely shared, copied, modified, and distributed, without permission or attribution, even for commercial purposes.
 For more information on the benefit of using CC0, and why CCBY is not always appropriate, see https://osc.universityofcalifornia.edu/2016/09/cc-by-and-data-not-always-a-good-fit/.
-Short and full overviews of CCBY and CC0 can be found at [@cc0-short; @cc0-long].
+Short and full overviews of CC0 can be found at @cc0-short and @cc0-long.
 

### 2020-11-30 new-paper.Rmd
Address Miles' comments

@@
 Without a license, or with a non-standard one, these rules are unclear, and can lead to issues with use, attribution and citation.
-While there are a handful of licenses for data [@data_world_inc_common_2019], we recommend the CC0 license, as it provides simple rules for attribution, warranty, and use, and is very common.
-The CC0 means data owners waive all rights and data licensed as such can be freely shared, copied, modified, and distributed, without permission or attribution, even for commercial purposes.
+While there are a handful of licenses for data [@data_world_inc_common_2019], we recommend using the CC0 license, which means data owners waive all rights to the data, so it can be freely shared, copied, modified, and distributed, without permission or attribution, even for commercial purposes.
+
+We recommend the CC0 as it provides simple rules for attribution, warranty, and use, and is very common.
+If data that is CC0 is used in scientific publications, you need to cite where you got the data from.
+There might be some concern that it is possible to conduct fraud, but it will be found out eventually if the original data is archived properly.
+
 For more information on the benefit of using CC0, and why CCBY is not always appropriate, see https://osc.universityofcalifornia.edu/2016/09/cc-by-and-data-not-always-a-good-fit/.
@@
 Authors can take an extra step by including their own CITATION file to give credit to a paper where the data and associated methods are described in detail.
+When citing data, you should cite the preferred CITATION file in the first instance, else, if there is a data publication, (for example in _Scientific Data_), cite this, and if none of those apply, cite the repository where the data is deposited (e.g., a Zenodo repository).
 

### 2020-11-30 new-paper.Rmd
added note about dependencies
Consistent type

--- 2020-09-23 new-paper.Rmd
@@
 In addition to locating software, one still has to install the right versions.
+Dependency management is a controversial topic in programming communities with no easy solutions.
+While languages like JavaScript have solved this with package managers like npm, it is much harder to accomplish in others like Python and R which are currently supported by a patchwork of solutions.
 Although this problem is far from solved, technologies like Docker have alleviated some of these concerns.
@@
 
-Tools to prepare datasets and metadata (*the infrastructure*) and minimally acceptable practices (guidelines) for reuse are sparse [@Arslan2019; @Roche2015].
+Tools to prepare datasets and metadata (*the infrastructure*) and minimally acceptable practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015].
 Unlike code, datasets are far more diverse in terms of heterogeneity, size, and formats.

### 2020-12-01 new-paper.Rmd
minor changes from meeting today with Karthik

--- 2020-11-30 new-paper.Rmd
@@
 ---
-title: Meaningful data sharing
+title: Elements of meaningful data sharing
 author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)
@@
 
+- Meaningful data sharing
 - Data sharing practices to realign incentives and enable positive/meaningful culture change\\
@@
 We also discuss considerations for sharing of data of various sizes (Section \@ref(size)).
+
+We limit the scope of this paper to small to medium sized tabular data as it covers a vast majority of use cases that researchers encounter day to day.
+For larger datasets that involve complex workflows (e.g., bioinformatics and medical imaging), the kinds of challenges they face are different and accordingly deserve to be written up separately.
 

### 2020-12-02 new-paper.Rmd
added paragraph on workflows to address Jeremy's comments

--- 2020-11-30 new-paper.Rmd
@@
 
+For more complex analyses that involve workflow management systems (see https://github.com/pditommaso/awesome-pipeline for a comprehensive list), a description of how to run the workflow should be included in the documentation, preferably in the README file.
+Depending on whether the workflow is GUI driven (e.g.
+Galaxy) or text driven (e.g.
+Nextflow, Snakemake), the workflow itself can be checked into the version control repository.
+As these communities work closely and use Common Workflow Language as a shared representation, these files can be widely used to recreate the steps involved in producing the final results.
+
 # Citation: How you want your data to be cited {#cite}

### 2020-12-02 new-paper.Rmd

--- 2020-12-01 new-paper.Rmd
@@
 Authors should run the scripts in a clean environment to ensure everything runs and that the code doesn't rely on other local artifacts, such as their own folder layout not available to others.
+
+For more complex analyses that involve workflow management systems (see https://github.com/pditommaso/awesome-pipeline for a comprehensive list), a description of how to run the workflow should be included in the documentation, preferably in the README file.
+Depending on whether the workflow is GUI driven (e.g.
+Galaxy) or text driven (e.g.
+Nextflow, Snakemake), the workflow itself can be checked into the version control repository.
+As these communities work closely and use Common Workflow Language as a shared representation, these files can be widely used to recreate the steps involved in producing the final results.
 

### 2021-03-29 new-paper.Rmd
add a note on metadata reminding using rectangular data again.
change no journals to few journals

@@
 With code, we have slowly progressed through these stages.
-While no journals require that authors share code, the venues to share, along with the necessary training have paved the way for incentives and credit.
+While few journals require that authors share code, the venues to share, along with the necessary training have paved the way for incentives and credit.
 All of the necessary foundation to make code sharing required are in place.
@@
 Use metadata with your data.
+As stated at the start of the paper, we refer to tabular data.
 There are three benefits to a dataset having metadata: (1) Provide context by describing variables to humans; (2) Avoid errors by describing and preserving variable types for machine readability; (3) Make data easier to find online by facilitating indexing [@google-data-search].

### 2021-03-29 new-paper.Rmd
re-add comment on privact concerns and the benefits of data sharing.

@@
 For larger datasets that involve complex workflows (e.g., bioinformatics and medical imaging), the kinds of challenges they face are different and accordingly deserve to be written up separately.
+
+Assuming that the data have no privacy concerns (e.g., human participants, locations of critically endangered species), or that the act of sharing does not put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
 

### 2021-03-29 new-paper.Rmd
Made a handful of revisions

@@
 To make data more reusable, we need to make it easier to generate metadata, standardise good practices within and across communities, and make reusable data sharing a recognized, normalised, part of research.
-Here we present a set of minimally acceptable practices to enable this change.
+Here we present a set of minimally acceptable practices to facilitate this change.
 These include documentation (Section \@ref(readme)), licensing (Section \@ref(licenses)), citation (Section \@ref(cite)), raw and analysis ready data (Section \@ref(rawdata)), code (Section \@ref(scripts)), and sharing in packages (Section \@ref(pkgs)).
@@
 Partner with Dryad [@dryad] to link archived data with research papers.
-Dryad will also provide a person to help curate your data, improving usability.
+Dryad will also provide professionals to curate your data, thereby improving usability.
 
@@
 First, archive data in a long-term data repository (Section \@ref(avenues)), before including it in a package.
-This makes the data easily citable wherever data is referenced.
+The act of packaging data in software can shave off hours required to prepare it for use.
+The canocical citation can be baked into the package's citation file.
 
@@
 Although data citations do not currently impact promotion and tenure packets, there is anecdotal evidence that papers sharing usable data accrue more citations [@Piwowar2007].
-Over time, the more useful we make the data, the more likely incentive mechanisms will take hold.
+Over time, the more useful we make the data, it is more likely incentive mechanisms will take hold.
 
 The steps leading up to requiring data sharing are still vague and disincentivized.
-In comparison, although few journals require that code be submitted with publications, it is more likely to be adopted given the foundation we have put in place.
+In comparison, although few journals require that code be submitted with publications, it is more likely to be adopted given the foundation we have established with carrots rather than sticks.
 We have taught researchers how to write modular code via popular training initiatives such as The Carpentries.
-Commercial tools like GitHub have made code sharing fun and productive.
-For more advanced researchers, data science languages have made it easier to package and disseminate code as software.
+Commercial services like GitHub have made code sharing fun and productive.
+For more advanced researchers, data science languages have made it easier to package and disseminate code as reusable software.
 If we now require mandatory code sharing, it is quite likely that more of it will be useful than the data currently being shared.
@@
 To make data sharing truly useful, we need to make the process easier, improve tools and training, then reward and incentivize researchers who engage in such practices.
-Only then can data truly be useful.
+Only then can shared data truly be useful.
 
@@
 We'd also like to thank those who provided feedback on the paper when it went up for presubmission: Adrian Barnett, Raymond B Huey, Christian Thurn, Christina Maimone, and GitHub user, 'dschief001'.
-Finally, we thank the following organizations for their support ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Sloan Foundation.
+Finally, we thank the following organizations for their support ACEMS International Mobility Programme, the Helmsley Charitable Trust, the Gordon and Betty Moore Foundation, and the Alfred P.
+Sloan Foundation.
 

### 2021-03-30 new-paper.Rmd
Rewrote abstract

@@
 \begin{singlespace}
+New abstract: There are numerous arguments that support the practice of open science with societal and individual benefits.
+For individual researchers, sharing research artifacts such as data can improve trust and transparency, reproducibility of one's own work, and catalyze new collaborations.
+Despite the general appreciation of the benefits of data sharing, research data are available only to the original investigators.
+For data that are shared, lack of metadata and documentation make them challenging to reuse.
+In this commentary we argue that incentives and infrastructure for making data useful are the biggest barrier to creating a culture of widespread data sharing.
+We compare data to code, and computational environments in the context of reproducible research and provide some practical guidance on how one can improve the chances of their data being reusable.
 
-Data makes science possible.
+Old abstract: Data makes science possible.
 Sharing data improves visibility, and makes the research process transparent.
@@
 However, a large proportion of data from published research is often only available to the original authors.
-Despite the obvious benefits of sharing data, and scientists' advocating for the importance of sharing data, most advice on sharing data discusses its broader benefits, rather than the practical considerations of sharing.
+Despite the obvious benefits of make data widely available, and scientists' advocating for the importance of sharing data, most advice on sharing data discusses its broader benefits, rather than the practical considerations of sharing.
 This paper provides practical, actionable advice on how to actually share data alongside research with minimal barriers.

### 2021-03-31 new-paper.Rmd
More updates

@@
 ---
-title: Elements of meaningful data sharing
+title: Barriers to meaningful data sharing
 author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)
@@
 \begin{singlespace}
-New abstract: There are numerous arguments that support the practice of open science with societal and individual benefits.
+New abstract: There are numerous arguments that strongly support practice of open science, with societal and individual benefits.
 For individual researchers, sharing research artifacts such as data can improve trust and transparency, reproducibility of one's own work, and catalyze new collaborations.
 Despite the general appreciation of the benefits of data sharing, research data are available only to the original investigators.
-For data that are shared, lack of metadata and documentation make them challenging to reuse.
+For data that are shared, lack of useful metadata and documentation make them challenging to reuse.
 In this commentary we argue that incentives and infrastructure for making data useful are the biggest barrier to creating a culture of widespread data sharing.
-We compare data to code, and computational environments in the context of reproducible research and provide some practical guidance on how one can improve the chances of their data being reusable.
+We compare data to code, and computational environments in the context of reproducible research and provide some practical guidance on how one can improve the chances of their data being reusable and partially bridge the incentive gap.
 
@@
 
-Meaningful progress on computational reproducibility, the ability of an independent researcher to recreate figures, tables and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence.
-Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by organizations like The Carpentries, combined with the growing adoption of GitHub by scientists [@Ram2013; @Zagalsky2015].
-The bigger driver for this is more likely driven by the growing popularity of data science as a discipline distinct from statistics [@Donoho2017].
+Meaningful progress on most computational reproducibility, the ability of an independent researcher to recreate figures, tables and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence.
+Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by efforts like The Carpentries, combined with the growing adoption of GitHub by scientists [@Ram2013; @Zagalsky2015].
+The bigger driver for this is more likely due to the growing popularity of data science as a discipline distinct from statistics [@Donoho2017].
 This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools.
@@
 Modern scientists now have access to many interactive computational frameworks (e.g.
-Jupyter, RMarkdown, Stencila, Observable) that can interact with kernels from dozens of programming languages.
+Jupyter, RMarkdown, Stencila, Observable) that can interface with kernels from dozens of programming languages.
 Beyond notebooks to run the code, a reader will also require access to the same software dependencies.
 In a study of applied computational research, [@Collberg2014] and colleagues were only able to install and run less than 20% of the software described in publications.
-In addition to locating software, one still has to install the right versions.
+In addition to locating software, one still has to locate the right versions.
 Dependency management is a controversial topic in programming communities with no easy solutions.
@@
 Researchers can share Dockerfiles to recreate computational environments [@Boettiger2015].
-Services like Binder can then leverage Docker and make possible to share the computational environments in which these notebooks were run.
+Services like Binder can then leverage Docker and make it possible to share the computational environments in which these notebooks were run.
 The barriers to verification are now low enough that trivial reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].
@@
 Beyond motivation, skills, tools, and incentives are critical to enable such change.
-In a five step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, then improve the infrastructure to make it easy, then increase awareness to make it normative.
+In a five step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, then improve the infrastructure to make it easy, followed by increased awareness to make it normative.
 Once there is a critical mass, find ways to reward those who put it into practice.
-After these steps, it would then be possible to require this of everyone Figure 1, table 1).
+After these steps, it would then be possible to require this of everyone Figure 1).
 This framework provides a useful lens through which to examine adoption of code, computational environments and data for reproducibility.
@@
 But running them still requires expertise and time on the part of reviewers.
-More recent open services like Binder [@Jupyter2018] make it possible to run notebooks atop Docker images, but the technology is still new and slowly gaining adoption.
+More recent open services like Binder [@Jupyter2018] make it possible to run notebooks atop Docker images, but the technology is still new and experiencing growing pains.
 There is still a long way to go before sharing computational environments in academia becomes mainstream.
@@
 Assuming that the data have no privacy concerns (e.g., human participants, locations of critically endangered species), or that the act of sharing does not put the authors at a competitive disadvantage (data can be embargoed for reasonable periods of time), sharing data will always have a net positive benefit.
+
+Our advice here is not meant to be exhaustive by any means and are insufficient to fully bridge the gap, but provide a starting point.
 
@@
 Although data sharing is widely mandated by funders, journals, and academic societies, the amount of usable data remains appallingly low.
-The mere act of depositing data does not make it reusable or useful to a potential reader.
+The mere act of depositing data does not make it reusable or useful to a potential consumer.
 Although data citations do not currently impact promotion and tenure packets, there is anecdotal evidence that papers sharing usable data accrue more citations [@Piwowar2007].

### 2021-04-01 new-paper.Rmd
update aff

@@
 title: Barriers to meaningful data sharing
-author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)
+author: Nicholas J Tierney\* (1,2,3) & Karthik Ram\* (4)
 output:
@@
 2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)\\
-3 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
+3 = Telethon Kids Institute, Perth, Australia\\
+4 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
 * = Both authors contributed equally to the work\\

### 2021-04-01 new-paper.Rmd
change minimally acceptable practices to minimal set of practices

--- 2021-03-31 new-paper.Rmd
@@
 
-Tools to prepare datasets and metadata (*the infrastructure*) and minimally acceptable practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015].
+Tools to prepare datasets and metadata (*the infrastructure*) and minimal set of practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015].
 Unlike code, datasets are far more diverse in terms of heterogeneity, size, and formats.
@@
 To make data more reusable, we need to make it easier to generate metadata, standardise good practices within and across communities, and make reusable data sharing a recognized, normalised, part of research.
-Here we present a set of minimally acceptable practices to facilitate this change.
+Here we present a set of minimal set of practices to facilitate this change.
 These include documentation (Section \@ref(readme)), licensing (Section \@ref(licenses)), citation (Section \@ref(cite)), raw and analysis ready data (Section \@ref(rawdata)), code (Section \@ref(scripts)), and sharing in packages (Section \@ref(pkgs)).

### 2021-04-01 new-paper.Rmd
remove references

@@
 This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools.
-Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data [@rcore; @python; @julia].
+Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data.
 Each of these languages enjoys support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
@@
 For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) Code; 2) Computing environment, and 3) Data.
-There has been tremendous progress on the first two, despite some limitations of existing tools and practices [@Poisot2019].
+There has been tremendous progress on the first two, despite some limitations of existing tools and practices.
 While it took nearly two decades for Clarebout's vision to become reality, improvements in technology and motivations played a strong role in the culture change.
 Sharing code wasn't just easy at first [@Barnes2010], but became a social and normative part of academic culture.
-The next stage of this evolution, turning code into software, also became popular with increasing training opportunities [@Wilson2014; @Teal2015], tooling [@devtools], and recognition as a scholarly contribution [@Smith2018].
+The next stage of this evolution, turning code into software, also became popular with increasing training opportunities [@Wilson2014], tooling, and recognition as a scholarly contribution [@Smith2018].
 Getting researchers to share usable data has been more of a struggle.
@@
 
-Nosek [-@Nosek2019] describes the strategies for culture change, especially when the status quo is too entrenched.
+Nosek describes the strategies for culture change, especially when the status quo is too entrenched.
 Beyond motivation, skills, tools, and incentives are critical to enable such change.
@@
 
-Despite the increase in data being shared alongside publications [@Culina2020], the status quo of data sharing in science is not good.
+Despite the increase in data being shared alongside publications, the status quo of data sharing in science is not good.
 [@Rowhani-Farid2016] showed that 7 of 157 (4.5%) sampled papers in the British Medical Journal shared data, and only 14% of articles claiming open sharing of data actually shared their data.
@@
 There are numerous examples of similar problems in other fields, such as psychology, biology, and bioinformatics.
-The reality is that while data sharing is increasing in many fields such as ecology, the amount of data that can be reused is likely lower [@Culina2020] .
+The reality is that while data sharing is increasing in many fields such as ecology, the amount of data that can be reused is likely lower.
 To make shared data useful, we need to realign incentives for researchers.
@@
 This makes them particularly challenging to standardize or easily "install" where the code is running.
-There are some efforts working towards improving this process, such as gogetdata , the 'dat' project, and a private company, Quilt [@goget; @dat; @quilt].
+There are some efforts working towards improving this process, such as gogetdata , the 'dat' project, and a private company, Quilt.
 
@@
 Use a README file with your data project to communicate its purpose and contents.
-A README is often the first place users will look to learn more about a project [@Prana2019].
-The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/> and [@Prana2019].
+A README is often the first place users will look to learn more about a project.
+The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/> and.
 For example, a README in a data repository might explain which directories contain data, tidying scripts, and also why data was collected.
@@
 Without a license, or with a non-standard one, these rules are unclear, and can lead to issues with use, attribution and citation.
-While there are a handful of licenses for data [@data_world_inc_common_2019], we recommend using the CC0 license, which means data owners waive all rights to the data, so it can be freely shared, copied, modified, and distributed, without permission or attribution, even for commercial purposes.
+While there are a handful of licenses for data, we recommend using the CC0 license, which means data owners waive all rights to the data, so it can be freely shared, copied, modified, and distributed, without permission or attribution, even for commercial purposes.
 
@@
 For more information on the benefit of using CC0, and why CCBY is not always appropriate, see https://osc.universityofcalifornia.edu/2016/09/cc-by-and-data-not-always-a-good-fit/.
-Short and full overviews of CC0 can be found at @cc0-short and @cc0-long.
+Short and full overviews of CC0 can be found at https://creativecommons.org/.
 
@@
 Metadata can be encoded into plain text formats that follow a set standard, such as EML or SCHEMA-LD.
-A deep understanding of these formats isn't required to use them, for more details, and guidance on generating them see @dataspice and @Arslan2019.
+A deep understanding of these formats isn't required to use them, for more details, and guidance on generating them see @Arslan2019.
 
@@
 Data should be deposited into a relevant domain repository that generates an accession number such as a DOI.
-For example, gbif [@gbif] for species data, or genbank for genetics data [@GenBank].
+For example, gbif for species data, or genbank for genetics data.
 Information on other domain specific repositories can be found at https://www.nature.com/sdata/policies/repositories.
@@
 It was not possible to easily access or cite the data without an explicit authorship agreement among other conditions.
-Only recently were the complete dataset deposited in Dryad without such conditions [@condit-2019].
+Only recently were the complete dataset deposited in Dryad without such conditions.
 
@@
 Writing a data paper can be very useful for researchers, however, data papers lack guidelines on how to structure data for effective reuse and sharing.
-Partner with Dryad [@dryad] to link archived data with research papers.
+Partner with Dryad to link archived data with research papers.
 Dryad will also provide professionals to curate your data, thereby improving usability.
@@
 
-Data can be distributed alongside code using a format such as a package in R, Python, and Julia [@nycflights; @horst-penguins; @data-python-pkg; @r-data-julia].
+Data can be distributed alongside code using a format such as a package in R, Python, and Julia.
 A package or module structure provides an advantage as it can be "installed" and loaded just like code.
@@
 Users who work in other languages are unlikely to download and export data out of a package.
-Secondly, data size can be a limitation depending on the repository (CRAN, the R repository, for example, limits package sizes to 5Mb although there are workarounds [@Brooke_Anderson2017]).
+Secondly, data size can be a limitation depending on the repository (CRAN, the R repository, for example, limits package sizes to 5Mb although there are workarounds).
 Other fields, such as bioinformatics have a process of sharing data through repositories like Bioconductor, with larger data size.

### 2021-04-01 new-paper.Rmd
update affiliations

--- 2021-04-01 new-paper.Rmd
@@
 2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)\\
-3 = Telethon Kids Institute, Perth, Australia\\
+3 = Telethon Kids Institute, Perth Children's Hospital, Perth, Australia\\
 4 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
@@
 
-We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Anna Krystalli, Daniella Lowenberg, and Miles McBain.
+We'd like to thank the following people for helpful discussions and contributions to the paper as we were preparing it for publication: Rich FitzJohn, Anna Krystalli, Daniella Lowenberg, and Miles McBain.
 We'd also like to thank those who provided feedback on the paper when it went up for presubmission: Adrian Barnett, Raymond B Huey, Christian Thurn, Christina Maimone, and GitHub user, 'dschief001'.
@@
 Nicholas Tierney has an honours degree in Psychological Science, and a PhD in Statistics.
-He is a Lecturer at Monash University, teaching introduction to data analysis.
-His ranges from exploratory data analysis, to statistical modelling, calculating diagnostics, and how we draw inferences, and make decisions from data.
+He is a research software engineer at Telethon Kids Institute in Perth, Australia, working on maintaining and extending statistical software and establishing research software pipelines.
+His research ranges from exploratory data analysis, to statistical modelling, calculating diagnostics, and how we draw inferences, and make decisions from data.
 Crucial to this work is producing high quality software to accompany each research idea.
-Nick is a keen outdoorsman, and loves hiking, trail running, rock climbing, photography, and making coffee.
+Nick is a keen outdoorsman and photographer.
 

### 2021-04-01 new-paper.Rmd

--- 2021-04-01 new-paper.Rmd
@@
 2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)\\
-3 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
+3 = Telethon Kids Institute, Perth Children's Hospital, Perth, Australia\\
+4 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
 * = Both authors contributed equally to the work\\

### 2021-04-01 new-paper.Rmd
remove old abstract and add new possible titles.

@@
 - Enabling culture change for meaningful data sharing\\
-- Enabling culture change for meaningful data sharing: Aligning incentives and minimally acceptable practices\\
 - Principles of sharing data\\
 - Elements of meaningful data sharing\\
-- Minimally acceptable practices in data sharing to enable culture change\\
+- The biggest barrier to widespread data sharing are incentives and infrastructure.\\
+- Some practical guidance on improving data reuse.\\
 
@@
 \begin{singlespace}
-New abstract: There are numerous arguments that strongly support practice of open science, with societal and individual benefits.
+There are numerous arguments that strongly support practice of open science, with societal and individual benefits.
 For individual researchers, sharing research artifacts such as data can improve trust and transparency, reproducibility of one's own work, and catalyze new collaborations.
@@
 We compare data to code, and computational environments in the context of reproducible research and provide some practical guidance on how one can improve the chances of their data being reusable and partially bridge the incentive gap.
-
-Old abstract: Data makes science possible.
-Sharing data improves visibility, and makes the research process transparent.
-This increases trust in the work, and allows for independent reproduction of results.
-However, a large proportion of data from published research is often only available to the original authors.
-Despite the obvious benefits of make data widely available, and scientists' advocating for the importance of sharing data, most advice on sharing data discusses its broader benefits, rather than the practical considerations of sharing.
-This paper provides practical, actionable advice on how to actually share data alongside research with minimal barriers.
 

### 2021-04-01 new-paper.Rmd
update title

@@
 ---
-title: Barriers to meaningful data sharing
+title: Incentives are the biggest barriers to widespread data sharing.
 author: Nicholas J Tierney\* (1,2,3) & Karthik Ram\* (4)
@@
 \begin{singlespace}
-
-Other titles under consideration:
-
-- Meaningful data sharing
-- Data sharing practices to realign incentives and enable positive/meaningful culture change\\
-- Enabling culture change for meaningful data sharing\\
-- Principles of sharing data\\
-- Elements of meaningful data sharing\\
-- The biggest barrier to widespread data sharing are incentives and infrastructure.\\
-- Some practical guidance on improving data reuse.\\
-
 1 = Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia\\
@@
 ```
+
+<!-- Other titles under consideration: -->
+<!-- Barriers to meaningful data sharing -->
+<!-- - Meaningful data sharing -->
+<!-- - Data sharing practices to realign incentives and enable positive/meaningful culture change\\ -->
+<!-- - Enabling culture change for meaningful data sharing\\ -->
+<!-- - Principles of sharing data\\ -->
+<!-- - Elements of meaningful data sharing\\ -->
+<!-- - The biggest barrier to widespread data sharing are incentives and infrastructure.\\ -->
+<!-- - Some practical guidance on improving data reuse.\\ -->
+
 # Abstract {.unnumbered}

### 2021-04-09 new-paper.Rmd
add author contributions and competing interests

@@
 
+# Author Contributions
+
+Both authors contributed equally to the work, making substantial contributions to the conception, design, drafting and revision of the work.
+The authors have approved the submitted version, and have agreed both to be personally accountable for the author's own contributions and to ensure that questions related to the accuracy or integrity of any part of the work, even ones in which the author was not personally involved, are appropriately investigated, resolved, and the resolution documented in the literature.
+
+# Competing Interests
+
+The authors declare no competing interests.
+
 # Literature Cited {.unnumbered}

### 2021-04-09 new-paper.Rmd
remove bigger picture statement and bio. change literature cited to refences

@@
 
-# Literature Cited {.unnumbered}
+# References {.unnumbered}
 
@@
 :::
-
-\newpage
-
-# Bigger Picture Statement {.unnumbered}
-
-Without data, there is no science.
-Science needs to be reproducible so we can trust the results, and progress as a field.
-But without data, we do not have true reproducibility.
-Currently, the role of data in reproducibility is not adequately discussed.
-We believe this is in part because discussions around data are centered on both data quality and data sharing.
-However, these are in fact two separate (but related) topics.
-We provide immediately usable guidelines to sharing data alongside your research, with minimal barriers, that a researcher could pick up and use today.
-
-In the future, we hope that the culture around data sharing will change, and sharing data will be rewarded in science.
-We want the new normal to be an expectation of data to be submitted with every research publication, and for these datasets to be easily discovered, shared, and extended in other analysis.
-
-# Bio {.unnumbered}
-
-**Nicholas Tierney**
-
-Nicholas Tierney has an honours degree in Psychological Science, and a PhD in Statistics.
-He is a research software engineer at Telethon Kids Institute in Perth, Australia, working on maintaining and extending statistical software and establishing research software pipelines.
-His research ranges from exploratory data analysis, to statistical modelling, calculating diagnostics, and how we draw inferences, and make decisions from data.
-Crucial to this work is producing high quality software to accompany each research idea.
-Nick is a keen outdoorsman and photographer.
-
-**Karthik Ram**
-
-Karthik is a senior data scientist at the Berkeley Institute of Data Science and a co-founder of the rOpenSci project.
-Karthik is also a senior PI at the Berkeley Initiative for Global Change Biology.
-Prior to joining Berkeley, he earned his PhD in Ecology & Evolution from the University of California, Davis.
-Karthik's interests are focused on reproducible research, as it applies to global change.
-Much of his recent work focuses on building tools and services around open data and growing diverse data science communities.

### 2021-04-09 new-paper.Rmd
remove section on Hubbell

@@
 ---
-title: Incentives are the biggest barriers to widespread data sharing.
+title: Incentives are the biggest barriers to widespread data sharing
 author: Nicholas J Tierney\* (1,2,3) & Karthik Ram\* (4)
@@
 Do not deposit data in local file storage that does not provide DOI or accession numbers.
-For example, [@Hubbell1999] was hosted on a university server and only accessible via a FTP server since 1982.
-It was not possible to easily access or cite the data without an explicit authorship agreement among other conditions.
-Only recently were the complete dataset deposited in Dryad without such conditions.
 
@@
 
-# Author Contributions
+# Author Contributions {.unnumbered}
 
@@
 
-# Competing Interests
+# Competing Interests {.unnumbered}
 
@@
 
-::: {#refs}
-:::
+\singlespacing

### 2021-04-09 figures/carbon/create-carbon-dir-structure.R
remove unused figures

@@ -1,27 +1,0 @@
-library(carbonate)
-egg <- 
-  'proj-name
-├── README.md
-├── LICENSE
-├── CITATION
-├── metadata
-│   ├── access.csv
-│   ├── attributed.csv
-│   ├── biblio.csv
-│   └── creators.csv
-├── raw-data
-│   ├── raw-data.csv
-│   └── 01-read-tidy-raw.R
-└── data
-    └── final-data.csv
-'
-
-egg
-
-x <- carbon$new(readr::read_lines(egg))
-x$language <- "Plain Text"
-x$template <- "solarized light"
-x$add_window_control <- FALSE
-x$add_drop_shadow <- FALSE
-x$add_line_number <- TRUE
-x$browse()

### 2021-04-09 figures/carbon/create-carbon-json-ld.R

@@ -1,37 +1,0 @@
-library(carbonate)
-json_ld <- '{
-  "name": "demographics",
-  # here we list the data files in this dataset
-  "resources": [
-    {
-      "path": "demographics.csv",
-      "schema": {
-        "fields": [
-          {
-            "name": "age",
-            "type": "integer"
-          },
-          {
-            "name": "height",
-            "type": "number"
-          },
-          {
-            "name": "nationality",
-            "type": "string"
-          }
-          ]
-      }
-    }
-    ]
-}
-'
-
-json_ld
-
-x <- carbon$new(readr::read_lines(json_ld))
-x$language <- "JSON"
-x$template <- "solarized light"
-x$add_window_control <- FALSE
-x$add_drop_shadow <- FALSE
-x$add_line_number <- TRUE
-x$browse()

### 2021-04-09 figures/create-incentives-table.R

@@ -1,27 +1,0 @@
-library(gt)
-incentive_table <- tibble(
-  Mechanism = c("Policy", "Incentives", "Communities", "UI/UX", "Infrastructure"),
-  Incentive = c("Required", "Rewarding", "Normative", "Easy", "Possible"),
-  Data = c(100, 0, 0, 0, 100),
-  Code = c(0, rep(100, 4)),
-  `Computational Environment` = c(0, 0, 0, 50, 100)
-)
-
-incentive_table %>%
-  gt() %>%
-  data_color(
-    columns = vars(Data, Code, `Computational Environment`),
-    colors = scales::col_numeric(
-      # Using a function from paletteer to generate a vector of colors
-      # Note that you need to wrap paletteer_d outputs in as.character()
-      palette = colorspace::sequential_hcl(n = 3, palette = "Greens2", rev = TRUE),
-      # Domain is from scales::col_numeric
-      # Domain = The possible values that can be mapped
-      # We don't HAVE to set a range for this since
-      # we're using it inside data_color()
-      domain = NULL
-    )
-  ) %>%
-  gtsave("figures/gt-incentive-table.png")
-
-# knitr::include_graphics("figures/gt-incentive-table.png")

### 2021-08-02 new-paper.Rmd
minor change of sentence to "In a study of applied computational research, less than 20% of software described in publications was able to be installed and run [@Collberg2014]"

@@
 bibliography: references.bib
+citation_package: 'natbib'
 csl: nature-communications.csl
@@
 Beyond notebooks to run the code, a reader will also require access to the same software dependencies.
-In a study of applied computational research, [@Collberg2014] and colleagues were only able to install and run less than 20% of the software described in publications.
+In a study of applied computational research, less than 20% of software described in publications was able to be installed and run [@Collberg2014].
 In addition to locating software, one still has to locate the right versions.

### 2021-08-02 new-paper.Rmd
minor modification to abstract as suggested by reviewer
change from "trivial reproducibility" to "reproducibility"

@@
 For individual researchers, sharing research artifacts such as data can improve trust and transparency, reproducibility of one's own work, and catalyze new collaborations.
-Despite the general appreciation of the benefits of data sharing, research data are available only to the original investigators.
+Despite the general appreciation of the benefits of data sharing, research data are often only available to original investigators.
 For data that are shared, lack of useful metadata and documentation make them challenging to reuse.
@@
 Services like Binder can then leverage Docker and make it possible to share the computational environments in which these notebooks were run.
-The barriers to verification are now low enough that trivial reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].
+The barriers to verification are now low enough that reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].
 

### 2021-08-04 new-paper.Rmd
Updated to make two sections - one for introduction, one for "Improving the state of resuable data sharing"

@@
 
-## Improving the state of reusable data sharing
+# Improving the state of reusable data sharing
 
@@
 
-# Documentation (READMEs) {#readme}
+## Documentation (READMEs) {#readme}
 
@@
 
-# Licences {#licenses}
+## Licences {#licenses}
 
@@
 
-# Metadata {#metadata}
+## Metadata {#metadata}
 
@@
 
-# Raw data vs analysis ready data {#rawdata}
+## Raw data vs analysis ready data {#rawdata}
 
@@
 
-# Data cleaning scripts {#scripts}
+## Data cleaning scripts {#scripts}
 
@@
 
-# Citation: How you want your data to be cited {#cite}
+## Citation: How you want your data to be cited {#cite}
 
@@
 
-# Avenues to share data {#avenues}
+## Avenues to share data {#avenues}
 
@@
 
-# Data as part of open source packages {#pkgs}
+## Data as part of open source packages {#pkgs}
 
@@
 
-# Considerations for data size {#size}
+## Considerations for data size {#size}
 

### 2021-08-04 new-paper.Rmd
added some further clarity on audience and how our work is different to the other papers

@@
 
+We believe a key audience could be better served by the current literature on data sharing: researchers who have finished data analysis on tabular data, are ready to submit for publication.
+Many papers provide rules and guidelines for sharing data, such as providing metadata, citation, licenses, and lists of data repositories to use.
+These papers provide general advice from the ethical considerations of sharing data [@Moles2013], how to share many different data types (images, video, etc) [@Hart2016], to “Reward Colleagues Who Share Their Data Properly” [@Goodman2014].
+There is also specific advice like the importance of very clean and properly shaped data [@White2013], and to use open source software with changes tracked [@Boland2017].
+These papers do not discuss the approach of sharing data within software, which has seen a rise in the use of research compendiums (e.g, [@Marwick2018]).
+Across many of these articles there is common advice that is useful, but individually, their advice is too general or specific for our audience.
+
 As academic data scientists, we spend considerable time cleaning messy data and believe that preparing data for easy reuse can be a catalyst to improve the quality of data and pave the way for other incentives such as community adoptions, which will then create the culture change needed to credit data similar to publications.
-In this note, we describe a set of practices that researchers can adopt to make data more useful in the context of reproducibility.
+In this paper, we describe a set of common sense practices that researchers can adopt to make data more useful in the context of reproducibility.
 It should not be a huge burden to share data in a format that is readily usable by others.

### 2021-08-04 new-paper.Rmd
update abstract with information on target audience.

@@
 For data that are shared, lack of useful metadata and documentation make them challenging to reuse.
-In this commentary we argue that incentives and infrastructure for making data useful are the biggest barrier to creating a culture of widespread data sharing.
+In this paper we argue that incentives and infrastructure for making data useful are the biggest barrier to creating a culture of widespread data sharing.
 We compare data to code, and computational environments in the context of reproducible research and provide some practical guidance on how one can improve the chances of their data being reusable and partially bridge the incentive gap.
+Other papers have focussed on many general rules for sharing data.
+This paper focusses on common sense ideas for sharing tabular data, for a target audience of academics working in data science adjacent fields who are about to submit for publication.
 

### 2021-08-04 new-paper.Rmd
add a bit of extra detail into the metadata section

@@
 **Unit text** defines the type of variable recorded (e.g., string, date, numeric, integer), this avoids serious errors of misrepresented data (e.g., a column of a gene sequence being interpreted as a date).
+A simple metadata file that describes the variable names, variable description, and unit text is a great starting point, and is sometimes known as a codebook.
 For more details, see [@Broman2017; @Ellis2017; @Arslan2019].
-Metadata can be encoded into plain text formats that follow a set standard, such as EML or SCHEMA-LD.
+Metadata can be encoded into more formal plain text formats that follow a set standard, such as EML or SCHEMA-LD.
 A deep understanding of these formats isn't required to use them, for more details, and guidance on generating them see @Arslan2019.

### 2021-08-10 new-paper.Rmd
remove sentence: "just archive the data".
Addresses R1-4

@@
 For example, gbif for species data, or genbank for genetics data.
-Information on other domain specific repositories can be found at https://www.nature.com/sdata/policies/repositories.
-Data not falling into domain repositories can be submitted to research data repositories **zenodo** and **dryad**.
+A exhaustive searchable list is available at re3data (https://www.re3data.org/) but provides no guidance to researchers.
+Data that do not fit appropriate domain repositories can be submitted to research data repositories **Zenodo** and **Dryad** which are currently leading the way in best practices for software and data.
 In addition to minting DOIs, they also provide citation templates.
@@
 While this approach is useful while iterating on the paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc.).
-Just archive the data.
 Providing back up data is particularly relevant in light of the case in [@Schleunes2020; @Viglione2020], where data was suspected of being tampered.

### 2021-08-11 new-paper.Rmd
Updated reference to say data citations are still new
Addressed comment R1-8

@@
 
-Nosek describes the strategies for culture change, especially when the status quo is too entrenched.
+Nosek [@nosek_2021] describes the strategies for culture change, especially when the status quo is too entrenched.
 Beyond motivation, skills, tools, and incentives are critical to enable such change.
@@
 Lastly, the infrastructure to track and get credit for data is non-existent.
-Datasets currently do not accrue citation credit since they are not tracked.
+Datasets currently do not accrue citation credit since they are not tracked [@lowenberg_2019].
 

### 2021-08-11 new-paper.Rmd
R-23
remove paragraph on how we don't have recommendations for non tabular data

@@
 
-We don't have particular recommendations for non tabular of data, like binary images, but given the ubiquity of tables in papers, these recommendations will be broadly useful.
-For other data formats use a copy of the most robust common format and avoid the latest tools that may be more brittle and likely to change in the future.
-For example, for an image, a TIFF image in addition to a custom photoshop file.
-
 ## Data cleaning scripts {#scripts}
@@
 Thirdly, there is the cost of authors and data curators/custodians understanding intricacies of package development in order to share data.
+Nevertheless, the frictionless approach to sharing data in software packages makes reuse substantially faster than retrieving and correctly parsing files from a repository.
+We recommend the software approach only as an additive to proper data curation.
 

### 2021-08-11 new-paper.Rmd
add change from Karthik on walking a fine line between highly specific and overly general advice

@@
 There is also specific advice like the importance of very clean and properly shaped data [@White2013], and to use open source software with changes tracked [@Boland2017].
-These papers do not discuss the approach of sharing data within software, which has seen a rise in the use of research compendiums (e.g, [@Marwick2018]).
-Across many of these articles there is common advice that is useful, but individually, their advice is too general or specific for our audience.
+We walk a fine line between providing highly specific advice (e.g.
+how to specify NULL values) and the overly general (e.g.
+sharing data is good).
+Our goal is to provide practical advice for the researchers who are familiar with modern data science tools and develop their work like research compendia.
 

### 2021-08-11 new-paper.Rmd
add altered paragraph into metadata section

@@
 **Unit text** defines the type of variable recorded (e.g., string, date, numeric, integer), this avoids serious errors of misrepresented data (e.g., a column of a gene sequence being interpreted as a date).
+
 A simple metadata file that describes the variable names, variable description, and unit text is a great starting point, and is sometimes known as a codebook.
-For more details, see [@Broman2017; @Ellis2017; @Arslan2019].
 Metadata can be encoded into more formal plain text formats that follow a set standard, such as EML or SCHEMA-LD.
-A deep understanding of these formats isn't required to use them, for more details, and guidance on generating them see @Arslan2019.
+A deep understanding of these formats isn't required to use them.
+For more details, see [@Broman2017; @Ellis2017; @Arslan2019 @Leipzig2021].
 

### 2021-08-11 new-paper.Rmd
add citations for data sharing in biology and psychology

@@
 [@Colavizza2020] reflects similar findings for PLOS and BMC journals.
-There are numerous examples of similar problems in other fields, such as psychology, biology, and bioinformatics.
-The reality is that while data sharing is increasing in many fields such as ecology, the amount of data that can be reused is likely lower.
+There are numerous examples of similar problems in other fields, such as psychology [@vanpaemel2015] and biology [@schofield2009].
+The reality is that while data sharing is increasing in many fields such as ecology and genomics, the amount of data that can be reused is likely lower.
 To make shared data useful, we need to realign incentives for researchers.

### 2021-08-11 new-paper.Rmd
small changes to improve readability, a couple of tweaked sentences and moving the superscript citations around so they fit the sentences better.

@@
 \begin{singlespace}
-There are numerous arguments that strongly support practice of open science, with societal and individual benefits.
+There are numerous arguments that strongly support the practice of open science, with societal and individual benefits.
 For individual researchers, sharing research artifacts such as data can improve trust and transparency, reproducibility of one's own work, and catalyze new collaborations.
-Despite the general appreciation of the benefits of data sharing, research data are often only available to original investigators.
+Despite the general appreciation of the benefits of data sharing, research data are often only available to the original investigators.
 For data that are shared, lack of useful metadata and documentation make them challenging to reuse.
@@
 Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data.
-Each of these languages enjoys support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
+Each of these languages enjoys the support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
 Modern scientists now have access to many interactive computational frameworks (e.g.
@@
 Once there is a critical mass, find ways to reward those who put it into practice.
-After these steps, it would then be possible to require this of everyone Figure 1).
-This framework provides a useful lens through which to examine adoption of code, computational environments and data for reproducibility.
+After these steps, it would then be possible to require this of everyone (Figure \@ref(fig:incentable-table)).
+This framework provides a useful lens through which to examine adoption of code, computational environments, and data for reproducibility.
 With code, we have slowly progressed through these stages.
@@
 Despite the increase in data being shared alongside publications, the status quo of data sharing in science is not good.
+Rowhani-Farid et al.
 [@Rowhani-Farid2016] showed that 7 of 157 (4.5%) sampled papers in the British Medical Journal shared data, and only 14% of articles claiming open sharing of data actually shared their data.
-Similarly, [@Stodden2018] showed only 36% of papers in the journal *Science* shared data, despite the Journal's data sharing policy.
-[@Colavizza2020] reflects similar findings for PLOS and BMC journals.
+Similarly, Stodden et al.
+[@Stodden2018] showed only 36% of papers in the journal *Science* shared data, despite the Journal's data sharing policy.
+Similar findings have been found for PLOS and BMC journals [@Colavizza2020].
 There are numerous examples of similar problems in other fields, such as psychology [@vanpaemel2015] and biology [@schofield2009].
@@
 
-We believe a key audience could be better served by the current literature on data sharing: researchers who have finished data analysis on tabular data, are ready to submit for publication.
+We believe a key audience could be better served by the current literature on data sharing: researchers who have finished data analysis on tabular data, and are ready to submit for publication.
 Many papers provide rules and guidelines for sharing data, such as providing metadata, citation, licenses, and lists of data repositories to use.
@@
 To make data more reusable, we need to make it easier to generate metadata, standardise good practices within and across communities, and make reusable data sharing a recognized, normalised, part of research.
-Here we present a set of minimal set of practices to facilitate this change.
-These include documentation (Section \@ref(readme)), licensing (Section \@ref(licenses)), citation (Section \@ref(cite)), raw and analysis ready data (Section \@ref(rawdata)), code (Section \@ref(scripts)), and sharing in packages (Section \@ref(pkgs)).
-We also discuss considerations for sharing of data of various sizes (Section \@ref(size)).
+Here we present a set of common sense practices to facilitate this change.
+These include documentation (Section \@ref(readme)), licensing (Section \@ref(licenses)), metadata (Section \@ref(metadata)), raw and analysis ready data (Section \@ref(rawdata)), code (Section \@ref(scripts)), avenues to share data (Section \@ref(avenues)) and citation (Section \@ref(cite)).
+We also discuss considerations for sharing in packages (Section \@ref(pkgs)) and sharing of data of various sizes (Section \@ref(size)).
 
 We limit the scope of this paper to small to medium sized tabular data as it covers a vast majority of use cases that researchers encounter day to day.
-For larger datasets that involve complex workflows (e.g., bioinformatics and medical imaging), the kinds of challenges they face are different and accordingly deserve to be written up separately.
+Larger datasets with complex workflows (e.g., bioinformatics and medical imaging), face different kinds of challenges and accordingly deserve to be written up separately.
 
@@
 
-Our advice here is not meant to be exhaustive by any means and are insufficient to fully bridge the gap, but provide a starting point.
+Our advice here is not meant to be exhaustive by any means and is insufficient to fully bridge the gap, but instaed, to provide a starting point.
 
@@
 A README is often the first place users will look to learn more about a project.
-The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/> and.
+The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/>).
 For example, a README in a data repository might explain which directories contain data, tidying scripts, and also why data was collected.
-The README should be placed in the top level of the project, and there can be multiple README files in a project.
+The README should be placed in the top level of the project, and there can be multiple README files in a project (e.g., in different directories).
 In the context of datasets, READMEs are particularly useful when there are no reliable standards.
@@
 
-The date of use, and versions of software should be recorded in the README, so users can in the future can replicate similar computing environment or computing conditions.
+The date of use, associated paper, and versions of software should be recorded in the README, so users can in the future can replicate similar computing environment or computing conditions.
 A more ideal approach would be to clearly describe the computing environment (e.g., a Dockerfile) and packages used (e.g., requirements.txt).
@@
 We recommend the CC0 as it provides simple rules for attribution, warranty, and use, and is very common.
-If data that is CC0 is used in scientific publications, you need to cite where you got the data from.
+If data that is CC0 is used in scientific publications, you need to cite where you got the data from as part of scientific practice.
 There might be some concern that it is possible to conduct fraud, but it will be found out eventually if the original data is archived properly.
@@
 Use metadata with your data.
-As stated at the start of the paper, we refer to tabular data.
 There are three benefits to a dataset having metadata: (1) Provide context by describing variables to humans; (2) Avoid errors by describing and preserving variable types for machine readability; (3) Make data easier to find online by facilitating indexing [@google-data-search].
@@
 A deep understanding of these formats isn't required to use them.
-For more details, see [@Broman2017; @Ellis2017; @Arslan2019 @Leipzig2021].
+For more details, see [@Broman2017; @Ellis2017; @Arslan2019; @Leipzig2021].
 
@@
 
-## Citation: How you want your data to be cited {#cite}
-
-One critical advantage of depositing data in a trusted repository \@ref(avenues) is that it comes with many benefits such as preservation, indexing,  discoverability and most importantly, citation with persistent identifier.
-Authors can take an extra step by including their own CITATION file to give credit to a paper where the data and associated methods are described in detail.
-When citing data, you should cite the preferred CITATION file in the first instance, else, if there is a data publication, (for example in _Scientific Data_), cite this, and if none of those apply, cite the repository where the data is deposited (e.g., a Zenodo repository).
-
 ## Avenues to share data {#avenues}
@@
 While this approach is useful while iterating on the paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc.).
-Providing back up data is particularly relevant in light of the case in [@Schleunes2020; @Viglione2020], where data was suspected of being tampered.
+Providing back up data is particularly relevant in light of the cases where data is suspected of being tampered with [@Schleunes2020; @Viglione2020].
 The use of raw data allowed for issues to be identified, the absence of raw data not being available in other publications created difficulty in verifying the validity of the data.
 
-Journals provide data papers, where they provide a formal publication explaining data that is archived.
+Journals provide data papers in a formal publication explaining data that is archived.
 This is a familiar avenue for researchers to receive credit and citation.
@@
 Dryad will also provide professionals to curate your data, thereby improving usability.
+
+## Citation: How you want your data to be cited {#cite}
+
+One critical advantage of depositing data in a trusted repository (Section \@ref(avenues)) is that it comes with many benefits such as preservation, indexing,  discoverability and most importantly, citation with persistent identifier.
+Authors can take an extra step by including their own CITATION file to give credit to a paper where the data and associated methods are described in detail.
+When citing data, you should cite the preferred CITATION file in the first instance, else, if there is a data publication, (for example in _Scientific Data_), cite this, and if none of those apply, cite the repository where the data is deposited (e.g., a Zenodo repository).
 
@@
 Other fields, such as bioinformatics have a process of sharing data through repositories like Bioconductor, with larger data size.
-Thirdly, there is the cost of authors and data curators/custodians understanding intricacies of package development in order to share data.
+Thirdly, there is the cost of authors and data curators/custodians understanding the intricacies of package development in order to share data.
 Nevertheless, the frictionless approach to sharing data in software packages makes reuse substantially faster than retrieving and correctly parsing files from a repository.
-We recommend the software approach only as an additive to proper data curation.
+However, we recommend the software approach only as an additive to proper data curation.
 
-Packaged data and archival on venues like Zenodo are not mutually exclusive.
+Packaged data, and archival on venues like Zenodo are not mutually exclusive.
 First, archive data in a long-term data repository (Section \@ref(avenues)), before including it in a package.
@@
 
-Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software \@ref(pkgs).
+Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software (as in \@ref(pkgs)).
 For medium data, consider a venue like Zenodo for archiving.
@@
 Although data citations do not currently impact promotion and tenure packets, there is anecdotal evidence that papers sharing usable data accrue more citations [@Piwowar2007].
-Over time, the more useful we make the data, it is more likely incentive mechanisms will take hold.
+Over time, the more useful we make the data, the more likely incentive mechanisms will take hold.
 

### 2021-08-11 response-letter/response-letter.Rmd
write draft response letter

---
title: "Response to reviewer comments"
authors: Nicholas Tierney, Karthik Ram
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libs, include = FALSE}
library(tidyverse)
library(googlesheets4)
library(knitr)
library(kableExtra)
```

```{r read-sheet, include = FALSE}
our_table_id <- gs4_get(
  "1fIl8jy72XWD4liFZVlx6WvsnXp7ld8fs7eZvwVJpSy0"
  )

our_table <- read_sheet(ss = our_table_id,
                        .name_repair = janitor::make_clean_names) %>% 
  select(ID = id, 
         comment = reviewer_comment, 
         response)
```

Dear Editor in chief of Patterns,

Thank you for considering our commentary article “Incentives are the biggest barriers to widespread data sharing” for publication in Patterns.

We thank the reviewers for their have addressed comments from the reviewers, these are presented in the table below. The "ID" column value "R1-1" refers to reviewer 1 comment 1, and so on.

Regards,

Nicholas Tierney
Karthik Ram

## Response to Reviewers

```{r fix-long-url}

new_comment_line <- "The authors need to further clarify how their recommendations align with - and differ from - the slew of other recommendations for data management which exist (in articles - e.g https://doi.org/10.1371/journal.pcbi.1003542, and at the level of institutional/regional data stewardship - e.g. from the Research Data Alliance, or from libraries: https://www.monash.edu/library/researchers /n/researchdata/guidelines/sharing.) I think the key to differentiating this work is again making the community/audience explicit from the very beginning"

```

```{r show-table}
our_table %>% 
  mutate(comment = 
           case_when(
             ID == "R1-2" ~ new_comment_line,
             TRUE ~ comment
           )) %>% 
  kbl(booktabs = TRUE,
      longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position", 
                                  "repeat_header"),
                full_width = TRUE) %>% 
  column_spec(2, width = "20em") %>% 
  column_spec(3, width = "20em")
```

### 2021-08-11 response-letter/response-letter.tex
write draft response letter

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\title{Response to reviewer comments}
\author{}
\date{\vspace{-2.5em}}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Response to reviewer comments},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\begin{document}
\maketitle

Dear Editor in chief of Patterns,

Thank you for considering our commentary article ``Incentives are the
biggest barriers to widespread data sharing'' for publication in
Patterns.

We thank the reviewers for their have addressed comments from the
reviewers, these are presented in the table below. The ``ID'' column
value ``R1-1'' refers to reviewer 1 comment 1, and so on.

Regards,

Nicholas Tierney Karthik Ram

\hypertarget{response-to-reviewers}{%
\subsection{Response to Reviewers}\label{response-to-reviewers}}

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright\arraybackslash}p{20em}>{\raggedright\arraybackslash}p{20em}}
\toprule
ID & comment & response\\
\midrule
\endfirsthead
\multicolumn{3}{@{}l}{\textit{(continued)}}\\
\toprule
ID & comment & response\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
R1-1 & The authors have worked to clarify the type of data that the article is geared toward. However, they still need to be even more specific earlier on about who the community is that they are addressing, namely the community of computational data scientists working in academia. This is mentioned later in the paper (p. 4), but ideally this information could be worked into the title and definitely it should be mentioned in the abstract. & We have added the following to the end of the abstract to clarify the audience: Other papers have focussed on many general rules for sharing data. This paper focusses on common sense ideas for sharing tabular data, for a target audience of academics working in data science adjacent fields who are about to submit for publication.\\
R1-2 & The authors need to further clarify how their recommendations align with - and differ from - the slew of other recommendations for data management which exist (in articles - e.g https://doi.org/10.1371/journal.pcbi.1003542, and at the level of institutional/regional data stewardship - e.g. from the Research Data Alliance, or from libraries: https://www.monash.edu/library/researchers /n/researchdata/guidelines/sharing.) I think the key to differentiating this work is again making the community/audience explicit from the very beginning & We have added the following into section two, to help further describe our audience, and to describe how our paper is different to others:"We believe a key audience could be better served by the current literature on data sharing: researchers who have finished data analysis on tabular data, are ready to submit for publication.
Many papers provide rules and guidelines for sharing data, such as providing metadata, citation, licenses, and lists of data repositories to use. These papers provide general advice from the ethical considerations of sharing data [@Moles2013], how to share many different data types (images, video, etc) [@Hart2016], to “Reward Colleagues Who Share Their Data Properly” [@Goodman2014]. There is also specific advice like the importance of very clean and properly shaped data [@White2013], and to use open source software with changes tracked [@Boland2017]. We walk a fine line between providing highly specific advice (e.g. how to specify NULL values) and the overly general (e.g. sharing data is good). Our goal is to provide practical advice for the researchers who are familiar with modern data science tools and develop their work like research compendia.\\
R1-3 & Some of the recommendations themselves border on being too generic - e.g. 'use metadata with your data' - without specifying how metadata schemas/vocabularies can be chosen or what type of metadata are being considered. The examples that are provided (e.g. a list of variable names) are something that is perhaps more akin to including a codebook rather than structured metadata following a particular schema. Perhaps this overlaps with data documentation? Do the authors want to address more structured forms of metadata schemas which are often required by repositories? & We have added the following into the metadata section:

A simple metadata file that describes the variable names, variable description, and unit text is a great starting point, and is sometimes known as a codebook. Metadata can be encoded into more formal plain text formats that follow a set standard, such as EML or SCHEMA-LD. A deep understanding of these formats isn't required to use them. For more details, see [@Broman2017; @Ellis2017; @Arslan2019 @Leipzig2021].\\
R1-4 & Along the same lines, it is unclear why the authors recommend using Zenodo and Dryad specifically to share data, aside from issues of data size (and I am not completely convinced by this argument either). Usually recommendations point researchers toward lists of repositories, searchable by subject, (e.g. re3data.org) which can be used to select an appropriate repository. Is there a reason for not doing this? & While we agree with the reviewer in general, R3Data is merely a long list of repositories. It does not provide curation or advice on best practices to make the data reusable. In contrast, Dryad and Zenodo are both general purpose non-commercial repositories that are leading the way in best practices for data and software deposition. So we have revised this section to note that researchers should follow community advice on domain repositories when available. Barring those, then Dryad and Zenodo are best suited for the purposes of reproducible data science. Revised text reads: "Data should be deposited into a relevant domain repository that generates an accession number such as a DOI. For example, gbif for species data, or genbank for genetics data. A exhaustive searchable list is available at re3data (https://www.re3data.org/) but provides no guidance to researchers. Data that do not fit appropriate domain repositories can be submitted to research data repositories **Zenodo** and **Dryad** which are currently leading the way in best practices for software and data. In addition to minting DOIs, they also provide citation templates."\\
R1-5 & The structure of the paper could be improved. The article consists of two main parts: 1) the Introduction (which is really a section discussing the current state of incentivising data code and sharing for academic data scientists) and 2) Recommendations for data sharing. I recommend putting all of the recommendations and the current section 1.1 as subpoints in a second section devoted only to recommendations. & We have now changed this so the Introduction is section 1, and then the recommendations are section 2, with ssubsections 2.1 to 2.9\\
\addlinespace
R1-6 & The title of the paper does not seem to match the main gist of the paper. It takes some work for the reader to figure out that the recommendations are tied to incentivising data sharing. How exactly do the recommendations provide incentives for academic data scientists? Perhaps consider changing the title and/or spelling out the link between the two sections a bit more clearly. & We have altered the abstract to better capture out intended message. The title has...\\
R1-7 & Perhaps it would help to spell out who the 'you' is that is being addressed in the recommendations? & This has been addressed by clarifying the audience, which has been done in the abstract and introduction\\
R1-8 & The treatment of data citations on p. 4 overlooks many issues about data citation and efforts which are underway. The authors state that 'datasets currently do not accrue citation credit since they are not tracked.' Not only are citations not tracked - many researchers don't actually cite data in the first place. & We have cited the most current work on data citations, but the rest of the reviwer comment restates what we have in the paper already.\\
R1-9 & typo p.2 - around footnote \#8 - there seem to be some missing phrases here? & Removed the unneeded 'and', changed, "In a study of applied computational research, and colleagues were only able to install" to "In a study of applied computational research, less than 20\% of software described in publications was able to be installed and run."\\
R1-10 & p.2 - What is meant by ''trivial reproducibility?' & Removed 'trivial' - sentence changes from, "The barriers to verification are now low enough that trivial reproducibility is sometimes a click away" to "The barriers to verification are now low enough that reproducibility is sometimes a click away"\\
\addlinespace
R1-11 & p..6 - 'We don't have recommendations for non-tabular data.' The authors probably don't need to mention this, as this has already been made clear. & we can remove this sentence: https://github.com/karthik/ddd/blob/master/new-paper.Rmd\#L130\\
R1-12 & p. 9 - Reference 25 - Does the Piwowar paper really provide 'anecdotal' evidence? & NA\\
R1-13 & A sentence in the abstract currently states "research data are only available to original investigators." A more accurate statement would be ""research data are often only available to original investigators." & Sentence changed from, "Despite the general appreciation of the benefits of data sharing, research data are available only to the original investigators" to "Despite the general appreciation of the benefits of data sharing, research data are often only available to original investigators."\\
R2-1 & There are many other similar papers that have already been published on this topic, and I do not see that this work adds substantial new knowledge or advice. For instance: 10 Simple Rules for the Care and Feeding of Scientific Data (https://doi.org/10.1371/journal.pcbi.1003542), Nine simple ways to make it easier to (re)use your data (https://ojs.library.queensu.ca/index.php/IEE/issue/view/478), Ten Simple Rules for Digital Data Storage (https://doi.org/10.1371/journal.pcbi.1005097), Ten Simple Rules to Enable Multi-site Collaborations through Data Sharing (https://doi.org/10.1371/journal.pcbi.1005278), and The Turing Way: A Handbook for Reproducible Data Science(http://doi.org/10.5281/zenodo.3233853). & Addressed in earlier comment\\
R2-2 & The title of the paper makes it sound like this will be a review of current incentives \& barriers to data sharing and maybe some original research into this, but that is not what this paper is. This paper has an introduction that includes some information about how data isn't being shared, and then tips for what you should do when you share your data. As stated above, there are many other papers that share these same tips for sharing data, so in my opinion, this paper would be much stronger if it was indeed a more in depth discussion/research on what incentives are currently available and how they are working/not working. & We have changed the paper title and added information in the introduction to better articulate the audience we are addressing and that we think would benefit the most from reading it, and how it is different to the many other guides out there.  While we believe that a paper which, provides "a more in depth discussion/research on what incentives are currently available and how they are working/not working" - our paper has a different goal - to describe the misalignment of these incentives, and to provide researchers with common sense approaches to sharing data.\\
\addlinespace
R2-3 & Section 9 is confusing. The authors say we can share data within a software package, but then spend a lot of time detailing why we shouldn't do that. What is the recommendation here? & We have added the following sentence to address this comment: " Nevertheless, the frictionless approach to sharing data in software packages makes reuse substantially faster than retrieving and correctly parsing files from a repository. We recommend the software approach only as an additive to proper data curation."\\
R2-4 & The audience is a bit unclear to me. There is a mix of advice that is very general with little detail and then advice that is very specific. Who is the target audience for this advice? & we have clarified the audience in an earlier comment.\\
R2-5 & The entire manuscript could benefit from more proof reading. There are several missing words, especially missing article/resource names - perhaps the citation manager removed the text of several citations? Also, the introduction is very long and has some issues with logical flow (e.g. the second paragraph switches context between progress and problems with reproducibility without any transitions). This could likely be resolved with another proof reading. & NA\\
R2-6 & Figure 1 could use more explanation. The authors should introduce Nosek, and perhaps add a citation. & We have updated the citation.\\
R2-7 & "There are numerous examples of similar problems in other fields, such as psychology, biology, and bioinformatics." Please add citations for this. & We have updated the citation.\\
\addlinespace
R2-8 & Watch for writing style that might discourage readers from implementing these ideas - for instance, remove phrases that are passive aggressive or off putting like "Just archive the data." & We have read through the paper and removed the phrase the reviewer mentioned, as well as checked the paper for clarity\\*
\end{longtabu}

\end{document}

### 2021-08-11 response-letter/response-letter.Rmd
minor changes to get the table to print (trick was to include `column_spec(1)`)

@@
 ```{r libs, include = FALSE}
-library(tidyverse)
+library(dplyr)
 library(googlesheets4)
@@
 library(kableExtra)
+```
+
+```{r fix-long-url}
+
+new_comment_line <- "The authors need to further clarify how their recommendations align with - and differ from - the slew of other recommendations for data management which exist (in articles - e.g https://doi.org/10.1371/journal.pcbi.1003542, and at the level of institutional/regional data stewardship - e.g.
+from the Research Data Alliance, or from libraries: https://www.monash.edu/library/researchers /n/researchdata/guidelines/sharing.) I think the key to differentiating this work is again making the community/audience explicit from the very beginning"
+
 ```
@@
          comment = reviewer_comment, 
-         response)
+         response) %>% 
+  mutate(comment = 
+           case_when(
+             ID == "R1-2" ~ new_comment_line,
+             TRUE ~ comment
+           )) 
 ```
@@
 
-We thank the reviewers for their have addressed comments from the reviewers, these are presented in the table below. The "ID" column value "R1-1" refers to reviewer 1 comment 1, and so on.
+We thank the reviewers for time and comments.
+We have addressed comments from the reviewers and believe the paper is clearer and better as a result.
+Our response to reviewer comments are presented in the table below.
+The "ID" column value "R1-1" refers to reviewer 1 comment 1, and so on.
 
@@
 
-Nicholas Tierney
-Karthik Ram
-
-## Response to Reviewers
-
-```{r fix-long-url}
-
-new_comment_line <- "The authors need to further clarify how their recommendations align with - and differ from - the slew of other recommendations for data management which exist (in articles - e.g https://doi.org/10.1371/journal.pcbi.1003542, and at the level of institutional/regional data stewardship - e.g.
-from the Research Data Alliance, or from libraries: https://www.monash.edu/library/researchers /n/researchdata/guidelines/sharing.) I think the key to differentiating this work is again making the community/audience explicit from the very beginning"
-
-```
+Nicholas Tierney & Karthik Ram
 
@@
 our_table %>% 
-  mutate(comment = 
-           case_when(
-             ID == "R1-2" ~ new_comment_line,
-             TRUE ~ comment
-           )) %>% 
   kbl(booktabs = TRUE,
       longtable = TRUE) %>% 
-  kable_styling(latex_options = c("hold_position", 
+  kable_styling(latex_options = c("striped",
+                                  "hold_position", 
                                   "repeat_header"),
                 full_width = TRUE) %>% 
+  column_spec(1) %>% 
   column_spec(2, width = "20em") %>% 

### 2021-08-11 new-paper.Rmd
Paper restructure as described in email

@@
 We compare data to code, and computational environments in the context of reproducible research and provide some practical guidance on how one can improve the chances of their data being reusable and partially bridge the incentive gap.
-Other papers have focussed on many general rules for sharing data.
-This paper focusses on common sense ideas for sharing tabular data, for a target audience of academics working in data science adjacent fields who are about to submit for publication.
+While previous papers have focused on describing ideal best practices for data and code, we focus on  common sense ideas for sharing tabular data, for a target audience of academics working in data science adjacent fields who are about to submit for publication.
 
@@
 The idea of open and transparent science dates back to the founding principles of The Royal Society of London in 1660.
-The movement, which has seen shifting support over the years, has enjoyed growing popularity over the past decade [@Nielsen2020; @Peng2011] mainly among early career researchers [@McKiernan2016].
+The movement, which has seen shifting support over the years, has enjoyed growing popularity over the past decade [@Nielsen2020; @Peng2011] primarily among early career researchers [@McKiernan2016].
 A key motivation for practicing openness is the notion that sharing all research artifacts will allow others to reproduce and verify results, thereby improving trust and verification of those findings.
@@
 Jupyter, RMarkdown, Stencila, Observable) that can interface with kernels from dozens of programming languages.
-Beyond notebooks to run the code, a reader will also require access to the same software dependencies.
-In a study of applied computational research, less than 20% of software described in publications was able to be installed and run [@Collberg2014].
-In addition to locating software, one still has to locate the right versions.
-Dependency management is a controversial topic in programming communities with no easy solutions.
-While languages like JavaScript have solved this with package managers like npm, it is much harder to accomplish in others like Python and R which are currently supported by a patchwork of solutions.
-Although this problem is far from solved, technologies like Docker have alleviated some of these concerns.
-Researchers can share Dockerfiles to recreate computational environments [@Boettiger2015].
-Services like Binder can then leverage Docker and make it possible to share the computational environments in which these notebooks were run.
 The barriers to verification are now low enough that reproducibility is sometimes a click away, free from the misery of software installation [@Jupyter2018].
@@
 
-Despite the increase in data being shared alongside publications, the status quo of data sharing in science is not good.
+While there has ben a growing trend of sharing data alongside publications, the status quo of data sharing in science is not good.
 Rowhani-Farid et al.
@@
 
-To make data more reusable, we need to make it easier to generate metadata, standardise good practices within and across communities, and make reusable data sharing a recognized, normalised, part of research.
+To make data more reusable, we need to make it easier to generate metadata, standardize good practices within and across communities, and make reusable data sharing a recognized, normalized, part of research.
 Here we present a set of common sense practices to facilitate this change.
@@
 A package or module structure provides an advantage as it can be "installed" and loaded just like code.
-Despite the advantages of sharing data as a software package, there are a few major disadvantages.
+Obvious benefits of this approach is that it allows data to be immediately used in a programming environment without the additional steps of downloading, preprocessing, specifying column types etc.
+But there are some limitations to this approach.
 Firstly, packaging data inside a software package makes the data availability language centric.
@@
 
-Packaged data, and archival on venues like Zenodo are not mutually exclusive.
+Packaging data in software and archiving in data repositories are not mutually exclusive activities.
 First, archive data in a long-term data repository (Section \@ref(avenues)), before including it in a package.
 The act of packaging data in software can shave off hours required to prepare it for use.
-The canocical citation can be baked into the package's citation file.
+The canonical citation can be baked into the package's citation file.
 
@@
 Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software (as in \@ref(pkgs)).
-For medium data, consider a venue like Zenodo for archiving.
+For medium data, a general purpose data repository (e.g.
+Zenodo) would be ideal for archiving.
 Sharing medium data might require it to be broken into separate chunks for reading (e.g., `teaching-1.csv`, `teaching-2.csv`, etc.), and scripts provided to read it.
@@
 The mere act of depositing data does not make it reusable or useful to a potential consumer.
-Although data citations do not currently impact promotion and tenure packets, there is anecdotal evidence that papers sharing usable data accrue more citations [@Piwowar2007].
+Although data citations do not currently impact promotion and tenure packets, there is some evidence that papers sharing usable data accrue more citations [@Piwowar2007].
 Over time, the more useful we make the data, the more likely incentive mechanisms will take hold.

### 2021-08-13 new-paper.Rmd
minor changes to first two pages

@@
 title: Incentives are the biggest barriers to widespread data sharing
-author: Nicholas J Tierney\* (1,2,3) & Karthik Ram\* (4)
+author: Nicholas J. Tierney\* (1, 2, 3) & Karthik Ram\* (4)
 output:
@@
 4 = Berkeley Institute for Data Science, University of California, Berkeley, USA\\
-* = Both authors contributed equally to the work\\
+* = Both authors contributed equally to the work.\\
 Correspondence to:\\
@@
 \begin{singlespace}
-There are numerous arguments that strongly support the practice of open science, with societal and individual benefits.
-For individual researchers, sharing research artifacts such as data can improve trust and transparency, reproducibility of one's own work, and catalyze new collaborations.
-Despite the general appreciation of the benefits of data sharing, research data are often only available to the original investigators.
+Numerous arguments strongly support the practice of open science, which offers several societal and individual benefits.
+For individual researchers, sharing research artifacts such as data can increase trust and transparency, improve the reproducibility of one's own work, and catalyze new collaborations.
+Despite a general appreciation of the benefits of data sharing, research data are often only available to the original investigators.
 For data that are shared, lack of useful metadata and documentation make them challenging to reuse.
-In this paper we argue that incentives and infrastructure for making data useful are the biggest barrier to creating a culture of widespread data sharing.
-We compare data to code, and computational environments in the context of reproducible research and provide some practical guidance on how one can improve the chances of their data being reusable and partially bridge the incentive gap.
+In this paper, we argue that a lack of incentives and infrastructure for making data useful are the biggest barrier to creating a culture of widespread data sharing.
+We compare data to code, examine computational environments in the context of  their ability to facilitate the reproducibility of research, provide some practical guidance on how one can improve the chances of their data being reusable, and partially bridge the incentive gap.
 While previous papers have focused on describing ideal best practices for data and code, we focus on  common sense ideas for sharing tabular data, for a target audience of academics working in data science adjacent fields who are about to submit for publication.
@@
 
-> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (The Adventure of the Copper Beeches by Sir Arthur Conan Doyle)
+> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." -- Sherlock Holmes ("The Adventure of the Copper Beeches" by Sir Arthur Conan Doyle)
 
@@
 The movement, which has seen shifting support over the years, has enjoyed growing popularity over the past decade [@Nielsen2020; @Peng2011] primarily among early career researchers [@McKiernan2016].
-A key motivation for practicing openness is the notion that sharing all research artifacts will allow others to reproduce and verify results, thereby improving trust and verification of those findings.
-Jon Clarebout famously wrote in 1992 that reproducibility in computationally oriented fields had become so trivial that non experts could do the task.
-Clarebout's optimistic vision was a little ahead of its time as the tools and technologies available in the early nineties for literate programming presented too high of a barrier for most researchers to adopt without incurring a significant loss in productivity [@Barnes2010].
+A key motivation for practicing openness is the notion that sharing all research artifacts will allow others to reproduce and verify results, thereby improving trust in and verification of those findings.
+Jon Clarebout famously wrote in 1992 that reproducibility in computationally-oriented fields had become so trivial that nonexperts could do the task.
+Clarebout's optimistic vision was a little ahead of its time, as the tools and technologies available in the early 1990s for literate programming presented too high a barrier for most researchers to adopt without incurring a significant loss in productivity [@Barnes2010].
 
-Meaningful progress on most computational reproducibility, the ability of an independent researcher to recreate figures, tables and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence.
+Meaningful progress on most computational reproducibility, the ability of an independent researcher to recreate figures, tables, and key findings, became more of a practical reality in the late 2000s as new tools and services came into existence.
 Scientists started sharing more code and software, partly due to the rapid increase in training opportunities made possible by efforts like The Carpentries, combined with the growing adoption of GitHub by scientists [@Ram2013; @Zagalsky2015].
-The bigger driver for this is more likely due to the growing popularity of data science as a discipline distinct from statistics [@Donoho2017].
-This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open source software tools.
-Programming languages such as Python, R and Julia have helped scientists implement and share new methods to work with data.
+The bigger driver for this is more likely the growing popularity of data science as a discipline distinct from statistics [@Donoho2017].
+This rapid growth in data science has been catalyzed by a Cambrianesque explosion of open-source software tools.
+Programming languages such as Python, R, and Julia have helped scientists implement and share new methods for working with data.
 Each of these languages enjoys the support of thriving communities of researchers and software developers who contribute many of the building blocks that make them popular.
@@
 
-Notebooks however, are an incomplete solution for computational reproducibility.
-For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) Code; 2) Computing environment, and 3) Data.
+Notebooks, however, are an incomplete solution for computational reproducibility.
+For a piece of computational research to be minimally reproducible, it requires three distinct elements: 1) code; 2) computing environment, and 3) data.
 There has been tremendous progress on the first two, despite some limitations of existing tools and practices.
-While it took nearly two decades for Clarebout's vision to become reality, improvements in technology and motivations played a strong role in the culture change.
-Sharing code wasn't just easy at first [@Barnes2010], but became a social and normative part of academic culture.
-The next stage of this evolution, turning code into software, also became popular with increasing training opportunities [@Wilson2014], tooling, and recognition as a scholarly contribution [@Smith2018].
+While it took nearly two decades for Clarebout's vision to become reality, improvements in technology and motivations have played a strong role in the culture change.
+Sharing code wasn't easy at first [@Barnes2010], but became a social and normative part of academic culture.
+The next stage of this evolution, turning code into software, also became popular with increasing training opportunities [@Wilson2014], tooling, and recognition of code as a scholarly contribution [@Smith2018].
 Getting researchers to share usable data has been more of a struggle.

### 2021-08-13 new-paper.Rmd
small improvements to text of section 1 completed.

@@
 
-```{r incentable-table, fig.cap = "Table of the mechanism for behavior change, the incentive, and our assessment of where the elements of Data, Code, and Computational Environment rank in terms of completing these aspects.
-We note that Data is often required, but the preceding steps are not required.
-This is in contrast to Code having no policy."}
+```{r incentable-table, fig.cap = "Table of the mechanisms for behavior change, the incentives, and our assessment of where the elements of data, code, and computational environment rank in terms of completing these aspects.
+We note that data is often required, but the preceding steps are not, in contrast to code, which has no policy."}
 knitr::include_graphics("figures/incentives_pyramid.png")
@@
 Nosek [@nosek_2021] describes the strategies for culture change, especially when the status quo is too entrenched.
-Beyond motivation, skills, tools, and incentives are critical to enable such change.
-In a five step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, then improve the infrastructure to make it easy, followed by increased awareness to make it normative.
-Once there is a critical mass, find ways to reward those who put it into practice.
-After these steps, it would then be possible to require this of everyone (Figure \@ref(fig:incentable-table)).
+For change to occur, more than just motivation is needed: Skills, tools, and invenctives are critical.
+In a five-step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, improve the infrastructure to make it easy, and then increase awareness to make it normative.
+Once there is a critical mass, ways must be found to reward those who put it into practice.
+Once change has become normative and is widely practices, it is possible to require it of everyone (Figure \@ref(fig:incentable-table)).
 This framework provides a useful lens through which to examine adoption of code, computational environments, and data for reproducibility.
 With code, we have slowly progressed through these stages.
-While few journals require that authors share code, the venues to share, along with the necessary training have paved the way for incentives and credit.
-All of the necessary foundation to make code sharing required are in place.
+While few journals require that authors share code, the provision of venues within which code can be shared, along with the necessary training, have paved the way for incentives and credit.
+All of the necessary foundations to make code sharing required are in place.
 For computational environments, Docker makes it possible to share them in most common use cases [@Boettiger2015].
 But running them still requires expertise and time on the part of reviewers.
-More recent open services like Binder [@Jupyter2018] make it possible to run notebooks atop Docker images, but the technology is still new and experiencing growing pains.
+More recent open services, like Binder [@Jupyter2018], make it possible to run notebooks atop Docker images, but the technology is still new and experiencing growing pains.
 There is still a long way to go before sharing computational environments in academia becomes mainstream.
+
 Finally, for data, we have implemented these steps out of order.
-The FAIR data principles were introduced in 2016 with good intentions [@Wilkinson2016] and describe the desired attributes for research data.
-That they be findable (good metadata), accessible (deposited somewhere persistent), interoperable (easily combined with other datasets), and be reusable.
-We as a community started out by making data required before demonstrating how this would directly benefit those who share data.
+The FAIR data principles, introduced in 2016 with good intentions, [@Wilkinson2016] describe the desired attributes for research data: that they be findable (good metadata), accessible (deposited somewhere persistent), interoperable (easily combined with other datasets), and reusable.
+We, as a community, started out by making data required before demonstrating how this would directly benefit those who share data.
 Researchers are asked to deposit their data in one of many institutional or commercial repositories.
 Many journals even require it as a condition for publication via policy.
-However, the cost of preparing data for reuse remains high, and the career incentives to do so are almost non-existent.
+However, the cost of preparing data for reuse remains high, and career incentives to do so are almost nonexistent.
 Until we make data reuse easier and more rewarding, no level of policy is going to improve the state of data sharing.
 
-While there has ben a growing trend of sharing data alongside publications, the status quo of data sharing in science is not good.
+While there has been a growing trend of sharing data alongside publications, the status quo of data sharing in science is not good.
 Rowhani-Farid et al.
-[@Rowhani-Farid2016] showed that 7 of 157 (4.5%) sampled papers in the British Medical Journal shared data, and only 14% of articles claiming open sharing of data actually shared their data.
+[@Rowhani-Farid2016] showed that 7 of 157 (4.5%) sampled papers in the *British Medical Journal* shared data, and only 14% of articles claiming open sharing of data actually shared their data.
 Similarly, Stodden et al.
-[@Stodden2018] showed only 36% of papers in the journal *Science* shared data, despite the Journal's data sharing policy.
-Similar findings have been found for PLOS and BMC journals [@Colavizza2020].
+[@Stodden2018] showed that only 36% of papers published in the journal *Science* shared data, despite the journal's data sharing policy.
+Similar findings have been made for PLOS and BMC journals [@Colavizza2020].
 There are numerous examples of similar problems in other fields, such as psychology [@vanpaemel2015] and biology [@schofield2009].
-The reality is that while data sharing is increasing in many fields such as ecology and genomics, the amount of data that can be reused is likely lower.
+The reality is that while data sharing is increasing in many fields, such as ecology and genomics, the amount of data that can be reused is likely lower.
 To make shared data useful, we need to realign incentives for researchers.
 
-Tools to prepare datasets and metadata (*the infrastructure*) and minimal set of practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015].
-Unlike code, datasets are far more diverse in terms of heterogeneity, size, and formats.
+Tools that can be used to prepare datasets and metadata (*the infrastructure*) and minimal sets of practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015].
+Unlike code, datasets are far more diverse in terms of heterogeneity, size, and format.
 This makes them particularly challenging to standardize or easily "install" where the code is running.
-There are some efforts working towards improving this process, such as gogetdata , the 'dat' project, and a private company, Quilt.
+There have been some efforts to improve this process, such as gogetdata, the Dat Project, and the formation of a private company, Quilt.
 
-Depositing data with rich metadata is laborious making it a pro forma activity outside of speciality repositories like Dryad that employ curators.
-Lastly, the infrastructure to track and get credit for data is non-existent.
+Depositing data with rich metadata is laborious, making it a pro forma activity, outside of speciality repositories like Dryad that employ curators.
+Finally, the infrastructure to track and get credit for data is does not yet exist.
 Datasets currently do not accrue citation credit since they are not tracked [@lowenberg_2019].

### 2021-08-13 new-paper.Rmd
finish the rest of the minor changes.

@@
 We believe a key audience could be better served by the current literature on data sharing: researchers who have finished data analysis on tabular data, and are ready to submit for publication.
-Many papers provide rules and guidelines for sharing data, such as providing metadata, citation, licenses, and lists of data repositories to use.
-These papers provide general advice from the ethical considerations of sharing data [@Moles2013], how to share many different data types (images, video, etc) [@Hart2016], to “Reward Colleagues Who Share Their Data Properly” [@Goodman2014].
-There is also specific advice like the importance of very clean and properly shaped data [@White2013], and to use open source software with changes tracked [@Boland2017].
+Many papers provide rules and guidelines for sharing data, such as providing metadata, citations, licenses, and lists of data repositories to use.
+These papers provide general advice that ranges from ethical considerations raised by sharing data [@Moles2013], to how to share many different data types (images, video, etc) [@Hart2016], to tips such as "Reward Colleagues Who Share Their Data Properly." [@Goodman2014] There is also specific advice, such as the importance of very clean and properly shaped data [@White2013], and how to use open-source software with changes tracked [@Boland2017].
 We walk a fine line between providing highly specific advice (e.g.
-how to specify NULL values) and the overly general (e.g.
-sharing data is good).
-Our goal is to provide practical advice for the researchers who are familiar with modern data science tools and develop their work like research compendia.
+how to specify null values) and  overly general suggestions (e.g., sharing data is good).
+Our goal is to provide practical advice for researchers who are familiar with modern data science tools and who already develop their work including research compendia.
 
-As academic data scientists, we spend considerable time cleaning messy data and believe that preparing data for easy reuse can be a catalyst to improve the quality of data and pave the way for other incentives such as community adoptions, which will then create the culture change needed to credit data similar to publications.
+As academic data scientists, we spend considerable time cleaning messy data and believe that preparing data for easy reuse can be a catalyst to improve the quality of data and pave the way for other incentives, such as community adoptions, which will then create the culture change needed to credit data in a way similar to the citations used to credit publications.
 In this paper, we describe a set of common sense practices that researchers can adopt to make data more useful in the context of reproducibility.
 It should not be a huge burden to share data in a format that is readily usable by others.
-And similarly, this burden should be reduced by those who want to use another's data.
+And, similarly, this burden should be reduced by and for those who want to use another's data.
 
@@
 Here we present a set of common sense practices to facilitate this change.
-These include documentation (Section \@ref(readme)), licensing (Section \@ref(licenses)), metadata (Section \@ref(metadata)), raw and analysis ready data (Section \@ref(rawdata)), code (Section \@ref(scripts)), avenues to share data (Section \@ref(avenues)) and citation (Section \@ref(cite)).
-We also discuss considerations for sharing in packages (Section \@ref(pkgs)) and sharing of data of various sizes (Section \@ref(size)).
+These include documentation (Section \@ref(readme)), licensing (Section \@ref(licenses)), metadata (Section \@ref(metadata)), raw and analysis-ready data (Section \@ref(rawdata)), code (Section \@ref(scripts)), avenues to share data (Section \@ref(avenues)) and citation (Section \@ref(cite)).
+We also discuss considerations for sharing in packages (Section \@ref(pkgs)) and sharing data of various sizes (Section \@ref(size)).
 
-We limit the scope of this paper to small to medium sized tabular data as it covers a vast majority of use cases that researchers encounter day to day.
-Larger datasets with complex workflows (e.g., bioinformatics and medical imaging), face different kinds of challenges and accordingly deserve to be written up separately.
+We limit the scope of this paper to small-to medium-sized tabular data, which includes a vast majority of use cases that researchers encounter day to day.
+Larger datasets with complex workflows (e.g., bioinformatics and medical imaging) face different kinds of challenges and accordingly deserve to be treated separately.
 
@@
 
-Our advice here is not meant to be exhaustive by any means and is insufficient to fully bridge the gap, but instaed, to provide a starting point.
+Our advice is not meant to be exhaustive, by any means, and is insufficient to fully bridge the gap, but we hope instead, to provide a starting point.
 
@@
 A README is often the first place users will look to learn more about a project.
-The README is a concise text file containing guidance on how to read and interpret the directory - the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/>).
+The README is a concise text file containing guidance on how to read and interpret the directory -- the who, what, when, where, and how (<https://betta.io/blog/2017/02/07/developer-experience-github-readmes/>).
 For example, a README in a data repository might explain which directories contain data, tidying scripts, and also why data was collected.
@@
 
-The date of use, associated paper, and versions of software should be recorded in the README, so users can in the future can replicate similar computing environment or computing conditions.
-A more ideal approach would be to clearly describe the computing environment (e.g., a Dockerfile) and packages used (e.g., requirements.txt).
+The date of use, associated paper, and versions of software should be recorded in the README, so users can, in the future, replicate similar computing environments or conditions.
+A more ideal approach would be to clearly describe the computing environment (e.g., in a Dockerfile) and packages used (e.g., requirements.txt).
 
@@
 A license for data establishes the rules for its use, and whether it can be modified or shared.
-Without a license, or with a non-standard one, these rules are unclear, and can lead to issues with use, attribution and citation.
-While there are a handful of licenses for data, we recommend using the CC0 license, which means data owners waive all rights to the data, so it can be freely shared, copied, modified, and distributed, without permission or attribution, even for commercial purposes.
+Without a license, or with a nonstandard one, these rules are unclear, and can lead to issues with use, attribution, and citation.
+While there are a handful of licenses for data, we recommend using the CC0 license, which means that data owners waive all rights to the data, so it can be freely shared, copied, modified, and distributed, without permission or attribution, even for commercial purposes.
 
 We recommend the CC0 as it provides simple rules for attribution, warranty, and use, and is very common.
-If data that is CC0 is used in scientific publications, you need to cite where you got the data from as part of scientific practice.
-There might be some concern that it is possible to conduct fraud, but it will be found out eventually if the original data is archived properly.
+If CC0-licensed is used in scientific publications, you need to cite the location from which you obtained the data as part of scientific practice.
+There might be some concern that it is possible to conduct fraud, but any such fraud is likely to be found out eventually if the original data is archived properly.
 
-For more information on the benefit of using CC0, and why CCBY is not always appropriate, see https://osc.universityofcalifornia.edu/2016/09/cc-by-and-data-not-always-a-good-fit/.
+For more information on the benefits of using CC0, and why CCBY is not always appropriate, see https://osc.universityofcalifornia.edu/2016/09/cc-by-and-data-not-always-a-good-fit/.
 Short and full overviews of CC0 can be found at https://creativecommons.org/.
@@
 
-Use metadata with your data.
-There are three benefits to a dataset having metadata: (1) Provide context by describing variables to humans; (2) Avoid errors by describing and preserving variable types for machine readability; (3) Make data easier to find online by facilitating indexing [@google-data-search].
+Include metadata with your data.
+There are three benefits to a dataset that includes metadata: (1) It provides context by describing variables to humans, (2) It avoids errors by describing and preserving variable types for machine readability, and (3) It makes data easier to find online by facilitating indexing [@google-data-search].
 Three pieces of metadata should be considered: (1) **variable names**, (2) **variable descriptions**, and (3) **unit text**.
-These should be curated by a person.
-Metadata created should be clearly labelled in a folder or in files named METADATA.
+These should be curated by a human.
+Metadata created should be clearly labeled in a folder or in files named METADATA.
 
-**Variable names** are short, descriptive names with no spaces or special characters, such as "job_position" or "date".
-**Variable description** describes the measured variables, e.g., "University Job Position" or "Date Started", and should also define *variable codes* and *missing data*.
-**Unit text** defines the type of variable recorded (e.g., string, date, numeric, integer), this avoids serious errors of misrepresented data (e.g., a column of a gene sequence being interpreted as a date).
+**Variable names** should be short, descriptive names with no spaces or special characters, such as "job_position" or "date." **Variable description** describes the measured variables, e.g., "university job position" or "date started," and should also define *variable codes* and *missing data*.
+**Unit text** defines the type of variable recorded (e.g., string, date, numeric, integer), in order to avoid serious errors of misrepresented data (e.g., a column of a gene sequence being interpreted as a date).
 
@@
 
-## Raw data vs analysis ready data {#rawdata}
+## Raw data versus analysis-ready data {#rawdata}
 
-Make raw and analysis ready data available.
+Make both raw and analysis-ready data available.
 Raw data is the first format of data provided before any cleaning has taken place (e.g., data in some binary or proprietary format).
 It should be placed in its own folder.
-Data used in analysis should be provided, especially if computationally intense to derive.
-It should be in its own folder, and to maximize interoperability, recorded in a plain text (e.g., csv, tsv) and if possible a binary format (e.g., .rds, .sav, .xlsx, Apache Parquet).
+Data used in analysis should be provided, especially if it was computationally intense to derive.
+It should be in its own folder, and to maximize interoperability, recorded in a plain text (e.g., csv, tsv) and if possible, in a binary format (e.g., .rds, .sav, .xlsx, Apache Parquet).
 The binary format allows for immediate reuse in the intended state, while the interoperability of plain text versions remain stable over time.
-Ideally data should be in "Tidy Data" form, where variables are in columns, observations in rows, and each cell has only one value [@Wickham2014].
+Ideally, data should be in "tidy data" form, where variables are in columns, observations in rows, and each cell has only one value [@Wickham2014].
 
@@
 
-Share code used to transform data from its raw format into analysis ready form.
+Share code used to transform data from its raw format into analysis-ready form.
 This should be kept in the same place as raw data.
-Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc), but any non-scripted data preparation steps should be recorded in a plain text file.
-Authors should run the scripts in a clean environment to ensure everything runs and that the code doesn't rely on other local artifacts, such as their own folder layout not available to others.
+Ideally, this transformation should involve only scripted languages (e.g, R, Python, Matlab, etc.), but any unscripted data preparation steps should be recorded in a plain text file.
+Authors should run the scripts in a clean environment to ensure that everything runs and that the code doesn't rely on other local artifacts, such as their own folder layout not available to others.
 
 For more complex analyses that involve workflow management systems (see https://github.com/pditommaso/awesome-pipeline for a comprehensive list), a description of how to run the workflow should be included in the documentation, preferably in the README file.
-Depending on whether the workflow is GUI driven (e.g.
-Galaxy) or text driven (e.g.
+Depending on whether the workflow is GUI-driven (e.g.
+Galaxy) or text-driven (e.g.
 Nextflow, Snakemake), the workflow itself can be checked into the version control repository.
@@
 
-Data should be deposited into a relevant domain repository that generates an accession number such as a DOI.
-For example, gbif for species data, or genbank for genetics data.
-A exhaustive searchable list is available at re3data (https://www.re3data.org/) but provides no guidance to researchers.
-Data that do not fit appropriate domain repositories can be submitted to research data repositories **Zenodo** and **Dryad** which are currently leading the way in best practices for software and data.
+Data should be deposited into a relevant domain repository that generates an accession number such as a DOI (for example, gbif for species data, or genbank for genetics data).
+An exhaustive searchable list is available at re3data (https://www.re3data.org/) but provides no guidance to researchers.
+Data that do not fit appropriate domain repositories can be submitted to research data repositories **Zenodo** and **Dryad**, which are currently leading the way in best practices for software and data.
 In addition to minting DOIs, they also provide citation templates.
 
-If using a research institute's library service to share data, ensure a DOI is provided so it can be accurately identified and cited.
+If using a research institute's library service to share data, ensure that a DOI is provided so it can be accurately identified and cited.
 Do not deposit data in local file storage that does not provide DOI or accession numbers.
 
-Although it is fine to share data among collaborators using temporary cloud storage (like Dropbox, Google Drive), once the paper is public, the data needs to be hosted somewhere persistent.
-While this approach is useful while iterating on the paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc.).
-Providing back up data is particularly relevant in light of the cases where data is suspected of being tampered with [@Schleunes2020; @Viglione2020].
-The use of raw data allowed for issues to be identified, the absence of raw data not being available in other publications created difficulty in verifying the validity of the data.
+Although it is fine to share data among collaborators using temporary cloud storage (like Dropbox or Google Drive), once the paper is public, the data needs to be hosted somewhere persistent.
+Although a cloud-based storage approach is useful while iterating on a paper, it does not scale to a larger readership and is highly prone to disappearing (lack of permissions, etc.).
+Providing back-up data is particularly relevant in light of cases in which data is suspected of being tampered with [@Schleunes2020; @Viglione2020].
+The use of raw data allows for issues to be identified; the absence of raw data in publications creates difficulties in verifying the validity of the data.
 
@@
 This is a familiar avenue for researchers to receive credit and citation.
-Writing a data paper can be very useful for researchers, however, data papers lack guidelines on how to structure data for effective reuse and sharing.
+Writing a data paper can be very useful for researchers, but, data papers lack guidelines on how to structure data for effective reuse and sharing.
 Partner with Dryad to link archived data with research papers.
@@
 
-One critical advantage of depositing data in a trusted repository (Section \@ref(avenues)) is that it comes with many benefits such as preservation, indexing,  discoverability and most importantly, citation with persistent identifier.
-Authors can take an extra step by including their own CITATION file to give credit to a paper where the data and associated methods are described in detail.
-When citing data, you should cite the preferred CITATION file in the first instance, else, if there is a data publication, (for example in _Scientific Data_), cite this, and if none of those apply, cite the repository where the data is deposited (e.g., a Zenodo repository).
+One critical advantage of depositing data in a trusted repository (Section \@ref(avenues)) is that it comes with many benefits, such as preservation, indexing,  discoverability, and most importantly, citation with a persistent identifier.
+Authors can take an extra step by including their own CITATION file, in which the data and associated methods are described in detail.
+When citing data, you should cite the preferred CITATION file in the first instance, else, if there is a data publication, (for example in _Scientific Data_), cite this, and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository).
 
@@
 Data can be distributed alongside code using a format such as a package in R, Python, and Julia.
-A package or module structure provides an advantage as it can be "installed" and loaded just like code.
-Obvious benefits of this approach is that it allows data to be immediately used in a programming environment without the additional steps of downloading, preprocessing, specifying column types etc.
+A package or module structure provides an advantage, as it can be "installed" and loaded just like code.
+An obvious benefit of this approach is that it allows data to be immediately used in a programming environment without the additional steps of downloading, preprocessing, specifying column types, etc.
 But there are some limitations to this approach.
-Firstly, packaging data inside a software package makes the data availability language centric.
+First, packaging data inside a software package makes the data availability language-centric.
 Users who work in other languages are unlikely to download and export data out of a package.
-Secondly, data size can be a limitation depending on the repository (CRAN, the R repository, for example, limits package sizes to 5Mb although there are workarounds).
-Other fields, such as bioinformatics have a process of sharing data through repositories like Bioconductor, with larger data size.
-Thirdly, there is the cost of authors and data curators/custodians understanding the intricacies of package development in order to share data.
+Second, data size can be a limitation for some repositories (CRAN, the R repository, for example, limits package sizes to 5Mb, although there are workarounds).
+Other fields, such as bioinformatics, have a process of sharing data through repositories like Bioconductor, which can accommodate a larger data size.
+Third, there is the cost of authors and data curators/custodians understanding the intricacies of package development in order to share data.
 Nevertheless, the frictionless approach to sharing data in software packages makes reuse substantially faster than retrieving and correctly parsing files from a repository.
@@
 First, archive data in a long-term data repository (Section \@ref(avenues)), before including it in a package.
-The act of packaging data in software can shave off hours required to prepare it for use.
+The act of packaging data in software can shave hours off the time required to prepare it for use.
 The canonical citation can be baked into the package's citation file.
@@
 How data is shared can change depending on its size.
-Defining data size will always be somewhat ill-defined, since storage capacity and norms constantly change.
+Defining data size will always be somewhat difficult to do, since storage capacity and norms constantly change.
 We consider data sizes in reference to a moderately priced laptop, where small data fits on computer memory (RAM), medium data does not fit on RAM, and big data does not fit on the hard disk.
-No matter the size, the data should be archived with an accession number or DOI.
+No matter the size, data should be archived with an accession number or DOI.
 
-Small data has more options in how it is shared - you can bundle it with your analysis code (as in \@ref(rawdata)), or in software (as in \@ref(pkgs)).
-For medium data, a general purpose data repository (e.g.
-Zenodo) would be ideal for archiving.
-Sharing medium data might require it to be broken into separate chunks for reading (e.g., `teaching-1.csv`, `teaching-2.csv`, etc.), and scripts provided to read it.
-Large data will require liaising with services from \@ref(avenues), or local research institute library to make special arrangements, to ensure there is some way to gain access, and also maintain a DOI.
+Small data has more options for how it can be shared -- you can bundle it with your analysis code (as in \@ref(rawdata)), or in software (as in \@ref(pkgs)).
+For medium data, a general purpose data repository (e.g., Zenodo) would be ideal for archiving.
+Sharing medium data might require it to be broken into separate chunks for reading (e.g., `teaching-1.csv`, `teaching-2.csv`, etc.), and scripts might need to be provided to read it.
+Large data will require liaising with services from \@ref(avenues), or making special arrangements with a local research institute library, to ensure that there is some way to gain access, and to maintain a DOI.
 
@@
 Although data citations do not currently impact promotion and tenure packets, there is some evidence that papers sharing usable data accrue more citations [@Piwowar2007].
-Over time, the more useful we make the data, the more likely incentive mechanisms will take hold.
+Over time, the more useful we make the data, the more likely it becomes that incentive mechanisms will take hold.
 
 The steps leading up to requiring data sharing are still vague and disincentivized.
-In comparison, although few journals require that code be submitted with publications, it is more likely to be adopted given the foundation we have established with carrots rather than sticks.
+In comparison, although few journals require that code be submitted with publications, such a norm is more likely to be adopted if a foundation like the one we have described is established, with carrots rather than sticks.
 We have taught researchers how to write modular code via popular training initiatives such as The Carpentries.
@@
 
-To make data sharing truly useful, we need to make the process easier, improve tools and training, then reward and incentivize researchers who engage in such practices.
+To make data sharing truly useful, we need to make the process easier, improve tools and training, and reward and incentivize researchers who engage in such practices.
 Only then can shared data truly be useful.

### 2021-08-13 response-letter/response-letter.Rmd
Updated response letter

@@
 
-Dear Editor in chief of Patterns,
+Dear Dr. Callaghan,
 
@@
 
-We thank the reviewers for time and comments.
+We thank the reviewers for time and detailed comments.
 We have addressed comments from the reviewers and believe the paper is clearer and better as a result.
-Our response to reviewer comments are presented in the table below.
-The "ID" column value "R1-1" refers to reviewer 1 comment 1, and so on.
+Our response to individual reviewer comments are presented in the table on page 2 of this letter.
 
@@
 Nicholas Tierney & Karthik Ram
+
+\newpage
 

### 2021-08-17 response-letter/response-letter.Rmd
change paper title

@@
 
-Thank you for considering our commentary article “Incentives are the biggest barriers to widespread data sharing” for publication in Patterns.
+Thank you for considering our commentary article, "Common sense approaches to sharing tabular data alongside publication" (previously titled "Incentives are the biggest barriers to widespread data sharing"), for publication in Patterns.
 

### 2021-09-14 new-paper.Rmd
change paper title
address comment R1-2

@@
 ---
-title: Incentives are the biggest barriers to widespread data sharing
+title: "Common sense approaches to sharing tabular data alongside publication"
 author: Nicholas J. Tierney\* (1, 2, 3) & Karthik Ram\* (4)
@@
 Dryad will also provide professionals to curate your data, thereby improving usability.
+While writing a data paper can be very useful for researchers, journals that publish data papers lack guidelines on how to structure data for effective reuse and sharing.
 

### 2021-09-14 new-paper.Rmd
work on addressing R2-2

@@
 If we now require mandatory code sharing, it is quite likely that more of it will be useful than the data currently being shared.
+Researchers will need to be familiar with existing tools and approaches for sharing data.
+There might be potential for researchers to reinvent the wheel, for example by creating new platforms for data sharing, when services such as zenodo and Dryad already exist.
+So it is worth mentioning that researchers should not reinvent new tools without an exhaustive search.
 

### 2021-09-17 new-paper.Rmd
address R2-5
added bib and fixed R1-1 and R1-3

@@
 While there has been a growing trend of sharing data alongside publications, the status quo of data sharing in science is not good.
-Rowhani-Farid et al.
+In 2016, Rowhani-Farid et al.
 [@Rowhani-Farid2016] showed that 7 of 157 (4.5%) sampled papers in the *British Medical Journal* shared data, and only 14% of articles claiming open sharing of data actually shared their data.
-Similarly, Stodden et al.
+Similarly, in 2018, Stodden et al.
 [@Stodden2018] showed that only 36% of papers published in the journal *Science* shared data, despite the journal's data sharing policy.
@@
 Partner with Dryad to link archived data with research papers.
-Dryad will also provide professionals to curate your data, thereby improving usability.
+We recommend partnering with best practice data repositories that emphasize curation, like Dryad.
+This way the data will have a long-term archiving and will also have professional curation to ensure that the published data are usable and have enough metadata to be understood.
 While writing a data paper can be very useful for researchers, journals that publish data papers lack guidelines on how to structure data for effective reuse and sharing.
+Data journals should routinely be partnering with best practice repositories to best support the underlying data.
 
@@
 One critical advantage of depositing data in a trusted repository (Section \@ref(avenues)) is that it comes with many benefits, such as preservation, indexing,  discoverability, and most importantly, citation with a persistent identifier.
-Authors can take an extra step by including their own CITATION file, in which the data and associated methods are described in detail.
-When citing data, you should cite the preferred CITATION file in the first instance, else, if there is a data publication, (for example in _Scientific Data_), cite this, and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository).
+Authors should add a CITATION file when possible but the real burden lies with publishers to ensure that these citations are marked up appropriately to allow for the building of open data metrics [@lowenberg_2021, @lowenberg_daniella_2021b].
+
+When citing data, you should cite the preferred CITATION file in the first instance, else, a data publication, (for example in _Scientific Data, Ecology_), and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository).
 

### 2021-09-17 new-paper.Rmd
Fixed R1-3
Updated text
Fixed R2-3

@@
 This makes them particularly challenging to standardize or easily "install" where the code is running.
-There have been some efforts to improve this process, such as gogetdata, the Dat Project, and the formation of a private company, Quilt.
+There have been some efforts to improve this process, such as gogetdata, the Dat Project, Frictionless Data [@Fowler_Barratt_Walsh_2018], and the formation of a private company, Quilt.
 
@@
 
-Journals provide data papers in a formal publication explaining data that is archived.
+Journals provide data papers in a formal publication explaining data that is archived (e.g.
+Ecology, Scientific Data).
 This is a familiar avenue for researchers to receive credit and citation.
 Writing a data paper can be very useful for researchers, but, data papers lack guidelines on how to structure data for effective reuse and sharing.
-Partner with Dryad to link archived data with research papers.
 We recommend partnering with best practice data repositories that emphasize curation, like Dryad.

### 2021-09-18 new-paper.Rmd
Clarified text a bit further that journals dont provide guidelines, not data papers themselves

@@
 This is a familiar avenue for researchers to receive credit and citation.
-Writing a data paper can be very useful for researchers, but, data papers lack guidelines on how to structure data for effective reuse and sharing.
+Writing a data paper can be very useful for researchers, but, journals that publish data papers lack guidelines on how to structure data for effective reuse and sharing.
 We recommend partnering with best practice data repositories that emphasize curation, like Dryad.

### 2021-09-18 new-paper.Rmd
Fixed two typos

@@
 We recommend partnering with best practice data repositories that emphasize curation, like Dryad.
-This way the data will have a long-term archiving and will also have professional curation to ensure that the published data are usable and have enough metadata to be understood.
+This way the data will have long-term archiving and will also have professional curation to ensure that the published data are usable and have enough metadata to be understood.
 While writing a data paper can be very useful for researchers, journals that publish data papers lack guidelines on how to structure data for effective reuse and sharing.
@@
 
-When citing data, you should cite the preferred CITATION file in the first instance, else, a data publication, (for example in _Scientific Data, Ecology_), and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository).
+When citing data, you should cite the preferred CITATION file in the first instance, else, a data publication, (for example in _Scientific Data, Ecology_), and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository) in the references section of related papers.
 

### 2021-09-20 companion-files.Rmd
push companion files and make sure the paper compiles to word

---
title: "companion files"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# eTOC blurb

> The eTOC blurb is a short summary that describes the context and significance of the findings for the broader readership. Please see the "In Brief" links in the table of contents for examples. The blurb should be 80 words or fewer.

Despite a general appreciation of the benefits of data sharing, research data are often only available to the original investigators. We argue the lack of incentives and infrastructure for making data useful are the biggest barriers to widespread data sharing. We compare data sharing to code sharing, in the context of facilitating reproducible research. We provide a common sense guide for sharing tabular data alongside research. Our target audience is academics in data science adjacent fields who are about to submit for publication.

# Bigger Picture Statement {-}

> https://www.cell.com/patterns/authors

Without data, there is no science. Science needs to be reproducible so we can trust the results, and progress as a field. Although academia generally appreciates the benefits of data sharing, research data are usually only available to the original researchers. If data is shared, it often lacks the documentation to make it easy to re-use. There is not a strong culture of data sharing, and we believe this is due to a lack of incentives and infrastructure for making data useful. While past papers focus on general practices for sharing data and code, we focus on a specific audience: academics working in data science adjacent fields who are about to submit for publication. We provide immediately usable guidelines to share tabular data alongside research that a researcher could pick up and use today. In the future, we hope the culture around data sharing will change, and be rewarded in science. A new normal where data is submitted with every research publication, and datasets are easily discovered, shared, and extended in other analysis.

# Bio {-}

**Nicholas Tierney**

Nicholas Tierney has an honours degree in Psychological Science, and a PhD in Statistics. He is a research software engineer at the Telethon Kids Institute, creating scientific software for statistical modelling, focussing on infectious disease. Prior to this, Nick was a lecturer at Monash University, teaching introduction to data analysis. Nick is a keen outdoorsman, and loves hiking, trail running, rock climbing, photography, and making coffee.

**Karthik Ram**

Karthik is a senior data scientist at the Berkeley Institute of Data Science and a co-founder of the rOpenSci project. Karthik is also a senior PI at the Berkeley Initiative for Global Change Biology. Prior to joining Berkeley, he earned his PhD in Ecology &amp; Evolution from the University of California, Davis. Karthik's interests are focused on reproducible research, especially as it applies to global change. Much of his recent work focuses on building tools and services around open data and growing diverse data science communities.

### 2021-09-20 new-paper.Rmd
push companion files and make sure the paper compiles to word

--- 2021-09-17 new-paper.Rmd
@@
 output:
+  bookdown::word_document2:
   bookdown::pdf_document2:
@@
 This makes them particularly challenging to standardize or easily "install" where the code is running.
-There have been some efforts to improve this process, such as gogetdata, the Dat Project, Frictionless Data [@Fowler_Barratt_Walsh_2018], and the formation of a private company, Quilt.
+There have been some efforts to improve this process, such as gogetdata, the Dat Project, and the formation of a private company, Quilt.
 
@@
 
-Journals provide data papers in a formal publication explaining data that is archived (e.g.
-Ecology, Scientific Data).
+Journals provide data papers in a formal publication explaining data that is archived.
 This is a familiar avenue for researchers to receive credit and citation.
 Writing a data paper can be very useful for researchers, but, data papers lack guidelines on how to structure data for effective reuse and sharing.
-We recommend partnering with best practice data repositories that emphasize curation, like Dryad.
-This way the data will have a long-term archiving and will also have professional curation to ensure that the published data are usable and have enough metadata to be understood.
+Partner with Dryad to link archived data with research papers.
+Dryad will also provide professionals to curate your data, thereby improving usability.
 While writing a data paper can be very useful for researchers, journals that publish data papers lack guidelines on how to structure data for effective reuse and sharing.
-Data journals should routinely be partnering with best practice repositories to best support the underlying data.
 
@@
 One critical advantage of depositing data in a trusted repository (Section \@ref(avenues)) is that it comes with many benefits, such as preservation, indexing,  discoverability, and most importantly, citation with a persistent identifier.
-Authors should add a CITATION file when possible but the real burden lies with publishers to ensure that these citations are marked up appropriately to allow for the building of open data metrics [@lowenberg_2021, @lowenberg_daniella_2021b].
-
-When citing data, you should cite the preferred CITATION file in the first instance, else, a data publication, (for example in _Scientific Data, Ecology_), and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository).
+Authors can take an extra step by including their own CITATION file, in which the data and associated methods are described in detail.
+When citing data, you should cite the preferred CITATION file in the first instance, else, if there is a data publication, (for example in _Scientific Data_), cite this, and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository).
 

### 2021-09-20 response-letter/response-letter.Rmd
update response letter rmd

--- 2021-08-13 response-letter/response-letter.Rmd
@@
 
-```{r fix-long-url}
-
-new_comment_line <- "The authors need to further clarify how their recommendations align with - and differ from - the slew of other recommendations for data management which exist (in articles - e.g https://doi.org/10.1371/journal.pcbi.1003542, and at the level of institutional/regional data stewardship - e.g.
-from the Research Data Alliance, or from libraries: https://www.monash.edu/library/researchers /n/researchdata/guidelines/sharing.) I think the key to differentiating this work is again making the community/audience explicit from the very beginning"
-
-```
-
 ```{r read-sheet, include = FALSE}
 our_table_id <- gs4_get(
-  "1fIl8jy72XWD4liFZVlx6WvsnXp7ld8fs7eZvwVJpSy0"
+  "14-qpuM3DaeaeWt4D6E7dLuvkT3L4QTe2gDr3ObxveOI"
   )
@@
          response) %>% 
-  mutate(comment = 
-           case_when(
-             ID == "R1-2" ~ new_comment_line,
-             TRUE ~ comment
-           )) 
+  tidyr::drop_na()
 ```
@@
 
-Thank you for considering our commentary article “Incentives are the biggest barriers to widespread data sharing” for publication in Patterns.
+Thank you for accepting our commentary article, "Common sense approaches to sharing tabular data alongside publication", for publication in Patterns.
 
 We thank the reviewers for time and detailed comments.
-We have addressed comments from the reviewers and believe the paper is clearer and better as a result.
-Our response to individual reviewer comments are presented in the table on page 2 of this letter.
+We have addressed the final comments from the reviewers, our response to individual reviewer comments are presented in the table on page 2 of this letter.
 

### 2021-09-21 new-paper.Rmd
Fixed zenodo capitalization

@@
 Researchers will need to be familiar with existing tools and approaches for sharing data.
-There might be potential for researchers to reinvent the wheel, for example by creating new platforms for data sharing, when services such as zenodo and Dryad already exist.
+There might be potential for researchers to reinvent the wheel, for example by creating new platforms for data sharing, when services such as Zenodo and Dryad already exist.
 So it is worth mentioning that researchers should not reinvent new tools without an exhaustive search.

### 2021-09-21 arxiv/figures/fig-cartesian-plane.R
get rid of arxiv folder

@@ -1,39 +1,0 @@
-library(ggplot2)
-
-# There are two Axis:
-#
-#   X axis: <- Easy --- time consuming ->
-#   Y axis: <- fragile --- robust ->
-
-xy_data <- data.frame(x_line = -1:1,
-                      y_line = -1:1)
-
-xy_data
-#>   x_line y_line
-#> 1     -1     -1
-#> 2      0      0
-#> 3      1      1
-
-ggplot(xy_data) +
-  geom_hline(yintercept = 0) +
-  geom_vline(xintercept = 0) +
-  coord_cartesian(xlim = c(-1.2,1.2),
-                  ylim = c(-1.2,1.2)) +
-  annotate("text",
-           x = -0.1,
-           y = -1,
-           label = "Easy") +
-  annotate("text",
-           x = -0.4,
-           y = 1,
-           label = "Time Consuming") +
-  annotate("text",
-           x = -1,
-           y = 0.1,
-           label = "Fragile") +
-  annotate("text",
-           x = 1,
-           y = 0.1,
-           label = "Robust") +
-  theme_minimal() +
-  theme(aspect.ratio = 1)

### 2021-09-21 arxiv/figures/fig-robust-difficult-graph.R

@@ -1,39 +1,0 @@
-library(ggplot2)
-
-# There are two Axis:
-#   X axis: <- Robustness (brittle - robust)
-#   Y axis: <- Difficulty (easy - hard)
-
-library(tibble)
-
-dat_prep <- tibble(prep = c(15,20,25,95, 96),
-               understanding = c(15,20,95,95,96),
-                  label = c("Data Dump",
-                            "Most studies",
-                            "Sensor",
-                            "50 hectare",
-                            "RCT"
-                            ))
-
-dat_prep
-
-library(ggrepel)
-
-plot_effort_understanding <-
-ggplot(dat_prep,
-       aes(x = prep,
-           y = understanding,
-           label = label)) +
-  geom_point() +
-  geom_label_repel(
-    segment.size = 0.2,
-    segment.color = "grey50") +
-  theme_minimal() +
-  labs(x = "Effort to Prepare",
-       y = "Ease of Understanding") +
-  xlim(0,100) + 
-  ylim(0, 100)
-
-ggsave(filename = "../ddd/figures/fig-effort-understanding.png",
-       plot = plot_effort_understanding,
-       device = "png")

### 2021-09-21 arxiv/figures/table-schema-example.R

@@ -1,48 +1,0 @@
-example_table_schema <- data.frame(age = c(12, 21, 37),
-                                   height = c(161.5, 181.2, 178.3),
-                                   nationality = c("Australian",
-                                                   "American",
-                                                   "New Zealand"))
-
-knitr::kable(example_table_schema) %>%
-  clipr::write_clip()
-
-# pkg::pkg_install("ropenscilabs/dataspice")
-# install_github("ropenscilabs/dataspice")
-
-# is there a way to go from:
-# 
-# | age| height|nationality |
-# |---:|------:|:-----------|
-# |  12|  161.5|Australian  |
-# |  21|  181.2|American    |
-# |  37|  178.3|New Zealand |
-#   
-#   ```
-# {
-#   "name": "demographics",
-#   # here we list the data files in this dataset
-#   "resources": [
-#     {
-#       "path": "demographics.csv",
-#       "schema": {
-#         "fields": [
-#           {
-#             "name": "age",
-#             "type": "integer"
-#           },
-#           {
-#             "name": "height",
-#             "type": "number"
-#           },
-#           {
-#             "name": "nationality",
-#             "type": "string"
-#           }
-#           ]
-#       }
-#     }
-#     ]
-# }
-# ```
-

### 2021-09-21 response-letter/response-letter.Rmd

@@
 
+\renewcommand{\arraystretch}{2}
 ```{r show-table}
@@
   kbl(booktabs = TRUE,
-      longtable = TRUE) %>% 
+      longtable = TRUE,
+        linesep = "\\addlinespace") %>% 
   kable_styling(latex_options = c("striped",

### 2021-09-22 companion-files.Rmd
make into word doc, remove extra details on what the blurbs are

@@
 title: "companion files"
-output: html_document
+output:
+  word_document: default
+  html_document: default
 ---
@@
 
-> The eTOC blurb is a short summary that describes the context and significance of the findings for the broader readership.
-Please see the "In Brief" links in the table of contents for examples.
-The blurb should be 80 words or fewer.
-
 Despite a general appreciation of the benefits of data sharing, research data are often only available to the original investigators.
@@
 # Bigger Picture Statement {-}
-
-> https://www.cell.com/patterns/authors
 
